### 2021-01-06 _posts/welcome/welcome.Rmd
Initial commit

---
title: "Welcome to My Blog"
description: |
  Welcome to our new blog, My Blog. We hope you enjoy 
  reading what we have to say!
author:
  - name: Nora Jones 
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: 01-06-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

### 2021-01-06 _posts/2020-12-22-submissionsepidemiclaw/submissionsepidemiclaw.Rmd
add post submissions

---
title: "Similarity of public submissions to Austria's amendment of the epidemic law"
description: |
  An analysis of public submissions to bill seeking to amend Austria's epidemic law.

date: 12-22-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - Corona virus
  - OCR
  - stringr
  - web scraping

preview: submissionsepidemiclaw_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

<details closed>

<summary>Code: Load packages</summary>

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(extrafont)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
hrbrthemes::update_geom_font_defaults(
  family = "Roboto Condensed",
  size = 3.5,
  color = "grey50"
)
library(scales)
library(knitr)
library(paletteer)
library(ggtext)
library(glue)
library(pdftools)
library(rvest)
library(janitor)
library(patchwork)
library(svglite)
library(countrycode)
library(tictoc)
library(furrr)
library(gt)
library(reactable)
plan(multisession, workers = 3)

```

</details>

<details closed>

<summary>Code: Define rmarkdown options</summary>

```{r setup, echo = T}
# knit_hooks$set(wrap = function(before, options, envir) {
#   if (before) {
#     paste0("<", options$wrap, ">")
#   } else {
#     paste0("</", options$wrap, ">")
#   }
# })

knitr::opts_chunk$set(
 	fig.align = "left",
	message = FALSE,
	warning = FALSE,
 	dev = "svglite",
#	dev.args = list(type = "CairoPNG"),
	dpi = 300,
 	out.width = "100%"
)
options(width = 180, dplyr.width = 150)
```

</details>

<details closed>

<summary>Code: Define plot theme, party colors, caption</summary>

```{r echo=TRUE}

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()

theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      #panel.border = element_rect(colour = plot_bg_color, fill=NA),
      #plot.border = element_rect(colour = plot_bg_color, fill=NA),
      plot.margin = margin(l = 0, 
                           t = 0.25,
                           unit = "cm"),
      plot.title = element_markdown(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 13
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 11
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(
        angle = 0,
        color = "grey50",
        hjust = 1
      ),
      axis.text.x = element_text(
        size = 9,
        color = "grey50"
      ),
      axis.title.y = element_blank(),
      axis.text.y = element_text(
        size = 9,
        color = "grey50"
      ),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 7
      ),
      legend.text = element_text(
        size = 7,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

data_date <- format(Sys.Date(), "%d %b %Y")

```

</details>

# Context

Austria's government tabled about three months ago an amendment to the country's epidemic law. The amendment was largely a reaction to the shortcomings of the law dating from 1950 when confronted with the current Covid crisis as well as a recent ruling of Austria's Constitutional Court which declared several measures passed by the government as violating the constitution. Hence, the government proposed the amendment.

While there would be a lot to say about the latent tension between civic rights and a state's obligation to curtail an epidemic or the thin line between legitimate restrictions and excessive infringements, this post will deal with the public submissions to the government's amendment bill. In short and crude terms, the legislative process in Austria provides the opportunity for citizens, NGOs, expert bodies etc to file submissions in which they can raise their concerns as to the (draft) version of the bill before it will be debated in parliament. So, at least in theory, it's an avenue for the government to solicit input from the public at large.

What caught my eye, or better my ears, was a related radio news report which mentioned something along the lines that the submissions to the amendment reached a) a record number and b) that a considerable number of submissions were similar in their wording.

I guess the former is indicative for the fundamental issues which the amendment touches and the overall somewhat 'edgy', not to say polarized atmosphere when it comes to Corona. The latter point, though, the similarity of submissions' wording, puzzled me a bit. What the similarity of the wording effectively means is the use of some kind of template by those filing a submission. In this regard, the news report mentioned that there have been some pertaining calls on social media channels to oppose the bill, possibly also including the provision of a template text.

It was this context which made me curious about the extent of the matter. At least curious enough to have an 'empirical' look at it with R.

So I went to the a) parliament's website which provides a record of received submissions, b) extracted the weblinks to all those submissions of which the text is public, c) download the pertaining pdfs, d) extracted their content, and e) finally checked for their similarity. As for the last point, I can gladly report that I learned something here and my approach changed a bit as I moved along. Hence, the blog post somewhat mirrors this process.

```{r include=FALSE}
submission_pattern <- "Tür und Tor"
```

When it comes to similarity, my first attempt was to simply look at the the recurrence of a distinct phrase which I repeatedly noticed while randomly glancing through some of the submissions. The formulation is *`r submission_pattern`* which means something like '(to open) the (flood)gates'. I guess it's safe to assume that those submissions articulated severe misgivings about the amendment.

As I'll show below, while this approach is already somewhat informative since it reveals a pervasive use of a rather distinct phrase and hence points towards the use of text template, it is rather inductive and likely to suffer from omissions of other, possibly even more re-current phrases. Furthermore, simply looking at one distinct phrase is likely to be a rather 'fragile' indicator. Only a minor, effectively non-substantive modification of the wording would result in missing out on similar submissions.

To make a long story short, this issue introduced me to the world of quantitative text analysis as provided by the powerful `quanteda` package. Admittedly, I am only starting to scratch the surface here, but hopefully deep enough to legitimately include it in this post. So, enough of context and waffling. In media R(es).

# Get data

## Links to submission subpages

The list of submissions filed for the amendment is provided [here](https://www.parlament.gv.at/PAKT/VHG/XXVII/ME/ME_00055/index.shtml#tab-Stellungnahmen){target="_blank"} at the Parliament's website. To extract the links to the sub-pages which include the links to individual texts, I make us of the provided RSS feed. From there I extract the relevant elements and combine them to one dataframe containing the name of the person/institution filing a submission, the submission's date, and the link to the sub-page.

<details open>

<summary>Code: get links to submission sub-pages</summary>

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
link_rss_all_submissions <- "https://www.parlament.gv.at/PAKT/VHG/XXVII/ME/ME_00055/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP=XXVII&ITYP=ME&INR=55&SUCH=&listeId=142&FBEZ=FP_142"

data <- xml2::read_xml(link_rss_all_submissions)

#get link to subpages with link to submissions
df_submission_pages_link <- data %>% 
  xml2::xml_find_all("//link") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="link_single_submission_page") %>% 
  mutate(link_single_submission_page=str_squish(link_single_submission_page)) %>% 
  filter(id>2) #removes first two rows which don't include data on submissions

#get title
df_submission_pages_name <- data %>% 
  xml2::xml_find_all("//title") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="title") %>% 
  mutate(name=str_extract(title, regex("(?<=\\>).*(?=\\<)"))) %>% 
  filter(id>2) %>% 
  select(-title)

#get publication date
df_submission_pages_pub_date <- data %>% 
  xml2::xml_find_all("//pubDate") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="date") %>% 
  mutate(date=date %>% str_squish() %>% lubridate::dmy_hms(., tz="Europe/Vienna"))

#combine to one dataframe
df_submission <- bind_cols(
  df_submission_pages_name,
  df_submission_pages_pub_date,
  df_submission_pages_link,
  ) %>% 
  select(-contains("id"))
```

</details>

```{r eval=FALSE, include=FALSE}
readr::write_excel_csv2(df_submission, file=here::here("blog_data", "epidemic_law_submissions", "df_submission_rss.csv"))
```

```{r include=FALSE}
df_submission <- readr::read_csv2(file="submissionsepidemiclaw_files/df_submission_rss.csv")

```

As it turns out, there were `r nrow(df_submission) %>% scales::comma()` submissions to the amendment in total.

```{r echo=FALSE, paged.print=TRUE}
reactable(df_submission, 
          defaultPageSize = 4, 
          columns = list(
            link_single_submission_page = colDef(
              cell = function(value) {
                url <- value
                htmltools::tags$a(href = url, 
                                  target = "_blank", 
                                  value)
                },
              width = 150,
              html=TRUE
              )),
          filterable = TRUE,
          bordered=TRUE,
          compact = TRUE,
          style = list(fontSize = "10px"),
          theme = reactableTheme(
            backgroundColor = "#f0eff0",
                filterInputStyle = list(
                  color="green",
                  backgroundColor = "#f0eff0")))
```

Notice that there are several submissions which state 'Keine öffentliche Stellungnahme' (no public statement) as a name. Hence, we will not be able to analyse the text of all submissions, but only of those which are made public. How many are there? Let's see:

<details closed>

<summary>Code: identify public and non-public submissions</summary>

```{r message=FALSE, warning=FALSE}
df_submission <- df_submission %>% 
  mutate(public=case_when(str_detect(name, regex("Keine öffentliche Stellungnahme",
                                                 ignore_case = T)) ~ "not public",
                          TRUE ~ as.character("public")))

table(df_submission$public)
```

</details>

```{r echo=FALSE}
table(df_submission$public)
```

Out of `r nrow(df_submission) %>% scales::comma()` submissions, the text of `r df_submission %>% filter(public=="public") %>% nrow() %>% scales::comma()` is publicly available (no idea why some are published and others not, probably it's up to the submitter to decide). While the high number of non-public submissions is obviously quite a caveat for the further analysis, we still have the text of more than 50 % of all submissions to look into.

## Links to pdfs

Now let's use the links to the submissions' sub-pages and search in each sub-page for a link which leads to the actual pdf document of the submission. By using the browser's developer tools (F12 key), we can see the html code behind the page. Clicking on the link of one submission reveals that the link contains the word 'imfname'. I use this string to identify the links to the submissions' pdfs and neglect all other links and pdfs which are also on the site.

![](submissionsepidemiclaw_files/identify_link.gif)

To do so, below the function which is applied to all (!) links leading to the sub-pages with the individual submissions. It reads the html code, extracts the links (href), converts the results into a dataframe, and subsequently filters out the links which include 'imfname'. I then complement the extracted link with the missing part of the web address (www.parlament....). Since there are several submission (sub-pages) which state that the text is not public, and don't include a link to a pdf, I insert in these cases a 'missing' instead of the link. To accelerate the process I apply the function in parallel by using \`furrr::future\_map\`.

<details open>

<summary>Code: read html of sub-pages, extract pdf-link</summary>

```{r eval=FALSE, include=TRUE}
# function to identify pdf link
fn_get_doc_links <- function(x){

  link_doc <- x %>% 
  read_html() %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  enframe(name=NULL,
          value="link_part") %>% 
  filter(str_detect(link_part, "imfname")) %>% 
  mutate(link_doc=paste0("https://www.parlament.gv.at", link_part)) %>% 
        pull(link_doc)
  ifelse(length(link_doc)>0, link_doc, "missing")

}

#apply function to all links
df_submission <- df_submission %>% 
  mutate(link_doc=future_map_chr(link_single_submission_page, 
                          possibly(fn_get_doc_links,
                                   otherwise="missing"),
                            .progress = T))
```

</details>

```{r eval=FALSE, include=FALSE}
#check
df_submission %>% 
  filter(link_doc=="missing" & public=="public") %>% 
  nrow()
#0 = ok; only links to non-public submissions are missing;

#readr::write_excel_csv2(df_submission, path=here::here("data", "df_submission.csv"))
```

## Download submissions

Now with all pdf links available we can start to download them. To do so, I use the `walk2` function of the `purrr` package. Note that I also use `magrittr`'s `%$%`operator to expose/pipe into the the `walk2` function.

```{r include=FALSE}
# load saved links to pdf_files
df_submission <- readr::read_csv2(file="submissionsepidemiclaw_files/df_submission.csv")

```

<details open>

<summary>Code: download files</summary>

```{r eval=FALSE, include=TRUE}
#define pdf's file name and download destination
df_submission<- df_submission %>% 
  mutate(file_name=str_extract(link_doc, regex("\\d+.pdf"))) %>% 
  mutate(pdf_destination=glue::glue("{here::here('blog_data','epidemic_law_submissions','pdf_files')}/{file_name}") %>% 
           as.character() %>% 
           str_trim(., side=c("both")))

#download
df_submission %>% 
  filter(link_doc!="missing") %$% #exposes names
  walk2(
    link_doc,
    pdf_destination,
    download.file, 
    mode = "wb")
```

</details>

After the steps above we now have the all the available submission files at our disposal. Finally time to start working on the actual texts.

## Extract text from pdfs

When it comes to extracting text by means of OCR, `pdftools` by Jeroen Ooms is the package of choice. In preparation, we need to get the German language data. Extracting text from pdfs can be quite a time consuming matter. Going through the more than 3700 submissions initially took me more than 7 hours. Reducing the dpi value from 300 to 150 accelerated the procedure quite considerably, down to approx. 60 minutes, and I didn't notice any relevant decrease of the output's quality.

<details open>

<summary>Code: extract text from pdfs</summary>

```{r eval=FALSE, include=TRUE}
tesseract::tesseract_download(lang = "deu")
tesseract_info()

df_submission <- df_submission %>% 
   mutate(doc_text=map(pdf_destination, possibly(~pdftools::pdf_ocr_text(.,
                                                                      language = "deu",
                                                                      dpi=150),
                                              otherwise="missing")))
```

</details>

```{r include=FALSE}
df_submission <- readr::read_rds(file="submissionsepidemiclaw_files/df_submission_150.rds")

```

Let's clean up the imported text, i.e. collapse multi-page submissions into one single string text, remove leading and trailing white space etc.

<details open>

<summary>Code: clean text</summary>

```{r}
df_submission <- df_submission %>% 
  mutate(date=date %>% 
           str_squish() %>% 
           str_trim() %>% 
           lubridate::dmy_hms()) %>% 
  mutate(doc_text=map_chr(doc_text, paste, collapse=" ")) %>% #collapse multipage docs
  mutate(doc_text=str_squish(doc_text)) %>% 
  mutate(doc_text=stringr::str_trim(doc_text, side=c("both"))) %>% 
  mutate(name_first=str_extract(name, regex("(?<=,\\s)[:alpha:]*(?=($|[^:alpha:]|\\s))")), 
         .after="name") 
```

</details>

Let's pause for a moment and take stock: There are `r nrow(df_submission) %>% scales::comma()` submissions, out of which `r df_submission %>% filter(public=="public") %>% nrow() %>% scales::comma()` are public. From the latter we were able to extract `r df_submission %>% filter(doc_text!="missing") %>% nrow() %>% scales::comma()`. Hence, `r df_submission %>% filter(public=="public") %>% nrow() - df_submission %>% filter(doc_text!="missing") %>% nrow()` texts could not be retrieved.

# Analysis

## Identify submissions with key formulation

Now with submissions' texts cleaned etc, let's identify those which contain the formulation which appeared quite frequently when randomly browsing through some of the texts: *`r submission_pattern`*.

<details open>

<summary>Code: identify submission containing key phrase</summary>

```{r include=T, eval=F}
submission_pattern <- "Tür und Tor"
```

```{r}
df_submission <- df_submission %>% 
  mutate(pattern_indicator=stringr::str_detect(doc_text,
                                                regex(pattern=submission_pattern,
                                                      dotall=T,
                                                      ignore_case=T,
                                                      multiline=T))) %>%
  mutate(overall_indicator=case_when(
      public=="not public" ~ "not public",
      pattern_indicator==TRUE ~ str_trunc(submission_pattern, 15),
      TRUE ~ as.character("other"))) %>% 
  mutate(doc_length=nchar(doc_text))
```

</details>

<details closed>

<summary>Code: table with identified texts</summary>

```{r}
tb_text <- reactable(df_submission %>%
            filter(pattern_indicator==TRUE) %>% 
            select(name, date, doc_text),
          defaultPageSize = 4, 
          filterable = TRUE,
          style = list(fontSize = "10px"),
          columns=list(
            name=colDef(width=100),
            date=colDef(width=100),
            doc_text=colDef(minWidth=200)),
          bordered=TRUE,
          compact = TRUE,
          theme = reactableTheme(
            backgroundColor = "#f0eff0",
                filterInputStyle = list(
                  color="green",backgroundColor = "#f0eff0")
          )
          )

```

</details>

```{r echo=FALSE}
tb_text
```

## Frequency of key formulation

<details closed>

<summary>Code: table</summary>

```{r message=FALSE, warning=FALSE}
# table -------------------------------------------------------------------
df_tb_submission_1 <- df_submission %>% 
  group_by(overall_indicator) %>% 
  summarise(n_abs=n()) %>% 
  mutate(n_rel=n_abs/sum(n_abs)) %>% 
  arrange(desc(n_abs)) %>% 
  mutate(indicator_tbl=case_when(overall_indicator=="not public" ~ "Keine öffentliche Stellungnahmen",
                                 overall_indicator==str_trunc(submission_pattern, 15) ~ glue::glue("enthalten Formulierung '{submission_pattern}'") %>% as.character(),
                                 overall_indicator=="other" ~ "andere öffentliche Stellungnahmen"),
         .before=1) %>% 
  select(-overall_indicator)
```

```{r message=FALSE, warning=FALSE}
df_tb_submission_2 <- df_submission %>% 
  group_by(overall_indicator) %>% 
  filter(overall_indicator!="not public") %>% 
  summarise(n_abs=n()) %>% 
  mutate(n_rel=n_abs/sum(n_abs)) %>% 
  arrange(desc(n_abs)) %>% 
  mutate(indicator_tbl=case_when(overall_indicator=="not public" ~ "Keine öffentliche Stellungnahmen",
                                 overall_indicator==str_trunc(submission_pattern, 15) ~ glue::glue("enthalten Formulierung '{submission_pattern}'") %>% as.character(),
                                 overall_indicator=="other" ~ "andere öffentliche Stellungnahmen"),
         .before=1) %>% 
  select(-overall_indicator, -n_abs, n_rel_public=n_rel)
```

```{r message=FALSE, warning=FALSE}
df_tb_submission <- df_tb_submission_1 %>% 
  left_join(., df_tb_submission_2)
```

```{r message=FALSE, warning=FALSE}
tb_submission <- df_tb_submission %>%   
  gt() %>% 
  tab_header(
    title = md("**Stellungnahmen zur Novelle des Epidemiegesetzes**"),
    subtitle = md(glue::glue(md(" Von den {df_tb_submission %>% filter(!str_detect(indicator_tbl, 'Keine')) %>% summarise(n_abs=sum(n_abs)) %>% pull(n_abs) %>% scales::comma()} Stellungnahmen von denen der Text öffentlich ist enthalten rund {df_tb_submission %>% filter(str_detect(indicator_tbl, 'Formulierung')) %>% pull(n_rel_public) %>% scales::percent()} die Formulierung *'{submission_pattern}'*.")))) %>% 
  tab_options(heading.align = "left",
              row_group.font.weight = "bold",
              table.background.color=plot_bg_color) %>% 
   tab_style(
    style = list(
      cell_fill(color = "orange"),
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = vars(n_abs),
      rows = n_abs == 2204) 
    )%>% 
    #  tab_style(
    # style = list(
    #   cell_fill(color = "orange"),
    #   cell_text(weight = "bold")
    #   ),
    # locations = cells_body(
    #   columns = vars(n_rel),
    #   rows = n_rel > 0.32 & n_rel < .34) 
    # )%>% 
       tab_style(
        style = list(
          cell_fill(color = "orange"),
          cell_text(weight = "bold")
        ),
        locations = cells_body(
        columns = vars(n_abs, n_rel, n_rel_public),
        rows = str_detect(indicator_tbl, "Formulierung")) 
    )%>% 
  cols_label(indicator_tbl="Stellungnahme",
             n_abs="Anzahl",
             n_rel="Anteil alle St.",
             n_rel_public="Anteil öff. St.") %>% 
  fmt_missing(columns=vars(n_rel_public), missing_text = "--") %>% 
  fmt_number(columns=vars(n_abs), decimals=0) %>% 
  fmt_percent(columns = vars(n_rel, n_rel_public), decimals = 1) %>% 
  tab_source_note(source_note = md("data: www.parlament.gv.at, 'Ministerialentwurf betreffend Bundesgesetz, mit dem das Epidemiegesetz 1950, das Tuberkulosegesetz und das COVID-19-Maßnahmengesetz geändert werden.'<br>analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes"))

```

</details>

As the table below shows, in `r df_tb_submission %>% filter(str_detect(indicator_tbl, "Formulierung")) %>% pull(n_abs) %>% scales::comma()` submissions the formulation '*`r submission_pattern`*' appears. *This number amounts to `r df_tb_submission %>% filter(str_detect(indicator_tbl, "Formulierung")) %>% pull(n_rel) %>% scales::percent()` of all submissions, and to `r (df_tb_submission %>% filter(str_detect(indicator_tbl, "Formulierung")) %>% pull(n_abs)/(df_tb_submission %>% filter(str_detect(indicator_tbl, "Formulierung")) %>% pull(n_abs)+df_tb_submission %>% filter(str_detect(indicator_tbl, "andere")) %>% pull(n_abs))) %>% scales::percent()` of all submissions for which we can access the text.* I think that's quite remarkable in the sense that it is suggest for quite pervasive use of a submission template.

```{r echo=FALSE, message=FALSE, warning=FALSE}
tb_submission

```

## Distribution of length

Let's approach the similarity of submissions from another angle now: The length of the text. The graph below shows the distribution of submissions' length in terms of number of characters. One density curve shows the distribution for submissions which include the key phrase '`r str_trunc(submission_pattern, 20)`', the other one shows the distribution of those which don't include it (and for which the text is available).

<details closed>

<summary>Code: distribution of length</summary>

```{r message=FALSE, warning=FALSE}

# distribution of length --------------------------------------------------

my_caption <- glue::glue("data: www.parlament.gv.at; Stellungnahmen zum Ministerialentwurf betreffend Bundesgesetz, mit dem das Epidemiegesetz 1950, das Tuberkulosegesetz \nund das COVID-19-Maßnahmengesetz geändert werden.\nanalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes")

pl_length <- df_submission %>% 
  filter(public!="not public") %>%
  filter(doc_length<1000) %>% #remove outliers
  ggplot(., aes(y=pattern_indicator,
             x=doc_length,
             fill=pattern_indicator,
             color=pattern_indicator)
         )+
  labs(title=paste("BEGUTACHTUNGSVERFAHREN EPIDEMIEGESETZ-NOVELLE:<br>","Verteilung der Länge von Stellungnahmen mit u. ohne Formulierung ","'",str_trunc(submission_pattern, 20),"'"),
       subtitle="Zur besseren Sichtbarkeit wurden Stellungnahmen mit mehr als 1000 Zeichen ausgeschlossen.",
       x="Länge Stellungnahme in Zeichen",
       caption=my_caption)+
  ggridges::geom_density_ridges(scale=5)+
  scale_fill_manual(values=c("FALSE"="darkolivegreen4",
                                   "TRUE"="orange"),
                    labels=c("FALSE"=glue::glue("ohne '{str_trunc(submission_pattern, 20)}'"),
                             "TRUE"=glue::glue("mit '{str_trunc(submission_pattern, 20)}'")))+
  scale_color_manual(values=c("FALSE"="darkolivegreen4",
                             "TRUE"="orange"),
                     labels=c("FALSE"=glue::glue("ohne '{str_trunc(submission_pattern, 20)}'"),
                             "TRUE"=glue::glue("mit '{str_trunc(submission_pattern, 20)}'")))+
  scale_y_discrete(expand=expansion(mult=c(0, 0.1)))+
  theme_post()+
  theme(
    plot.title.position = "plot",
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.caption = element_text(hjust=0),
    legend.position = "top",
    legend.justification = "left",
   legend.direction = "horizontal",
    legend.title=element_blank(),
    axis.title.y = element_blank(),
    axis.text.y=element_blank())+
  guides(fill=guide_legend(reverse=T),
         color=guide_legend(reverse=T))

```

</details>

```{r echo=FALSE, dev='svg'}
pl_length
```

What the graph shows is that the length of submissions which contain the formulation '`r str_trunc(submission_pattern, 25)`' is concentrated at around 350 characters, while the length of other submissions is more dispersed. Hence, when it comes to the length of documents, those containing the distinct formulation are also pretty similar in term of their length. I take this as another indicator for the similarity of the submissions with our search phrase. As for the density curve of those not containing the search phrase, note that there is still a little 'bump' around 350. This could potentially reflect that we are missing out on some submissions which may originate from the same submission template, but do not include the exact wording of our search phrase.

## Submissions over time

I was also wondering whether there is any 'temporal' pattern as to the date of the submission. If the submissions were the result of a social media campaign, are the timings of their submissions clustered? This could be read as a reaction to e.g. a social media post providing a template and asking to submit it.

<details closed>

<summary>Code: submissions per day</summary>

```{r message=FALSE, warning=FALSE}
# bar graph ---------------------------------------------------------------

df_submission_pattern <- df_submission %>% 
  mutate(overall_indicator=fct_infreq(overall_indicator)) %>% 
  group_by(date, overall_indicator, .drop=F) %>% 
  summarise(n=n()) %>% 
  ungroup() %>% 
  mutate(date=as.Date(date))

levels(df_submission_pattern$overall_indicator)

label_submission_pattern <- as.character(glue::glue("Enthält '{str_trunc(submission_pattern,15)}'"))

colors_indicators=c("grey70", "orange", "darkolivegreen4")
names(colors_indicators) <- c("not public", str_trunc(submission_pattern, 15), "other")

pl_submission_pattern <- df_submission_pattern %>% 
  filter(date<=as.Date("2020-09-19")) %>% 
  ggplot()+
  labs(title=paste("BEGUTACHTUNGSVERFAHREN EPIDEMIEGESETZ-NOVELLE:<br>","Zeitpunkt der Einrechung der Stellungnahmen, mit u. ohne Formulierung"),
       subtitle="",
       caption=my_caption,
       x="Datum",
       y="Anzahl Einreichungen")+
  geom_bar(aes(x=date,
               y=n,
               fill=overall_indicator),
           position=position_dodge2(),
           color=NA,
           stat="identity",
           key_glyph = "dotplot")+
  scale_x_date(labels=scales::label_date_short(),
               date_breaks="1 days")+
  scale_y_continuous(expand=expansion(mult=c(0,0.1)),
                     labels = scales::label_comma())+
  scale_fill_manual(values=colors_indicators,
                    labels=c("not public"="keine öffentliche Stellungnahme",
                             submission_pattern=label_submission_pattern,
                                  "other"="andere öffentliche Stellungnahmen"))+
  theme_post()+
  theme(
    plot.title.position = "panel",
    plot.subtitle = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.caption = element_text(hjust=0),
    legend.position = "top",
    legend.box.margin = margin(l=0, unit="cm"),
    legend.margin = margin(l=0, unit="cm"),
    legend.key.size = unit(1, units="cm"),
    legend.justification = "left",
    legend.title=element_blank(),
    axis.title.x = element_blank()
  )+
  guides(fill=guide_legend(size=15))
```

</details>

```{r echo=FALSE, dev='svg'}
pl_submission_pattern
```

I think the bar chart above doesn't suggest any temporal clustering of submissions including '`r submission_pattern`'. But considering that the speed with which the amendment was introduced and the very short period to file submissions doesn't really allow for much clustering.

# Analysis with 'quanteda'

As already indicated in the introduction, the above analysis - while informative to some extent - has its limitations. Limiting the search to the presence/absence of the key phrase '`r submission_pattern`' is prone to miss out on other wordings which might be even more prevalent in some of the submissions. Furthermore, only marginal changes to the wording could result in the omission of submissions which are otherwise very similar.

To overcome these limitations and approach the issue of text similarity in a methodologically more rigors manner, let's now take up some of the features of the `quanteda` package. I never worked with the package before but from what I can tell it provides a very comprehensive and powerful set of tools for quantitative text analysis (for an overview see [here](https://quanteda.io/articles/pkgdown/comparison.html){target="_blank"}).

To be able to use the offered tools, we have to convert our collection of submission texts into a 'corpus'. Before doing so, I further clean the extracted texts so that we only have the actual text of the submission and not any header, footer, or meta-data. Then let's create a `corpus`.

<details open>

<summary>Code: Clean text, create corpus</summary>

```{r}
library(quanteda)
df_submission_quanteda <- df_submission %>% 
  filter(doc_text!="missing") %>% 
  select(file_name, doc_text) %>% 
  mutate(doc_text=str_remove(doc_text,regex("^.*Eingebracht am: \\d{2}\\.\\d{2}\\.\\d{4}\\s*"))) %>% 
  mutate(doc_text=str_remove(doc_text,regex("\\s?www.parlament.gv.at\\s*$"))) %>% 
  mutate(doc_text=str_remove(doc_text,regex("^.?Betr\\.?\\:?\\s?Änderung(en)? des Ep(i|e)demie Gesetzes\\.?"))) %>% 
  mutate(doc_text=doc_text %>% str_squish() %>% str_trim()) %>% 
  ungroup()

#create vector
vec_submission <- df_submission_quanteda %>% 
  deframe()

#create corpus
corp_submission <- corpus(vec_submission, docvars=data.frame(submission=names(vec_submission)))
```

</details>

The corpus is essentially a named vector with the names of the sumbission files the names of the vector's elements.

```{r echo=FALSE}
head(corp_submission)
```

## Collocations

With this as the basis, we can now identify all collocations e.g. which comprise `r stringi::stri_count_words(submission_pattern)` words as e.g. in '`r submission_pattern`'. The `textstat_collocation` function provides us with the frequencies for all (!) collocations of the requested length. For performance reasons I limited the results to those collocations which result in more than 500 hits. For a detailed description of the function see [here](https://quanteda.io/reference/textstat_collocations.html){target="_blank"}.

<details open>

<summary>Code: Get collocations </summary>

```{r}
#collocations

n_words <- str_count(submission_pattern, regex("\\S+"))
n_words

collocs <- textstat_collocations(
  corp_submission,
  method = "lambda",
  size = n_words,
  min_count = 500,
  smoothing = 0.5,
  tolower = T) %>% 
  arrange(desc(count))

df_collocs <- collocs %>% 
  as_tibble()

```

</details>

And as it turns out, our used inductively selected key phrase '`r submission_pattern`' is close to the most frequent one. The collocation '`r df_collocs %>% slice_head(n=1) %>% pull(collocation)`' appears `r df_collocs %>% slice_head(n=1) %>% pull(count) %>% scales::comma()`.[^1]

[^1]: Note that the frequency of `r submission_pattern` quanteda result is different than the frequency I detected earlier. I am not entirely sure why this is, but I assume it's due to different approaches. While our result above is the actual

```{r echo=FALSE}
reactable(df_collocs,
          bordered=TRUE,
          compact = TRUE,
          style = list(fontSize = "10px"),
          filterable = TRUE,
          theme = reactableTheme(
            backgroundColor = "#f0eff0",
                filterInputStyle = list(
                  color="green",
                  backgroundColor = "#f0eff0")))
```

## Cosine similarity

Quantitative text analysis offers also the concept of **cosine similarity** when it comes to evaluating the similarity of two or more documents. I won't dig into the details here, but want to refer you to the video below which provides in my view a clear and accessible overview.

<iframe width="560" height="315" src="https://www.youtube-nocookie.com/embed/5lvS8078ykA" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<br>

One of the convenient properties of the concept is that it allows us to compare documents of any lengths along a scale of 0 (not similar at all) to 1 (identical).

In order to control for different document length when it comes to check for similarity, we have to create a weighted document feature matrix which accounts for the relative frequency of words (see [here](https://quanteda.io/reference/dfm_weight.html?q=dfm%20_%20weight){target="_blank"}).

<details open>

<summary>Code: create document feature matrix (dfm) </summary>

```{r}
#create document-feature matrix

vec_submission_wo_pattern <- map_chr(vec_submission, ~str_remove_all(., submission_pattern))

corp_submission_wo_term <- corpus(vec_submission_wo_pattern, docvars=data.frame(submission=names(vec_submission_wo_pattern)))

dfmat_submission <- dfm(corp_submission_wo_term, 
                        remove = stopwords("de"),
                        stem = TRUE, 
                        remove_punct = TRUE)

dfmat_prop_submission <- dfm_weight(dfmat_submission,
                                    scheme="prop")
```

</details>

With this we can calculate the cosine similarity of texts.

<details open>

<summary>Code: create similarity matrix (dfm) </summary>

```{r}
#create similarity matrix
mat_similarity <- textstat_simil(dfmat_prop_submission, dfmat_prop_submission, 
                            margin = "documents", 
                            method = "cosine")

#convert to dataframe
df_similarity <- mat_similarity %>% 
  as.data.frame(., 
                upper=FALSE,
                diag=FALSE,
  ) %>% 
  mutate(sim_interval=cut(cosine, c(seq(0,1,.1)),
                          right=F,
                          include.lowest=T))
```

</details>

### Document pairwise comparison

What we obtain now is a comparison of all submission pairs (!) as to their cosine similarity. Overall, these are `r df_similarity %>% nrow() %>% scales::comma()` pairs (duplicates removed, i.e. document 1 vs document 2 = document 2 vs document 1). Seems like text analysis can quickly get a bit large. To keep the table manageable I only show the first thousand.

```{r echo=FALSE}
reactable(head(df_similarity, n = 10^3),
          columns=list(
            cosine=colDef(format=colFormat(digits=4),
                          align="left")),
          bordered=TRUE,
          compact = TRUE,
          style = list(fontSize = "10px"),
          filterable = TRUE,
          theme = reactableTheme(
            backgroundColor = "#f0eff0",
                filterInputStyle = list(
                  color="green",
                  backgroundColor = "#f0eff0")))

```

To be able to group these pairs according to their similarity, I cut their cosine similarity values into 10 intervals (`sim_interval`). We can now count the number of submission pairs per interval. The barplot below presents the relative result, i.e. the share of submission pairs  from the total number of pairs.

<details closed>

<summary>Code: Frequency of pairs per cosine interval </summary>

```{r}
df_similarity <- df_similarity %>% 
  left_join(., df_submission %>% 
              select(file_name, pattern_indicator),
            by=c("document1"="file_name")) %>% 
  rename(document1_indicator=pattern_indicator) %>% 
  left_join(., df_submission %>% 
              select(file_name, pattern_indicator),
            by=c("document2"="file_name")) %>% 
  rename(document2_indicator=pattern_indicator) %>% 
  mutate(indicator=case_when(document1_indicator==T & document2_indicator==T ~ "both submissions with key phrase",
                             document1_indicator==F & 
                               document2_indicator==F ~ "both submissions without key phrase",
                             TRUE ~ as.character("one submission without key phrase")))

vec_color_submissions <- c("#FFC125", "#104E8B", "grey50")
names(vec_color_submissions) <- c("both submissions with key phrase",
"one submission without key phrase",
"both submissions without key phrase")

#barplot
df_bar_sim <- df_similarity %>% 
  group_by(sim_interval, indicator) %>% 
  summarise(interval_n=n()) %>% 
  ungroup() %>% 
  mutate(interval_rel=interval_n/sum(interval_n)) %>% 
  ungroup() %>% 
  mutate(law="EpidemieG") %>% 
  mutate(indicator=fct_relevel(indicator, names(vec_color_submissions)))

pl_bar_sim <- df_bar_sim %>% 
  ggplot() +
  labs(
    title="PUBLIC CONSULTATION PROCEDURE ON EPIDEMIC LAW:<br>
    Distribution of all submission pairs' cosine similarity",
    subtitle="",
    y="% of all submission pairs",
    x="cosine similarity",
    caption=my_caption)+
  geom_bar(aes(x=sim_interval,
               y=interval_rel,
               fill=indicator),
           position=position_stack(),
           stat="identity")+
  geom_text(
    # data=. %>%
    #           filter(sim_interval %in% c("[0.7,0.8)", "[0.8,0.9)", "[0.9,1]")) %>%
    #                    filter(str_detect(indicator, "both submissions with key phrase")),
data=. %>%  mutate(interval_rel_label=case_when(
                                                indicator=="both submissions without key phrase" ~ NA_real_,
  interval_rel>0.05 ~ interval_rel,
                                                TRUE ~ NA_real_)),
    aes(x=sim_interval,
    y=interval_rel,
    label=interval_rel_label %>% scales::percent(., accuracy = 0.01),
    group=indicator),
position=position_stack(vjust=0.5, reverse = F),
color="white")+
  # stat_summary(fun.y = sum, 
  #              aes(x=sim_interval,
  #                  y=interval_rel+.1,
  #                  label = ..y.. %>% scales::percent(., accuracy = 0.01), 
  #                  group = sim_interval), 
  #              # position=position_stack(),
  #              # vjust=2,
  #              geom = "text")+
  geom_text(
    aes(x=sim_interval,
        y=interval_rel,
        label = ..y.. %>% scales::percent(., accuracy = 0.01),
        group = sim_interval), 
    stat = 'summary', 
    fun = sum, 
    vjust = -1
  )+
  scale_y_continuous(labels=scales::label_percent(accuracy = 1),
                     expand=expansion(mult=c(0, 0.1)))+
  scale_fill_manual(values=vec_color_submissions)+
  theme_post()+
  theme(axis.title.y = element_text(angle = 90,
                                    size=9,
                                    color="grey50",
                                    hjust=1),
        legend.direction = "horizontal",
        legend.position="top",
        plot.subtitle = element_blank(),
        plot.title=element_markdown(margin=unit(0.5, "cm")),
        legend.title=element_blank(),
        legend.text = element_text(size=9))

```

</details>

```{r echo=FALSE, dev='svg'}
pl_bar_sim
```
<br>
If we take 0.7 as a cut-off point, we can conclude that `r scales::percent(df_similarity %>% filter(cosine>=.7) %>% nrow()/nrow(df_similarity), accuracy=0.01)` of all submission pairs (!) were very similar. Note that among these submission pairs, there is no pair in which not at least one document contains our key phrase *'`r submission_pattern`'*. On the other end of the scale, we see that the largest junk of non-similar pairs (`r df_bar_sim %>% filter(sim_interval=="[0,0.1)" & str_detect(indicator, "one submission without")) %>% pull(interval_rel) %>% scales::percent(accuracy=0.01)` in the interval [0, 0.1)) consists of pairs of one document with the key phrase and the other document without it.

Note that I removed the key phrase before calculating the cosine similarity. Hence, the high degree of similarity of submission pairs containing the key phrase does not originate from the key phrase, but from the rest of the texts!

<details closed>
<summary>Code: Plot composition per cosine interval </summary>
```{r}
df_bar_sim %>% 
  group_by(sim_interval, indicator) %>% 
              summarise(sum_interval=sum(interval_rel, na.rm=T)) %>% 
              mutate(sum_indidicator_rel=sum_interval/sum(sum_interval))
```

Another way to approach the issue is to look into the composition of each interval. The plot below provides the pertaining overview. As it becomes quite clear, the the share of submission pairs with both documents containing our search phrase becomes higher the more similar the documents are. Again, let's note that the detected similarity does not originate from both documents containing '`r submission_pattern`' since I removed these words before calculating pairs' similarity. I take plot below as another indication of the use of a submission template.

```{r message=FALSE, warning=FALSE}
pl_bar_sim_fill <- df_bar_sim %>% 
  ggplot() +
  labs(
    title="PUBLIC CONSULTATION PROCEDURE ON EPIDEMIC LAW:<br>
    Composition of submission pairs' cosine similarity intervals",
    # subtitle="",
    y="% of all submission pairs",
    x="cosine similarity",
    caption=my_caption)+
  geom_bar(aes(x=sim_interval,
               y=interval_rel,
               fill=indicator),
           position="fill",
           stat="identity")+
  # geom_text(aes(x=sim_interval,
  #               y=interval_rel,
  #               group=indicator,
  #               label=interval_rel %>% scales::percent(., accuracy = 0.01)),
  #           position=position_fill(reverse=FALSE, vjust=0.5),
  #           color="white")+
  geom_text(data=. %>% 
              group_by(sim_interval, indicator) %>% 
              summarise(sum_interval=sum(interval_rel, na.rm=T)) %>% 
              mutate(sum_indidicator_rel=sum_interval/sum(sum_interval)) %>% 
              mutate(sum_indidicator_rel=case_when(sum_indidicator_rel<0.05 ~ NA_real_,
                                                   TRUE ~ as.numeric(sum_indidicator_rel))),
            aes(x=sim_interval,
                y=sum_indidicator_rel,
                group=indicator,
                label=sum_indidicator_rel %>% scales::percent(., accuracy=0.1)),
            color="white",
            position=position_fill(vjust=0.5))+
  geom_text(data=df_similarity %>% 
              group_by(sim_interval) %>% 
              summarise(n_obs=n()),
            aes(x=sim_interval,
                y=1.1,
                label=n_obs %>% scales::comma()),
            color="grey40",
            stat="identity")+
  geom_text(label="number of pairs:",
            x=.5,
            y=1.17,
            color="grey40",
            hjust=0,
            check_overlap = T)+
  scale_y_continuous(labels=scales::label_percent(accuracy=1),
                     expand=expansion(mult=c(0, 0.1)),
                     breaks=seq(0,1,.25))+
  scale_fill_manual(values=vec_color_submissions)+
  theme_post()+
  theme(axis.title.y = element_text(angle = 90,
                                    size=9,
                                    color="grey50",
                                    hjust=1),
        legend.position="top",
        legend.direction = "horizontal",
        legend.text=element_text(size=9),
        plot.subtitle = element_blank(),
        plot.title=element_markdown(margin=unit(0.5, "cm")),
        legend.title=element_blank())
```
</details>

```{r echo=FALSE, dev='svg'}
pl_bar_sim_fill
```

### Groupwise comparison

Finally, to further highlight the similarity of documents with `r submission_pattern` in contrast to other submission pairs, let's contrast the distribution of of the cosine similarity values per group with a violin plot.

<details closed>
<summary>Code: Groupwise comparison </summary>
```{r}
#create indicator for each document
df_submission <- df_submission %>%
  mutate(search_term=str_detect(doc_text, regex(submission_pattern,
                                                ignore_case = T,
                                                multiline = T,
                                                dotall = T)))

#add indicator to pairwise similarity comparison for first document
df_similarity <- df_similarity %>% 
  left_join(.,
            df_submission %>% 
              select(file_name, search_term),
            by=c("document1"="file_name")) %>% 
  rename(document1_term=search_term)

#add indicator to pairwise similarity comparison for first document
df_similarity <- df_similarity %>% 
  left_join(.,
            df_submission %>% 
              select(file_name, search_term),
            by=c("document2"="file_name")) %>% 
  rename(document2_term=search_term)

#create one, both, none groups
df_similarity <- df_similarity %>% 
  mutate(group_id=case_when(
    document1_term==TRUE & document2_term==TRUE ~ "both",
    document1_term==FALSE & document2_term==FALSE ~ "none",
    document1_term==TRUE & document2_term==FALSE ~ "one",
    document1_term==FALSE & document2_term==TRUE ~ "one",
    TRUE ~ as.character("missing")))
```

```{r}
pl_sim_groups <- df_similarity %>% 
  mutate(indicator=fct_relevel(indicator, names(vec_color_submissions))) %>% 
  ggplot()+
  labs(
    title="PUBLIC CONSULTATION PROCEDURE ON EPIDEMIC LAW:<br>Cosine similarity of submission pairs per group",
    subtitle=str_wrap(glue::glue("Grouped by presence/absence of key phrase '{submission_pattern}'. Key phrase was removed before calculating cosine similarity."),110),
    y="cosine similarity",
       caption=my_caption)+
  geom_violin(aes(x=indicator,
                   y=cosine,
                  fill=indicator),
              color=NA)+
  theme_post()+
  theme(legend.position = "none",
        axis.title.x = element_blank())+
  scale_fill_manual(values=vec_color_submissions)

```
</details>

```{r echo=FALSE, dev='svg'}
pl_sim_groups
```

I think the result shows quite clearly that those submissions which include the key phrase are much more similar to each other than all other groups of submission pairs even after having removed the key phrase. The graph also shows that those submission which do not contain the wording are hardly similar to each other.

## Comparision with other bills

But how does this result contrast with e.g. submissions to other bills. To put the results for the epidemic bill into context, I repeated the analysis from above for another bill, the 2018/19 Fundamental Law on Social Welfare ('[Sozialhilfegrundgesetz](https://www.parlament.gv.at/PAKT/VHG/XXVI/ME/ME_00104/index.shtml#tab-Stellungnahmen){target="_blank"}').[^2] The plot below contrasts the distribution of cosine similarities between submission pairs.

[^2]: I don't reproduce the code for the analysis here, but it is available on my github.

```{r echo=FALSE}
df_sozial_sim <- readr::read_csv2(file="submissionsepidemiclaw_files/df_bar_sim_sozial.csv") %>% 
  mutate(law="Sozialhilfegrundgesetz")

df_sim_comb <- bind_rows(df_sozial_sim, df_bar_sim) %>% 
  group_by(law, sim_interval) %>% 
  summarise(interval_rel=sum(interval_rel, na.rm=T))

pl_sim <- df_sim_comb %>% 
  ungroup() %>% 
  ggplot() +
  labs(title="PUBLIC CONSULTATION PROCEDURE:<br>Cosine similarity of submissions to 
        <span style=color:#FFC125>Epidemic Law</span> vs 
       <span style=color:firebrick '>Law on Social Welfare</span>",
       subtitle="Epidemic law = Epidemiegesetz; Law on Social Welfare = Sozialhilfegrundgesetz.",
    y="% of all submission pairs",
    x="cosine similarity",
    caption=glue::glue(str_wrap("data: www.parlament.gv.at; Stellungnahmen zum 'Ministerialentwurf betreffend Bundesgesetz, mit dem das Epidemiegesetz 1950, das Tuberkulosegesetz 
und das COVID-19-Maßnahmengesetz geändert werden' und 'Ministerialentwurf betreffend Bundesgesetze, mit dem ein Bundesgesetz betreffend Grundsätze für die Sozialhilfe (Sozialhilfe-Grundsatzgesetz) und ein Bundesgesetz über die bundesweite Gesamtstatistik über Leistungen der Sozialhilfe (Sozialhilfe-Statistikgesetz) erlassen werden.'", 140), "\n\nanalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes")) +
  geom_bar(aes(x=sim_interval,
               fill=law,
               y=interval_rel),
           position=position_dodge2(),
           stat="identity")+
  scale_y_continuous(labels=scales::label_percent(),
                     expand=expansion(mult=c(0, 0.1)))+
  # scale_fill_paletteer_d("ggsci::default_jama")[2:3]  +
  scale_fill_manual(values=c("#FFC125", "firebrick"))+
  theme_post()+
  theme(axis.title.y = element_text(angle = 90,
                                    size=9,
                                    color="grey50",
                                    hjust=1),
        plot.title.position = "panel",
        legend.position = "none",
        legend.direction = "horizontal",
        legend.title = element_blank())
```

```{r echo=FALSE, dev='svg'}
pl_sim
```
<br>
I think the result is quite telling. In contrast to  epidemic law, the values pertaining to the Law on Social Welfare are almost normally distributed and centered on the [0.3, 0.4) interval. In other words, there almost no submissions which are particularly similar or dissimilar. I would think that this is a distribution typical for submissions without the (prevalent) presence of a text template.

# Wrap up

So that's it (for now). Again, this got more lengthy than initially intended. A next step could be to see whether the similarity to submissions grew since the ability to file submissions via email. After all, copying a template, modifying it a bit before submission is not too difficult. Maybe as a final word, this is not to say that copy-and-paste submissions are of some sort of diminished democratic value. As always, if you have any comment or question regarding the post, feel free to contact me via DM on [twitter.](https://twitter.com/zoowalk){target="_blank"}.

### 2021-01-06 _posts/2020-12-19-vienna-elections-2020-analysis-of-parties-ads-on-facebook/welcome.Rmd
add post facebooks

---
title: "Vienna Elections 2020: Analysis of parties' ads on facebook"
description: |
    Analysis of the facebook ads placed in the run-up to the 2020 Vienna city elections.

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    # affiliation: home 
    # affiliation_url: https://orcid.org/0000-0002-6897-1693
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - social media
  - election
date: 12-19-2020
preview: welcome_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
---

# Setup

<details>

<summary>Code: Load packages</summary>

```{r setup, include=T, echo=T}
knitr::opts_chunk$set(echo = T)

# load packages -----------------------------------------------------------
library(tidyverse)
library(tictoc)
library(lubridate)
library(scales)
library(ggiraph)
library(tidytext)
library(ragg)
library(extrafont)
#extrafont::font_import()
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
hrbrthemes::update_geom_font_defaults(
  family = "Roboto Condensed",
  size = 3.5,
  color = "grey50"
)
library(ggtext)
library(gt)
library(reactable)
library(readxl)
library(htmltools)
library(paletteer)
library(lemon)
library(plotly)
options(scipen = 999)
library(kableExtra)
```

</details>

<details closed>

<summary>Code: themes, color codes </summary>

Do I see this text

```{r}
# DEFINITIONS -------------------------------------------------------------

# blog theme --------------------------------------------------------------

theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      #panel.border = element_rect(colour = plot_bg_color, fill=NA),
      #plot.border = element_rect(colour = plot_bg_color, fill=NA),
      plot.margin = margin(l = 0, 
                           t = 0.25,
                           unit = "cm"),
      plot.title = element_markdown(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 11
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 10
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(
        angle = 0,
        color = "grey50",
        hjust = 1
      ),
      axis.text.x = element_text(
        size = 9,
        color = "grey50"
      ),
      axis.title.y = element_blank(),
      axis.text.y = element_text(
        size = 9,
        color = "grey50"
      ),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 7
      ),
      legend.text = element_text(
        size = 7,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()

```

```{r}
# reactable heading -------------------------------------------------------

fn_reactable_headings <- function(title, subtitle, body, caption) {
  div(class="reactable-table",
      div(
        class="reactable-title",
        title
      ),
      div(
        class="reactable-subtitle",
        subtitle
      ),
      # div(
        #class="reactable-body",
        body
        # )
      ,
      div(
        class="reactable-caption",
        caption
      )
  )
}

rt_theme <- reactableTheme(backgroundColor = plot_bg_color,
                           style=list(fontFamily="Karla",
                                      fontSize="12px"),
                           filterInputStyle = list(
                             color="green",
                             backgroundColor = plot_bg_color))
```

```{r}
# caption -----------------------------------------------------------------

my_caption <- c("data: Facebook Ad Library API\nanalysis: Roland Schmidt | @zoowalk | http://werk.statt.codes")

labeller_page <- function(x) {str_extract(x, regex("\\w+$"))}
```

```{r}
# party colors -----------------------------------------------------

fpoe <- "#005DA8"
neos <- "#EA5290"
oevp <- "#5DC2CC"
spoe <- "#FC0204"
greens <- "#A3C630"
strache <- "White"
bier <- "yellow"

df_party_colors <- tibble(fpö=fpoe, neos=neos, övp=oevp, spö=spoe, greens=greens, strache=strache, bier=bier) %>% 
  pivot_longer(everything(), values_to = "party_colors", names_to = "party") %>% 
  mutate(party=str_to_upper(party))

vec_party_colors <- df_party_colors$party_colors
names(vec_party_colors) <- df_party_colors$party

```

and this one

</details>

<details closed>

<summary>Code: main candidates</summary>

```{r}
# main candidates ---------------------------------------------------------

main_candidates <- c("Blümel", "Michael Ludwig", "Strache", "Hebein", "Nepp","Wiederkehr")
vec_main_candiates <- str_c(main_candidates, collapse="|")

main_parties <- c("SPÖ", "ÖVP", "GREENS", "NEOS", "FPÖ", "STRACHE", "BIER")
vec_main_parties <- str_c(main_parties, collapse="|")

levels_party <- c("SPÖ", "ÖVP", "GREENS", "NEOS", "FPÖ", "STRACHE", 
                  "LINKS",
                  "BIER", 
                  "SÖZ", 
                  "VOLT", 
                  "PDA", 
                  "WANDEL",
                  "all others"
                  )
```

</details>

<details closed>

<summary>Code: time range</summary>

```{r}
# define time range -------------------------------------------------------

date_observation_start <- lubridate::ymd("2020-07-11")
date_observation_end <- lubridate::ymd("2020-10-11")
date_observation_start_format <- format.Date(date_observation_start, "%d %b %Y")
date_observation_end_format <- format.Date(date_observation_end, "%d %b %Y")

```

</details>

# Context

Here's a post which I probably should have written right before Vienna's local elections, and not after. But better late than never: a descriptive go at candidates' ads on Facebook in the run-up to the elections of 11 Oct 2020. For a previous blog post dealing with Facebook ads in the run-up to the European Parliamentary Elections 2019 see [here](https://werk.statt.codes/post/european-parliament-elections-2019-facebook-ads-of-austrian-candidates/){target="_blank"}.

To provide some general context, for a while now, Facebook has been offering access to details on electoral ads placed on the company's social networking site. The service was introduced as some kind of response to the allegations of manipulative and illegitimate interference in the 2016 US presidential elections or the Brexit referendum via its platform. By Facebook's own account, the library seeks to 'provide advertising transparency by offering a comprehensive, searchable collection of all ads currently running from across Facebook apps and services, including Instagram.' In addition to aggregated reports, the pertaining API allows to conduct 'customized keyword searches of active and inactive ads about social issues, elections or politics.'[^1] Hence, and notwithstanding some limitations, the library provides a critical peep-hole into electoral campaigning. In this case, it's the elections to the Vienna city council and its lower-level district councils.

[^1]: For details see [here](https://www.facebook.com/ads/library/api){target="_blank"}

# Scope conditions

A basic yet critical step in the analysis of Facebook ads is to delineate which Facebook pages and ads to include. How to identify those pages and ads which actually relate to the Vienna 2020 elections? Since Facebook's ad library does not group ads according to a specific campaign or cause this task is part of every analysis.

Here, I came up with the following main criteria which all have to be fulfilled.

-   *political parties*: I'll include only those ads which were placed by parties which eventually won a mandate to the city council. An exception is the party of HC Strache. The latter I include simply because Strache is the former Vice-Chancellor of Austria and former leader of the FPÖ.

-   *time*: ads have to be placed within the three months prior to the election date (11 July to 11 Oct).

-   *ads' content*: the text of the ad has to contain the word 'Wien' (Vienna). Admittedly, this is a somewhat crude condition, but it's simply the most reasonable indicator I came up with for my time available for this post. The condition is particularly important when considering the ads placed by the ÖVP's main candidate, Gernot Blümel. Blümel is also Austria's finance minister and hence his ads may not only relate to the Vienna electoral campaign, but also to issues related to his federal position. The latter category of ads has to be excluded.

-   *regional link*: the Facebook page should have a clear link to Vienna. E.g. ads placed by a party's branch from another state were not included.

-   *minimum number of ads*: to keep the number of pages somehow manageable, I included only pages which placed at least three ads within the period under consideration.

The above scope conditions might not be 100 % water tight, and there might be edge cases which erroneously fall out or into the set of ads eventually included. Overall though I would think to get with these criteria a sufficiently robust base to come to valid findings (and keep the effort for a blog post at a reasonable level). In any case it's important to highlight that all subsequent results of the analysis will be contingent on these scoping decisions. That's something to bear in mind.

If you're primarily interested in the results of the analysis and can't be bothered by all the R fuzz, just go directly [here](#number-and-value-of-ads).

# Getting the data

Note that there is by now also an R package which seeks to make data retrieval from Facebook's API easier ([Rfacebook](https://github.com/pablobarbera/Rfacebook){target="_blank"}). Since I had accessed Facebook's API already [previously](https://werk.statt.codes/post/european-parliament-elections-2019-facebook-ads-of-austrian-candidates/){target="_blank"} via the `httr` package, I stuck to this approach.

<details open>

<summary>Code: Get data from Facebook API</summary>

```{r eval=FALSE, include=TRUE}
#token from Facebook api app
your_fb_api_token <- "#########"

#define fields of interest
search_fields=c("ad_creation_time", "ad_delivery_start_time", "ad_delivery_stop_time",
                "ad_creative_body", 
                "page_id",
                "page_name",
                "currency",
                "spend",
                "demographic_distribution",
                "funding_entity",
                "impressions",
                "potential_reach",
                "publisher_platforms",
                "region_distribution") %>% 
  stringr::str_c(., collapse=", ")

my_link<- "https://graph.facebook.com"

#define GET request
page_one_response <- GET(my_link,
                         path = "/ads_archive",
                         query = list(access_token = my_token,
                                      limit=100,
                                      ad_active_status="ALL",
                                      search_terms="''",
                                      fields=search_fields,
                                      ad_reached_countries="AT"))

#get first page of results
page_one_content<- content(page_one_response)

df_ads_all <- tibble(data=page_one_content$data)

df_ads_all <- x %>% 
  unnest_wider(data)

#extract link for next results page
next_link <- page_one_content$paging$`next`

#initiate loop until there is no link to a next page
page_i=2
while(length(next_link)>0) {

  next_response <- GET(next_link)
  next_content<- content(next_response)

  y <- tibble(data=next_content$data)

  df_next <- y %>% 
    unnest_wider(data)

  df_ads_all <- bind_rows(df_ads_all, df_next)

  next_link <- next_content$paging$`next`

  page_i <- page_i+1
  print(page_i)

  print(df_next %>% 
    nrow())

}

#save results
df_ads_all$time_api_download <- Sys.time()
time_stamp <- format(Sys.time(), "%Y%m%d-%H%M")

write_rds(df_ads_all, path=paste0(wdr,"/data/scraped_ads",time_stamp,".rds"))
```

</details>

```{r include=FALSE}
df_ads <- readr::read_rds(file=here::here("blog_data", 
                                          "vienna_elections_2020_facebook",
                                          "df_ads.rds")) 
```

The code above provides us with all (!) ads placed in Austria. Overall, the result had more than 1 gig, and retrieving the data took a while. There are certainly ways to limit the search from the outset, but I thought it can't harm to have the entire set, also for possible future analysis.

## Apply scope conditions

Now let's apply the scope conditions from above. Applying the filter for ads' *time of creation* and whether ads *include the word 'Wien'* is straight forward:

```{r eval=T, include=T}
# SCOPE TIME ------------------------------------------------------------
date_observation_start <- lubridate::ymd("2020-07-11")
date_observation_end <- lubridate::ymd("2020-10-11")

df_ads <- df_ads %>% 
  filter(ad_creation_time>=date_observation_start & 
         ad_creation_time<=date_observation_end)

```

```{r include=FALSE}
# define time range -------------------------------------------------------

date_observation_start_format <- format.Date(date_observation_start, "%d %b %Y")
date_observation_end_format <- format.Date(date_observation_end, "%d %b %Y")
```

```{r eval=T, include=T}
df_ads <- df_ads %>% 
  filter(str_detect(ad_creative_body, "Wien(?!.*Neustadt)")) #to also exclude ads which refer to Wiener Neustadt

```

Based on this result I 'manually' checked every remaining facebook page whether a party affiliation and a regional link with Vienna exists. If interested in these 'coding decisions' you can see the table by unfolding the section below.

```{r include=FALSE}
pages_of_interest <- readxl::read_xlsx(path=here::here("blog_data", "vienna_elections_2020_facebook", "df_pages_of_interest.xlsx")) %>% 
  # filter(include=="yes") %>% 
  distinct(page_name, party, page_type, include) %>% 
  mutate(party=stringr::str_to_upper(party, locale="de"))
```

<details closed>

<summary> Coding of pages of interest </summary>

```{r echo=FALSE}
rt_pages <- reactable(pages_of_interest,
          columns=list(
            page_name=colDef(name="page")
          ),
          filterable = T,
          pagination = F,
          height = 250,
          outlined = T,
          bordered = F,
          compact=T,
          theme = rt_theme)

fn_reactable_headings(title="WIEN-WAHL 2020: Facebook pages within scope and affiliated party", subtitle="", body = rt_pages, caption=my_caption)

```

</details>

Let's consider only ads which were posted by those pages, add the party affiliations, limit the selection to those parties in which we are interested, and keep only those pages with at least three ads posted.

```{r message=FALSE, warning=FALSE}
### add party affiliations

df_ads <- df_ads %>%
  filter(page_name %in% pages_of_interest$page_name[pages_of_interest$include=="yes"]) %>% 
  left_join(., pages_of_interest %>% 
              select(page_name, party, page_type), 
            by=c("page_name"="page_name")) %>% 
  mutate(name_party=map_chr(page_name, labeller_page) %>% 
           paste0(., " (", party, ")")) %>% 
  mutate(name_party=case_when(page_type=="institutional" ~ page_name,
                              TRUE ~ as.character(name_party)))

### only main parties
main_parties <- c("SPÖ", "ÖVP", "GREENS", "NEOS", "FPÖ", "STRACHE")
vec_main_parties <- str_c(main_parties, collapse="|")

df_ads <- df_ads %>% 
  filter(str_detect(party, vec_main_parties))

### remove pages with less than 3 ads
df_pages_small_n <- df_ads %>% 
  group_by(page_name) %>% 
  summarise(n_ads=n()) %>% 
  filter(n_ads<3)

df_ads <- df_ads %>% 
  anti_join(., df_pages_small_n,
            by="page_name")
```

With our dataset now put into scope we have details on `r nrow(df_ads) %>% scales::comma()` ads by `r length(unique(df_ads$page_name))` distinct Facebook pages.

```{r message=FALSE, warning=FALSE}
### rearrange columns for better overview
df_ads <- df_ads %>% 
  select(starts_with("ad"), contains("page"), contains("party"),  contains("spend"), currency, funding_entity, contains("impression"), contains("demographic"), potential_reach, publisher_platforms)

glimpse(df_ads)

```

## Arrange data

Now with our dataset eventually put into scope, let's bring it into a format with which we can actually work.

Note that the data on ads' prices (`spend`), the audience's `demographic distribution` (gender and age) as well as ads' `impressions` are contained as lists. We will have to bring them into an other format for our subsequent analysis.

Let's first un-nest the data on spending. Note that Facebook's API does not provide a specific price of an ad, but instead a price category (interval) in which each ad falls. Hence when calculating i.e. the total amount spent by a candidate, we are bound to operate within possible minimum and maximum totals. As a means of simplification, I also use each price segment's mid point for further comparison.

<details open>

<summary>Code: data mugging, unnest lists </summary>

```{r}
# unnest spending --------------------------------------------------------

df_ads_wide <- df_ads %>% 
  unnest_wider(spend, names_sep = "_") %>% #adds "spend prefix"
  mutate(across(.cols=contains("spend"), .fns=as.numeric)) %>% 
  mutate(funding_entity=forcats::fct_infreq(funding_entity)) %>% 
  # mutate(spend_lower_bound_rev=case_when(spend_lower_bound==0 ~ 1, #otherwise aggregation distorting
  #                                  TRUE ~ spend_lower_bound)) %>% 
  mutate(spend_lower_bound_rev=spend_lower_bound) %>% 
  mutate(spend_mid=(spend_upper_bound-spend_lower_bound_rev)/2+spend_lower_bound) %>% 
  mutate(across(.cols=contains("bound"), .fns=list(char=~map_chr(.x, comma)))) %>% 
  unite("spend_interval", c(spend_lower_bound_rev_char, spend_upper_bound_char), sep="-", remove=F) %>% 
  mutate(spend_interval=forcats::fct_reorder(spend_interval, spend_lower_bound))

# # _nested df demographic distribution (age & gender)-----------------------------------
df_ads_wide <- df_ads_wide %>% 
  mutate(demographic_distribution_df=map(demographic_distribution, bind_rows))

# #_unnest impressions, create category, and mid-value --------------------
df_ads_wide <- df_ads_wide %>% 
  unnest_wider(impressions, names_sep="_") %>% 
  unite("impressions_category", c("impressions_lower_bound","impressions_upper_bound"), 
        sep="-",
        remove=F) %>% 
  mutate_at(vars("impressions_lower_bound", 
                 "impressions_upper_bound"),
            as.numeric) %>% 
  mutate(impressions_mid=(impressions_upper_bound-impressions_lower_bound)/2+impressions_lower_bound) %>% 
  mutate(impressions_category=as.factor(impressions_category) %>% 
           forcats::fct_reorder(., impressions_mid))

# _variable types ----------------------------------------------------------

df_ads_wide<- df_ads_wide %>% 
  mutate(across(.cols=contains("_time"), .fns=~ymd(.x))) %>% 
  mutate_at(vars(contains("time_api_download")), lubridate::ymd_hms)

# _create ad_id ----------------------------------------------------------
df_ads_wide <- df_ads_wide %>% 
  group_by(page_name) %>% 
  mutate(ad_id=paste(stringr::str_remove_all(page_name, "[:space:]"), 
                     lubridate::date(ad_creation_time),
                     row_number(), sep="-")) %>% 
  ungroup()

# _funding entity missing -------------------------------------------------

df_ads_wide <- df_ads_wide %>% 
  mutate(funding_entity=case_when(is.na(funding_entity) ~ "missing",
                                  TRUE ~ as.character(funding_entity)))

# rearranging
df_ads_wide <- df_ads_wide %>% 
  select(contains("page"), contains("party"), starts_with("ad"), contains("spend"), currency, funding_entity, contains("impression"), contains("demographic"), potential_reach, publisher_platforms)
```
</details>

```{r eval=FALSE, include=FALSE}
tb_all <- reactable(df_ads_wide,
          filterable = T,
          compact = T,
          outlined = T,
          bordered = F,
          wrap=F,
          theme=rt_theme)

fn_reactable_headings(title="WIEN-WAHL 2020: All Facebook ads",
                      subtitle=glue::glue("Ads posted between {date_observation_start_format} and {date_observation_end_format}"),
                      body = tb_all,
                      caption=my_caption)
```

We now have a dataframe with all ads of interest and most of the data rearrangement already implemented.

```{r eval=FALSE, include=FALSE}
df_impression_check <- df_ads_wide %>% 
  select(contains("page"), party, contains("impression"))

readr::write_csv2(df_impression_check,file=here::here("blog_data","vienna_elections_2020_facebook","df_impression_check.csv"))
```

```{r echo=FALSE}
glimpse(df_ads_wide)
```

# Number and value of ads {#number-and-value-of-ads}

Let's first look at the number and value of ads. The table below provides the aggregate number and amount spent on ads per party. These numbers comprise the expenditures by all Facebook pages included in the analysis. To drill further down and see the pertaining contribution of each page, click on the little arrow sign.

## Overview Table

<details closed>

<summary>Code: Table on parties' number of ads and total spending </summary>

The imho neat thing about this `reactable` table is that it comprises three tables with nested information: The aggregate values on the party level, the individual values for each Facebook page; and the funding entities for each page. To make this work, one has to subset the data of interest with the corresponding `[index]` value. See the code below:

```{r message=FALSE, warning=FALSE}
# Number and values of ads

df_party<- df_ads_wide %>% 
  group_by(party) %>% 
  summarise(ads_n=n(),
            across(.cols=contains("spend_") & where(is.numeric),
                   .fns=sum)) %>% 
  select(party, ads_n, spend_lower_bound_rev, spend_mid, spend_upper_bound) %>% 
  ungroup() %>% 
  arrange(-ads_n)

#nested pages df
df_pages <- df_ads_wide %>% 
  group_by(party, page_name) %>% 
  summarise(ads_n=n(),
            across(.cols=contains("spend_") & where(is.numeric),
                   .fns=sum)) %>% 
  select(party, page_name, ads_n, spend_lower_bound_rev, spend_mid, spend_upper_bound) %>% 
  ungroup() %>% 
  arrange(-ads_n)

#nested funding entities
df_funding <- df_ads_wide %>% 
  group_by(party, page_name, funding_entity) %>% 
  summarise(ads_n=n(),
            across(.cols=contains("spend_") & where(is.numeric),
                   .fns=sum)) %>% 
  select(party, page_name, funding_entity, ads_n, spend_lower_bound_rev, spend_mid, spend_upper_bound) %>% 
  ungroup() %>% 
  arrange(-ads_n)

#funding table
rt_page_funding <- reactable(df_pages,
          details=function(index) {
            funding_data <- df_funding[df_funding$page_name==df_pages$page_name[index],]
            htmltools::div(style = "padding: 16px",
                           reactable(funding_data, 
                                     columns=list(
                                       party=colDef(show=F),
                                       page_name=colDef(show=F)),
                                     sortable = T,
                                     filterable = T,
                                     outlined = TRUE))})

orange_pal <- function(x) rgb(colorRamp(c(plot_bg_color, "#ff9500"))(x), maxColorValue = 255)
blue_pal <- function(x) rgb(colorRamp(c(plot_bg_color, "#0C6D82"))(x), maxColorValue = 255)

#assemble table
rt_party_pages <- reactable(df_party,
                            columnGroups = list(
                              colGroup(name="spending", 
                                        columns=c("spend_lower_bound_rev", 
                                                  "spend_mid",
                                                  "spend_upper_bound"))
                            ),
                            columns=list(
                              party=colDef(name="party"),
                              ads_n=colDef(name="number of ads",
                                           format=colFormat(separators = T),
                                           style=function(value){ 
                                             normalized <- (value-min(df_party$ads_n))/(max(df_party$ads_n)-min(df_party$ads_n))
                                             color <- orange_pal(normalized)
                                             list(background=color)
                                             }),
                              spend_lower_bound_rev=colDef(name="min",
                                                           format=colFormat(separators = T,
                                                                            digits=0)),
                              spend_mid=colDef(name="mid",
                                               format=colFormat(separators = T,
                                                                digits=0),
                                               style=function(value){ 
                                             normalized <- (value-min(df_party$spend_mid))/(max(df_party$spend_mid)-min(df_party$spend_mid))
                                             color <-blue_pal(normalized)
                                             list(background=color)
                                             }),
                              spend_upper_bound=colDef(name="max",
                                                       format=colFormat(separators = T,
                                                                        digits=0))),
                            sortable = T,
                            filterable = F,
                            outlined = TRUE,
                            theme=reactableTheme(
                              backgroundColor = plot_bg_color,
                              filterInputStyle = list(
                                color="green",
                                backgroundColor = plot_bg_color)
                            ),

                            details = function(index) {
            page_data <- df_pages[df_pages$party == df_party$party[index], ]
            htmltools::div(style = "padding: 16px;",
                 reactable(page_data,
                           columns=list(
                             party=colDef(show=F),
                             page_name=colDef(name="fb page"),
                             ads_n=colDef(name="number of ads",
                                          format=colFormat(separators = T)),
                             spend_lower_bound_rev=colDef(name="min",
                                                          format=colFormat(separators = T)),
                             spend_mid=colDef(name="mid",
                                              format=colFormat(separators = T)),
                             spend_upper_bound=colDef(name="max", 
                                                      format=colFormat(separators = T))),
                           theme=reactableTheme(backgroundColor = plot_bg_color,
                                                filterInputStyle = list(
                                                  color="green",
                                                  backgroundColor = plot_bg_color) ),
                           sortable = T,
                           style="margin-left:60px;",
                           fullWidth = T,
                           filterable = T,
                           borderless = T,
                           outlined = TRUE,
                           details = function(index) {
                             funding_data <- df_funding[df_funding$page_name==page_data$page_name[index]&
                                                          df_funding$party == page_data$party[index],]
                             htmltools::div(style = "padding: 2px;",
                                            reactable(funding_data,
                                                      columns=list(
                                                        party=colDef(show=F),
                                                        page_name=colDef(show=F),
                                                        funding_entity=colDef(name="funding entity"),
                                                        ads_n=colDef(name="number of ads"),
                                                        spend_lower_bound_rev=colDef(name="min"),
                                                        spend_mid=colDef(name="mid"),
                                                        spend_upper_bound=colDef(name="max")),
                                                      theme=reactableTheme(backgroundColor = plot_bg_color),
                                                      sortable = T,
                                                      borderless = T,
                                                      style="margin-left:60px",
                                                      filterable = F,
                                                      outlined = TRUE))}
                           )
            )
            }
            )
```

</details>

```{r echo=FALSE}
fn_reactable_headings(title="WIEN-WAHL 2020: Number and total amount spent by party",
                      subtitle=glue::glue("Click to see details for indiviudal Facebook pages. Ads placed between {date_observation_start_format} and {date_observation_end_format}."),
                      body=rt_party_pages,
                      caption=my_caption)

```

<br> As it becomes quite clear, the `r df_party %>% slice_max(ads_n) %>% pull(party)` has bought the largest number of ads with `r df_party %>% slice_max(ads_n) %>% pull(ads_n) %>% scales::comma()`, followed by the `r df_party %>% slice_max(ads_n, n=2) %>% slice_tail(n=1) %>% pull(party)` with `r df_party %>% slice_max(ads_n, n=2) %>% slice_tail(n=1) %>% pull(ads_n) %>% scales::comma()` ads. The gap is quite considerable. If we unfold the `r df_party %>% slice_max(ads_n) %>% pull(party)` row, we can see that the largest chunk of its ads was bought for `r df_pages %>% filter(party==df_party %>% slice_max(ads_n) %>% pull(party)) %>% slice_max(ads_n) %>%  pull(page_name)` with `r df_pages %>% filter(party==df_party %>% slice_max(ads_n) %>% pull(party)) %>% slice_max(ads_n) %>%  pull(ads_n) %>% scales::comma()` ads.

When it comes to the total sum spent on Facebook ads, the picture is less specific. As already mentioned above, Facebook's ad library API provides only each ad's *price category with its lower and upper bounds*. Hence, when seeking to obtain a total sum of expenditures, we are limited to sum up the lower and upper bounds within which the overall total is located. This limitation is particularly consequential when a party bought many ads from the lowest price category which includes the value 0 (0-99). As we will see later, this is particularly the case with the ÖVP.

Bearing this caveat in mind and as an 'educated guess', let's focus on the aggregated mid-values of each ad. Again, the `r df_party %>% slice_max(spend_mid) %>% pull(party)` comes out on top with a total of `r df_party %>% slice_max(spend_mid) %>% pull(spend_mid) %>% scales::comma(, prefix="€")` spent. This number exceeds quite considerably the mid-values obtained for the `r df_party %>% slice_max(spend_mid, n=3) %>% slice(2) %>% pull(party)` (`r df_party %>% slice_max(spend_mid, n=3) %>% slice(2) %>% pull(spend_mid ) %>% scales::comma(, prefix="€")`) and the `r df_party %>% slice_max(spend_mid, n=3) %>% slice(3) %>% pull(party)` (`r df_party %>% slice_max(spend_mid, n=3) %>% slice(3) %>% pull(spend_mid) %>% scales::comma(, prefix="€")`).

## Number of ads

For an overview let's plot the pertaining numbers for the top 5 pages of each party (in terms of number of ads). Hover with the mouse over the bars to get details.

<details closed>

<summary>Code: Graph on parties' number of ads </summary>

```{r message=FALSE, warning=FALSE}
df_pl_pages <- df_ads_wide %>% 
  group_by(party) %>% 
  mutate(page_name_lum=fct_lump_n(page_name, n=5,
                                  other_level="all other")) %>% 
  group_by(party, page_name_lum) %>% 
  summarise(ads_n=n(),
            across(.cols=contains("spend_") & where(is.numeric),
                   .fns=sum)) %>% 
  select(party, page_name_lum, ads_n, spend_lower_bound_rev, spend_mid, spend_upper_bound) %>%   ungroup() %>% 
  arrange(-ads_n)

pl_pages <- df_pl_pages %>% 
  mutate(party=fct_relevel(party, levels_party)) %>% 
  ggplot()+
  labs(title="WIEN-WAHL 2020: Number of Facebook ads",
       subtitle=str_wrap(glue::glue("Top 5 Facebook pages per party in terms of number of ads and one 'lump' category for all other pages. Ads placed between {date_observation_start_format} and {date_observation_end_format}."), 100),
       caption=my_caption)+
  geom_bar_interactive(aes(x=ads_n,
               y=tidytext::reorder_within(x=page_name_lum,
                                          by = ads_n,
                                          within = party),
               tooltip=paste("Number of ads:", ads_n %>% scales::comma(., 
                                                                       accuracy=1)),
               fill=party),
           stat="identity")+
  scale_y_discrete(labels=function(x) str_extract(x, regex(".*(?=___)")))+
  scale_x_continuous(expand=expansion(mult=c(0.01, 0.05)),
                     breaks=seq(0, 1200, 300),
                     limits=c(0, 1250),
                     labels=scales::label_comma())+
  scale_fill_manual(values=vec_party_colors)+
  lemon::facet_rep_wrap(vars(party),
                        ncol=2,
                        repeat.tick.labels = T,
                        scales = "free_y")+
  theme_post()+
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color="grey50", size = 0.1),
        panel.spacing.x=unit(0, units="cm"),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7),
        axis.title.x = element_blank())
```

</details>

```{r echo=FALSE, fig.height=4, dev='svg', preview=T}
girafe(ggobj=pl_pages,
                             options=list(
                               opts_hover_inv(css = "opacity:0.1;"),
                               opts_hover(css = "fill:#FFA500;color:#FFA500"),
                               opts_tooltip(css = "background-color:#323E4F;
                               color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000)
                               ),
       height_svg=4)
```

A few details stood out for me: The ÖVP and Blümel's lead in terms of number of ads is quite remarkable. With `r df_pl_pages %>% filter(str_detect(page_name_lum, "Blümel")) %>% pull(ads_n) %>% scales::comma()` ads, Gernot Blümel's tally exceeds the one of Vienna's (previous and new) major Michael Ludwig with `r df_pl_pages %>% filter(str_detect(page_name_lum, "Ludwig")) %>% pull(ads_n) %>% scales::comma()` ads many times over. But even if we would put Blümel's record aside, and only focus on the number of ads bought by his party and colleagues, the ÖVP's strong record remains fully in order. The page of the ÖVP Vienna placed with `r df_pl_pages %>% filter(str_detect(page_name_lum, "Neue Volkspartei Wien")) %>% pull(ads_n) %>% scales::comma()` ads more than the pages of the SPÖ Wien, Michael Ludwig, and the SPÖ Inner Stadt (1st district) together (`r df_pl_pages %>% filter(str_detect(page_name_lum, "SPÖ Wien|Michael Ludwig|SPÖ Innere Stadt")) %>% pull(ads_n) %>% sum(.) %>% scales::comma()`). While I expected the ÖVP to be ahead, I didn't see the gulf to be so wide.

As for the Greens, what stood out to me was the small number of ads placed by the party's main candidate, Birigt Hebein. With only `r df_pl_pages %>% filter(str_detect(page_name_lum, "Hebein")) %>% pull(ads_n) %>% sum(.) %>% scales::comma(., accuracy=1)` ads, her campaign - at least on Facebook's platforms - left the main floor to the party's institutional pages, first and foremost 'Die Grüne Wien' with `r df_pl_pages %>% filter(str_detect(page_name_lum, "Die Grünen Wien")) %>% pull(ads_n) %>% sum(.) %>% scales::comma(., accuracy=1)` ads, but also to the pages of the district branches Hietzing and Penzing with `r df_pl_pages %>% filter(str_detect(page_name_lum, "Die Grünen Hietzing|Grüne Penzing")) %>% pull(ads_n) %>% sum(.) %>% scales::comma(., accuracy=1)` ads in total. (Tellingly, Hebein was effectively removed from the party leadership by her colleagues in the wake of the elections, despite the party's all-time best result).

Finally, what also struck me was the very low number of ads placed by Team Strache. With as little as `r df_pl_pages %>% filter(str_detect(party, regex("Strache", ignore_case=T))) %>% pull(ads_n) %>% sum(.) %>% scales::comma(., accuracy=1)` ads in total, the party had a very limited campaign on Facebook. Even when considering that the party was newly founded, for the former leader of the FPÖ who was renown for its strong social media presence, this was all in all quite a dramatic decline (even if the number of ads and followers are obviously not the same).

## Spending

Let's now look at the money spent on ads. As already mentioned above, the analysis is somewhat complicated by the fact that Facebook does not provide the specific price of an ad, but only the price category of an ad, with its lower and upper bounds. In the graph above, the point indicates the mid-value of the interval, the line's ends the lower and upper bounds.

<details closed>

<summary>Code: Graph on expenditures </summary>

```{r message=FALSE, warning=FALSE}
df_pl_pages_spend <- df_ads_wide %>% 
  group_by(party, page_name) %>% 
  summarise(ads_n=n(),
            across(.cols=contains("spend_") & where(is.numeric),
                   .fns=sum)) %>% 
  ungroup() %>% 
  select(party, page_name, ads_n, spend_lower_bound_rev, spend_mid, spend_upper_bound) %>%   
  group_by(party) %>% 
  mutate(page_name_lum=fct_lump_n(page_name, n=5, 
                                  w=spend_mid,  #weight frequencies by value
                                  other_level="all other")) %>% 
  group_by(party, page_name_lum) %>% 
  summarise(across(.cols=contains("spend_") & where(is.numeric),
                   .fns=sum)) %>% 
  ungroup()

pl_pages_spend <-  df_pl_pages_spend %>% 
  mutate(party=fct_relevel(party, levels_party)) %>% 
  ggplot()+
  labs(title="WIEN-WAHL 2020: Amount spent of Facebook ads",
       subtitle=str_wrap(glue::glue("Top 5 Facebook pages per party. Ads placed between {date_observation_start_format} and {date_observation_end_format}. Data only for ad's min/max price category available."), 100),
       caption=my_caption)+
  geom_linerange_interactive(aes(xmin=spend_lower_bound_rev,
                    xmax=spend_upper_bound,
                    y=reorder_within(x=page_name_lum,
                                     by = spend_mid,
                                     within=party),
                    color=party,
                    tooltip=paste("min:", spend_lower_bound_rev %>% scales::comma(), "\n",
                                  "max:", spend_upper_bound %>% scales::comma())))+
  geom_point_interactive(aes(x=spend_mid,
                 y=reorder_within(x=page_name_lum,
                                     by = spend_mid,
                                     within=party),
                 color=party,
                 tooltip=paste("mid:", spend_mid %>% scales::comma())
                 ))+
  scale_y_discrete(labels=function(x) str_extract(x, regex(".*(?=___)")))+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)),
                     breaks=seq(0, 150000, 50000),
                     minor_breaks = NULL,
                     labels=scales::number_format(scale=.001,
                                                  prefix="€",
                                                  suffix="k"),
                     limits=c(0, 150000))+
  scale_color_manual(values=vec_party_colors)+
  lemon::facet_rep_wrap(vars(party),
                        ncol=2,
                        repeat.tick.labels = T,
                        scales = "free_y")+
  theme_post()+
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color="grey50", size = 0.1),
        panel.spacing.x=unit(0, units="cm"),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7),
        axis.title.x = element_blank())

```

</details>

```{r echo=FALSE, fig.height=4, dev='svg'}
girafe(ggobj=pl_pages_spend,
       options=list(
         opts_hover_inv(css = "opacity:0.1;"),
         opts_hover(css = "fill:#FFA500;color:#FFA500"),
         opts_tooltip(css = "background-color:#323E4F;
         color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000)
                               ),
       height_svg=4)
```

With a few small differences, the general result on spending is congruent with the preceding analysis on the number of ads. The ÖVP and Blümel come out on top. The institutional page of the Greens in Vienna also spent considerably. Note that the top five pages in terms of money spent on ads are not always the top five pages in terms of number of ads bought.

# Impressions

Eventually though, the number and amount spent on ads doesn't tell us anything on how often an ad has been seen. Facebook's ad library provides data on each ad's 'impressions'. An impression is defined as the number of times an ad entered a person's screen.[^2] However, note that these figures do not tell us anything on how many unique individual have actually seen an ad. Furthermore, Facebook only provides interval data for an ad's impression (lower/upper bound). Again, I use the mid-value of these interval as an 'educated guess' and aggregate lower and upper bounds for totals. Hover with the mouse cursor over the graph to get details.

[^2]: See [here.](https://www.facebook.com/business/help/675615482516035){target="_blank"}

<details closed>

<summary>Code: Impression of ads </summary>

```{r message=FALSE, warning=FALSE}
df_ads_impressions <- df_ads_wide %>% 
  select(party, page_name, contains("impressions")) %>% 
  group_by(party, page_name) %>% 
  summarize(across(.cols=(contains("impressions") & where(is.numeric)), .fns=~sum(., na.rm = T))) %>% 
  group_by(party) %>% 
  mutate(page_name_lum=fct_lump_n(page_name, n=5, 
                                  w=impressions_mid,
                                  other_level = "all others")) %>% 
  group_by(party, page_name_lum) %>% 
  summarise(across(.cols=where(is.numeric), .fns=~sum(., na.rm=T),
                   .names="sum_{.col}")) %>% 
  ungroup()

pl_df_impression <-  df_ads_impressions %>% 
  mutate(party=fct_relevel(party, levels_party)) %>% 
  ggplot()+
  labs(title="WIEN-WAHL 2020: Total number of impressions of Facebook ads",
       subtitle=str_wrap(glue::glue("Top 5 Facebook pages per party. Ads placed between {date_observation_start_format} and {date_observation_end_format}. Impressions: Number of ads' appearances on a screen."), 100),
       caption=my_caption)+
  geom_linerange_interactive(aes(xmin=sum_impressions_lower_bound,
                    xmax=sum_impressions_upper_bound,
                    y=reorder_within(x=page_name_lum,
                                     by = sum_impressions_mid,
                                     within=party),
                    color=party))+
                    # tooltip=paste("min:", sum_impressions_lower_bound %>% scales::comma(), "\n",
                    #               "max:", sum_impressions_upper_bound %>% scales::comma())))+
  geom_point_interactive(aes(x=sum_impressions_mid,
                 y=reorder_within(x=page_name_lum,
                                     by = sum_impressions_mid,
                                     within=party),
                 color=party,
                 tooltip=paste("min:", sum_impressions_lower_bound %>% scales::comma(),"\n",
                               "mid:", sum_impressions_mid %>% scales::comma(),"\n",
                 "max:", sum_impressions_upper_bound %>% scales::comma())))+
  scale_y_discrete(labels=function(x) str_extract(x, regex(".*(?=___)")))+
  scale_x_continuous(expand=expansion(mult=c(0.01, 0.05)),
                     breaks=seq(0, 15*10^6, 5*10^6),
                     minor_breaks = NULL,
                     labels=scales::number_format(scale=.000001,
                                                  accuracy = 1,
                                                  suffix="m")
                     )+
  scale_color_manual(values=vec_party_colors)+
  lemon::facet_rep_wrap(vars(party),
                        ncol=2,
                        repeat.tick.labels = T,
                        scales = "free_y")+
  theme_post()+
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color="grey50", size = 0.1),
        panel.spacing.x=unit(0, units="cm"),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7),
        axis.title.x = element_blank())

```

</details>

```{r echo=FALSE, fig.height=4, dev='svg'}
girafe(ggobj=pl_df_impression,
       options=list(
         opts_hover_inv(css = "opacity:0.1;"),
         opts_hover(css = "fill:#FFA500;color:#FFA500"),
         opts_tooltip(css = "background-color:#323E4F;
         color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000)
                               ),
       opts_toolbar(saveaspng = FALSE), 
       height_svg=4)
```

What stands out is the value obtained for the institutional Facebook page of the Greens. With a total value of impressions ranging between `r df_ads_impressions %>% filter(str_detect(party, regex("Greens", ignore_case=T))) %>% slice_max(., n=1, order_by=sum_impressions_mid) %>% pull(sum_impressions_lower_bound) %>% comma()` and `r df_ads_impressions %>% filter(str_detect(party, regex("Greens", ignore_case=T))) %>% slice_max(., order_by=sum_impressions_mid, n=1) %>% pull(sum_impressions_upper_bound) %>% comma()`, the page's ads exceed the values by the pages of the ÖVP despite the latter's higher number of ads.

## Distribution of ads over impression categories

To get a better idea where the Green's high impression number comes from, let's disaggregate the total number of ads by their impression category. The table below shows for each impression interval, a) the number of ads per page, b) each interval's share of the page's total number of ads, and c) the contribution of the ads of each impression category to the total number of impressions. Note that these latter values are calculated on the basis of the mid values of each impression interval (e.g. impression category \<1K = 0 - 999 impressions; mid value = 499.5)!

<details closed>

<summary>Code: Distribution of ads over impression categories</summary>

```{r message=FALSE, warning=FALSE}
df_impression_cat <- df_ads_wide %>% 
  group_by(party) %>% 
  mutate(page_name_lum=fct_lump_n(page_name, n=5,
                                  other_level="all other")) %>% 
  group_by(party, page_name_lum, impressions_category, .drop=T) %>% 
  summarise(n_abs=n()) %>% 
  group_by(party, page_name_lum) %>% 
  mutate(n_total=sum(n_abs)) %>% 
  mutate(n_rel=n_abs/n_total) %>% 
  arrange(-n_rel, .by_group=T) %>% 
  ungroup() %>% 
  mutate(indicator=case_when(str_detect(page_name_lum, "Die Grünen Wien$") & 
                               impressions_category=="10000-14999"
~ "yes",
                         TRUE ~ "no"))
```

```{r message=FALSE, warning=FALSE}
fn_impression_label <- function(x) {
  str_extract(x, ".*(?=-)") %>% 
    as.numeric(.) %>% 
    scales::comma(.,
                  scale=0.001,
                  accuracy=1,
                  suffix="K")

}
```

</details>

<details closed>
<summary>Code: Table </summary>

```{r message=FALSE, warning=FALSE}
df_tb_impression_cat <- df_ads_wide %>% 
  group_by(party) %>% 
  mutate(page_name_lum=fct_lump_n(page_name, n=5,
                                  other_level="all other")) %>% 
  group_by(party, page_name_lum, impressions_category, .drop=T) %>% 
  summarise(n_abs=n()) %>% 
  group_by(party, page_name_lum) %>% 
  mutate(n_total=sum(n_abs)) %>% 
  mutate(n_rel=n_abs/n_total) %>% 
  arrange(-n_rel, .by_group=T) %>% 
  ungroup() %>% 
  mutate(indicator=case_when(str_detect(page_name_lum, "Die Grünen Wien$") &
                               impressions_category=="10000-14999"
~ "yes",
                         TRUE ~ "no")) %>% 
  #impression_mid_point
  separate(impressions_category, 
           into=c("impression_min", "impression_max"),
           sep = "-",
           remove=F) %>% 
  mutate(across(.cols=c("impression_min", "impression_max"),
                .fns=as.numeric)) %>% 
  mutate(impression_mid=(impression_max-impression_min)/2+impression_min) %>% 
  mutate(across(.cols=starts_with("impression") & !contains("category"), 
                .fns=function(x) x*n_abs,
                .names = "abs_{.col}"))

tb_impression_cat <- df_tb_impression_cat %>% 
  filter(str_detect(page_name_lum, "Die Grünen Wien$|Gernot Blümel|Neue Volkspartei Wien")) %>% 
  select(page_name_lum, impressions_category, n_abs, n_rel, starts_with("abs_impression")) %>% 
  pivot_wider(id_cols="impressions_category",
              names_from="page_name_lum",
              names_glue="{page_name_lum}_{.value}", #first name then value
              values_from=c(n_abs, n_rel, starts_with("abs")),
              values_fill = 0,
              names_sort=T) %>% 
  select(impressions_category, starts_with("Die Grünen"), starts_with("Gernot"), starts_with("Neue")) %>% 
  janitor::clean_names() %>% 
  ungroup() %>% 
  # mutate(xx=sum(die_grunen_wien_abs_impression_min, na.rm = T))
  mutate(across(.cols=contains("_abs_impression_"), 
                .fns=~sum(.,na.rm=T),
                .names="{.col}_sum")) %>% 
  mutate(die_grunen_abs_impression_mid_rel=die_grunen_wien_abs_impression_mid/die_grunen_wien_abs_impression_mid_sum,
         gernot_blumel_abs_impression_mid_rel=gernot_blumel_abs_impression_mid/gernot_blumel_abs_impression_mid_sum,
         neue_volkspartei_wien_abs_impression_mid_rel=neue_volkspartei_wien_abs_impression_mid/neue_volkspartei_wien_abs_impression_mid_sum) %>% 
    select(impressions_category, contains("grunen"), contains("blumel"), contains("volkspartei"))

```

```{r}
orange_pal <- function(x) rgb(colorRamp(c(plot_bg_color, "#ff9500"))(x), maxColorValue = 255)
firebrick_pal <- function(x) rgb(colorRamp(c(plot_bg_color, "#B22222"))(x), maxColorValue = 255)

rt_impressions <- reactable(tb_impression_cat %>% 
            mutate(impressions_category=str_extract(impressions_category, regex("(?<=-).*")) %>% 
                     as.numeric()+1 ) %>% 
            mutate(impressions_category = impressions_category / 1000) %>%                 #%>% #keep numeric
              arrange(impressions_category) %>% 
            select(-ends_with("_sum"), -ends_with("_min"), -ends_with("_max"),
                   -ends_with("_mid")),   
          columns=list(
            "impressions_category"=colDef(name="impression\ncategory", 
                                          width=100,
                                          sortable = T,
                                          format=colFormat(prefix="<",
                                                           suffix="K"),
                                          style = list(borderRight = "1px solid rgba(0, 0, 0, 0.1)")),
            "die_grunen_wien_n_abs"=colDef(name="Die Grünen Wien", minWidth=50),
            "die_grunen_wien_n_rel"=colDef(name="Die Grünen Wien",
                                           format = colFormat(percent=T,
                                                              digits=1), 
                                           style=function(value){

                                          color <- orange_pal(value)
                                          list(background=color)
                                        },
                                          minWidth=50),
            "die_grunen_abs_impression_mid_rel"=colDef(name="Die Grünen Wien", 
                                        format=colFormat(percent=T,
                                                                  digits = 1),
                                        style=function(value){

                                          color <- firebrick_pal(value)
                                          list(background=color)
                                        },
                                          minWidth=50),
            "gernot_blumel_abs_impression_mid_rel"=colDef(name="Gernot Blümel", 
                                        format=colFormat(percent=T,
                                                                  digits = 1),
                                        style=function(value){

                                          color <- firebrick_pal(value)
                                          list(background=color)
                                        },
                                        minWidth=50),
            "neue_volkspartei_wien_abs_impression_mid_rel"= colDef(name="Neue Volkspartei Wien", 
                                        format=colFormat(percent=T,
                                                                  digits = 1),
                                        style=function(value){

                                          color <- firebrick_pal(value)
                                          list(background=color)
                                        },
                                          minWidth=50),
            "gernot_blumel_n_abs"=colDef(name="Gernot Blümel", minWidth=50),
            "gernot_blumel_n_rel"=colDef(name="Gernot Blümel",
                                         format=colFormat(percent=T,
                                                          digits=1),
                                          style=function(value){

                                          color <- orange_pal(value)
                                          list(background=color)
                                        },
                                         minWidth=50),

            "neue_volkspartei_wien_n_abs"=colDef(name="Neue Volkspartei Wien", minWidth=50),
            "neue_volkspartei_wien_n_rel"=colDef(name="Neue Volkspartei Wien",
                                                 format=colFormat(percent=T,
                                                                  digits = 1),
                                                  style=function(value){

                                          color <- orange_pal(value)
                                          list(background=color)
                                        },
                                                 minWidth=50)),
            columnGroups = list(
              colGroup(name="number of page's ads ",
                       align="left",
                       columns=str_subset(names(tb_impression_cat), regex("n_abs$"))),
              colGroup(name="% of page's total ads",
                       align="left",
                       columns=str_subset(names(tb_impression_cat), regex("n_rel$"))),
              colGroup(name="% of page's total impressions",
                       align="left",
                       columns=str_subset(names(tb_impression_cat), regex("mid_rel$")))),
 sortable = T,
 filterable = F,
 compact = T,
 fullWidth = T,
 defaultPageSize = 15,
 theme=rt_theme
)

```

</details>

```{r echo=FALSE, message=FALSE, warning=FALSE}
fn_reactable_headings(title="Impressions",
                      subtitle="",
                      body = rt_impressions,
                      caption = my_caption)
```

<br>
Let's start with the lowest impression category, ads with fewer than 1000 impressions (<1K). As the table shows, Gernot Blümel ran `r df_impression_cat %>% filter(str_detect(page_name_lum, "Blümel")) %>% filter(impressions_category=="0-999") %>% pull(n_abs)` ads in this category. This number represents `r df_impression_cat %>% filter(str_detect(page_name_lum, "Blümel")) %>% filter(impressions_category=="0-999") %>% pull(n_rel) %>% percent(.,accuracy=.1)` of all ads bought by his page. Remarkably though, despite this high number, these ads only contributed `r tb_impression_cat %>% filter(impressions_category=="0-999") %>% pull(gernot_blumel_abs_impression_mid_rel) %>% percent(.,accuracy=.1)` of the total number of impressions generated by all his ads. This discrepancy is quite remarkable, however, the situation is not much different for the other two pages with the overall largest number of pages. The `r tb_impression_cat %>% filter(impressions_category=="0-999") %>% pull(die_grunen_wien_n_abs)` ads of the page of the Wiener Grünen amounted to `r tb_impression_cat %>% filter(impressions_category=="0-999") %>% pull(die_grunen_wien_n_rel) %>% percent(., accuracy=.1)` of all their ads, but contributed as little as `r tb_impression_cat %>% filter(impressions_category=="0-999") %>% pull(die_grunen_abs_impression_mid_rel) %>% percent(.,accuracy=.1)` of the total number of impressions. For the page of the Neue Wiener Volkspartei the `r tb_impression_cat %>% filter(impressions_category=="0-999") %>% pull(die_grunen_wien_n_abs)` ads triggered only `r tb_impression_cat %>% filter(impressions_category=="0-999") %>% pull(neue_volkspartei_wien_abs_impression_mid_rel) %>% percent(.,accuracy=.1)` of the overall impression.

Hence, the question - where does the larger chunks of impressions come from?

You can sort the columns by clicking on their headings. If you sort, the pertaining column of the Viennese Greens, you'll see that `r tb_impression_cat %>% filter(str_detect(impressions_category, "699999")) %>% pull(die_grunen_abs_impression_mid_rel) %>% percent(.,accuracy=.1)` of their total impressions originated from as little as `r tb_impression_cat %>% filter(str_detect(impressions_category, "699999")) %>% pull(die_grunen_wien_n_abs)` ads in the category of \<700K impressions, or `r tb_impression_cat %>% filter(str_detect(impressions_category, "699999")) %>% pull(die_grunen_wien_n_rel) %>% percent(., accuracy=.1)` of its total number of ads. Another `r tb_impression_cat %>% filter(str_detect(impressions_category, "899999")) %>% pull(die_grunen_abs_impression_mid_rel) %>% percent(.,accuracy=.1)` come from from only `r tb_impression_cat %>% filter(str_detect(impressions_category, "899999")) %>% pull(die_grunen_wien_n_abs)` ads from the \<900K category. Hence, in short, the total number of impressions seems rather concentrated on a few ads. As with the page of Gernot Blümel, the picture is somewhat similar. `r tb_impression_cat %>% filter(str_detect(impressions_category, "124999")) %>% pull(gernot_blumel_abs_impression_mid_rel) %>% percent(.,accuracy=.1)` of all impressions generated by his ads originate from only `r tb_impression_cat %>% filter(str_detect(impressions_category, "124999")) %>% pull(gernot_blumel_n_abs)` ads in the \<125K impressions category.

Another way to visualize the origin of the difference in impressions is the graph below. The x-axis shows the impression categories/intervals of ads in increasing order. The y-axis shows the total number of impressions if we accumulate the impressions over the different impression categories. The graph reveals that the Green's strong impression numbers are largely due to their ads placed in the highest impression categories (larger than 400K impressions). Neither Blümel nor Nepp had placed any ads in these categories.

<details closed>

<summary>Code: Impressions per ad category</summary>

```{r}

df_pl_impression <- df_tb_impression_cat %>% 
  mutate(impressions_category_num=str_extract(impressions_category, regex("(?<=-).*")) %>% 
                     as.numeric()+1 ) %>% 
  mutate(impressions_category_num = impressions_category_num / 1000) %>% 
  mutate(impressions_category_fac = as_factor(impressions_category_num)) %>%  
  mutate(impressions_category=fct_reorder(impressions_category, impressions_category_num)) %>% 
  arrange(impressions_category) %>% 
  group_by(page_name_lum) %>% 
  mutate(across(.cols=starts_with("abs_impression"),
                .fns=list(cumsum=cumsum),
                .names="{.fn}.{.col}")) %>% 
  ungroup() %>% 
  filter(str_detect(page_name_lum, regex("Blümel|Nepp|Die Grünen Wien$")))

df_pl_impression %>% 
  group_by(page_name_lum) %>%
  slice_tail(n=1) %>% 
  ungroup() %>% 
  filter(str_detect(page_name_lum, "Grünen"))

pl_impression <- df_pl_impression %>% 
  arrange(impressions_category_num) %>% 
  ggplot()+
  labs(title="WIEN-WAHL 2020: Total impressions by impression category",
       subtitle="Top 3 pages with most total impressions.",
       y="impressions, accumulated",
       x="impression category of ads",
       caption=my_caption)+
  geom_vline(xintercept=seq(0.5, length(df_pl_impression$impressions_category_fac)+.5),
             linetype="dotted",
             color="grey50")+
  geom_pointrange(aes(x=impressions_category_fac,
                      y=cumsum.abs_impression_mid,
                      ymin=cumsum.abs_impression_min,
                      ymax=cumsum.abs_impression_max,
                      color=party))+
  geom_line(aes(x=impressions_category_fac,
                y=cumsum.abs_impression_mid,
                group=page_name_lum,
                color=party))+
  geom_text(data=. %>% 
              group_by(page_name_lum) %>%
              slice_tail(n=1) %>% 
              ungroup() %>% 
              filter(!str_detect(page_name_lum, "Grünen")),
            aes(x=impressions_category_fac,
                y=cumsum.abs_impression_mid,
                label=page_name_lum, 
                color=party),
            hjust=0,
            nudge_x=1)+
    geom_text(data=. %>% 
              group_by(page_name_lum) %>%
              slice_tail(n=1) %>% 
              ungroup() %>% 
              filter(str_detect(page_name_lum, "Grünen")),
            aes(x=impressions_category_fac,
                y=cumsum.abs_impression_mid,
                label=page_name_lum, 
                color=party),
            hjust=1,
            fontface="bold",
            nudge_x=-1)+
  scale_y_continuous(labels=scales::label_comma(scale=10^-6,
                                                accuracy=1,
                                                suffix="m"),
                     expand=expansion(mult=c(0,0)))+
  scale_x_discrete(label=function(x) paste0("<",x,"K"),
                   guide= guide_axis(n.dodge = 2),
                   expand=expansion(ad=c(0.5, 3)))+
  scale_color_manual(values=vec_party_colors)+
  theme_post()+
  theme(axis.title.y=element_markdown(angle=90,
                                      hjust=1,
                                      color="grey50"),
        axis.text.x = element_markdown(color="grey50"),
        panel.grid.minor.x = element_line(size=1),
        legend.position = "none")
```

</details>

```{r echo=FALSE, message=FALSE, warning=FALSE}
pl_impression
```

# Demography

## Age

Now let's turn to some of the demographic data provided by Facebook's ad library API. One demographic property is age. For this we have to unnest again the pertaining field. A critical detail here is to calculate an ad's audience share for every age group. E.g. if an ad has an audience share of 100 % in the age group of 25 to 34, we still want to have the observations for all the other age group's (with the resulting values of 0 %). Otherwise, with these observations missing, the subsequent calculation of i.e. the median per each age group would be distorted. These observations are created with the `complete` and `nesting` function of the `tidyr` package.

<details closed>

<summary>Code: Unnest age list column</summary>

```{r message=FALSE, warning=FALSE}
df_age <- df_ads_wide %>% 
  select(name_party, party, page_name, ad_id, ad_creative_body, demographic_distribution_df, impressions_mid) %>% 
  unnest(cols=c(demographic_distribution_df)) %>% 
  mutate(percentage=as.numeric(percentage)) %>% 
  group_by(party) %>% 
  mutate(name_party_lum=fct_lump(name_party, n=3, other_level="all others")) %>% 
  group_by(party, name_party_lum, page_name, ad_id, age, impressions_mid, .drop=T) %>% 
  summarize(age_share=sum(percentage)) %>% #sum across all genders for one ad
  ungroup() %>% 
  complete(age, nesting(ad_id, name_party_lum, party, page_name, impressions_mid), fill=list(age_share=0)) %>% 
  mutate(party=fct_relevel(party, levels_party)) %>% 
  group_by(age, name_party_lum) %>% 
  mutate(median_share=median(age_share, na.rm=T)) %>% 
  ungroup()

#every adid should appear in each age segment; => each adid should have 7 occurances;
check <- df_age %>% 
  group_by(ad_id) %>% 
  summarise(n=n())
#summary(check$n)

# levels_main_pages <- c("Michael Ludwig","Gernot Blümel","Birgit Hebein","Christoph Wiederkehr","Dominik Nepp","Heinz-Christian Strache / HC Strache","Die Bierpartei")
vec_main_pages <- c("Gernot Blümel|Michael Ludwig|Birgit Hebein|Christoph Wiederkehr|Dominik Nepp|Heinz-Christian Strache / HC Strache|Die Bierpartei")

df_age_stats <-  df_age %>% 
  filter(str_detect(page_name, regex(vec_main_pages))) %>%
  mutate(name_party_lum=fct_drop(name_party_lum)) %>% 
  group_by(name_party_lum, age) %>% 
  summarise(category_median=median(age_share, na.rm=T),
            category_mean=stats::weighted.mean(x=age_share,  #weighted mean
                                        w=impressions_mid,
                                        na.rm=T),
            category_sd=sd(age_share, na.rm=T)) %>% 
  group_by(age) %>% 
  mutate(category_median_max=max(category_median, na.rm=T)) %>% 
  mutate(category_median_max_indic=case_when(category_median_max == category_median ~ "max",
                                             TRUE ~ as.character("not max"))) %>% 
  ungroup()
```

</details>

```{r include=FALSE}
sample_ad <- "MichaelLudwig-2020-09-10-195"

df_example <- df_age %>% 
  filter(ad_id==sample_ad) %>% 
  left_join(., df_ads_wide %>% 
              select(ad_id, ad_creative_body),
            by="ad_id") %>% 
  select(page_name, ad_id, ad_creative_body, age, age_share)

df_example %>% 
  slice(1) %>% 
  select(page_name, ad_creative_body)
```

As an illustration, the ad with the text

>`r kableExtra::text_spec(df_example %>% slice(1) %>% pull(ad_creative_body) %>% str_remove(., "\n") %>% str_replace(., "\n", " "), italic="true", font_size=14)`

run by the page of `r df_example %>% slice(1) %>% pull(page_name)` has the following age distribution:

```{r echo=FALSE}
df_example %>% 
  select(starts_with("age"))

```

The graph below plots one dot for each age share for every ad. To better compare the audience shares between candidates, I added a boxplot to the graph. It provides us with a summary of the distribution of all ads' audience shares for each candidate in each age group. The vertical line within the box is the median value. Hover with the mouse over one dot to get details on the ad and highlight the ad's impression share in the other age categories.

<details closed>
<summary>Code: Plot age profile of ads</summary>

```{r message=FALSE, warning=FALSE}
pl_demographics <- df_age %>% 
  mutate(ad_id=str_replace_all(ad_id, "'"," ")) %>% 
  filter(str_detect(page_name, regex(vec_main_pages))) %>%
  mutate(name_party_lum=fct_drop(name_party_lum)) %>% 
  ggplot()+
  labs(title="WIEN-WAHL 2020: Age profile of Facebook ads' audience",
       subtitle=str_wrap(glue::glue("Each dot represents one ad's audience share in a specific age group. Each ad has one dot in every age group. Over all age groups, each ad has an audience of 100 %. Only top candidates. Ads placed between {date_observation_start_format} and {date_observation_end_format}."), 110),
x="Share of age group in ad's overall audience",
y=NULL,
caption=my_caption)+
  #geoms
  geom_jitter_interactive(aes(
    y=reorder_within(x=name_party_lum,
                     by=median_share,
                     within = age),
                  x=age_share,
                  tooltip=glue::glue("Ad id: {ad_id},
                                     Age group share: {scales::percent(age_share)}"),
                  data_id=ad_id),
                  fill="transparent",
                  color="grey50",
                  alpha=0.9,
                  height = 0.1, 
                  size=.3,
                  width = 0.05)+
  geom_boxplot(aes(
    y=reorder_within(x=name_party_lum,
                     by=median_share,
                     within = age),
                   x=age_share,
    #weight=impressions_mid,
    color=party),
    outlier.shape = NA,
    fill="transparent")+
#headings
 geom_text(aes(y=length(unique(name_party_lum))+1,
            x=1.2),
            label="median", 
            size=2.5,
            hjust=1,
            check_overlap = T)  +
  geom_text(aes(y=length(unique(name_party_lum))+1,
            x=1.4),
            label="SD", 
            size=2.5,
            hjust=1,
            check_overlap = T)  +
  #median values
  geom_text(data=df_age_stats,
            aes(x=1.2,
                y=reorder_within(x=name_party_lum,
                                               by=category_median,
                                               within = age
                                                 ),
                label=c(category_median %>% scales::percent(.,
                                                            suffix="",
                                                            accuracy=0.1))),
            fontface=ifelse(df_age_stats$category_median_max_indic=="max",
                         "bold", "plain"),
            size=2.5,
            hjust=1)+
  #sd values
  geom_text(data=df_age_stats,
            aes(x=1.4,
                y=reorder_within(x=name_party_lum,
                                               by=category_median,
                                               within = age
                                                 ),
                label=c(category_sd %>% scales::percent(.,
                                                        suffix="",
                                                        accuracy=0.1))),
            hjust=1,
            size=2.5)+
  scale_y_discrete(labels=function(x) str_extract(x, regex(".*(?=\\(|___)")),
                   expand=expansion(mult=c(0, 0.25)),
  drop=T)+
  scale_x_continuous(labels=scales::percent,
                     expand=expansion(mult=c(0,0.2)),
                     breaks=seq(0,1,.5))+
  scale_color_manual(values=vec_party_colors)+
  lemon::facet_rep_wrap(vars(age),
                        labeller=as_labeller(function(x) paste("age group:", x)),
                        ncol=2,
                        scales="free_y",
                        repeat.tick.labels = T)+
  theme_post()+
  theme(legend.position = "none",
        axis.title.y = element_markdown(size=6, 
                                        angle=90,
                                        color = "grey50",
                                        hjust = 1),
        axis.text.x = element_text(size=7),
        axis.title.x = element_markdown(size=7, hjust=0),
        axis.text.y=element_text(size=9),
        panel.spacing.x = unit(0, units="cm"),
        panel.grid.major.y = element_blank())

ggsave("pl_demographics.png",
       plot = pl_demographics,
       dpi=300,
       device="png")

```

</details>

```{r echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, dev='svg'}
girafe(ggobj=pl_demographics,
                             options=list(
                               opts_hover_inv(css = "opacity:0.0;"),
                               opts_hover(css = "fill:orange;colour:orange;
                                          opacity:1;r:3pt;"),
                               opts_tooltip(css = "background-color:#323E4F;
                               color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000),
                               opts_toolbar(saveaspng = FALSE) 
                               ),
       height_svg=7)
```

I won't go through all details, but to understand the graph let's have a look at the age group 25-34. Ads placed by the Green's main candidate Hebein have a median audience share of `r df_age_stats %>% filter(str_detect(name_party_lum, "Hebein")) %>% filter(age=="25-34") %>% pull(category_median) %>% scales::percent()` in this age group. No other main candidate had such high median audience share in this or any other age group. On the other hand, if we look at higher/older age groups, we can see that Hebein's medians were the lowest from the age 45 up. In contrast, Mayor Ludwig featured comparably high values in the older age groups. More generally, across age groups candidates have higher values in the age groups 25-54. Gernot Blümel has relatively evenly distributed audience shares. However, note that Blümel features in almost every age group a few (outlier) ads which have an audience share of 100% in the respective age group. These were ads which were particularly targeted at this specific age group.

## Gender

Another demographic detail provided by Facebook's ad library is the gender composition of an ad's audience. Going back to our previous sample ad by Michael Ludwig, we obtain the following data:

```{r echo=FALSE}
df_sample_age_gender <- df_ads_wide %>% 
  select(name_party, party, page_name, ad_id, demographic_distribution_df) %>% 
  unnest(cols=c(demographic_distribution_df)) %>% 
  mutate(percentage=as.numeric(percentage)) %>% 
  group_by(party) %>% 
  mutate(name_party_lum=fct_lump(name_party, n=3, other_level="all others")) %>% 
  ungroup() %>% 
  complete(age, gender, nesting(ad_id, name_party_lum, party, page_name), fill=list(percentage=0)) %>%
  filter(ad_id==sample_ad) %>% 
  select(age, gender, percentage) %>% 
  pivot_wider(id_cols=age,
              names_from = gender,
              values_from = percentage)

df_sample_age_gender
```

Let's aggregate these percentages to get subtotals for each gender.

```{r echo=FALSE}
df_sample_gender <- df_sample_age_gender %>% 
  select(where(is.numeric)) %>% 
  mutate(across(.cols=where(is.numeric), .fns=sum)) %>% 
  slice(1)

df_sample_gender
```

Hence, in this specific case, `r df_sample_gender %>% pull(female) %>% scales::percent()` of the ad's audience were women.

We can apply this analysis to all ads of an candidate and contrast the obtained average with those of other candidates. The results are presented in the graph below.

Each gray dot represents one ad's audience share of the respective gender group. Again, the vertical line inside the boxplot indicates the median value of all ads' audience shares of a candidate in one gender group.

<details closed>

<summary>Code: Gender of ads' audience</summary>

```{r message=FALSE, warning=FALSE}
df_gender <- df_ads_wide %>% 
  select(name_party, party, page_name, ad_id, demographic_distribution_df) %>% 
  unnest(cols=c(demographic_distribution_df)) %>% 
  mutate(percentage=as.numeric(percentage)) %>% 
  group_by(party) %>% 
  mutate(name_party_lum=fct_lump(name_party, n=3, other_level="all others")) %>% 
  complete(gender, nesting(ad_id, name_party_lum, party, page_name), fill=list(percentage=0)) %>%
  group_by(party, name_party_lum, page_name, ad_id, gender) %>%
  summarise(gender_sum=sum(percentage, na.rm = T, .drop=F)) %>% ##nesting needed
  ungroup() %>% 
  mutate(party=fct_relevel(party, levels_party)) %>% 
  mutate(name_party_lum=fct_reorder(name_party_lum, as.numeric(party)))

# 3 observations for each ad (male/female/unknown)
# check <- df_gender %>% 
#   group_by(ad_id) %>% 
#   summarise(n_ads=n())
# summary(check)

df_gender_stats <- df_gender %>% 
  group_by(name_party_lum, page_name, gender) %>% 
  summarise(
    gender_mean=mean(gender_sum, na.rm=T),
    gender_median=median(gender_sum, na.rm=T),
    gender_sd=sd(gender_sum, na.rm=T),
            ) %>% 
  filter(str_detect(page_name, vec_main_pages)) %>% 
  filter(gender!="unknown")

pl_gender <- df_gender %>% 
  filter(str_detect(page_name, vec_main_pages)) %>% 
  filter(gender!="unknown") %>% 
  ggplot()+
  labs(title="WIEN-WAHL 2020: Gender profile of Facebook ads' audience",
       subtitle=str_wrap(glue::glue("Each dot represents one ad's audience share in a specific gender group. Each ad has one dot in every gender group. Over all age groups, each ad has an audience of 100 %. Only top candidates. Ads placed between {date_observation_start_format} and {date_observation_end_format}. Category 'unknown' not shown."),110),
       x="Share of gender group in ad's overall audience",
       caption=my_caption)+
  geom_jitter_interactive(aes(y=gender,
                  x=gender_sum,
                  tooltip=glue::glue("Ad id: {ad_id},
                                     Share: {scales::percent(gender_sum)}"),
                  data_id=ad_id),
                  color="grey50",
                  alpha=0.9,
                  height = 0.2, 
                  size=.3,
                  width = 0.01)+
  geom_boxplot(aes(y=gender,
                   x=gender_sum,
                   color=party),
               outlier.shape = NA,
               fill="transparent")+
  #labels
  geom_text(aes(y=length(unique(gender))+1,
            x=1.2),
            label="median", 
            size=2.5,
            hjust=1,
            check_overlap = T)  +
  geom_text(aes(y=length(unique(gender))+1,
            x=1.3),
            label="SD", 
            size=2.5,
            hjust=1,
            check_overlap = T)  +
  #median values
  geom_text(data=df_gender_stats,
            aes(x=1.2,
                y=gender,
                label=gender_mean %>% scales::percent(.,
                                                            suffix="",
                                                            accuracy=0.1)),
            # fontface=ifelse(df_age_stats$category_median_max_indic=="max",
            #              "bold", "plain"),
            size=2.5,
            hjust=1)+
  #sd values
  geom_text(data=df_gender_stats,
            aes(x=1.3,
                y=gender,
                label=gender_sd %>% scales::percent(.,
                                                        suffix="",
                                                        accuracy=0.1)),
            hjust=1,
            size=2.5)+
  #scales
  scale_color_manual(values=vec_party_colors)+
  scale_x_continuous(labels=scales::label_percent())+
  scale_y_discrete(expand=expansion(add=c(0, 1.5)), drop=T)+
  facet_wrap(vars(name_party_lum),
             ncol=1)+
  theme_post()+
  theme(legend.position = "none",
        axis.text.y=element_text(size=9),
        panel.spacing.x = unit(0, units="cm"),
        panel.grid.major.y = element_blank(),
                axis.text.x = element_text(size=7),
        axis.title.x = element_markdown(size=7, hjust=0))

ggsave("pl_gender.png", 
       plot = pl_gender,
       dpi=300,
       device="png")

```

</details>

```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, dev='svg'}
girafe(ggobj=pl_gender,
                             options=list(
                               opts_hover_inv(css = "opacity:0.0;"),
                               opts_hover(css = "fill:orange;colour:orange;opacity:1;r:3pt;"),
                               opts_tooltip(css = "background-color:#323E4F;
                               color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000),
                               opts_toolbar(saveaspng = FALSE) 
                               ),
       height_svg=5)
```

What we can see is that for all main candidates the median value of the male audience share is larger than that of the female audience share. The differences between the median values are particularly apart for Strache. But even for the Greens with their female candidate Hebein, ads were seen on average more often by men than by women.

This result made me wonder whether there were in fact any major Facebook pages during the campaign whose ads were seen on average by more women than men.

<details closed>

<summary>Code: Pages with female audience</summary>

```{r message=FALSE, warning=FALSE}
df_female <- df_gender %>% 
  select(party, page_name, gender, gender_sum) %>% 
  group_by(party, page_name, gender) %>% 
  mutate(n_ads=n()) %>% 
  group_by(party, page_name, gender, n_ads) %>% 
  summarise(across(.cols=gender_sum,
                   .fns = list(mean=mean,
                               median=median,
                               sd=sd))) %>% 
  filter(gender=="female" & gender_sum_median >.5) %>% 
  arrange(desc(gender_sum_median)) %>% 
  ungroup()

rt_female <- reactable(df_female,
          columns =list(
            party=colDef(name="Party"),
            page_name=colDef(name="page"),
            n_ads=colDef(name="# ads"),
            gender=colDef(show=F),
            gender_sum_mean=colDef(name="mean",
                                   format = colFormat(percent=T,
                                                      digits = 1)),
            gender_sum_median=colDef(name="median",
                                   format = colFormat(percent=T,
                                                      digits = 1)),
            gender_sum_sd=colDef(name="SD",
                                   format = colFormat(percent=T,
                                                      digits = 1))),
          columnGroups = list(
            colGroup(name="female audience share",
                     columns=c("gender_sum_mean", "gender_sum_median","gender_sum_sd"))
          ),
          theme=rt_theme,
          defaultPageSize=11,
          sortable = T,
          borderless = T,
          filterable = T,
          outlined = TRUE)
```

</details>

```{r echo=FALSE, message=FALSE, warning=FALSE}
fn_reactable_headings(title="WIEN-WAHL 2020: Accounts which placed ads with a median female audience share of >50%", subtitle="", body = rt_female, caption=my_caption)

```
<br>
Indeed, as the table above shows there were indeed a few pages with an average female audience share larger than 50 %. What is noteworthy, these pages belong almost exclusively to the Greens. There were only `r df_female %>% filter(party=="SPÖ") %>% nrow()` pages which belonged to the SPÖ, and `r df_female %>% filter(party=="ÖVP") %>% nrow()` to the ÖVP. There was not a single page from the FPÖ, Strache and, more surprisingly, the Neos.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df_female %>% 
  count(party)
```

# Fin

So, again a blog post which got much longer than initially planned. If there's anything you want to know more about, something that's not clear, feel free to send me a DM via [twitter.](https://twitter.com/zoowalk){target="_blank"}

### 2021-01-06 _posts/2020-12-22-submissionsepidemiclaw/submissionsepidemiclaw.Rmd
post submission: removed files from subfolder => now all _blog_data

@@
 ```{r include=FALSE}
-df_submission <- readr::read_csv2(file="submissionsepidemiclaw_files/df_submission_rss.csv")
+# df_submission <- readr::read_csv2(file="submissionsepidemiclaw_files/df_submission_rss.csv")
+df_submission <- readr::read_csv2(file=here::here("_blog_data",                              "epidemic_law_submissions", "df_submission_rss.csv"))
 
@@
 # load saved links to pdf_files
-df_submission <- readr::read_csv2(file="submissionsepidemiclaw_files/df_submission.csv")
+# df_submission <- readr::read_csv2(file="submissionsepidemiclaw_files/df_submission.csv")
+df_submission <- readr::read_csv2(file=here::here("_blog_data", "epidemic_law_submissions","df_submission.csv"))
 
@@
 ```{r echo=FALSE}
-df_sozial_sim <- readr::read_csv2(file="submissionsepidemiclaw_files/df_bar_sim_sozial.csv") %>% 
-  mutate(law="Sozialhilfegrundgesetz")
+df_sozial_sim <- readr::read_csv2(file=here::here("_blog_data", "epidemic_law_submissions", "df_bar_sim_sozial.csv")) %>%  mutate(law="Sozialhilfegrundgesetz")
 

### 2021-01-06 _post/2020-12-19-vienna-elections-2020-analysis-of-parties-ads-on-facebook/welcome.Rmd
changed post directory; applied ful site.yml

--- 2021-01-06 _posts/2020-12-19-vienna-elections-2020-analysis-of-parties-ads-on-facebook/welcome.Rmd
@@
 
-write_rds(df_ads_all, path=paste0(wdr,"/data/scraped_ads",time_stamp,".rds"))
+#write_rds(df_ads_all, path=paste0(wdr,"/data/scraped_ads",time_stamp,".rds"))
 ```
@@
 ```{r include=FALSE}
-df_ads <- readr::read_rds(file=here::here("blog_data", 
+df_ads <- readr::read_rds(file=here::here("_blog_data", 
                                           "vienna_elections_2020_facebook",
@@
 ```{r include=FALSE}
-pages_of_interest <- readxl::read_xlsx(path=here::here("blog_data", "vienna_elections_2020_facebook", "df_pages_of_interest.xlsx")) %>% 
+pages_of_interest <- readxl::read_xlsx(path=here::here("_blog_data", "vienna_elections_2020_facebook", "df_pages_of_interest.xlsx")) %>% 
   # filter(include=="yes") %>% 
@@
 
-readr::write_csv2(df_impression_check,file=here::here("blog_data","vienna_elections_2020_facebook","df_impression_check.csv"))
+#readr::write_csv2(df_impression_check,file=here::here("blog_data","vienna_elections_2020_facebook","df_impression_check.csv"))
 ```

### 2021-01-06 _post/2020-11-25-static-maps-with-mapbox-background-images-snapbox/static-maps-with-mapbox-background-images-snapbox.Rmd
add snapbox post

---
title: "Static maps with mapbox background images (snapbox)"
description: |
  How to create ggplot maps with mapbox images as background
date: 11-25-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693

categories:
  - maps
  - Austria
  - sf
  - snapbox
  - mapbox

preview: submissionsepidemiclaw_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

# CONTEXT

Partly triggered by my [last post](https://werk.statt.codes/post/recreating-nyt-election-map/){target="_blank"} on replicating NYT's electoral maps, I came across the nascent/experimental `snapbox` package ([link](https://github.com/anthonynorth/snapbox){target="_blank"}) by Anthony North and Miles McBain. The package provides a convenient way to add mapbox tiles as basemaps to maps created with ggplot (did I mention 'maps'?`).

The package is quite straightforward and not much to explain about. What caused me some hassles though was to figure out how to combine the basemap with my data, plotted with `geom_segment` or `geom_sf`. The critical issue here is to ensure that the maps projection of the coordiante system are the same. Hence, this post is somewhat a note to myself, graphs are not polished and I don't dig into substantive matters. For the purpose of demonstration, I'll use electoral data from the city of Vienna and want to highlight how the number of eligible voters per precinct ('Wahlsprengel') changed between 2015 and 2020.

# SETUP

<details closed>
<summary>Code: Load packages </summary>
```{r message=FALSE, warning=FALSE}
library(snapbox)
library(sf)
library(tidyverse)
#library(magick)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop) #allows to crop white frame from ggplots.
#for details see https://bookdown.org/yihui/rmarkdown-cookbook/crop-plot.html
plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()
```
</details>

# ELECTORAL DATA

Get the election results from 2015 and 2020, retrieve the number of eligible voters per precinct (`wber``), and calculate the change.

<details open>
<summary>Code: Get electoral data + change in number of elegible voters </summary>
```{r message=FALSE, warning=FALSE}
# https://www.data.gv.at/katalog/dataset/stadt-wien_wahleninwiengemeinderatswahlen/resource/a526df98-c307-4dee-8233-43b02856d1d7
# https://www.wien.gv.at/politik/wahlen/hilfetext/rohdaten.html
res20 <- readr::read_csv2(file = "https://www.wien.gv.at/politik/wahlen/ogd/gr201_99999999_9999_spr.csv")
res15 <- readr::read_csv2(file = "https://www.wien.gv.at/politik/wahlen/ogd/gr151_99999999_9999_spr.csv")

res <- bind_rows(elections=list('2020'=res20, '2015'=res15), .id="election") %>% 
  janitor::clean_names() %>% 
  janitor::remove_empty("cols") %>% 
  mutate(election=as.numeric(election)) %>% 
  mutate(indicator_sprengel=bz*1000+spr, .after=spr) %>% 
  select(election:wber)

# change of eligible voters ----------------------------------------------------------

df_wber_diff <- res %>% 
  filter(wber>0) %>% #remove 'technical' precincts for mail & absentee voting 
  select(-contains("nuts")) %>% 
  group_by(indicator_sprengel) %>% 
  arrange(desc(election), .by_group=T) %>% 
  mutate(wber_diff=wber-lead(wber)) %>% 
  filter(election==2020) %>% 
  mutate(wber_diff_rel=wber_diff/wber) %>% 
  ungroup() %>% 
  mutate(change_indicator=case_when(wber_diff_rel>0 ~ "increase",
                                    wber_diff_rel<0 ~ "decrease",
                                    TRUE ~ as.character("stable")))
```
</details>

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
df_wber_diff %>% 
  select(-election, -district_code, -wber, -c(t:spr))
```

Let's have a quick look at the distribution of the change rates:

<details closed>
<summary>Code: Distribution of changes</summary>
```{r message=FALSE, warning=FALSE}
pl_dist <- df_wber_diff %>% 
  ggplot()+
  ggridges::geom_density_line(aes(x=wber_diff_rel))+
  labs(x="% change number of eligible voters 2020 vs 2015")+
  theme(axis.text.y=element_blank())

table(df_wber_diff$change_indicator)

```
</details>

```{r echo=FALSE, message=FALSE, warning=FALSE}
table(df_wber_diff$change_indicator)

pl_dist
```

So, while the purpose of this post is not to dig into substantive matters, I was a bit surprised to see that a large majority of precincts features a decreased number of eligible voters.

# MAP

## Get boundaries of precincts

Get the shapefile of the precincts' boundaries.

<details open>
<summary>Code: Get shape with boundaries of precincts </summary>
```{r}
# https://www.data.gv.at/katalog/dataset/79c1030d-5cf6-4d58-ade6-02f66fb4dffb
precincts <- st_read(here::here("_blog_data","static_map_snapbox","shapefile","WAHLSPRGR2020OGDPolygon.shp"),
              quiet=T) %>% 
  janitor::clean_names() %>% 
  mutate(sprengel=as.numeric(sprengel))

ggplot()+
  geom_sf(data=precincts)
```
</details>

Now let's calculate the center of each precinct. Note that I transform the coordination projection from WGS84 to 3857. The latter is 'a Spherical Mercator projection coordinate system popularized by web services such as Google and later OpenStreetMap.'^[See [here](https://wiki.openstreetmap.org/wiki/EPSG:3857#:~:text=EPSG%3A3857%20is%20a%20Spherical,as%20Google%20and%20later%20OpenStreetMap.){target="_blank"}.] 
After the transformation I take the x and y values from the centroids' geometry. The function `st_centroid` gives us the centre of each polygon.

<details open>
<summary>Code: Get centroids for precincts</summary>
```{r message=FALSE, warning=FALSE}
st_crs(precincts)

cent_precincts <- precincts %>% 
  st_transform(., crs=3857) %>% 
  st_centroid() %>% 
  mutate(point_x=map_dbl(geometry, pluck, 1),
         point_y=map_dbl(geometry, pluck, 2))

ggplot()+
  geom_sf(data=precincts)+
  geom_sf(data=cent_precincts,
          size=0.1)
```
</details>

## Add change indicators (arrows)

Calculate the length of the arrow indicating the change: Here the arrows should have an angle of 20 degrees; positive changes point towards the right, negative changes towards the left. With the sin and cos function we calculate the horizontal and vertical distance from the centroids.  In addition, to make the arrows sufficiently long to be visible, I multiply them with a scale factor. Note that I add `coord_sf()` to the plot which 'ensures that all layers use a common CRS.'^[See [here.](https://ggplot2.tidyverse.org/reference/ggsf.html){target="_blank"}]

<details open>
<summary>Code: Calculate length of arrows indicating change </summary>
```{r message=FALSE, warning=FALSE}
scale_factor <- 10^3

df_change <- cent_precincts %>% 
  left_join(., df_wber_diff,
            by=c("sprengel"="indicator_sprengel")) %>% 
  mutate(x_change=wber_diff_rel*sin(20),
         y_change=abs(wber_diff_rel)*cos(20)) %>% 
  mutate(point_x_end=point_x+(x_change*scale_factor), .after=point_x) %>% 
  mutate(point_y_end=(y_change*scale_factor)+point_y, .after=point_y)

pl_change <- ggplot()+
  geom_sf(data=precincts,
          size=0.1,
          fill="transparent")+
  geom_segment(data=df_change,
               aes(x=point_x,
                   xend=point_x_end,
                   y=point_y,
                   yend=point_y_end,
                   color=change_indicator),
               linejoin="mitre",
               arrow=arrow(length=unit(0.05, "cm"),
                           angle=15,
                           type="closed"),
               size=0.5)+
  scale_color_manual(values=c("increase"="darkgreen",
                              "decrease"="maroon",
                              "stable"="orange"))+
  coord_sf(crs=3857)+
  theme_void()

```
</details>

```{r echo=FALSE}
pl_change
```

## Add mapbox basemap

Now, let's add a mapbox basemap by using the `layer_mapbox` function of `snapbox`. To so, we first have to define the scope or boundary of the basemap. I do this by extracting the minimum and maximum x and y values from our map of the precincts. And then create a bbox object. Note that this boundary box is has the 4326 projection. The basemap for this boundary box is then added to the plot by using the layer_mapbox function. Note that you need an [API-key](https://docs.mapbox.com/help/how-mapbox-works/access-tokens/){target="_blank"} to access basemaps of mapbox.

Et voila.

<details open>
<summary>Code: Add mapbox basemap</summary>
```{r message=FALSE, warning=FALSE}
#get boundaries of plot

#map boundaries: get max/min x, y values
d <- precincts$geometry
d <- attributes(d)$bbox
my_xmin <- d[[1]]
my_ymin <- d[[2]]
my_xmax <- d[[3]]
my_ymax <- d[[4]]
my_xmin
my_ymin
my_xmax
my_ymax

#get basemap
background_image <- st_bbox(
  c(xmin = my_xmin, ymin = my_ymin, xmax = my_xmax, ymax = my_ymax),
  crs=4326)

#combine
pl_combined <- ggplot()+
  layer_mapbox(background_image,
               map_style = mapbox_galley_pencil(),
               mapbox_api_access_token = Sys.getenv("MAPBOX_ACCESS_TOKEN"),
               scale_ratio = 1)+
  geom_sf(data=precincts,
          size=0.2,
          fill="transparent")+
  geom_segment(data=df_change,
               aes(x=point_x,
                   xend=point_x_end,
                   y=point_y,
                   yend=point_y_end,
                   color=change_indicator),
               linejoin="mitre",
               arrow=arrow(length=unit(0.05, "cm"),
                           angle=15,
                           type="closed"),
               size=0.5)+
  scale_color_manual(values=c("increase"="darkgreen",
                              "decrease"="maroon",
                              "stable"="orange"))+
  coord_sf(crs=3857)+
  theme_void()+
  theme(legend.position = "none")

```
</details>

```{r echo=FALSE, message=FALSE, warning=FALSE, crop=T}
pl_combined
```

You can quickly change the background image by specifying a different `map_style` value, e.g. below 
I use the *decimal style*. And also combine with a different aesthetic, e.g. using fill to color the precincts according to their respective change.

<details closed>
<summary>Code: Gradient fill of precincts</summary>
```{r message=FALSE, warning=FALSE, crop=T}
df_fill <- precincts %>% 
  left_join(., df_change %>% select(sprengel, wber_diff_rel) %>% as.data.frame(),
            by=c("sprengel"="sprengel"))

pl_combined2<- ggplot()+
  layer_mapbox(background_image,
               map_style = mapbox_gallery_north_star(),
               mapbox_api_access_token = Sys.getenv("MAPBOX_ACCESS_TOKEN"),
               scale_ratio = 1)+
  geom_sf(data=df_fill,
          aes(fill=wber_diff_rel),
          size=0.2)+

  scale_fill_gradient2(low="red", high="green", mid="white",
                       breaks=scales::pretty_breaks(n=4)(min(df_fill$wber_diff_rel, na.rm = T):
                                                   abs(min(df_fill$wber_diff_rel, na.rm=T))),
                                                 guide = "colourbar",
                       name="% change eligible voters 2015-2020",
                       limits=c(min(df_fill$wber_diff_rel, na.rm=T), abs(min(df_fill$wber_diff_rel, na.rm=T))))+

  coord_sf(crs=3857)+
  theme_void()+
  theme(plot.background = element_rect(fill=plot_bg_color, color=plot_bg_color),
        legend.title = element_text(size=7, vjust=1),
        legend.text = element_text(size=5),
        legend.position = "bottom",
        legend.justification = "left")+
  guides(fill = guide_colourbar(barwidth = 3, barheight = .2, title.position = "top"))
```
</details>

```{r echo=FALSE, message=FALSE, warning=FALSE, crop=T, dev="CairoPNG", dpi=300, fig.align="left"}
pl_combined2
```

### 2021-01-06 about.Rmd
add snapbox post

---
title: "About this blog"
description: |
  Yet another blog...
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Not that the world would have been waiting for another blog… but there you go. This (work in progress?) blog is a collection of my ‘thoughts in R’ with varying degrees of seriousness. Posts will mostly relate to my substantive interests (comparative political science, human rights) or are simply notes to myself on a specific R function. Not claiming any final truth. One step at a time.

Roland Schmidt #IStandWithCEU #COYG

Codes are collected at [www.github.com/werkstattcodes](https://github.com/werkstattcodes){target="_blank"}
Tweets are at @zoowalk.

### 2021-01-06 _post/2020-10-10-vienna-elections-2020-age-profile/vienna-elections-2020-age-profile.Rmd
add post vienna election age profile

---
title: "Vienna Elections 2020: Age profile of electoral candidates"
description: |
  A short description of the post.
date: 10-10-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693

categories:
  - Austria
  - elections
  - OCR
  - regex
  - reactable
  - gt

preview: 2020-10-10-vienna-elections-2020-age-profile/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

# Setup
<details closed> 
<summary>Code: Load packages</summary> 
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(extrafont)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
hrbrthemes::update_geom_font_defaults(
  family = "Roboto Condensed",
  size = 3.5,
  color = "grey50"
)
library(scales)
library(knitr)
library(paletteer)
library(ggtext)
library(glue)
library(pdftools)
library(svglite)
library(tictoc)
library(tidytext)
library(gt)
library(reactable)
library(ggforce)
library(ggiraph)
library(htmltools)

```
</details>

<details closed> 
<summary>Code: Define rmarkdown options</summary> 
```{r setup, echo = T}
knit_hooks$set(wrap = function(before, options, envir) {
  if (before) {
    paste0("<", options$wrap, ">")
  } else {
    paste0("</", options$wrap, ">")
  }
})

knitr::opts_chunk$set(
 	fig.align = "left",
	message = FALSE,
	warning = FALSE,
 	dev = "svglite",
#	dev.args = list(type = "CairoPNG"),
	dpi = 300,
 	out.width = "100%"
)
options(width = 180, dplyr.width = 150)
```
</details>

<details closed> 
<summary>Code: Define plot theme, party colors, caption</summary> 
```{r echo=TRUE}

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()

caption_table <- "Source:\ndata: https://www.wien.gv.at/politik/wahlen/grbv/2020/ analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes"

fn_reactable_headings <- function(header, subtitle, table, caption) {
  div(class="reactable-table",
      div(
        class="reactable-title",
        header
        ),
      div(
        class="reactable-subtitle",
        subtitle
        ),
      table,
      div(class="reactable-caption",
      caption
      )
  )
  }

theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      #panel.border = element_rect(colour = plot_bg_color, fill=NA),
      #plot.border = element_rect(colour = plot_bg_color, fill=NA),
      plot.margin = margin(l = 0, 
                           t = 0.25,
                           unit = "cm"),
      plot.title = element_markdown(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 11
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 10
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(
        angle = 0,
        color = "grey50",
        hjust = 1
      ),
      axis.text.x = element_text(
        size = 9,
        color = "grey50"
      ),
      axis.title.y = element_blank(),
      axis.text.y = element_text(
        size = 9,
        color = "grey50"
      ),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 7
      ),
      legend.text = element_text(
        size = 7,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

data_date <- format(Sys.Date(), "%d %b %Y")

my_caption <- glue::glue("data: https://www.wien.gv.at/politik/wahlen/grbv/2020/\nanalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes")

```

# Context
Elections in Vienna are today and while glancing through the electoral lists I couldn’t help but paying attention to candidates’ birth years. Maybe that’s an age thing...  This got me thinking that I haven’t seen any more systematic analysis of parties’/candidates’ age profile. So as a modest contribution to this end, here are my two cents. Again, I'll focus mainly on the pertaining steps in R and related number crunching. Due to a lack of time and not being an expert on Vienna's electoral system, I'll be brief when it comes to substantive matters. But the presented results hopefully provide sufficient material to dig into.

As always, if you see any glaring error or have any constructive comment, feel free to let me now (best via twitter DM).

# Data
Again, as so often, the trickiest part is to get the data ‘liberated’ from the format it is provided in. The entire list of candidates is published in [this pdf](https://www.wien.gv.at/politik/wahlen/grbv/pdf/amtsblatt2020.pdf){target=”_blank”}. Note that there are three lists: One for the city council ('Gemeinderat'; composed on the basis of the results in 18 multi-member electoral districts), one for the 23 district councils ('Bezirksrat'; one in each district), and the ‘city election proposal’ ('Stadtwahlvorschlag'; admittedly a somewhat clumsy translation). The latter doesn't constitute a body in itself, but serves to allocate mandates which remained unassigned after counting the votes for the city council ('zweites Ermittlungsverfahren'/similar to the d'Hondt procedure).

When it comes to extracting the data from the linked pdf, a difficulty may arise due to the two-column format of the document. Hence, simple row-wise extraction doesn’t help much since it would put candidates together which could be from different parties. Similarly, simply isolating the two columns and extracting candidates would also not do the trick since breaks betwee parties, districts etc run over two, and not one column. To illustrate this, I resorted to cutting edge technology and drew the two arrows below:

![](/post/2020-10-10-vienna-elections-2020-age-profile_files/Inkedtext_flow_LI_Moment.jpg){width=50% height=50%}

Luckily, the [`tabulizer` package](https://docs.ropensci.org/tabulizer/){target="_blank"} is not only very powerful when it comes to extracting text/data from a pdf, it is also sophisticated enough to take into consideration the text flow highlighted above. I am not familiar with the underlying heuristic, but I assume it is contingent on consistently formatted section headings. Hence, empowered with this tool, retrieving the text becomes rather effortless. The subsequent steps are a battery of regular expressions to extract the specific data we are interested in. To see the code, unfold the snippets below.

<details open> 
<summary>Code: Extract text from pdf</summary> 
```{r eval=FALSE, include=T}
df_raw <- tabulizer::extract_text(file=here::here("_blog_data", 
                                                  "vienna_elections_2020",
                                                  "amtsblatt2020.pdf"),
                        pages=c(1:115),
                        encoding="UTF-8") %>%  #capital letters of UTF!
  enframe(name=NULL,
          value="text_raw")
```
</details>

<details open> 
<summary>Code: Extract data of interest (regex)</summary> 
```{r eval=FALSE, include=T}
df_clean <- df_raw %>% 
  ungroup() %>% 
  mutate(text_split=str_split(text_raw, regex("\r\n\\s*(?=\\d+\\.)"))) %>% 
  unnest_longer(text_split) %>% 
  mutate(text_split=text_split %>% str_squish() %>% str_trim()) %>% 
  mutate(text_split=str_split(text_split, ".(?=Zustellungsbevollmächtigte(r)? Vertreter(in)?)")) %>% 
  unnest_longer(text_split)

# get Listenplatz ------------------------------------------------------------

df_clean <- df_clean %>% 
  mutate(listenplatz=str_extract(text_split, regex("^\\d+\\.?\\s+(?!Bezirk)")) %>%
           str_extract(., "\\d*") %>% as.numeric())

# get elections -----------------------------------------------------------

df_clean <- df_clean %>%   
  mutate(election=text_raw %>% str_extract(., regex("(?<=[A-Z]\\.)\\s*[A-z]+wahl(en)?", dotall = T)) %>% 
           str_trim(., side=c("both"))) %>% 
  tidyr::fill(election, .direction="down")

# electoral district --------------------------------------------------------------
df_clean <- df_clean %>%   
    mutate(wahlkreis=case_when(election=="Bezirksvertretungswahlen" ~ str_extract(text_split, "\\d{1,2}\\. Bezirk"),
                             election=="Gemeinderatswahl"~ str_extract(text_split, regex("Wahlkreis.*?(?=[:upper:]{2,}?)",
                                                                                     dotall = T,
                                                                                     multiline = T)),
                             election=="Stadtwahl" ~ as.character("Stadtwahl"),
                             TRUE ~ as.character("missing"))) %>% 
  mutate(wahlkreis=str_trim(wahlkreis, side=c("both"))) %>% 
  tidyr::fill(wahlkreis, .direction="down") %>% 
  mutate(wahlkreis=str_remove(wahlkreis, "Wahlkreis ") %>% 
           str_remove(., regex("\\(.*\\)")) %>% 
           str_trim(., side=c("both")))

# other -------------------------------------------------------------------

df_clean <- df_clean %>% 
  mutate(page=text_raw %>% str_extract(., regex("Seite \\d+")) %>% str_extract(., "\\d+") %>% 
           as.numeric()) %>% 
  mutate(name=str_extract(text_split, regex("(?<=\\d\\.\\s?).*?(?=,\\s?\\d{4},)")) %>% 
           str_trim(., side=c("both"))) %>% 
  mutate(first_name=text_split %>% str_extract(., regex("[:alpha:]*(?=,\\s?\\d+)"))) %>% 
  mutate(year_birth=text_split %>% str_extract(., regex("\\d{4}")) %>% 
           as.numeric()) %>% 
  mutate(year_interval=cut(year_birth, seq(1930, 2005, 5))) %>% 
  mutate(plz=text_split %>% str_extract(., regex("\\d{4}\\s(?=Wien)")) %>% 
           str_trim(., side=c("both")))

# get party ---------------------------------------------------------------

df_clean <- df_clean %>%
  mutate(party=text_split %>% str_extract(., regex("(?<=^Zustellung)[:alpha:]*$"))) %>%
  mutate(party=case_when(lead(listenplatz==1) ~ str_extract(text_split, regex("\\w+$")),
                         TRUE ~ NA_character_)) %>% 
  tidyr::fill(party, .direction = "down") %>% 
  mutate(party=party %>% 
             as_factor() %>% 
             fct_relevel(., sort) %>% 
             fct_relevel(., "SPÖ", "FPÖ", "GRÜNE", "ÖVP", "NEOS"))

# wrap up -----------------------------------------------------------------

df_clean <- df_clean %>% 
  mutate(wahlkreis_plz=str_extract(wahlkreis, regex("\\d+")) %>% 
           as.numeric()+100) %>% 
  mutate(wahlkreis_plz=wahlkreis_plz %>% as.character() %>% paste0(., "0")) %>% 
  mutate(wahlkreis_plz=case_when(str_detect(wahlkreis, "Zentrum") ~ "1010, 1040, 1050, 1060",
                                 str_detect(wahlkreis, "Innen") ~ "1070, 1080, 1090",
                                 str_detect(wahlkreis, "Leopoldstadt") ~ "1020",
                                 str_detect(wahlkreis, "Landstraße") ~ "1030",
                                 str_detect(wahlkreis, "Favoriten") ~ "1100",
                                 str_detect(wahlkreis, "Simmering") ~ "1110",
                                 str_detect(wahlkreis, "Meidling") ~ "1120",
                                 str_detect(wahlkreis, "Hietzing") ~ "1130",
                                 str_detect(wahlkreis, "Penzing") ~ "1140",
                                 str_detect(wahlkreis, "Rudolf") ~ "1150",
                                 str_detect(wahlkreis, "Ottakring") ~ "1160",
                                 str_detect(wahlkreis, "Hernals") ~ "1170",
                                 str_detect(wahlkreis, "Währing") ~ "1180",
                                 str_detect(wahlkreis, "Döbling") ~ "1190",
                                 str_detect(wahlkreis, "Brigittenau") ~ "1200",
                                 str_detect(wahlkreis, "Floridsdorf") ~ "1210",
                                 str_detect(wahlkreis, "Donaustadt") ~ "1220",
                                 str_detect(wahlkreis, "Liesing") ~ "1230",
                                 TRUE ~ as.character(wahlkreis_plz))) %>% 
  mutate(residence=case_when(
    str_detect(wahlkreis_plz, plz) ~ "inside",
    !str_detect(wahlkreis_plz, plz) ~ "outside",
    TRUE ~ as.character("missing"))) #%>%

df_clean <- df_clean %>%   
  select(-text_raw) %>% 
  filter(!is.na(listenplatz))

```
</details>

```{r include=FALSE}
 # readr::write_excel_csv2(df_clean, path=here::here("_blog_data",
 #                                                  "vienna_elections_2020",
 #                                                  "df_clean.csv"))

df_clean <- readr::read_csv2(file=here::here("_blog_data",
                                             "vienna_elections_2020",
                                             "df_clean.csv")) %>% 
  mutate(party=party %>% 
             as_factor() %>% 
             fct_relevel(., sort) %>% 
             fct_relevel(., "SPÖ", "FPÖ", "GRÜNE", "ÖVP", "NEOS")) %>% 
  mutate(election=str_remove(election, regex("s?wahl"))) %>% 
  mutate(election=fct_relevel(election, "Gemeinderat", "Bezirksvertretungen", "Stadt"))
```

After these few steps we have a searchable/sortable table of all candidates (or better candidatures since one person can be candidate on multiple districts/lists). There have been `r df_clean %>% nrow() %>% scales::comma()` candidatures by `r df_clean %>% distinct(name) %>% nrow() %>% scales::comma()` individuals.

The table essentially provides all necessary data for the subsequent analysis. While the pdf does not include the exact birth date of each candidate, it provides us with their birth years which we can take as a proxy for age. Note that I also extracted candidates' residence zip code to see how often place of residence and candidature actually overlap (see below).

## Result

<details closed> 
<summary>Code: Table with all candidates</summary> 
```{r}

tb_all <- reactable(df_clean %>% 
            select(election, wahlkreis, party, name, listenplatz, year_birth, plz),
          columns=list(election=colDef(name="Wahl", width=130),
                       wahlkreis=colDef(name="Wahlbezirk", width=100),
                       party=colDef(name="Partei", width=50),
                       name=colDef(name="KandidatIn"),
                       listenplatz=colDef(name="Listenplatz", 
                                          width=70,
                                          align="center"),
                       year_birth=colDef(name="Geburtsjahr", width=90),
                       plz=colDef(name="PLZ Wohnort", width=90)),
          bordered=F,
          compact = TRUE,
          highlight = TRUE,
          style = list(fontSize = "10px"),
          filterable = TRUE,
          defaultPageSize = 23,
          theme = reactableTheme(
                borderColor = "#7f7f7f",
                borderWidth = 1,
            backgroundColor = "#f0eff0",
                filterInputStyle = list(
                  color="green",
                  backgroundColor = plot_bg_color)))

html_tb_all <- div(class="reactable-table",
  div(
    div(
     class="reactable-title", 
     "WIEN-WAHL 2020: Liste aller KandidatInnen"
     ),
    ),
  tb_all,
  div(class="reactable-caption",
      caption_table)
  )

```
</details>

```{r echo=FALSE}
html_tb_all
```

# Analysis

## Oldest and youngest candidates

```{r eval=FALSE, include=FALSE}
df_old<- df_clean %>% 
  slice_min(., order_by=year_birth, n=5) %>% 
  select(
    name,
    year_birth,
    party,
    election,
    wahlkreis,
    listenplatz) %>% 
  mutate(age_group="Ältesten")

df_young<- df_clean %>% 
  slice_max(., order_by=year_birth, n=5) %>% 
  select(
    name,
    year_birth,
    party,
    election,
    wahlkreis,
    listenplatz)%>% 
  mutate(age_group="Jüngsten")

df_old_young <- bind_rows(df_old, df_young) %>% 
  group_by(age_group) %>% 
  arrange(year_birth, name, .by_group=T) %>% 
  mutate(index=dplyr::min_rank(year_birth)) %>% 
  mutate(index_name=paste(index, ". ",name)) %>% 
  select(-index)

#table
tb_old_young <- df_old_young %>% 
  select(-name) %>% 
  gt(groupname_col = "age_group",  rowname_col  = "index_name") %>% 
  tab_header(title=md("**WIEN-WAHLEN 2020:<br>Ältesten u jüngsten KandidatInnen**"),
             subtitle = md("Top 5 nach Geburtsjahr")) %>% 
  gt::cols_label(year_birth="Geburtsjahr",
                 party="Partei",
                 election="Wahl",
                 wahlkreis="Wahlkreis",
                 listenplatz="Listenplatz") %>% 
  cols_width(
      vars(index_name) ~ px(100),
      vars(year_birth, party, wahlkreis, listenplatz) ~ px(50),
      vars(election) ~ px(100)
               ) %>% 
  tab_options(heading.align = "left",
                table.background.color = plot_bg_color,
                table.width=pct(100),
                table.font.size = "11px",
                row_group.font.weight = "bold",
                data_row.padding = px(0),
                table.align = "left")

tb_old_young

```

Let's now look at the overall youngest and oldest candidates. We could retrieve this information already from the main table provided above (sort column birth year). Here, however, let's nest candidates' different candidatures for the sake of clarity.

### Youngest candidates
<br>
<details closed> 
<summary>Code: Youngest candidates</summary> 
```{r}
df_main <- df_clean %>% 
  distinct(name, year_birth, party) %>% 
  slice_max(.,order_by=year_birth, n=10) %>% 
  arrange(name, desc(year_birth)) %>% 
  mutate(index=dplyr::min_rank(year_birth)) %>% 
  mutate(index_name=paste(index, ". ",name), .before=1) %>% 
  select(-index)

tb_young <- reactable(df_main,
                          columns=list(index_name=colDef(name="KandidatIn", 
                                                         width=130),
                                       name=colDef(show=F),
                                       year_birth=colDef(name="Geburtsjahr",
                                                         width=70),
                                       party=colDef(name="Partei",
                                                         width=50)),
                          pagination = FALSE,
                          onClick = "expand",
                          bordered=F,
                          compact = TRUE,
                          highlight = TRUE,
                          rowStyle = list(cursor = "pointer"),
                          style = list(fontSize = "10px"),
                          theme = reactableTheme(
                            borderWidth = 1,
                            borderColor = "#7f7f7f",
                            backgroundColor = plot_bg_color,
                            filterInputStyle = list(
                              color="green",
                              backgroundColor = plot_bg_color)),

            details=function(index){
              df_nested <- df_clean %>% 
                slice_max(.,order_by=year_birth, n=10) %>% 
                select(name, election, wahlkreis, listenplatz) %>% 
                filter(name==df_main$name[index]) %>% 
                select(-name)

              tbl_nested <- reactable(df_nested, 
                                      columns = list(
                                        election=colDef(name="Wahl"),
                                        wahlkreis=colDef(name="Wahlbezirk"),
                                        listenplatz=colDef(name="Listenplatz")
                                      ),
                                        outlined = TRUE, 
                                        highlight = TRUE, 
                                        fullWidth = TRUE,
                                      theme = reactableTheme(
                                        backgroundColor = "#ab8cab"))

              htmltools::div(style = list(margin = "12px 45px"), tbl_nested)}

            )

# html_tb_young <- div(
#   div(
#     div(
#      class="reactable-title", 
#      "WIEN-WAHL 2020: Jüngesten KandidatInnen"
#      ),
#     div(
#       class="reactable-subtitle",
#       "Geburtsjahr lt. Wahlvorschlag als Basis. Top 10."
#       ),
#     ),
#   tb_young_old)

html_tb_young <- fn_reactable_headings(header="WIEN-WAHL 2020: Jüngesten KandidatInnen", 
                                      subtitle="Geburtsjahr lt. Wahlvorschlag als Basis. Top 10.",
                                      table=tb_young,
                                      caption=caption_table)

```
</details>
```{r echo=FALSE}
html_tb_young
```

As becomes clear from the table, there are overall `r df_main %>% nrow()` candidates who were all born in `r df_main %>% distinct(year_birth) %>% pull()`.

### Oldest candidates
<br>
<details closed> 
<summary>Code: oldest candidates</summary> 
```{r}
df_main <- df_clean %>% 
  distinct(name, year_birth, party) %>% 
  slice_min(., order_by=year_birth, n=10) %>% 
  arrange(year_birth, name) %>% 
  mutate(index=dplyr::min_rank(year_birth)) %>% 
  mutate(index_name=paste(index, ". ", name), .before=1) %>% 
  select(-index)

tb_old <- reactable(df_main,
                          columns=list(index_name=colDef(name="KandidatIn", 
                                                         width=130),
                                       name=colDef(show=F),
                                       year_birth=colDef(name="Geburtsjahr",
                                                         width=70),
                                       party=colDef(name="Partei",
                                                         width=50)),
                          pagination = FALSE,
                          onClick = "expand",
                          bordered=F,
                          compact = TRUE,
                          highlight = TRUE,
                          rowStyle = list(cursor = "pointer"),
                          style = list(fontSize = "10px"),
                          theme = reactableTheme(
                            borderWidth = 1,
                            borderColor = "#7f7f7f",
                            backgroundColor = plot_bg_color,
                            filterInputStyle = list(
                              color="green",
                              backgroundColor = plot_bg_color)),

            details=function(index){
              df_nested <- df_clean %>% 
                slice_min(.,order_by=year_birth, n=10) %>% 
                select(name, election, wahlkreis, listenplatz) %>% 
                filter(name==df_main$name[index]) %>% 
                select(-name)

              tbl_nested <- reactable(df_nested, 
                                      columns = list(
                                        election=colDef(name="Wahl"),
                                        wahlkreis=colDef(name="Wahlbezirk"),
                                        listenplatz=colDef(name="Listenplatz")
                                      ),
                                        outlined = TRUE, 
                                        highlight = TRUE, 
                                        fullWidth = T,
                                      theme = reactableTheme(
                                        backgroundColor = "#ab8cab"))

              htmltools::div(style = list(margin = "12px 45px"), tbl_nested)}

            )

html_tb_old <- fn_reactable_headings(header="WIEN-WAHL 2020: Ältesten KandidatInnen",
                                     subtitle = "Geburtsjahr lt. Wahlvorschlag als Basis. Top 10.",
                                     table=tb_old,
                                     caption=caption_table)
```
</details>

```{r echo=FALSE}
html_tb_old

```
<br>

The oldest candidate is `r df_main  %>% slice_head(.,n=1) %>% pull(name)`, who was born in `r df_main %>% slice_head(., n=1) %>% pull(year_birth)`. Not bad.

## Avgerage birth year per election and party

Let's now look at the average year of birth of parties' candidates on each of the different electoral levels. The table below provides the median, mean and standard deviation for each party. The thin white line in the density plots on the right indicates the median.

<details closed> 
<summary>Code: Median age of lists</summary> 
```{r}
# summarize data (median, mean, sd)
df_list_age <- df_clean %>% 
  group_by(election, party) %>% 
  summarize(year_median=median(year_birth, na.rm = T),
         year_mean=mean(year_birth, na.rm=T),
         year_sd=sd(year_birth, na.rm=T)) %>% 
  group_by(election) %>% 
  arrange(desc(year_median), .by_group=T) #order as gt table

#create graphs for table

## define function creating plot
fn_plot <- function(data){
  data %>% 
    ggplot()+
    ggridges::geom_density_ridges(aes(x=year_birth,
                     y=0),
                 fill="firebrick",
                 color=plot_bg_color,
                 quantile_lines=T,
                 quantiles=2,
                 panel_scaling = F,
                 size=12)+
    scale_x_continuous(limits=c(min(df_clean$year_birth),
                                max(df_clean$year_birth)),
                       expand=expansion(mult=0))+
    scale_y_discrete(expand=expansion(mult=0))+
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      plot.margin = margin(0, unit="cm"),
      axis.text = element_blank(),
      axis.title = element_blank()
    )

  }

## apply function, dataframe with plots
box_plot <- df_clean %>% 
  select(election, party, year_birth) %>% 
  group_by(election, party) %>% 
  mutate(year_median=median(year_birth, na.rm = T)) %>% 
  #ungroup() %>% 
  nest(year_birth_nest=c(year_birth)) %>% 
  mutate(plot=map(year_birth_nest, fn_plot)) %>% 
  group_by(election) %>% 
  arrange(desc(year_median), .by_group=T) #order as gt table

#create gt table & insert df with plots; 
tb_list_age <- df_list_age %>% 
  mutate(year_mean=round(year_mean, digits = 2)) %>% 
  select(election, party, contains("year")) %>% 
  group_by(election) %>% 
  arrange(desc(year_median), .by_group=T) %>% #order as plots
  mutate(index=min_rank(-year_median), .before=1) %>% 
  mutate(index_party=paste0(index,". ", party)) %>% 
  ungroup() %>% 
  select(-index, -party) %>% 
  mutate(boxplot=NA) %>% 
  gt(groupname_col = "election",  rowname_col  = "index_party") %>% 
    tab_header(title=md("**WIEN-WAHLEN 2020:<br>Durchschnittliches Geburtsjahr der KandidatInnen**")) %>% 
  gt::cols_label(index_party="Partei",
                 year_median="Median",
                 year_mean="arith. Mittel",
                 year_sd="Std. Abw.",
                 boxplot="Verteilung") %>% 
  gt::fmt_number(columns=vars(year_sd),
                              decimals=2,
                  suffixing=F) %>% 
  gt::text_transform(
    locations=cells_body(vars(boxplot)),
    fn=function(x){
      map(box_plot$plot, ggplot_image, height=px(11), aspect_ratio=4)}
    ) %>% 
    cols_width(
      vars(boxplot) ~ px(200),
      vars(index_party) ~ px(100)
               ) %>% 
    tab_options(heading.align = "left",
                table.background.color = plot_bg_color,
                table.width=pct(100),
                table.font.size = "11px",
                row_group.font.weight = "bold",
                data_row.padding = px(0),
                table.align = "left") %>% 
   tab_footnote(
    footnote = "Vertical white line indicates median.",
    locations = cells_column_labels(
      columns = vars(boxplot))
  ) %>% 
   tab_source_note(
    source_note = my_caption)

# gtsave(tb_list_age, "tb_list_age.png", path = here::here("_blog_data"))

```
</details>

```{r echo=FALSE}
tb_list_age
```

I won't dig into every details, so only a few general remarks:
On the city council level ('Gemeinderat') the established parties (SPÖ, FPÖ, Greens, ÖVP, Neos) feature (on average) somewhat older candidates  than other, newer parties. On the district level ('Bezirksvertretungen'), the former are occupying more of a middle ground. Among them, Neos' candidates are on average younger, FPÖ candidates older. Funnily, when ranking parties according their median age, FPÖ and Team HC Strache (=former leader of FPÖ) are always adjacent.

```{r eval=FALSE, include=FALSE}
pl_list_median_age <- df_list_median_age %>%   
  ggplot()+
  labs(title="Wienwahl 2020: Durchschnittsalter per Liste",
       caption=my_caption
       )+
  scale_x_continuous(breaks=seq(1940, 2000, 10))+
  geom_jitter(aes(y=tidytext::reorder_within(party, by=year_median, within = election),
                   x=year_birth,
                   group=party,
                   color=party),
              alpha=0.1)+
  geom_boxplot_interactive(aes(y=tidytext::reorder_within(party, by=year_median, within = election),
                   x=year_birth,
                   group=party,
                   tooltip=glue::glue("Partei: {party},
                                      Median Geb.Jahr: {year_median}"),
                   color=party),
               fill=NA,
               outlier.shape = NA)+
  theme_post()+
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())+
  tidytext::scale_y_reordered()+
  ggforce::facet_col(vars(election),
             scales = "free_y",
             space = "free")
```
</details>

```{r eval=FALSE, fig.height=12, dev='svg', include=FALSE}
girafe(ggobj =pl_list_median_age,
                             options=list(
                               opts_hover_inv(css = "opacity:0.1;"),
                               opts_hover(css = "fill:#FFA500;color:#FFA500"),
                               opts_tooltip(css = "background-color:#323E4F;
                               color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000)
                               ),
       height_svg=6)

```
<br>

## Average birth year per electoral districts

Let's now take up a 'geographical perspective' and see how birth years are distributed within electoral districts and the difference between the latter.

```{r eval=FALSE, include=FALSE}
pl_district_age <- df_clean %>% 
  group_by(election, wahlkreis) %>% 
  mutate(year_median=median(year_birth, na.rm=T)) %>% 
  ungroup() %>% 
  ggplot()+
  labs(title="WIEN-WAHLEN 2020:<br>Durchschnittsalter der Kandidatinnen pro Wahlbezirk",
       subtitle="Geburtsjahr als proxy für Alter.",
       x="Geburtsjahr",
       caption=my_caption)+
  geom_boxplot_interactive(aes(y=reorder_within(wahlkreis,
                                    by=year_median,
                                    within=election),
               x=year_birth,
               group=wahlkreis,
              # color=wahlkreis,
               tooltip=glue::glue("Median Geb.Jahr: {year_median}"),
               data_id=paste0(election, wahlkreis)),
               fill=NA)+
  #tidytext::scale_y_reordered()+
  scale_y_discrete(labels=function(x) {stringr::str_remove(x, regex("___.*")) %>% 
      str_trunc(., width=10, side="right", ellipsis = "...")})+
  theme_post()+
  theme(legend.position="none",
      #  axis.text.y = element_markdown(size=3),
        panel.grid.major.y = element_blank()
        # axis.text.y = element_text_interactive(
        #        group=wahlkreis,
        #   data_id="axis.text.y",
        #  tooltip = "Description for y axis",
        #   hover_css = "fill:red;stroke:none;")
        )+
  facet_col(vars(election),
            scale="free_y",
            space="free")

giraph_options=list(opts_hover_inv(css = "opacity:0.5;"),
                    opts_hover(css = "fill:#FFA500;color:#FFA500"),
                    opts_tooltip(css = "background-color:#323E4F;
                               color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000)
                               )

giraph_height=7

girafe(ggobj=pl_district_age,
       options=giraph_options,
       height_svg=giraph_height)
```

<details closed> 
<summary>Code: birth years by electoral districts </summary> 
```{r}

df_pl <- df_clean %>% 
  group_by(election) %>% 
  mutate(year_median_election=median(year_birth, na.rm=T)) %>%
  group_by(election, wahlkreis) %>% 
  mutate(year_median_wahlkreis=median(year_birth, na.rm=T)) %>% 
  mutate(year_mean_wahlkreis=mean(year_birth, na.rm=T)) %>% 
  mutate(year_sd_wahlkreis=sd(year_birth, na.rm=T)) %>% 
  ungroup() %>% 
  mutate(wahlkreis_fct=fct_reorder(wahlkreis, year_median_wahlkreis))

ls_pl <- df_pl %>% 
  group_split(election) %>% 
    imap(~ggplot(data=.x)+
           labs(title=glue::glue("WIEN-WAHLEN 2020 <span style=color:firebrick>{unique(.x$election)}</span>:<br>Verteilung des Geburtsjahres von KandidatInnen pro Wahlbezirk"),
      subtitle="Median bei vertikaler weißer Linie.",
       x="Geburtsjahr",
       caption=my_caption)+
  ggridges::geom_density_ridges(aes(y=wahlkreis_fct,
                                     x=year_birth),
                                 fill="firebrick",
                                 scale=2,
                                rel_min_height = 0.01,
                                quantile_lines=T,
                                quantiles=2,
                                color="white")+
  geom_text(aes(y=length(unique(.x$wahlkreis)),
            x=2010),
            label="Median", 
            nudge_y=+1.5,
            size = 3,
            check_overlap = T,
            hjust=0)  +

  geom_text(aes(y=length(unique(.x$wahlkreis)),
            x=2020),
            label="Mean",
            check_overlap = T,
            nudge_y=+1.5,
            size = 3,
            hjust=0)  +

  geom_text(aes(y=length(unique(.x$wahlkreis)),
            x=2030),
            label="SD", 
            nudge_y=+1.5,
            size = 3,
            check_overlap = T,
            hjust=0)  +

  geom_text(aes(y=wahlkreis_fct,
                x=2010,
                label=glue::glue("{year_median_wahlkreis}")),
            check_overlap = T,
            size = 3,
            color = "grey50",
            hjust=0,
            nudge_y = 0.5
            )+
  geom_text(aes(y=wahlkreis_fct,
                x=2020,
                label=glue::glue("{round(year_mean_wahlkreis, digits=1)}")),
            check_overlap = T,
            size = 3,
            color = "grey50",
                        size=9,

            hjust=0,
            nudge_y = 0.5
            )+
          geom_text(aes(y=wahlkreis_fct,
                x=2030,
                label=glue::glue("{round(year_sd_wahlkreis, digits=2)}")),
            check_overlap = T,
            size = 3,
            color = "grey50",
            hjust=0,
            nudge_y = 0.5
            )+
  scale_y_discrete(labels=function(x) {stringr::str_remove(x, regex("___.*")) %>% 
      str_trunc(., width=10, side="right", ellipsis = "...")},
      expand=expansion(mult=c(0,.1)))+
  scale_x_continuous(limits=c(1925, 2030),
                     breaks=seq(1950, 2000, 10),
                     expand=expansion(mult=c(0,.10)))+
  theme_post()+
  theme(legend.position="none",
        axis.title.x=element_text(hjust=0.7),
        axis.text.y=element_text(vjust=0)))

```
</details>

```{r echo=FALSE, dev='svg'}
ls_pl[[1]]
```

There is quite a difference in the median value between Ottakring (the youngest, 1979) and Donaustadt (the oldest, 1970.5). But then again the median value is just an aggregate value and conveys only a bit of the wider picture. I think the distribution curves are more telling. Whether that's useful information or not, I am not entirely sure. At least it's something, I have not seen before.

<details closed> 
<summary>Plot: District Councils</summary> 
```{r echo=FALSE, dev='svg'}
ls_pl[[2]]

```
</details>

<details closed> 
<summary>Plot: City electoral proposal</summary> 
```{r echo=FALSE, dev='svg'}
ls_pl[[3]]
```
</details>
<br>

## Average birth year per electoral district and party

Now let's dig deeper and see for the distribution of birth years by electoral districts and parties, again for each election (city council, district councils).

### Graph
<br>
Hover with the mouse over the box plot and dots to obtain the pertaining data.
<details closed> 
<summary>Code: graph</summary> 
```{r}
df_age_list_district_graph <- df_clean %>% 
  filter(!str_detect(election, "Stadt")) %>% 
  group_by(election, wahlkreis, party) %>% 
  mutate(year_median=median(year_birth, na.rm=T),
    year_mean=mean(year_birth, na.rm=T),
    year_se=sd(year_birth, na.rm=T),
    values_nested=list(year_birth=year_birth)) %>% 
  group_by(election) %>% 
  mutate(median_election=median(year_birth, na.rm = T)) %>% 
  group_by(election, wahlkreis) %>% 
  mutate(median_wahlkreis=median(year_birth, na.rm=T)) %>% 
  ungroup() %>% 
  mutate(wahlkreis_num=str_extract(wahlkreis, regex("\\d+")) %>% as.numeric()) %>% 
  mutate(wahlkreis_fac=fct_reorder(wahlkreis, wahlkreis_num)) %>% 
  mutate(party=fct_rev(party))

ls_age_list_district_graph <- df_age_list_district_graph %>% 
  mutate(election=fct_relevel(election, "Gemeinderat", "Bezirksvertretungen")) %>%   group_split(election) %>% 
  imap(~ggplot(data=.x)+
         labs(title=glue::glue("WIEN-WAHL 2020: <span style=color:firebrick>{unique(.x$election)}</span><br>Verteilung des Geburtsjahres pro Wahlbezirk und Partei"),
              #subtitle="Geburtsjahr als proxy für Alter.",
              x="Geburtsjahr",
              caption=my_caption)+
  geom_jitter_interactive(aes(y=tidytext::reorder_within(
    x=party,
    by=year_median,
    within=wahlkreis_fac
  ),
                              x=year_birth,
                              tooltip=glue::glue("{name}
                                                 {party}
                                                 {year_birth}"),
                              data_id=paste(name)),
              color="grey50",
              size=.5,
              alpha=0.5)+
  geom_boxplot_interactive(aes(
    y=tidytext::reorder_within(
    x=party,
    by=year_median,
    within=wahlkreis_fac
  ),
                               x=year_birth,
                               tooltip=glue::glue("{party}
                                                  Median: {year_median}
                                                  arith Mittelw.: {round(year_mean, 2)}
                                                  Std.Abw.: {round(year_se,2)}"),
                               data_id=paste(wahlkreis, party)),
               outlier.shape = NA,
               size=.2,
               color="#374E55FF",
               fill=NA)+
  geom_vline(aes(xintercept=median_election,
                 color="color_vline",
                 group=wahlkreis,
                 name=""),
             #key_glyph=c("vline"),
             linetype="dotted"
             )+
  scale_color_manual(labels=c(color_vline="median of all candidates"),
                     values=c(color_vline="firebrick"))+
  scale_x_continuous(labels=function(x) str_sub(x, start=3),
                     breaks=seq(1940, 2000, 20))+
  tidytext::scale_y_reordered()+
  facet_wrap(vars(wahlkreis_fac),
             scales="free",
             ncol=4,
             labeller=as_labeller(function(x) str_trunc(x, 
                                                        width=15, 
                                                        side=c("right"),
                                                        ellipsis="...")
                                  )
             )+
  theme_post()+
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_markdown(size=6),
    axis.text.x = element_markdown(size=6),
    axis.title.x = element_markdown(size=6),
    legend.position = "top",
    legend.justification = "left",
    legend.title=element_blank()
  ))

```
</details>
<br>

```{r echo=FALSE}
giraph_options=list(# opts_hover_inv(css = "opacity:0.1;"),
                               opts_hover(css = "fill:#FFA500;color:#FFA500"),
                               opts_tooltip(css = "background-color:#323E4F;
                               color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000)
                               )
giraph_height=9

girafe(ggobj=ls_age_list_district_graph[[1]],
       options=giraph_options,
       height_svg=giraph_height)
```

<br>
<details closed> 
<summary>Graph: Bezirksvertretung</summary> 
```{r echo=FALSE}
girafe(ggobj=ls_age_list_district_graph[[2]],
       options=giraph_options,
       height_svg=giraph_height)

```
</details>

<br>

### Table

Below essentially the same information as provided in the graph above but in a tabular form.

<details closed> 
<summary>Code: Avg. birth year of lists by district</summary> 
```{r}

df_age_list_district <- df_clean %>% 
  #filter(!str_detect(election, "Stadt")) %>% 
  group_by(election, wahlkreis, party) %>% 
  summarize(year_median=median(year_birth, na.rm=T),
    year_mean=mean(year_birth, na.rm=T),
    year_se=sd(year_birth, na.rm=T),
    values_nested=list(year_birth=year_birth)) %>% 
  ungroup() %>% 
  mutate(values_df=map(values_nested, as_tibble_col, column_name ="birth_year"))

tb_age_list_district <- df_age_list_district %>% 
  reactable(.,
            columns=list(election=colDef(name="Wahl"),
                         wahlkreis=colDef(name="Wahlkreis"),
                         party=colDef(name="Partei",
                                      width=70),
                         year_median=colDef(name="Median",
                                            format=colFormat(digits=1)),
                         year_mean=colDef(name="arith. Mittel", 
                                          format=colFormat(digits=1)),
                         values_nested=colDef(show = F),
                         values_df=colDef(show = F),
                         year_se=colDef(name="Std.Abw.",
                                        format=colFormat(digits=1))),
            bordered=F,
            compact = TRUE,
            highlight = TRUE,
            style = list(fontSize = "10px"),
            filterable = TRUE,
            theme = reactableTheme(
                            borderWidth = 1,
                            borderColor = "#7f7f7f",
                            backgroundColor = plot_bg_color,
                            filterInputStyle = list(
                              color="green",
                              backgroundColor = plot_bg_color))
            )

html_age_list_district <- fn_reactable_headings(header ="WIEN-WAHL 2020: Durchschnittliches Alter pro Wahlbezirk", subtitle = "Geburtsjahr lt. Wahlvorschlag als Basis.",
                                                table=tb_age_list_district,
                                                caption=caption_table)

```
</details>
```{r echo=FALSE}
html_age_list_district
```
<br>

## Birth year and position on electoral list

Another question which came to my mind: How is a candidate's birth year/age related to his or her position on the electoral list. Are more senior candidates more likely to be on more attractive, i.e. top positions? Or is youth related to competitive positions? Are there any differences between the parties or elections (city vs district level) when it comes to this relation? 
Hover with the mouse over the individual dots to obtain data on indivdual candidates.

<details closed> 
<summary>Code: Age and position on electoral list</summary> 
```{r}
ls_pl_age_rank_district <- df_clean %>% 
  group_split(election) %>% 
  imap(~ggplot(data=.x)+
  labs(
    title=glue::glue("WIEN-WAHL 2020: <span style  = color:firebrick>{unique(.x$election)}</span><br>Listenplatz und Geburtsjahr"),
    subtitle="Loess regression.",
       x="Geburtsjahr",
       y="Listenplatz",
       caption=my_caption)+
  geom_jitter_interactive(aes(x=year_birth,
                              y=listenplatz,
                              tooltip=glue::glue("{name}
                                                 Geb.Jahr: {year_birth}
                                                 Listenplatz: {listenplatz}
                                                 Wahlkreis: {wahlkreis}
                                                 ")),
              color="grey50",
              alpha=0.2)+
  geom_smooth(aes(x=year_birth,
                  y=listenplatz),
                  method='loess', 
              formula= y~x,
              size=0.5,
              color="firebrick")+
    scale_x_continuous(labels=function(x) str_sub(x, start=3),
                     breaks=seq(1940, 2000, 20))+
  facet_wrap(vars(party))+
  theme_post()+
  theme(axis.title.x=element_text(size=6),
        axis.title.y=element_text(
          angle=90,
          size=6,
          color="grey50",
          hjust=1),
        axis.text.y = element_markdown(size=6),
        axis.text.x = element_markdown(size=6)))

```
</details>

<details open> 
<summary>Plot: City Council - Age and position on electoral list</summary> 
```{r echo=FALSE}
giraph_options=list(# opts_hover_inv(css = "opacity:0.1;"),
                               opts_hover(css = "fill:#FFA500;color:#FFA500"),
                               opts_tooltip(css = "background-color:#323E4F;
                               color: white;
                               font-family: Roboto;",
                               offx = 30, offy = -30,
                               delay_mouseout = 5000)
                               )
giraph_height=4

girafe(ggobj=ls_pl_age_rank_district[[1]],
       options=giraph_options,
       height_svg=giraph_height)
```
</details>

Interestingly, there are some subtle differences between the regression lines' slope. Note how it increases for higher birth years for the SPÖ and the FPÖ. In other words, younger candidates tend to be on higher (less attractive) positions on the electoral list for these two parties. As for the other major parties, there is no comparable relation, at least not as visible as in the two previous cases. Obviously, this is purely descriptive.

To see the results for the other electoral levels unfold the sections below.

<details closed> 
<summary>Plot: District Councils - Age and position on electoral list</summary> 
```{r echo=FALSE}
girafe(ggobj=ls_pl_age_rank_district[[2]],
       options=giraph_options,
       height_svg=6)
```
</details>

<details closed> 
<summary>Plot: City - Age and position on electoral list</summary> 
```{r echo=FALSE}
girafe(ggobj=ls_pl_age_rank_district[[3]],
       options=giraph_options,
       height_svg=giraph_height)
```
</details>
<br>

## Non-resident candidates

When extracting data from the pdf document with candidates’ details, I also retrieved the zip code of their places of residence. The idea was to see whether candidates actually live in the electoral districts where they are running for office. For the lack of a better term, I call these candidates ‘non-resident candidates.’

I could imagine that ‘non-residency’ conflates quite a number of factors, and the reasons and consequences of ‘non-residency’ are likely to be quite complex. A possible hypotheses could be that parties with high number of non-resident candidates are unevenly institutionalized across the city and hence have to bring in members from outside of the electoral district. But again, this is purely speculative and I haven’t thought, let alone read about it systematically.  Nevertheless,  as some kind of tentative inquiry, I thought it’s worth looking at the numbers. 
Three specific questions came to my mind: 1) Are there parties which feature particularly frequently such non-resident candidates? 2) Are there districts where non-resident candidates are particularly frequently running for office? 3) Do the shares of non-resident candidates differ between elections for the city council and for the district councils? I would think that the relation between a candidate and her electoral district, i.e. place of residence, is more of a significance in the latter case.

### Share per party
<br>
<details closed> 
<summary>Code: Out-of-residence candidates per party and election</summary> 
```{r}
#party with most out-of-residence candidates
df_residence_party <- df_clean %>% 
  filter(!str_detect(election, "Stadt")) %>% 
  group_by(election, party, residence) %>% 
  summarize(n_residence=n()) %>% 
  mutate(rel_residence=n_residence/sum(n_residence, na.rm = T)) %>% 
  filter(residence=="outside") %>% 
  arrange(desc(rel_residence)) %>% 
  ungroup()

#fn for barchart
bar_chart <- function(label, width = "100%", height = "16px", fill = "#00bfc4", background = NULL) {
  bar <- div(style = list(background = fill, width = width, height = height))
  chart <- div(style = list(flexGrow = 1, marginRight = "8px", background = background), bar)
  div(style = list(display = "flex", alignItems = "right"), chart, label)
}

#assemble table
tb_residence_party <- df_residence_party %>% 
  select(party, everything()) %>% 
  reactable(.,
            columns = list(
              party=colDef(name="Party"),
              election=colDef(name="Election"),
              residence=colDef(show=FALSE),
              n_residence = colDef(name = "Number",
                                   align = "left", 
                                   cell = function(value) {
    width <- paste0(value / max(df_residence_party$n_residence) * 100, "%")
    value <- format(value, trim=F, width=3, justify = "right")
    bar_chart(value, width = width, fill = "#00bfc4")}),

              rel_residence = colDef(name = "Share",
                                   align = "left",
                                   cell = function(value) {
    width <- paste0(value * 100, "%")
    value <- scales::percent(value)
    bar_chart(value, width = width, fill = "orange", background = "#e1e1e1")})),
    bordered=F,
    compact = TRUE,
    highlight = TRUE,
    style = list(fontSize = "10px"),
    filterable = TRUE,
    theme = reactableTheme(
      borderColor = "#7f7f7f",
      borderWidth = 1,
      backgroundColor = "#f0eff0",
      filterInputStyle = list(
                  color="green",
                  backgroundColor = plot_bg_color)
            ))

tb_residence_party <- fn_reactable_headings(header = "Parties' number and share of non-resident candidates per election",
                                            subtitle = "Non-resident candidates: Candidates not residing in their electoral district according to their residency zip code.",
                                            table=tb_residence_party,
                                            caption=caption_table)

```
</details>

```{r echo=FALSE}
tb_residence_party
```

### Share per electoral districts
<br>
<details closed> 
<summary>Code: Share of out-of-residence candidates per district</summary> 
```{r}
#share in districts
df_residence_district <- df_clean %>% 
  filter(!str_detect(election, "Stadt")) %>% 
  group_by(election, wahlkreis, residence) %>% 
  summarize(n_residence=n()) %>% 
  mutate(rel_residence=n_residence/sum(n_residence, na.rm = T)) %>% 
  filter(residence=="outside") %>% 
  ungroup() %>% 
  arrange(election, desc(rel_residence))

#fn for barchart
bar_chart <- function(label, width = "100%", height = "16px", fill = "#00bfc4", background = NULL) {
  bar <- div(style = list(background = fill, width = width, height = height))
  chart <- div(style = list(flexGrow = 1, marginRight = "8px", background = background), bar)
  div(style = list(display = "flex", alignItems = "right"), chart, label)
}

#assemble table
tb_residence_district <- df_residence_district %>% 
  reactable(.,
            columns = list(
              election=colDef(name="Election"),
              wahlkreis=colDef(name="Electoral district"),
              residence=colDef(show=FALSE),
              n_residence = colDef(name = "Number",
                                   align = "left", 
                                   cell = function(value) {
    width <- paste0(value / max(df_residence_district$n_residence) * 100, "%")
    value <- format(value, trim=F, width=3, justify = "right")
    bar_chart(value, width = width, fill = "#00bfc4")}),

              rel_residence = colDef(name = "Share",
                                   align = "left",
                                   cell = function(value) {
    # width <- paste0(value / max(df_residence_district$rel_residence) * 100, "%")
    width <- paste0(value * 100, "%")
    value <- scales::percent(value)
    bar_chart(value, width = width, fill = "orange", background = "#e1e1e1")})),
    bordered=F,
    compact = TRUE,
    highlight = TRUE,
    style = list(fontSize = "10px"),
    filterable = TRUE,
    theme = reactableTheme(
      borderColor = "#7f7f7f",
      borderWidth = 1,
      backgroundColor = "#f0eff0",
      filterInputStyle = list(
                  color="green",
                  backgroundColor = plot_bg_color)
            ))

tb_residence_district <- fn_reactable_headings(header = "Share of non-resident candidates per electoral district", subtitle="",
                                               table=tb_residence_district,
                                               caption=caption_table)

```
</details>

```{r echo=FALSE}
tb_residence_district
```
<br>

### City council vs district councils

Finally, let's contrast parties' non-resident candidate ratios for the election to the city council and the district councils.

<details closed> 
<summary>Code: City vs district councils </summary> 
```{r message=FALSE, warning=FALSE}
diff_share_non_residency <- df_clean %>% 
  filter(!str_detect(election, "Stadt")) %>% 
  group_by(election, party, residence) %>% 
  summarize(n_residence=n()) %>% 
  mutate(rel_residence=n_residence/sum(n_residence, na.rm = T)) %>% 
  filter(residence=="outside") %>% 
  ungroup() %>% 
  group_by(party) %>% 
  mutate(n_elections=n()) %>% 
  filter(n_elections==2)

diff_share_non_residency_wide <- diff_share_non_residency %>% 
  pivot_wider(id_cols=party, names_from=election,
              values_from=rel_residence) %>% 
  mutate(city_larger_district=case_when(Gemeinderat > Bezirksvertretungen ~ "GR größer",
                                        Gemeinderat <  Bezirksvertretungen ~ "GR kleiner"))
```

```{r}

pl_diff_share_non_residency<- diff_share_non_residency %>% 
  drop_na() %>% 
  ggplot()+
  labs(title=glue::glue("WIEN-WAHL 2020: Share of non-resident candidates per party.<br><span style=color:{paletteer_d('ggsci::default_jama')[2]}>City council (Gemeinderat)</span> vs <span style=color:{paletteer_d('ggsci::default_jama')[3]}>district councils (Bezirksräte)</span>"),
  subtitle="Non-resident candidates: Candidates with different residence zip-code than their electoral district",
  x="Share of non-resident candiates",
  caption=my_caption)+
  geom_segment(data=diff_share_non_residency_wide,
               aes(y=reorder(party, -Gemeinderat),
                   yend=reorder(party, -Gemeinderat),
                 x=Gemeinderat,
                 xend=Bezirksvertretungen),
               color="grey50")+
  geom_point(aes(y=reorder(party, -rel_residence),
                 x=rel_residence,
                 shape=election,
                 color=election),
             size=2)+
  scale_color_manual(values=c("Gemeinderat"=paletteer_d("ggsci::default_jama")[c(2)],
                          "Bezirksvertretungen"=paletteer_d("ggsci::default_jama")[c(3)]),
                          label=c("Gemeinderat"="City council",
                                  "Bezirksvertretungen"="District councils"))+
  scale_shape(guide="none")+
  scale_x_percent()+
  theme_post()+
  theme(legend.position = "none",
        legend.direction = "horizontal",
        legend.title=element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line())
```
</details>

```{r echo=FALSE, dev='svg'}
pl_diff_share_non_residency
```

The graph shows that almost all parties feature a smaller share of non-resident candidates in the elections to the city council than in the elections to the district councils. This difference is particularly strong for the FPÖ and for Team HC Strache. However, there are two notable exceptions. On is WIFF - Wir für Floridsdorf, a party with a clear programmatic focus for one specific district. Hence, it doesn't come as a surprise that the share of non-resident candidates for the district council is low. The second exceptions are the Greens. I am second guessing here, but I could imagine that this is indicative for a strong 'grass-root', locally based party organization and reflecting the engagement of candidates who actually live in their electoral districts. I find the graph in any case quite noteworthy.

# Wrap-up

So that's it for now. Again, this has been largely about retrieving and crunching numbers, but I guess some bits are worth entertaining more substantively..

### 2021-01-06 _post/2020-11-18-recreating-nyt-election-map/recreating-nyt-election-map.Rmd
add post vienna election age profile

---
title: "Recreating NYT election map"
description: |
  How to re-create election maps by the NYT with ggplot
date: 11-18-2020

draft: TRUE

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - Corona virus
  - OCR
  - stringr
  - web scraping

preview: recreating-nyt-election-map_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

# CONTEXT

While following the NYT's coverage of the US presidential elections, I became once again intrigued by their data visualizations. One type of visualization which I found particularly remarkable was the map with arrows indicting how far a state's counties shifted to democrats or republicans. See [here](https://www.nytimes.com/interactive/2020/11/07/us/elections/pennsylvania-counties-battleground-state.html){target="_blank"} on the NYT's website.

![](index_files/Capture.PNG)

This prompted me to give it a try with `ggplot` and see how far I could replicate the map. Below the relevant steps. Thanks also to the respondents of my pertaining [tweet](https://twitter.com/zoowalk/status/1325486039687782400){target="_blank"}.

I hope the code speaks largely for itself. Hence, only a few additional comments.

# SETUP

<details closed>

<summary>Load packages</summary>

Load the relevant packages.

```{r message=FALSE, warning=FALSE}
library(jsonlite)
library(ggtext)
library(extrafont)
loadfonts(device = "win", quiet = T)
library(sf)
library(patchwork)
library(knitr)
library(svglite)
library(tidyverse)

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()
```

<details closed>

<summary>Code: Define rmarkdown options</summary>

```{r setup, echo = T}
knit_hooks$set(wrap = function(before, options, envir) {
  if (before) {
    paste0("<", options$wrap, ">")
  } else {
    paste0("</", options$wrap, ">")
  }
})

knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)

knitr::opts_chunk$set(
 	fig.align = "left",
	message = FALSE,
	warning = FALSE,
 	dev = "svglite",
#	dev.args = list(type = "CairoPNG"),
	dpi = 300,
 	out.width = "100%"
)
options(width = 180, dplyr.width = 150)
```

</details>

# CREATE MAP

## Get electoral results

<details open>

<summary>Code: Get election results </summary>

```{r message=FALSE, warning=FALSE}

# 2020 results ------------------------------------------------------------

results <- "https://www.nbcnews.com/politics/2020-elections/pennsylvania-president-results?format=json"
df_results <- jsonlite::fromJSON(results, flatten=T)

Biden <- df_results$candidateCountyResults$countyResults[[1]]
Biden <- Biden %>%
  mutate(candidate="Biden",
         party="democrat")

Trump <- df_results$candidateCountyResults$countyResults[[2]]
Trump <- Trump %>%
  mutate(candidate="Trump",
         party="republican")

df_2020 <- bind_rows(Biden, Trump) %>% 
  mutate(year=2020) %>% 
  select(year, county=countyName, party, candidate, vote_share=percentOfVote,
         vote_absolute=votesReceived,
         -isLeading, county=countyName)

# 2016 results ------------------------------------------------------------
# source: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ#
df_2016 <- readxl::read_xlsx(here::here("_blog_data", "nyt_map","results2016.xlsx")) %>% 
  select(year, state, county, candidate, party, contains("votes")) %>% 
  filter(year==2016 & state=="Pennsylvania") %>% 
  group_by(county) %>% 
  mutate(vote_share=candidatevotes/totalvotes*100) %>% 
  select(year, county, party, candidate, vote_absolute=candidatevotes, vote_share) %>% 
  ungroup()

# combine -----------------------------------------------------------------

df_combined <- bind_rows(df_2016, df_2020) %>% 
  filter(party %in% c("democrat", "republican")) %>% 
  ungroup()

# Since there are also other candidates than Trump and Biden (yes, really), its not ok to simply use the difference between their 2020 and 2016 results. There are counties in which both candidates increased their vote share. Hence, whether a county became more democrat or republican has to be calculated on the basis of their joint votes (share).

#change
df_all <- df_combined %>% 
  group_by(county, party) %>% 
  arrange(desc(year), .by_group=T) %>% 
  mutate(change=vote_share-lead(vote_share)) %>% 
  group_by(year, county) %>% 
  mutate(vote_share_combined=sum(vote_share)) %>% 
  ungroup() %>% 
  mutate(vote_share_party=vote_share/vote_share_combined*100) %>% 
  group_by(county, party) %>% 
  arrange(desc(year), .by_group = T) %>% 
  mutate(nchange_parties=vote_share_party-lead(vote_share_party),
         change_vote_absolute=vote_absolute-lead(vote_absolute)) %>% 
  ungroup()
```

```{r}

df_change <- df_combined %>% 
  pivot_longer(cols=contains("vote"),
               names_to = "indicator") %>% 
  group_by(county, party, indicator) %>% 
  arrange(desc(year), .by_group = T) %>% 
  mutate(change=value-lead(value) %>% 
           round(., digits = 2)) %>% 
  ungroup()

df_change <- df_change %>% 
  filter(year=="2020") %>% 
  group_by(county, indicator) %>% 
  arrange(desc(party), .by_group = T) %>% 
  summarise(party_diff=change-lead(change)) %>% 
  filter(!is.na(party_diff)) %>% 
  pivot_wider(id_cols = "county",
              values_from = "party_diff",
              names_from="indicator",
              names_glue="diff_change_{indicator}") %>% 
  mutate(party=case_when(diff_change_vote_share < 0 ~ "democrat",
                         diff_change_vote_share > 0 ~ "republican"))

```

</details>

## Create map

<details open>

<summary>Code: Create maps</summary>

```{r message=FALSE, warning=FALSE}
#import shapefile
#https://data-pennshare.opendata.arcgis.com/datasets/pennsylvania-county-boundaries
pa <- st_read(here::here("_blog_data", "nyt_map","shapefile",
                         "county_boundary.shp"),
              quiet=T) %>% 
  st_transform(32617)

#extract x and y positions from the geometry column
sf_cent <- st_centroid(pa) %>% 
  mutate(point_x=map_dbl(geometry, pluck, 1),
         point_y=map_dbl(geometry, pluck, 2))
```

</details>

Calculate center of each county for the starting position of the arrows (=centroids for each polygon) as well as the length of the arrows. The arrows of in the NYT's plot have an angle of 20 degrees (e.g. the measurement tool in Gimp allows you to get the angle). With this angle and the desired length (=difference of % changes of republicans and democrats) we can calculate the horizontal and vertical distance of the arrows' end from their starting position. The cos and sin functions are  our friends here (remember from high school..., if not see [here](https://en.wikipedia.org/wiki/Trigonometric_functions){target="_blank"}).

<details open>
<summary>Code: Calculate centroids</summary>

```{r message=FALSE, warning=FALSE}
# data --------------------------------------------------------------------

scale_factor <- 10^3.65

sf_cent2 <- sf_cent %>% 
  # left_join(., df_change %>% select(county, party, change_parties, change_vote_absolute) %>% 
  left_join(., df_change %>% 
              # select(county, party, change_parties) %>% 
              mutate(county=stringr::str_to_upper(county)),
            by=c("COUNTY_NAM"="county")) %>% 
  mutate(x_change=diff_change_vote_share*sin(20),
         y_change=abs(diff_change_vote_share)*cos(20)) %>% 
  mutate(point_x_end=point_x+(x_change*scale_factor),
         point_y_end=(y_change*scale_factor)+point_y)
```
</details>

Use the `googleway` package to get the long/lat details for Philly and Pittsburgh (see [here](https://werk.statt.codes/post/primary-schools-in-vienna/){target="_blank"} for another blog where I used the package quite heavily.)

<details open>
<summary>Code: Geocode cities</summary>
```{r message=FALSE, warning=FALSE}
# data for pittsburgh and philadelphia ------------------------------------

library(googleway)
cities <- c("Philadelphia", "Pittsburgh")
df_cities <- map_dfr(cities, ~google_geocode(address=.x, 
                                             key=Sys.getenv("googleway"),
                                             simplify = T))

df_cities_geom <- df_cities$results$geometry$location

df_cities_add <- df_cities$results$formatted_address %>% enframe(., name=NULL, value="city")

df_cities <- bind_cols(df_cities_geom, df_cities_add)

sf_cities <- st_as_sf(df_cities, coords = c("lng", "lat"), remove = FALSE, 
         crs = 4326, agr = "constant") %>% 
  st_transform(32617)
```
</details>

### Main map

Create the main plot with the map.

<details open>
<summary>Code: map</summary>
```{r message=FALSE, warning=FALSE}
# plot --------------------------------------------------------------------

pl_map <- ggplot()+
  labs(
    # title = 'The Vote in Battleground States<br>
    #    <span style="color:#606060;font-size:35pt;font-family:Times;">Pennsylvania</span>',
    caption="graph remake: Roland Schmidt|@zoowalk|https://werk.statt.codes\noriginal: https://nyti.ms/3lftRDe"
  )+
  geom_sf(data=pa, 
          size=.1,
          fill="white")+
  geom_sf(data = sf_cities, 
          size=1.5,
          color = 'grey50') +
  geom_sf_text(data = sf_cities,
                aes(label=city %>% 
                      str_extract(., regex("[:alpha:]*"))),
               nudge_y = -10^4.2,
               size=3.2,
               family="Arial",
               color="grey50")+
  geom_segment(data=sf_cent2,
               aes(x=point_x,
                   xend=point_x_end,
                   y=point_y,
                   yend=point_y_end,
                   size=abs(diff_change_vote_absolute),
                   color=party),
               linejoin="mitre",
               arrow=arrow(length=unit(0.2, "cm"),
                           angle=15,
                           type="closed"))+
  scale_x_continuous(expand=expansion(mult=c(.1,.3)))+
  scale_size_continuous(range=c(0,3))+
  scale_color_manual(values=c(democrat="#3f95b0", republican="#ba6b67"))+
  theme_void()+
  theme(plot.background = element_rect(fill=plot_bg_color),
   plot.title = element_markdown(hjust = .5,
                                 lineheight = 3,
                                 color="#606060"
                                 ),
   plot.margin = margin(0,0,0,0, unit="cm"),
   plot.caption = element_text(hjust=0, 
                               family = "Arial",
                               lineheight = 1,
                               color="grey60"),
        legend.position = "none")

```
</details>

```{r echo=FALSE, message=FALSE, warning=FALSE, dev.args=list(bg=plot_bg_color)}
pl_map
```

### Legend

Create a second plot which shows the legend in the graph.

<details open>
<summary>Code: legend inset</summary>
```{r message=FALSE, warning=FALSE}
# inset -------------------------------------------------------------------

df_inset <- data.frame(
  stringsAsFactors = FALSE,
             party = c("republican", "democrat"),
            x = c(.5L, -.5L),
              xend = c(3L, -3L),
            y = c(0L, 0L),
              yend = c(.5L, .5L)
)

pl_inset <-  df_inset %>% 
ggplot()+
  labs(title="Shift from 2016",
       subtitle="In counties that have reported\n almost all of their votes")+
  geom_segment(aes(x=x,
                   xend=xend,
                   y=y,
                   yend=yend,
                   color=party),
               size=1,
               linejoin="mitre",
               arrow=arrow(length=unit(0.1, "cm"),
                           angle=20,
                           type="closed")
               )+
  geom_text(x=.5, y=-.3, 
            label="More\nRepublican", 
            hjust=0, 
            lineheight=0.8,
            size=3,
            color="grey60", 
            family="Arial")+
  geom_text(x=-.5, y=-.3, label="More\nDemocratic", 
            lineheight=0.8,
            hjust=1, 
            size=3,
            color="grey60", 
            family="Arial")+
  scale_color_manual(values=c(democrat="#3f95b0", republican="#ba6b67"))+
  scale_x_continuous(expand=expansion(ad=c(1,1)), limits=c(-7,7))+
  scale_y_continuous(limits=c(-1,.5), expand=expansion(add=c(0.5, 0.2)))+
  theme_void()+
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5,
                                  family = "Arial",
                                  face="bold",
                                  color="#606060",
                                  margin = margin(b=0, unit="cm"),
                                  size=13),
        plot.subtitle = element_text(hjust=0.5,
                                     margin = margin(t=0.1, unit="cm"),
                                     size=10,
                                     color="grey60",
                                     family = "Arial"))
```
</details>

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, dev.args=list(bg=plot_bg_color)}
pl_inset

```

### Result

Insert the 'legend plot' as an inset by using the latest release of the `patchwork` package (see details [here](https://www.data-imaginist.com/2020/insetting-a-new-patchwork-version/){target="_blank"}).

<details open>
<summary>Code: Combine plots</summary>
```{r message=FALSE, warning=FALSE}
# combine plots -----------------------------------------------------------

pl_combined <- pl_map+inset_element(pl_inset, top=0.9, bottom=0.6, left=0.75, right=1, align_to="full")

```
</details>

Here's the result.

```{r echo=FALSE, fig.asp=.6, fig.width=9, message=FALSE, warning=FALSE, dev.args=list(bg=plot_bg_color), out.height="110%"}
pl_combined

```

And here the original for comparison:

![](index_files/Capture.PNG)

# PENDING DETAILS

For my liking, the plot comes pretty close to the one of the NYT, but a few details are still off. 
*Width of arrows*:  I assumed that the width would be weighted by the absolute number of votes, but apparently that's not what the NYT used. Without any further details available (which I think should be actually mentioned on their site), I can only guess here. *Length of arrows*: There are also a few counties in which the length doesn't really mirror the arrows in the NYT's plot. I am not entirely clear why this is. I strongl assume that's due to differences in the underlying election results, meaning different points in time.

On a purely aesthetic front, I couldn't find a way how to *remove the white frame around the plot*, despite trying out a few suggested solutions; see [here](https://community.rstudio.com/t/ggsave-aspect-ratio-whitespace-use-case-favicon-for-blogdown/6793/7){target="_blank"}, [here](https://www.pmassicotte.com/post/removing-borders-around-ggplot2-graphs/){target="_blank"}, and [here](https://stackoverflow.com/questions/58723151/remove-white-margin-surrounding-geom-sf-objects-in-r-shiny-ggplot-with-renderp){target="_blank"}. The [`crop` chunk option](https://bookdown.org/yihui/rmarkdown-cookbook/crop-plot.html){target="_blank"} can do the trick, but in this case, with the output an svg, the final result looked quite blurred.

Overall tough, while the graph is not 100 % identical with the NYT's versions, its sufficiently close for my purpose and available time. And more importantly to me, the basic concept how to create such graphs is now clear to me.

### 2021-01-08 _post/2020-06-08-update-on-mobility-patterns/update-on-mobility-patterns.Rmd
add covid mobility update post

---
title: "Update on Google mobility reports on Austria"
description: |
  This is a short update on my previous post dealing with Google’s mobility reports on Austria during the Covid-19 crisis.
date: 06-08-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - climate
  - Corona virus
  - shiny

# preview: /preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

# Context

This is a short update on my [previous post](https://werk.statt.codes/post/google-s-mobility-reportsmd/) dealing with Google's mobility reports on Austria. Since the last post was published, the (at least preliminary) peak of the epidemic has passed, restrictions are lifted one after the other , and the country is gradually 're-booted'. Against this background, I thought it might be instructive to see how this is reflected in Google's mobility reports.

To recap, Google's mobility reports present data on the change of numbers of individuals staying at particular types of places (e.g. parks, groceries, transit stations etc) by using the pertaining data of Android's location history (which users knowingly or unknowingly have activated). An exception is the location type 'residential' which captures the change of duration of people's stays. Changes are calculated by *contrasting each weekday's value with the median value of the same weekdays from a pre-Covid19 baseline period*. For details see Google's pertaining [info](https://support.google.com/covid19-mobility?hl=en#topic=9822927){target="_blank"}.

## R Code and Shiny
Most of the relevant R code for this post is folded in chunks in the text. The repo including with the full code for this post is available at my [github](https://github.com/werkstattcodes){target="_blank"}.
This post is also my first use of RStudio's powerful `shiny` package, which allows you to produce e.g. responsive graphs. Have a look at [this](https://shiny.rstudio.com/tutorial/){target="_blank"}, [this](http://zevross.com/blog/2016/04/19/r-powered-web-applications-with-shiny-a-tutorial-and-cheat-sheet-with-40-example-apps/){target="_blank"} and [this](https://deanattali.com/blog/building-shiny-apps-tutorial/#implement-server-logic-to-create-outputs){target="_blank"} tutorial to get going.

As always, if there's any glaring error etc. pls don't hesitate and let me know via Twitter [dm](https://twitter.com/zoowalk){target="_blank"}.

# Setup
<details closed> 
<summary>Load packages</summary> 
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(extrafont)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
hrbrthemes::update_geom_font_defaults(family="Roboto Condensed")
# hrbrthemes::update_geom_font_defaults(family="Plex Sans")
library(scales)
library(grid)
library(knitr)
library(geofacet)
library(paletteer)
library(ggtext)
library(glue)
library(gt)
```
</details>

<details closed> 
<summary>Define rmarkdown options</summary> 
```{r echo = T}
# knit_hooks$set(wrap = function(before, options, envir){
#   if (before){
#     paste0('<', options$wrap, '>')
#   } else {
#     paste0('</', options$wrap, '>')
#   }
# })

knitr::opts_chunk$set(
	echo = TRUE,
	fig.align = "left",
	# fig.height = 5,
	# fig.width = 7,
	out.width = "100%",
	message = FALSE,
	warning = FALSE,
	dpi = 300
)

options(width=180, dplyr.width = 150)
```
</details>

<details closed> 
<summary>Define plot theme</summary> 
```{r echo=TRUE}
theme_post <- function(){ 
  hrbrthemes::theme_ipsum_rc() +
  theme(
    plot.margin=margin(l=0, unit="cm"),
    plot.title = element_markdown(color="grey20",
                                  face="bold",
                                  margin=margin(l=0, unit="cm"),
                                  size=13),
    plot.title.position = "plot",
    plot.subtitle = element_text(color="grey50",
                                 margin=margin(t=0.1, b=0.3, unit="cm"),
                                 size=9),
    plot.caption = element_text(color = "grey50", 
                                size=7,
                                hjust=c(0, 1)),
    plot.caption.position = "panel",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 9),
    axis.title.y=element_blank(),
    axis.text.y = element_text(size = 9),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.spacing = unit(0.25, "cm"),
    panel.spacing.y = unit(0.25, "cm"),
    strip.text = element_text(angle = 0,
                              size= 9,
                              vjust = 1,
                              face="bold"),
    legend.title = element_text(color="grey30",
                                face="bold",
                                vjust=1,
                                size=7),
    legend.text = element_text(size=7, 
                               color="grey30"),
    legend.justification = "left",
    legend.box = "horizontal",  #arrangement of multiple legends
    legend.direction = "vertical",
    legend.margin = margin(l=0, t=0, unit="cm"),
    legend.spacing.y = unit(0.07, units="cm"),
    legend.text.align = 0,
    legend.box.just = "top",
    legend.key.height = unit(0.2, "line"),
    legend.key.width = unit(0.5, "line"),
    text=element_text(size=5))
  }

```
</details>

<details closed> 
<summary>Define weekend and holiday dates</summary> 
```{r echo=T}
# set dates ---------------------------------------------------------------

#dates on lockdown
date.start.lockdown <- tibble(date=as.Date("2020-03-14"), event="start of lockdown")
date.opening.shops.small <- tibble(date=as.Date("2020-04-14"), event="small shops reopen")
date.opening.shops.all <- tibble(date=as.Date("2020-05-04"), event="all shops reopen")
date.schools <- tibble(date=as.Date("2020-05-18"), event="schools reopen")
date.restaurants <- tibble(date=as.Date("2020-05-15"), event="restaurants reopen")

df_dates <- bind_rows(date.start.lockdown, date.opening.shops.all, date.opening.shops.small,
                      date.schools, date.restaurants) %>%
  mutate(label=paste0(event, " (", format(date, "%d %b"), ")")) %>% 
  mutate(label.2=paste0(format(date, "%d %b"), ", ", event)) %>% 
  arrange(date) %>% 
  mutate(label=fct_inorder(label))

#dates on holidays in Austria
df_holidays <- jsonlite::fromJSON("https://date.nager.at/api/v2/publicholidays/2020/AT ") %>% 
  mutate(date=lubridate::ymd(date))

```
</details>

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
df_dates %>% gt()
df_holidays %>% gt()

```

# Get and rearrange data

<details closed> 
<summary>Import data from Google and rearrange </summary> 
```{r, fig.height=7}

# import csv --------------------------------------------------------------
# file_link <- "https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv"
# df_AT_mobility <- readxl::read_xlsx(path=here::here("_blog_data", "update_mobility_reports",
#                                                         "Global_Mobility_Report_AT.xlsx")) %>% 
#   filter(country_region_code=="AT")
#   
# readr::write_csv2(df_AT_mobility, file=here::here("_blog_data", "update_mobility_reports",
#                                                   "AT_mobility_report.csv"))

df_AT_mobility <- readr::read_csv2(file=here::here("_blog_data", "update_mobility_reports", "AT_mobility_report.csv"))

df_mobility_AT <- df_AT_mobility %>% 
  mutate(sub_region_1=case_when(is.na(sub_region_1) ~ "State level",
                   TRUE ~ as.character(sub_region_1)))

df_AT_long <- df_mobility_AT %>% 
  pivot_longer(cols=contains("baseline"),
                      names_to="type",
                      values_to="value") %>% 
  mutate(week.day=lubridate::wday(date, label=T), .after=date) %>% 
  mutate(sub_region_1=forcats::as_factor(sub_region_1))

df_AT_long <- df_AT_long %>% 
  mutate(place.description=case_when(str_detect(type, "retail") ~ "‘places like restaurants, cafes, shopping centers, theme parks, museums, libraries, and movie theaters.’",
                                     str_detect(type, "grocery") ~ "’places like grocery markets, food warehouses, farmers markets, specialty food shops, drug stores, and pharmacies.’",
                                     str_detect(type, "park") ~ "‘places like local parks, national parks, public beaches, marinas, dog parks, plazas, and public gardens.’",
                                     str_detect(type, "transit") ~ "‘places like public transport hubs such as subway, bus, and train stations.’",
                                     str_detect(type, "work") ~ "‘places of work’",
                                     str_detect(type, "residential") ~ "‘places of residence’",
                                     TRUE ~ as.character("missing"))) %>% 
  mutate(day.indicator=case_when(week.day %in% c("Sat", "Sun") ~ "weekend",
                                     date %in% df_holidays$date ~ "public holiday",
                                     TRUE ~ as.character("business day"))) %>% 
  mutate(type=str_remove(type, "_percent.*$") %>% str_replace_all(., "_", " ") %>% 
           str_to_title(.))

```
</details>

# Regional differences

Below a shiny app which allows you to select the type of place and get the corresponding results displayed. Since this blogdown site is rendered with Hugo and hence a static website, shiny can't be run 'directly within' the site. To overcome this, shiny apps are hosted externally and subsequently embedded via an iframe. the code for the app is available [here](https://github.com/werkstattcodes/Shiny_Corona_Mobility_Reports){target="_blank"}.

```{r eval=FALSE}
#<iframe height="1000" width="100%" frameborder="no"  src="https://werkstattcodes.shinyapps.io/Shiny_Corona_Mobility_Reports/"> </iframe>
```
Select the type of place in the dropdown and the graph should subsequently update. 
<!-- https://github.com/rstudio/distill/issues/119 -->

<!-- <iframe  -->
<!-- src="https://werkstattcodes.shinyapps.io/Shiny_Corona_Mobility_Reports/"> class="l-screen-inset </iframe> -->

<iframe scrolling="no" frameborder="no" src="https://werkstattcodes.shinyapps.io/Shiny_Corona_Mobility_Reports/" class="l-page" style="height: 100vh;"></iframe>

<br>

Each dot represents a day. The location on the y-axis represent the % change relative to the median value for the same weekdays during the baseline period from 3 January to 6 February (= before Covid19 related restrictions were implemented). The color indicates whether it's a business, weekend or holiday. The difference is insofar of relevance, since individuals' mobility patterns are different during business, weekend and holidays and Covid19 restrictions hence have different consequences. The vertical lines indicated dates pertaining to the introduction and gradual withdrawal of the lockdown.

## Interpretation
### Grocery and Pharmacy
Groceries and pharmacies have experienced a  strong recovery if compared to other types of places. With the exception of Tyrol and Salzburg (absence of tourists?), values have returned to or slightly exceed  pre-Corona values now. After all, you somehow have to get your food and medicine. The reopening of all shops on 4 May seems to have triggered a particularly strong jump in the numbers.

### Parks
Mobility patterns for parks are particularly strongly fluctuating. By Google's [own account](https://support.google.com/covid19-mobility/answer/9825414?hl=en&ref_topic=9822927){target="_blank"}, numbers for 'visitors to parks are heavily influenced by the weather'. Acting on the maxim to never let a distraction go unused, let's contrast temperature data with mobility patterns for parks.

To account for temperature, I draw on the `rnoaa` [package](https://docs.ropensci.org/rnoaa/){target="_blank"} which facilitates access to the comprehensive dataset of the US National Oceanic and Atmospheric Administration (unfortuantely though, it doesn't provide data for all nine Austrian states.^[As a proxy for the weather I used the weather data for each state's captial.])

```{r include=FALSE}
options(noaakey = Sys.getenv("noaakey"))
```
<details closed> 
<summary>Get weather data, merge datasets</summary>
```{r eval=TRUE, fig.height=6}
library(rnoaa)
#get all stations in Austria
df_AT <- ncdc_stations(locationid='FIPS:AU')$data

#only those of interest
df_stations <- df_AT %>%  
  filter(str_detect(name, regex("Wien|Salzburg|Graz|Klagenfurt|Poelten", ignore_case = T)))

#get temperature data
df_res <- ncdc(datasetid = "GHCND", stationid = df_stations$id, datatypeid = "TAVG", startdate = "2020-01-01", enddate = "2020-06-01",
     add_units=T, limit=1000)

df_temps <- df_res$data %>% 
  mutate(date=lubridate::ymd_hms(date) %>% as.Date()) %>% 
  mutate(tavg=value/10) %>% #! tenth celcius
  left_join(., df_stations %>% 
              select(name, id) %>% 
              distinct(name, id),
            by=c("station"="id")) %>% 
  mutate(state=case_when(str_detect(name, regex("Wien", ignore_case = T)) ~ "Vienna",
                         str_detect(name, regex("Salzburg", ignore_case = T)) ~ "Salzburg",
                         str_detect(name, regex("Graz", ignore_case = T)) ~ "Styria",
                         str_detect(name, regex("Poelten", ignore_case = T)) ~ "Lower Austria",
                         str_detect(name, regex("Klagenfurt", ignore_case = T)) ~ "Carinthia"))

df_temp_mobility <- df_AT_long %>% 
  filter(str_detect(type, regex("park", ignore_case=T))) %>% 
  filter(!str_detect(sub_region_1, "State")) %>% 
  select(sub_region_1, date, google.value=value, day.indicator) %>% 
  inner_join(., df_temps %>% 
              select(date, state, tavg),
            by=c("date"="date", "sub_region_1"="state"))

df_temp_mobility <- df_temp_mobility %>% 
  mutate(lockdown.indicator=case_when(date>=as.Date("2020-03-14") ~ "lockdown",
                                      date<as.Date("2020-03-14") ~ "pre-lockdown")) %>% 
  mutate(lockdown.indicator=forcats::fct_inorder(lockdown.indicator))

```
</details>

```{r eval=FALSE, fig.height=6, include=FALSE}
plot_park <- df_temp_mobility %>% 
  ggplot()+
  labs(x="daily average temperature",
       y="change to pre-Covid19 baseline period",
       title="Changes in presences in parks, Covid19 lockdown measures and weather",
       subtitle="Average temperature of state's captial used",
       caption=c("data: NOAA, Google Mobility Reports", "graph: Roland Schmidt | @zoowalk | www.werk.statt.code"))+
  geom_smooth(aes(y=google.value,
                  x=tavg),
              size=0.5,
              color="black",
              method='lm', 
              formula= y~x)+
  geom_point(aes(y=google.value,
                 x=tavg,
                 color=day.indicator),
             size=.5)+
  # geom_smooth(aes(y=google.value,
                  # x=tavg))+
  scale_color_manual(values=c("business day"="steelblue",
                                        "public holiday"="green3",
                                        "weekend"="orange"))+
  scale_y_continuous(labels=scales::label_percent(accuracy = 1,
                                                            scale=1),
                               limits=c(-100, 200),
                     position="right")+
  scale_x_continuous(labels=scales::label_number(suffix="°",
                                                 accuracy = 1))+
  facet_grid(lockdown.indicator~sub_region_1,
             switch = "y")+
  theme_post()+
  theme(
        axis.title.x = element_text(size = 6,
                                    color="grey30",
                                    hjust=1),
        axis.title.y = element_text(size = 6,
                                    color="grey30",
                                    hjust=1),
        legend.position="bottom",
        strip.text.y = element_text(hjust=1)
  )+
  guides(color=guide_legend(title = "type of days",
                            title.position = "left",
                            direction = "horizontal"))

          # guides(fill=guide_legend(override.aes=list(size=2), 
          #                            title="type of days",
          #                            title.position = "left",
          #                            keyheight= 0.5,
          #                            label.hjust = 0),

```

```{r eval=FALSE, fig.height=4, message=FALSE, warning=FALSE, dpi=300, include=FALSE}
plot_park
```

<details closed> 
<summary>Create plot</summary>
```{r fig.height=4, message=FALSE, warning=FALSE, dpi=300}
library(ggbeeswarm)

plot_temp <- df_temp_mobility %>% 
  ggplot()+
    labs(x="daily average temperature",
       y="% change to pre-Covid19 baseline period",
       title="Changes in presences in parks, Covid19 lockdown measures and weather",
       subtitle="Average temperature of state's captial used",
       caption=c("data: NOAA, Google Mobility Reports", "graph: Roland Schmidt | @zoowalk | www.werk.statt.code"))+
  ggbeeswarm:: geom_quasirandom(aes(x=tavg,
                                y=google.value,
                                color=lockdown.indicator),
                                size=.5,
                                groupOnX = T)+
  scale_color_manual(values=c("pre-lockdown"="#008B00",
                              "lockdown"="#EE0000"),
                     name=NULL)+
  scale_y_continuous(labels=scales::label_percent(accuracy = 1,
                                                            scale=1),
                               limits=c(-100, 200),
                     position="right")+
  scale_x_continuous(labels=scales::label_number(suffix="°",
                                                 accuracy = 1))+
  theme_post()+
  theme(
        axis.title.x = element_text(size = 6,
                                    color="grey30",
                                    hjust=1),
        axis.title.y = element_text(size = 6,
                                    color="grey30",
                                    hjust=1),
        legend.position="bottom",
        legend.direction = "horizontal",
        strip.text.y = element_text(hjust=1))+
  lemon::facet_rep_grid(day.indicator~sub_region_1,
                        repeat.tick.labels = T,
                        switch = "y")

```
</details>

```{r echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, dpi=300}
plot_temp
```

Admittedly, this is an oversimplification, and I'll only focus here on business days. But what the graph shows is that numbers in parks dropped following the lockdown as long as temperatures remained relatively low. However, once temperature started to rise, numbers increased and exceeded pre-Covid19 levels (base line period). The latter is not surprising, considering that the base line period runs 3 January to 6 February, hence much colder days.

### Residential
A critical difference of the residential category is its unit of measurement. While the other categories measure changes in total visitors, the residential category shows a *'change in duration'*. According to Google's [documentation](https://support.google.com/covid19-mobility/answer/9825414?hl=en&ref_topic=9822927){targe="_blank"} of its mobility report 'because people already spend so much of the day at places of residence (even on workdays), the capacity for change isn't so large.' Even at the early stages of the lockdown, when individuals were most compliant and cautious, the pertaining values 'only' increased by roughly 25 %. Unsurprisingly, these changes were stronger for business days than for weekends. Gradually, even before any formal restrictions were lifted, the duration of presence has been declining. By the end of May, the pertaining values are between six and nine percent above pre-Covid19 levels.

This dynamic fits pretty well with my own, completely personal     observations. While most people in our neighborhood initially left their home in very few instances, people got eventually more relaxed, spend time outside, or even returned to their workplace. The remaining gap at the end of May is likely to be also indicative for the increased use of home-office/remote work and the thousands of newly unemployed.

### Workplace
The data for the workplace category is imho quite impressive since it shows in someway the scale of Covid19's economic impact. Even by the end of May, at a point in time when many restrictions were already lifted, the number of persons at their workplaces is still around 30 % lower than during pre-Covid19 levels.

A detail what is striking is the East-West difference when it comes to the dynamic during *weekends*. In the eastern states, numbers of weekends started to increase relatively early and in clear contrast to business days. Or to put it differently, the difference of the number of persons at their workplace at weekends before COVID and during the lockdown was higher in the west than in the east. Again, I can only speculate, but I wouldn't surprised if tourism plays a significant part in the explanations. In areas with particularly strong tourism during the base line period (=winter, particularly Tyrol and Salzburg), work during the weekend is far much common than in other areas of Austria. Shops can open also on Sundays; hotels etc obviously as well.

### Retail and Recreation
Places related to retail and recreation experienced a remarkable sudden drop following the introduction of the lockdown on 14 March. Subsequently, and at a closer look, values increased almost 'stepwise', following the re-opening of small shops on 14 April, the opening of all other shops on 4 May and more recently the opening of restaurants and school on 18 May. 
What stands out is the particularly strong gap in Tyrol and Salzburg. I can only speculate, but I could imagine that this has something to do with both state's strong tourism industry and the absence of tourists also after the removal of various restrictions.

### Transit station
As with most other results, the data on transit stations also reveals a steady increase, starting with the end of March. However, despite this growth there is still a considerable gap to the pre-Covid19 lockdown level. As of end of May, values are still down, ranging between -45% (Tyrol) and -16% (Carinthia). As for Vienna (-32 %), the low number ties in well with recent [news reporting](https://wien.orf.at/stories/3051627/){target="_blank"} on the city's public transport system.

I am not sure why the numbers related to Burgenland are missing from mid-April onward. Burgenland would be a state which - as far as I can tell - features particularly strong numbers of commuters (to Vienna). According to Google, its reports can contain gaps when the data doesn't meet the quality and privacy threshold (i.e. not enough data to ensure anonymity; see [here](){target="_blank"}). As for the latter, I would be surprised if the number of commuters dropped to such a low level that revealing the pertaining numbers would facilitate identifying individual Android users. Furthermore, the numbers only started to be missing from mid-April onward, hence from a point in time when the number of persons present on transit stations was likely to increase.

## Recap
Overall, I find Google's mobility reports quite informative about where people spend their time in the course of this epidemic and the interplay with various restrictions. While the data show that 'normality' has almost returned in some areas (grocery), it's still impressive how fundamental the impact remains in others (workplace, transport).

### 2021-01-08 _post/2020-07-21-austrian-parliament-mps-transparency-and-conflict-of-interest-report/austrian-parliament-mps-transparency-and-conflict-of-interest-report.Rmd
add covid mobility update post

---
title: "Austrian Parliament: MPs transparency and conflict of interest report"
description: |
  Analysis of MP's transparency and conflict of interest report.
date: 07-21-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - OCR
  - web scraping

preview: austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

# Context

Recently, at the end of June, the Austrian Parliament published its annual report on MPs' additional sources of revenue and pertaining activities (see [here](https://www.parlament.gv.at/POOL/SWBRETT/ZUSD/BezBegrBVGPar9-NR.pdf){target="_blank"}). The publication is mandatory by law, and basically seeks to curtail potential conflicts of interest and to increase transparency as to MPs' activities outside of parliament. While the list comes, unsurprisingly, with a number of caveats ^[For details see [here.](https://www.parlament.gv.at//POOL/SWBRETT/25020/0010/BezBegrBVGPar9Erkl-NR.pdf#page=1){target="_blank"}], and by itself can hardly constitute a watertight safeguard against the various 'undemocratic temptations', the report nevertheless seems to provide interesting data on MPs' financial statuses and professional machinations. All in all enough reason to give it a closer look, i.e. a blog post.^[Halfway through the blog, I realized that the transparency website meineabgeordneten.at had alreay published a similar, although non-technial blog on this matter. If you speak German, have a look [here](https://www.meineabgeordneten.at/News/detail/Nationalrat_Nebeneinkommen_Offenlegung_2019){target="_blank"}. In a previous version I also had wrongly identified the gender of one MP. Kindly, Marion Breitschopf from meineabgeordneten.at helped clarifying this.] Note that I will focus only on Austria's lower chamber (*Nationalrat*) in this post. If time permits, I'll follow up with one on the upper chamber (*Bundesrat*) at a later point.

If you're not that much interested in the R code, simply collapse the code chunks. The entire code for the analysis will be on my github account ([link](https://github.com/werkstattcodes){target="_blank"}). For any questions, comments etc. feel free to send me a direct message on twitter ([link](https://twitter.com/zoowalk){target="_blank"}).

# Setup
But step by step:
Let's first take care of the necessary R packages and blog/rmarkdown setup.

<details closed> 
<summary>Code: Load packages</summary> 
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(extrafont)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
hrbrthemes::update_geom_font_defaults(family = "Roboto Condensed",
                                      size=3)
library(scales)
library(knitr)
library(paletteer)
library(ggtext)
library(glue)
library(gt)
library(DT)
library(pdftools)
library(lemon)
library(ggiraph)

```
</details>

<details closed> 
<summary>Code: Define rmarkdown options</summary> 
```{r setup, echo = T}
# knit_hooks$set(wrap = function(before, options, envir) {
#   if (before) {
#     paste0("<", options$wrap, ">")
#   } else {
#     paste0("</", options$wrap, ">")
#   }
# })

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "left",
  # fig.height = 5,
  # fig.width = 7,
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  dpi = 300
)

options(width = 180, dplyr.width = 150)
```
</details>

<details closed> 
<summary>Code: Define plot theme</summary> 
```{r echo=TRUE}
theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.margin = margin(l = 0, unit = "cm"),
      plot.title = element_markdown(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 13
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 11
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(angle = 0,
                                         color="grey50",
                                         hjust=1),
      axis.text.x = element_text(size = 9, 
                                 color = "grey50"),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 9,
                                 color="grey50"),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 7
      ),
      legend.text = element_text(
        size = 7,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

my_caption=c("data: Austrian Parliament, Liste gemäß § 9 Bezügebegrenzungs-BVG, 2020. \nanalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes")
```
</details>

# Data extraction
As so often, one of the main challenges to actually analyze available data is to 'liberate' the information of interest from the format in which they are provided. The Parliament's report comes as a pretty 'untidy' pdf with subheadings etc.,  a format hardly conducive for further analysis. Here's one MP's declaration from the report:

![](/post/2020-07-03-austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/list_example - Copy.PNG){width=100%}

Essentially, the table provides four relevant pieces of information: *1) an income category*, which groups MPs' additional income (which is within the scope of the legal reporting obligations and in addition to their salary as an MP) according to their average monthly income in 2019;

The income categories are:
```{r echo=FALSE}
data.frame(
  category = c(1L, 2L, 3L, 4L, 5L),
  income.bracket = c(
    "1-1,000", "1,001-3,500",
    "3,501-7,000", "7,001-10,000", ">10,000")) %>% 
    gt() %>% 
  gt::cols_label(income.bracket=md("**income bracket (Euro)**"),
                 category=md("**category**")) %>% 
  gt::cols_align(align="right",
                 columns = vars(income.bracket))
```

And lists of **2) executive positions** in stock companies, private limited companies, foundations or saving banks held by each MP; **3) other activities** from which MPs generate income; and **4) honorary activities** which yield no income (but could potentially nevertheless lead to some conflict of interest).

Extracting data from pdfs has been the topic of at least one of my previous posts (see [here](https://werk.statt.codes/post/bih-voting-behavior-of-mps-in-house-of-representatives/){target="_blank"}), and again the primary tool for the job is the wonderful `pdftools` package. Rather than seeking to extract tables from the pdf (e.g. with `tabulizer` ), I convert the content of each page of the report to a long text string and subsequently extract the parts of interest by means of `stringr` and regex. However, as I'll explain below, this approach was only suited to do some part of the task at hand.

Let's start by retrieving the raw text.

<details open> 
<summary>Code: Extract text from pdfs </summary> 
```{r}
my_path <- here::here("_blog_data", "MP_activities", "BezBegrBVGPar9-NR.pdf")

df_all <- pdftools::pdf_text(pdf = my_path) %>%
  enframe(.,
    name = NULL,
    value = "text_raw"
  ) %>%
  mutate(text_raw = str_replace_all(text_raw, "\\h+", " "))
```
</details>

The code above provides us with a tibble containing as many rows as the report has pages (186). The expression `str_replace_all(text_raw, "\\h+", " ")` replaces all horizontal white space with a single space, but keeps line breaks (\\n and \\r). The latter will be needed at a later stage when it comes to identifying the location of data of interest and extract them.^[I am grateful to the respondents to my pertaining question on [stackoverflow](https://stackoverflow.com/questions/62678640/stringr-remove-multiple-spaces-but-keep-linebreaks-n-r/62679016#62679016){target="_blank"}]

Each row is one long string containing all the information from only one single page on an MP. However, note that the declarations of some MPs span over multiple pages, hence their data is contained in multiple strings. This is something we have to account for when crunching numbers during the actual analysis. Here one example for the data on one MP. Use the horizontal scrollbar to get an idea of the strings composition.

```{r echo=FALSE, tidy=FALSE, R.options=list(width=10)}
stringr::str_wrap(df_all$text_raw[1], 10)
```

## Get income category
As a first step, let's extract the name of the MPs and the income category from the raw strings obtained above. The code chunk below is the pertaining code, making use of `stringr` and `regex`. Note the `(?s)` term in the regex expressions. While `.` would match any pattern but not line breaks, the `(?s)` alters this limitation and forces `.` to match also `\n`  and `\r`.^[See [here](https://stackoverflow.com/questions/159118/how-do-i-match-any-character-across-multiple-lines-in-a-regular-expression/45981809#45981809){target="_blank"} for a very helpful answer on SO.]

<details open> 
<summary>Code: Extract MPs' names and income category </summary> 
```{r}
df_all <- df_all %>%
  mutate(name = str_extract(text_raw, regex("(?<=Einkommenskategorie)(?s).*(?=für das Jahr)", ignore_case = T)) %>%
    str_remove(., "\\(i\\)") %>%
    str_trim()) %>%
  mutate(income_category_19_raw = str_extract(text_raw, regex("(?<=2019:\\s?)\\S"))) %>% 
  mutate(income_category_19=case_when(income_category_19_raw=="-" ~ 0,
                                      TRUE ~ as.numeric(income_category_19_raw))) %>% 
    replace_na(., list(income_category_19="missing")) %>%
  select(-income_category_19_raw)

```
</details>

As already highlighted, since the declarations of some MPs run over multiple pages and are hence included several times in the above output, we have to limit the result to one row per MP. This is what `distinct()` does in the code chunk below.

<details open> 
<summary>Code: Account for multi-page declarations </summary>  
```{r}
df_income <- df_all %>%
  select(name, income_category_19) %>%
  distinct()
```

```{r}
nrow(df_all)
nrow(df_income)
```
</details>

## Missing income data
Let's have a first, quick look at the information obtained so far:

<details closed> 
<summary>Code: Number of MPs per income bracket </summary>  
```{r}
tbl_income <- df_income %>% 
  count(income_category_19) %>% 
  gt() %>% 
  gt::cols_label(income_category_19=md("**income category**"),
                 n=md("**number of MPs**")) 
```
</details>

```{r echo=FALSE}
tbl_income
```

Before digging into frequencies of the income categories etc, a note on the missing observations. In ten instances, MPs did not report their income for 2019. Here are the ten MPs with missing income data.

<details closed> 
<summary>Code: MPs with with missing income data</summary>  
```{r}
df_income_missing <- df_income %>% 
  filter(income_category_19=="missing") %>% 
  mutate(family_name=str_remove(name, ",.*$") %>% 
           str_extract(., "\\S*$"))

tbl_missing <- df_income_missing %>% 
  select(name, income_category_19) %>% 
  gt() %>% 
  cols_label(income_category_19=md("**income bracket**"),
             name=md("**MP**")) %>% 
  gt::cols_align(align="center",
                 columns = vars(income_category_19))
```
</details>
```{r echo=FALSE}
tbl_missing
```

One possible explanation for missing income data could be that these MPs have joined parliament only after the reporting period, i.e. in 2020 (I mean, there couldn't be any other possibility, no?). To check this, we need to obtain the dates of each MP's entry to parliament. Initially, I expected that such data should be readily available, but I must have been looking at the wrong places... Anyway, never waste an opportunity to scrap data.

The website of the Austrian Parliament provides a list of all (current and previous) MPs. This list is provided in an xml format from which we can obtain the links to each MP's  individual biographical page. From these bios, we can subsequently extract the data when MP's entered the parliament. Not a bad detour...or, as a saying in German goes, Umwege erhöhen die Ortskenntnisse...(something like 'detours increase your local knowledge'.)

<details closed>
<summary>Code: Get entry dates for MPs with missing income data</summary>
```{r eval=FALSE}
library(xml2)
library(rvest)
link <- "https://www.parlament.gv.at/WWER/PARL/J1918/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&NRBR=ALLE&GP=XXVII&R_WF=FR&FR=ALLE&W=W&M=M&listeId=8&FBEZ=FW_008"

#parse info from parliament's website
data <- xml2::read_xml(link)

#extract name of MPs
df_title <- data %>% 
  xml2::xml_find_all("//title") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="title")

#extract link to bio
df_link <- data %>% 
  xml2::xml_find_all("//link") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="link")

#combine both to one dataframe
df <- df_title %>% 
  left_join(., df_link,
            by="id")

#remove non-MP relevant rows
df <- df %>% 
  filter(str_detect(link, "PAD_")) %>% 
  mutate(link=str_trim(link))

#define function to extract MP's entry/exit dates from biography
fn_get_dates <- function(x) {
  read_html(x) %>% 
  html_nodes(xpath="/html/body/div[1]/div[5]/div[1]/div[1]/div[3]/div[2]/div[2]/ul[1]/li/div/text()[2]") %>% 
  html_text()
}

df

#apply funciton to all links
# df_mp_dates <- df %>% 
#   mutate(mp_dates=map(link, ~fn_get_dates(.)))

```

```{r include=FALSE}
# readr::write_rds(df_mp_dates, path=here::here("_blog_data", "MP_activities", "MP_dates.rds"))

df_mp_dates <- readr::read_rds(file=here::here("_blog_data", "MP_activities", "MP_dates.rds"))

```

```{r}
#extract family name for subsequent merger with income data
df_mp_dates <- df_mp_dates %>% 
  mutate(family_name=str_trim(title) %>% 
           str_extract(., regex("^\\S*")))

#merge with MPs whose income data is missing
df_income_missing <- df_income_missing %>% 
  left_join(., df_mp_dates %>%
              select(family_name, mp_dates),
            by=c("family_name"="family_name")) %>% 
  select(name, mp_dates) %>% 
  mutate(mp_dates=unlist(mp_dates) %>% 
           str_remove(., regex("(?<=\\d{4}).*")))

```
</details>

```{r echo=FALSE}
df_income_missing %>% 
  gt() %>% 
  gt::cols_label(mp_dates=md("**entry date**"),
                 name=md("**MP**"))
```

And indeed. From the result above, we see that all those MPs who haven't submitted any income information were not MPs in 2019 *for a full month* (this is the requirement stipulated in the pertaining regulation). Hence, for the subsequent analysis on income data, when can safely disregard these MPs.

<details closed>
<summary>Code: Remove MPs joining outside of reporting period </summary>
```{r echo=TRUE}
nrow(df_income)
df_income <- df_income %>% 
  filter(!name %in% df_income_missing$name)
nrow(df_income)
```
</details>

# MPs' gender

```{r include=FALSE}
gender_colors <- c(female="firebrick", male="grey", missing="#BEBEBE")

df_income_categories <- tibble::tribble(
  ~income_category_19, ~category_values, ~category_labels,
                   0L,              "0",              "0",
                   1L,        "1-1,000",             ">1",
                   2L,    "1,001-3,500",            ">1K",
                   3L,    "3,501-7,000",          ">3.5K",
                   4L,   "7,001-10,000",            ">7K",
                   5L,        ">10,000",           ">10K"
  )

income_categories <- df_income_categories %>% 
  select(1, 2) %>%
  deframe()

income_categories_short <- df_income_categories %>% 
  mutate(income_label=paste0(
    "[",income_category_19,"]",
    "\n",
    category_labels, "€")) %>% 
  select(1, 4) %>%
  deframe()

```

Let's have a gender-sensitive look at the income data. With the `gender`package we are able to infer with some acceptable margin of error MPs' gender from their first names. Names which were not (properly) recognized by the package I corrected manually in the code chunk below.

<details open>
<summary>Code: Get MPs' gender</summary>
```{r}
#get a clean name, without any academic or prefessional titles
df_income <- df_income %>% 
  mutate(name_family=str_remove(name, ",.*$") %>% 
           str_extract(., "\\S*$"),
         .after=name) %>% 
  mutate(name_clean=name %>% str_remove(., regex(",.*$")) %>% 
           str_remove_all(., regex("\\S*\\.")) %>% 
           str_remove_all(., regex("\\(.*\\)")) %>% 
           str_trim())

library(gender)
#extract first name and apply gender function to it
df_income <- df_income %>% 
  mutate(first_name=str_extract(name_clean, regex("^\\S*"))) %>% 
  mutate(name_gender=map(first_name, possibly(gender,
                                              otherwise="missing")))

# manually correct for some missing gender data
df_income<- df_income %>% 
  unnest_wider(name_gender,names_repair="universal") %>% 
  rename(name=name...1) %>% 
  select(-contains("year"), -contains("proportion"), -ends_with("7")) %>% 
  mutate(gender=case_when(str_detect(name, "Brandstötter")==TRUE ~ "female",
                          str_detect(name, "El-Nagashi")==TRUE ~ "female",
                          str_detect(name, "Sibylle")==TRUE ~ "female",
                          str_detect(name, "Himmelbauer")~"female",
                          str_detect(name, "Karlheinz")==TRUE ~ "male",
                          str_detect(name, "Jörg")==TRUE ~ "male",
                          str_detect(name, "Volker")==TRUE ~ "male",
                          str_detect(name, "Gertraud")==TRUE ~ "female",
                          str_detect(name, "Nurten")==TRUE ~ "female",
                          str_detect(name, "Süleyman")==TRUE ~ "male",
                          str_detect(name, "Ribo")==TRUE ~ "female",
                          TRUE ~ as.character(gender)))
```
</details>

## Female share per income group
What the subsequent analysis reveals is that the share of women per income category decreases the further we move up the income ladder. While there is almost gender parity in the income category 0, womens' shares decrease continuously up to category 4.

<details closed>
<summary>Code: Calculate gender ratios and create plots</summary>
```{r}
df_gender <- df_income %>% 
  group_by(income_category_19, gender) %>% 
  summarise(n_obs=n()) %>% 
  mutate(n_rel_category=n_obs/sum(n_obs)) %>% 
  group_by(gender) %>% 
  mutate(n_gender=sum(n_obs)) %>% 
  ungroup() %>% 
  mutate(n_rel_gender=n_obs/n_gender)

```

```{r}
pl_gender_category <- df_gender %>% 
  # filter(gender=="female") %>% 
  mutate(gender=fct_rev(gender)) %>% 
  ggplot()+
  labs(title=glue("Austrian Nationalrat: Share of <span style  = color:{gender_colors['female']}>female</span> and <span style = color:{gender_colors['male']}>male</span> MPs per income category"),
       subtitle=str_wrap("Income: average monthly income from professional positions in addition to MPs' salary. Absolute number of female MPs in brackets. Only lower chamber. Data for 2019.", 105),
       y="[income category] average monthly income",
       caption=my_caption)+
  geom_bar(aes(y=income_category_19,
               x=n_obs,
               fill=gender),
           stat="identity",
           position="fill")+
  geom_text(data=. %>% filter(gender=="female"),
            aes(y=income_category_19,
                x=n_rel_category-0.01,
                label=paste0(round(n_rel_category,2) %>% 
                  percent(scale=100),
                  "\n(",
                  n_obs,
                  ")")),
            color="white",
            lineheight=0.6,
            hjust=1,
            stat="identity")+
  geom_vline(xintercept = .5,
             color="white",
             linetype="dashed"
             )+
  scale_fill_manual(values=gender_colors)+
  scale_x_continuous(label=scales::label_percent(),
                     expand=expansion(mult=c(0, 0.05)))+
  scale_y_discrete(labels=income_categories_short,
                   expand=expansion(mult=c(0, 0)),
                   position = "left")+
  theme_post()+
  theme(legend.position="none",
        plot.title=element_markdown(),
        axis.title.x=element_blank(),
        axis.title.y.left = element_text(angle = 90,
                                         color="grey50",
                                         hjust=1),
        axis.text.y.left = element_text(),
        axis.text.y.right = element_blank(),
        axis.title.y.right = element_blank(),
        panel.grid.major.y  = element_blank())

```
</details>

```{r echo=FALSE}
pl_gender_category

```

## Share of income groups per gender group
Another gender sensitive perspective is to look at the distribution of the MPs across income categories within each gender group (in contrast to within each income group as in the previous graph). The graph highlights the difference between female and male MPs in terms of the income categories. While the distribution for female MPs is skewed towards lower income categories, that for male MPs is more concentrated on income category 2 and 3. Furthermore, only 4.4 % of all female MPs have additional incomes falling into the two highest categories, 4 and 5. In contrast, 19 % of all male MPs have income from these categories.

<details closed> 
<summary>Code: graph intra-gender distribution </summary>
```{r}
pl_gender_intra_dist <- df_gender %>% 
  ggplot()+
  labs(title="Austrian Nationalrat: Distribution of MPs over income groups per gender",
       subtitle=str_wrap("Income: average monthly income from professional positions in addition to MPs' salary. Absolute number of MPs in brackets. Only lower chamber. Data for 2019.", 105),
       x="[income category] average monthly income",
       caption=my_caption)+
  geom_bar(aes(x=income_category_19,
               y=n_rel_gender,
               fill=gender),
           stat="identity")+
  geom_text(aes(x=income_category_19,
                y=n_rel_gender+0.02,
                label=paste0(percent(n_rel_gender, accuracy = 0.1),
                             "\n",
                             "(", n_obs, ")")),
            lineheight=0.7,
            stat="identity")+
  scale_fill_manual(values=gender_colors, 
                    na.value = "grey30")+
  scale_x_discrete(labels=income_categories_short)+
  scale_y_continuous(label=scales::label_percent(accuracy=1),
                     expand=expansion(mult=c(0, 0.1)))+
  facet_wrap(vars(gender))+
  theme_post()+
  theme(legend.position = "none",
        axis.title.x = element_text(angle = 0,
                                         color="grey50",
                                         hjust=1),)
```
</details>

```{r echo=FALSE}
pl_gender_intra_dist
```

# Party affiliation

## Get MPs' party membership
As the next step to facilitate the analysis of the data, let's add MPs' party affiliation. I do this by importing the pertaining data from wikipedia's entry on Austria's current MP (the parliament's website was simply to messy for my liking). Here, the `rvest` package is our friend. Once the data on MP membership is retrieved, I merge it with MPs' income/transparency data. The key/identifier to merge both sets of data is an MP's name. To make this work, I had first to remove all academic/professional titles (it's Austria after all) and harmonize the sequence of first and family names.

<details closed> 
<summary>Code: Get party affiliation </summary>  
```{r}
wiki_link <- "https://de.wikipedia.org/wiki/Liste_der_Abgeordneten_zum_%C3%96sterreichischen_Nationalrat_(XXVII._Gesetzgebungsperiode)"

library(rvest)

df_nr <- wiki_link %>% 
  read_html() %>% 
  html_nodes("table") %>% 
  html_table(fill=TRUE) %>% 
  .[[1]] %>% 
  janitor::remove_empty(., which=c("cols")) %>% 
  janitor::clean_names() %>% 
  as_tibble()

df_nr <- df_nr %>% 
  mutate(name_family=str_extract(name, regex("^\\S*"))) %>% 
  select(contains("name"), party=wahl_partei) %>% 
  mutate(name_first=str_remove(name, regex("\\S*"))) %>% 
  mutate(name_clean=paste(name_first, name_family) %>% 
           str_trim() %>% str_squish()) %>% 
  mutate(name_clean=case_when(str_detect(name, "Künsberg")==TRUE ~ "Martina Künsberg Sarre",
                               TRUE ~ as.character(name_clean)))

#manually adjust inconsistently spelled names
df_income <- df_income %>% 
  mutate(name_clean=case_when(str_detect(name, "Nurten")==TRUE ~ "Nurten Yilmaz",
                               # str_detect(name, "Künsberg")==TRUE ~ " Martina Künsberg Sarre",
                              str_detect(name, "Herr")==TRUE ~ "Julia Herr",
                              str_detect(name, "Schellhorn")==TRUE ~ "Sepp Schellhorn",
                              str_detect(name, "Lercher")==TRUE ~ "Max Lercher",
                              str_detect(name, "Niss")==TRUE ~ "Maria Niss",
                              TRUE ~ as.character(name_clean)))

df_income <- df_income %>% 
  left_join(., df_nr %>% 
              select(name_clean, party),
            by=c("name_clean"="name_clean")) %>% 
  mutate(party=case_when(str_detect(name, "Strache")==T ~ "independent",
                         TRUE ~ as.character(party))) %>% 
  ungroup()

df_income <- df_income %>% 
  mutate(party=fct_infreq(party))

```
</details>

Below the result with all MPs, their party affiliation, and income data.

```{r echo=FALSE}
datatable(df_income %>% select(name, party, gender, income_category_19),
          colnames=c("MP name"="name", "income category"="income_category_19"),
          filter="top",
          options=list(pageLength=10,
                       columnDefs = list(list(className = 'dt-center', targets = 1:3,
                                         searchable=TRUE))),
          rownames = FALSE)
```

```{r include=FALSE}
party_colors <- c(FPÖ = "#005DA8", NEOS = "#EA5290", ÖVP = "#5DC2CC", SPÖ = "#FC0204", GRÜNE = "#A3C630",
                  independent="grey80")

```

## MPs' income category by party

With data on each MP's income category and party membership available, let's aggregate the data by parties. The graph below show the distribution of MPs over income categories per party in relative (and absolute) terms.

<details closed> 
<summary>Code: plots on income categories</summary>  
```{r}
df_agg <- df_income %>% 
  ungroup() %>%
  mutate(income_category_19_fac=as_factor(income_category_19) %>% 
           fct_inseq()) %>% 
  mutate(party=fct_infreq(party)) %>% 
  group_by(income_category_19, party, .drop = F) %>% 
  summarise(n_obs=n()) %>% 
  group_by(party) %>% 
  mutate(n_obs_total=sum(n_obs)) %>% 
  mutate(n_obs_rel=n_obs/n_obs_total)

pl_rel <- df_agg %>% 
  ggplot()+
  labs(title="Austrian Nationalrat: Relative number of MPs per income category and party",
    subtitle=str_wrap("Percentages are shares of each party's total number of members. Absolute numbers in brackets. Lower chamber only (Nationalrat). Data for 2019.", 105),
    x="[income category] average monthly income",
    caption=my_caption)+
  geom_label(data=. %>% filter(n_obs_rel >0),
            aes(x=party,
                y=n_obs_rel+.3,
                family="Roboto Condensed",
                group=party,
                label=paste0(scales::percent(n_obs_rel,
                                      accuracy=0.1),
                            "\n",
                            "(", n_obs, ")")
                ),
            fill="white",
            label.size=0,
            label.padding = unit(0.1, "lines"),
            lineheight=0.7,
            family="Roboto Condensed",
            position=position_dodge2(width = 1))+
    geom_bar(aes(x=party,
               y=n_obs_rel,
               group=income_category_19,
               fill=party),
           position=position_dodge2(),
           stat="identity"
           )+
  scale_fill_manual(values=party_colors)+
  scale_x_discrete(labels=abbreviate)+
  scale_y_continuous(label=scales::label_percent(),
                     # limits = c(0, 1.15),
                     expand=expansion(mult=c(0, 0.15)))+
  lemon::facet_rep_wrap(vars(income_category_19),                        labeller=labeller(income_category_19=map_chr(income_categories_short, ~str_remove_all(., regex("\\]|\\["))) %>% 
                                                                                             map_chr(., ~str_replace(., regex("\\\n"), ": ")) %>%
                                                                                             map_chr(., ~paste0("Income category ", .))),
                        ncol=2,
                        repeat.tick.labels = TRUE)+
  theme_post()+
  theme(legend.position="none",
        axis.title.x = element_blank())
```
</details>

```{r echo=FALSE}
pl_rel
```

The ÖVP features the smallest share of MPs with no additional income (income category 0, only 6.2 %) followed by the FPÖ (though with a substantially higher share of 26.7 %). In contrast, Greens and Neos feature with 45.5 % and 40 % the highest share of MPs in in this category (leaving aside, Philippa Strache, the one independent MP with no additional income). The ÖVP features the highest absolute numbers in income categories 3, 4 and 5. In contrast, there are no Green MPs with an income falling into category 4 or 5. Similarly, there is also not a single MP from the SPÖ with an income from category 5. What's noteworthy with Neos is that they feature a comparably large share of MPs from income category 0, but also the largest relative number of MPs from the highest income category (13.3 %; equal with that of the FPÖ).

# MPs' further jobs

The subsequent part of the analysis will focus on MPs' additional professional positions, as provided by the transparency report. As outlined in the introduction, it provides details on MPs' additional 1) executive positions in stock companies, private limited companies, foundations or saving banks; this section further distinguishes between those executive positions which are remunerated and those which aren't. 2) other professional activities from which MPs generate income; and 3) (unpaid) honorary activities.

## Extracting data
Extracting MPs' additional professional positions from the pdfs has been somewhat challenging. Initially - I was actually about to post the blog - I thought I could retrieve the data again with the `pdf_text` function as done above. However, when double checking some of the obtained results with the original pdf, I had to realize that some of my obtained numbers were inflated. After some digging and more digging, it turned out that the error originated from professional positions with names running over more than one line, i.e. which have a line break. Since I initially had counted the number of lines of positions to get the number of positions, some positions were counted multiple times. This problem arose since I was not able to distinguish between line breaks within cells (position with a long name) and those line breaks which indicated a new cell/row in the pdf's table (different professional position).

The answer to this challenge is provided by the `pdf_data` function, also from the `pdftools` package. As the output below shows, the function provides you with x and y coordinates for the actual, physical location of each textbox (for details see [here](https://ropensci.org/technotes/2018/12/14/pdftools-20/){target="_blank"}).

```{r echo=FALSE}
pdf_file <- pdftools::pdf_data(pdf = my_path) 
datatable(pdf_file[[8]],
          options=list(pageLength=10,
                       columnDefs = list(list(className = 'dt-center', targets = 0:4)),
                       list(className='dt_left', targets=5)),
          rownames=FALSE)
```

Taking a closer look at the rows in the results, I noticed that the difference in y values for linebreaks within a cell (long position name) is different than that between cells (different position). While 'normal' line breaks result in a y value change of 11, new cells result in a higher change of the y value. Hence, the difference in y values can be used as an indicator for a new cell and hence a different professional position.

To make use of this difference, I first calculate the difference of y values per row (`line_diff`), and subsequently and indicator which turns 1 if the y difference exceeds 11 (`row_jump`). Finally, I accumulate the value of the indicator to get an indicator for each row (`row_indicator`). For an example see the output below.

<details open>
<summary>Code: identifying new positions</summary>
```{r}
df_linebreaks <- pdf_file[[8]] %>%
  select(y, text) %>% 
    arrange(y) %>%
    mutate(line_diff = y - dplyr::lag(y, default = 0)) %>%
    mutate(row_jump = case_when(
      line_diff > 11 ~ 1,
      TRUE ~ 0
    )) %>%
    mutate(row_indicator = cumsum(row_jump)) 
```
</details>

```{r echo=FALSE}
datatable(df_linebreaks,
          options=list(pageLength=10,
                       columnDefs = list(list(className = 'dt-center', targets = "all"))),
          rownames=FALSE)
```

In the next step, I group the data by the `row_indicator` and combine the different text parts with the same row indicator into one string. As a result, we get a dataframe where each row contains one row of the table (including those parts after the line break). Hooray!

<details open>
<summary>Code: Table with one row per position</summary>
```{r paged.print=TRUE}
df_rows <- df_linebreaks %>% 
  group_by(row_indicator) %>%
    summarise(text_combined = paste(text, collapse = " ")) %>%
    mutate(text_combined = text_combined %>% str_trim() %>% str_squish())
```
</details>

```{r}
df_rows
```

From this new dataframe we have to filter out those rows which contain the different professional positions we are interested in. To do this, I first extract the row number of those rows which contain the subheading of each nested table (on executive, other, and honorary positions). With row number values of these rows, I can subsequently filter out those rows which are located between the headings (= professional positions).

<details closed>
<summary>Code: row numbers of sub-headings</summary>
```{r}
# headings
heading_executive <- df_rows %>%
    filter(str_detect(text_combined, regex("Leitende Stellung in", ignore_case = T))) %>%
    pull(row_indicator)

heading_other <- df_rows %>%
    filter(str_detect(text_combined, regex("SONSTIGE TÄTIGKEITEN", ignore_case = T))) %>%
    pull(row_indicator)

heading_honorary <- df_rows %>%
    filter(str_detect(text_combined, regex("ehrenamtlich", ignore_case = T))) %>%
    pull(row_indicator)

heading_executive
heading_other
heading_honorary
```
</details>

## Putting it all into a function
I put all the above outlined (and subsequent) steps into a function which provides me with the number of positions per category for each MP. Have a look in the code chunk below. In a sense, it's the key element of the overall analysis.

<details open>
<summary>Code: Function to extract positions </summary>
```{r}
fn_get_positions <- function(x) {
  d <- x %>%
    arrange(y) %>%
    mutate(line_diff = y - dplyr::lag(y, default = 0)) %>%
    mutate(row_jump = case_when(
      line_diff > 11 ~ 1,
      TRUE ~ 0
    )) %>%
    mutate(row_indicator = cumsum(row_jump)) %>%
    group_by(row_indicator) %>%
    summarise(text_combined = paste(text, collapse = " ")) %>%
    mutate(text_combined = text_combined %>% str_trim() %>% str_squish())

  # headings
  heading_executive <- d %>%
    filter(str_detect(text_combined, regex("Leitende Stellung in", ignore_case = T))) %>%
    pull(row_indicator)

  heading_other <- d %>%
    filter(str_detect(text_combined, regex("SONSTIGE TÄTIGKEITEN", ignore_case = T))) %>%
    pull(row_indicator)

  heading_honorary <- d %>%
    filter(str_detect(text_combined, regex("ehrenamtlich", ignore_case = T))) %>%
    pull(row_indicator)

  # executive position
  if (length(heading_executive)!=0 & length(heading_other!=0)) {
    df_executive <- d %>%
      filter(!str_detect(text_combined, regex("Rechtsträger Leitende Stellung"))) %>%
      filter(row_indicator > heading_executive) %>%
      filter(row_indicator < heading_other) %>%
      select(-row_indicator)

    n_executive <- ifelse(df_executive %>%
      filter(str_detect(text_combined, regex("^keine$"))) %>%
      nrow() == 0,
    nrow(df_executive),
    0
    )
  } else {
    df_executive <- tibble(text_combined = NA)
    n_executive <- NA
  }

  # non-remunerated positions
  n_executive_non_remunerated <- df_executive %>% 
    filter(str_detect(text_combined, regex("\\*"))) %>% 
    nrow()

  # other

  if (length(heading_other)!=0 & length(heading_honorary)!=0) {
    df_other <- d %>%
      filter(!str_detect(text_combined, regex("Dienstgeber/Rechtsträger/"))) %>%
      filter(row_indicator > heading_other) %>%
      filter(row_indicator < heading_honorary) %>%
      select(-row_indicator)

    n_other <- ifelse(df_other %>%
      filter(str_detect(text_combined, regex("^keine$"))) %>%
      nrow() == 0,
    nrow(df_other),
    0
    )
  } else {
    df_other <- tibble(text_combined = NA)
    n_other <- NA
  }

  # honorary
  if (length(heading_honorary!=0)) {
    df_honorary <- d %>%
      filter(!str_detect(text_combined, regex("Rechtsträger"))) %>%
      filter(!str_detect(text_combined, regex("\\(i\\) Näheres bei\\: Informationen"))) %>%
      filter(row_indicator > heading_honorary) %>%
      select(-row_indicator)

    n_honorary <- ifelse(df_honorary %>%
      filter(str_detect(text_combined, regex("^keine$"))) %>%
      nrow() == 0,
    nrow(df_honorary),
    0
    )
  } else {
    df_honorary <- tibble(text_combined = NA)
    n_honorary <- NA
  }

  # combine retreived info
  df <- tibble(
    executive_positions = list(df_executive),
    executive_n = n_executive,
    executive_non_remunerated_n=n_executive_non_remunerated,
    other_positions = list(df_other),
    other_n = n_other,
    honorary_positions = list(df_honorary),
    honorary_n = n_honorary
  )
  df
}

```
</details>

This function is applied to our previous results of the `pdf_data` function. At the end, I combine these results with MPs' names and aggregate those reports of MPs which run over multiple pages. Et voila! You can use the arrows to sort the table and get e.g the top-10 MPs as to their number of executive positions.

<details open>
<summary>Code: Apply function </summary>
```{r}
df_data <- pdf_file %>%
  map_dfr(., fn_get_positions)

df_positions <- bind_cols(df_all, df_data)

#rowwise & c_across to calculate rowwise sums
df_count <- df_positions %>% 
  group_by(name) %>% 
  summarize(across(ends_with("_n"), sum, na.rm=T)) %>% 
  rowwise() %>% 
  mutate(total_n=sum(c_across(ends_with("_n")&!contains("remunerated")), na.rm=T))
```
</details>

```{r echo=FALSE}
datatable(df_count,
          colnames=c("MP name"="name",
                     "# executive pos (total)"="executive_n",
                     "# executive pos not remunerated"="executive_non_remunerated_n",
                     "# other pos"="other_n",
                     "# honorary pos"="honorary_n",
                     "total # pos" = "total_n"),
          options=list(pageLength=10,
                       columnDefs = list(list(className = 'dt-center', targets = "all"))),
          rownames=FALSE)
```

Note that the table comprises 183 MPs, hence also includes those MPs which were not obliged to submit income data (see above, missing data).

```{r include=FALSE}
df_count2 <- df_count %>% 
    mutate(name_clean=name %>% str_remove(., regex(",.*$")) %>% 
           str_remove_all(., regex("\\S*\\.")) %>% 
           str_remove_all(., regex("\\(.*\\)")) %>% 
           str_trim()) %>% 
    mutate(first_name=str_extract(name_clean, regex("^\\S*"))) %>% 
    mutate(name_gender=map(first_name, possibly(gender,
                                                otherwise="missing"))) %>% 
  unnest_wider(name_gender, names_repair="universal") %>% 
  rename(name=name...1) %>% 
  select(-contains("year"), -contains("proportion"), -ends_with("7")) %>% 
  mutate(gender=case_when(str_detect(name, "Brandstötter")==TRUE ~ "female",
                          str_detect(name, "El-Nagashi")==TRUE ~ "female",
                          str_detect(name, "Sibylle")==TRUE ~ "female",
                          str_detect(name, "Himmelbauer")~"female",
                          str_detect(name, "Karlheinz")==TRUE ~ "male",
                          str_detect(name, "Jörg")==TRUE ~ "male",
                          str_detect(name, "Volker")==TRUE ~ "male",
                          str_detect(name, "Gertraud")==TRUE ~ "female",
                          str_detect(name, "Nurten")==TRUE ~ "female",
                          str_detect(name, "Süleyman")==TRUE ~ "female",
                          str_detect(name, "Ribo")==TRUE ~ "female",
                          TRUE ~ as.character(gender))) %>% 
   mutate(name_clean=case_when(str_detect(name, "Nurten")==TRUE ~ "Nurten Yilmaz",
                              str_detect(name, "Herr")==TRUE ~ "Julia Herr",
                              str_detect(name, "Schellhorn")==TRUE ~ "Sepp Schellhorn",
                              str_detect(name, "Lercher")==TRUE ~ "Max Lercher",
                              str_detect(name, "Niss")==TRUE ~ "Maria Niss",
                              TRUE ~ as.character(name_clean))) %>% 
  left_join(., df_nr %>% 
              select(name_clean, party),
            by=c("name_clean"="name_clean")) %>%
  mutate(party=case_when(str_detect(name, "Strache")==T ~ "independent",
                         TRUE ~ as.character(party)))

df_position_long <- df_count2 %>% 
  select(name_clean, party, gender, ends_with("_n")) %>% 
  pivot_longer(cols=ends_with("_n"), 
               names_to="position",
               values_to="number") %>% 
  mutate(party=fct_infreq(party))
```

## Paid positions and income category
Let's have a look at MPs' number of paid positions and income category.

Paid are the positions stated in the category 'other positions' and 'executive position'. As for the latter, however, the list also includes unpaid executive positions which are marked with an asterisk (\*). Hence these positions have to be removed when calculating the total number of paid positions. For details see [below](#unpaid_executive).

Hover over the points in the plot below to get details on each individual MP. Note that the points are located randomly *within* each square to avoid overplotting.

<details closed>
<summary>Code: Number of paid positions and income category</summary>
```{r}
df_income_positions <- df_count2 %>% 
  left_join(., df_income %>% select(gender, income_category_19, name),
            by="name") %>% 
  select(name_clean, name, contains("executive"), contains("other"), party, income_category_19) %>% 
  mutate(positions_paid=(executive_n-executive_non_remunerated_n)+other_n)

df_income_positions <- df_income_positions %>% 
  filter(income_category_19!="missing") %>% 
  mutate(across(.cols=c(positions_paid, income_category_19), list(fct=~as_factor(.) %>% fct_inseq)))

income.key <- levels(factor(df_income_positions$income_category_19))
position.key <- seq(0, max(df_income_positions$positions_paid), 1) %>% 
  as.character()

pl_income_positions <- df_income_positions %>% 
  mutate(income_category_19_fct=fct_rev(income_category_19_fct)) %>% 
  mutate(positions_paid_fct=fct_expand(positions_paid_fct, position.key) %>% 
           fct_inseq()) %>% 
  mutate(party=fct_infreq(party)) %>% 
  ggplot()+
  labs(x="number of paid positions",
       y="[income category] average monthly income",
       title="Austrian Nationalrat: MPs' number of paid positions and income category",
       subtitle="Members of lower chamber only (Nationalrat). Data for 2019.\nMove mouse over dots to get info on MPs.",
       caption=my_caption)+
  geom_jitter_interactive(aes(x=positions_paid,
                                       y=income_category_19,
                                       color=party,
                                       tooltip=paste0(name_clean, " ", party, "\n", "paid positions: ", positions_paid, "\n", "income category: ", income_category_19)))+
  scale_color_manual(values=party_colors)+
  scale_y_discrete(labels=income_categories_short)+
  facet_grid(income_category_19_fct ~ positions_paid_fct,
             drop=FALSE,
             scales="free",
             switch="x")+
  theme_post()+
  theme(panel.border = element_rect(color = "black", fill = NA, size = .5),
        panel.grid.major = element_blank(),
        panel.spacing.y = unit(0, "cm"),
        panel.spacing.x = unit(0, "cm"),
        legend.position = "bottom",
        legend.title=element_blank(),
        legend.direction = "horizontal",
        axis.text.x = element_blank(),
        axis.title.y=element_text(color="grey50",
                                  hjust=1,
                                  angle=90),
        strip.text.x = element_text(hjust=0.5, color="grey50",
                                    face="plain"),
        strip.text.y = element_blank())+
  guides(color=guide_legend(nrow=1))
```
</details>

```{r echo=FALSE, fig.height=15}
girafe(ggobj = pl_income_positions)

```

Overall, there seems to be a positive association between the number of positions and income category. However, bear in mind that we do not know how much income is generate from each specific position. We only know about the sum and shouldn't be tempted to infer any average salaries per position. E.g. while Herbert Fuchs' four paid positions yield in sum an income of more than 10,000 Euros per average month (income category 5), it could be that the bulk of the income sum originates from one particularly position only. This is particularly relevant if we think about comparing MPs within the same income category but different number of paid positions.

A further qualifying note on the lower end of the income scale: Note that there are five curious cases of MPs having at least one paid position, but zero additional income. How can that be? Let's look at these five cases more closely.

<details closed>
<summary>Code: MPs with no income but paid positions</summary>
```{r}
df_mp5 <- df_income_positions %>% 
  filter(income_category_19==0) %>% 
  filter(positions_paid>0) %>% 
  select(contains("name"), party, executive_n, positions_paid) %>% 
  left_join(., df_positions %>% select(name, executive_positions, other_positions), by=c("name"="name")) %>% 
  pivot_longer(cols=c(contains("positions"), -positions_paid),
               names_to="positions",
               values_to="positions_name") %>% 
  mutate(positions_chr=map(positions_name, deframe) %>% 
           map_chr(., ~paste(., collapse=";\n "))) %>% 
  select(-positions_name) %>% 
  pivot_wider(id_cols=c(name, party),
              names_from=positions,
              values_from=positions_chr)

tb_mp5 <- df_mp5 %>% 
  gt() %>% 
  tab_header(
    title=md("**MPs with zero income, but paid positions**")) %>% 
   tab_footnote(footnote = "Honorary positions exclude since unpaid.",
    locations = cells_title(group="title")) %>% 
  cols_label("name"=md("**MP**"),
             "party"=md("**party**"),
             "executive_positions"=md("**executive positons**"),
             "other_positions"=md("**other positions**"))

```
</details>

```{r echo=FALSE}
tb_mp5
```

As for Neos MPs' Hoyos-Trauttmansdorff and Scherak a look at the parliament's list reveals that their declaration features an additional note indicating that both assumed their position only from April 2020 onward. Hence, there was no pertaining income for the 2019 category. Scherak's second executive position ('Margaretha Lupac - Stiftung') is unpaid as indicated by the \*. Hence, these two cases are clear.

The other three cases are less clear to me. Eva-Maria Himmelbauer has an executive position in the Cloudcompany and there is no \* sign which would indicate that the position was without remuneration. What's furthermore noteworthy, if you look at the table above, it states 'Ablauf der Meldefrist: Geschäftsführerin 23.11.2019'. Remarkably, this text is not visible in the original pdf on the parliament's website. I have no idea why and how, but the  used `pdf_data` function extracted also 'hidden' text from the pdf. Whether this text was a comment to an e.g. Excel cell or a hidden row/column, I don't know. But irrespective of this, there is nothing in the additional text what would suggest - at leat to me - that Himmelbauer did not receive at least for one month a salary during the reporting period. 
Similarly, Markus Koza and Meri Dioski had both paid positions for their own parities up until 31.10.2019. Nevertheless, this income was not reported as indicated by the income category 0.

To be clear, I have no reason to think that there is any malice etc. behind these inconsistencies, and I assume there is some formal, regulatory explanation which hasn't been made explicit in the parliament's list. If anyone has an answer to this I would be keen to hear it.

## Share of MPs with at least one additional position

To move on and, to get some contrast between parties, the barplot below indicates the percentage of MPs per party which do have at least one additional professional position.

<details open>
<summary>Code: MP share w at least one add. position</summary>
```{r}
df_bar_position <- df_position_long %>% 
  filter(!position %in% c("executive_non_remunerated_n",
                        "total_n")) %>% 
  group_by(party, position) %>% 
  summarize(n_zero=length(number[number==0]),
            n_non_zero=length(number[number!=0])) %>% 
  ungroup() %>% 
  mutate(perc=n_non_zero/(n_non_zero+n_zero)) %>% 
  mutate(position=fct_relevel(position, "honorary_n", after = 2))

pl_bar_position <- df_bar_position %>% 
  ggplot()+
  labs(title="Austrian Nationalrat: Percentage of MPs with at least one additional position per party",
       subtitle=str_wrap("Percentages are shares of each party's total number of members. Absolute numbers in brackets. Lower chamber only (Nationalrat). Data for 2019.", 105),
       caption=my_caption)+
  geom_bar(aes(x=party,
               y=perc,
               fill=party),
           stat="identity")+
  geom_text(aes(x=party,
               y=perc,
               group=party,
               label=paste0(perc %>% scales::percent(),
                           "\n",
                           "(", n_non_zero,")")),
            nudge_y = 0.05,
            lineheight=0.7,
           stat="identity")+
  scale_fill_manual(values=party_colors)+
  scale_color_manual(values=party_colors)+
  scale_y_continuous(label=scales::label_percent(accuracy = 1),
                     expand=expansion(mult=c(0, 0.15)),
                     breaks=seq(0,1,.25))+
  scale_x_discrete(label=abbreviate)+
  facet_wrap(vars(position),
             labeller = labeller(position=c("executive_n"="executive functions","other_n"="other activities",
                                            "honorary_n"="honorary positions")))+
  theme_post()+
  theme(legend.position="none",
        axis.title.x = element_blank())

```
</details>

```{r echo=FALSE}
pl_bar_position
```

As far as *executive positions* are concerned, the relative numbers of Neos stand out. More than half of Neos' MPs have an executive position. In absolute numbers, the ÖVP with 26 MPs leads in this category.

When it comes to *honorary positions*, there is a considerable gap between the three larger parties (ÖVP, SPÖ, FPÖ) and the two smaller ones (Greens, Neos). One possible explanation (which I have not looked into though) could be that larger parties have more party affiliated associations in which their MPs take up honorary positions. Maybe, and I am purely speculating here, the number of honorary positions is a function of a party's institutional width.

## Executive positions {#unpaid_executive}
Below I further disaggregate executive positions into those which entail a salary and those which don't. In the parliament’s report, executive positions without salary were marked with an asterisk (\*). When retrieving the data above, I already accounted for this difference (`str_detect(., "\\*")`). The graph below now shows the percentage of MPs per party without executive positions, those with at least one executive position without salary (and no position with salary), and those with at least one remunerated executive position.

<details open>
<summary>Code: different types of executive positions</summary>
```{r}
df_executive <- df_count2 %>% 
    select(party, contains("executive")) %>% 
    mutate(executive_remunerated_n=executive_n-executive_non_remunerated_n) %>% 
  group_by(party) %>% 
  summarize(executive_no=length(executive_n[executive_n==0]),
    remunerated_yes=length(executive_remunerated_n[executive_remunerated_n>0]),
    remunerated_no=length(executive_remunerated_n[executive_remunerated_n==0 &
                            executive_n>0])) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(total=sum(c_across(where(is.numeric)), na.rm=T))

pl_executive <- df_executive %>% 
  pivot_longer(cols=-c(party, total),
               values_to="number",
               names_to="remunerated") %>% 
  mutate(party=fct_reorder(party, total) %>% fct_rev) %>%   mutate(perc=number/total) %>% 
  mutate(remunerated=case_when(str_detect(remunerated, regex("executive_no"))==TRUE ~"% MPs without executive positions",                               str_detect(remunerated, regex("remunerated_no"))==TRUE ~ "% MPs with non-remunerated executive positions",
                               str_detect(remunerated, regex( "remunerated_yes"))==TRUE ~ "% MPs with remunerated executive positions",
                               TRUE ~ as.character(remunerated))) %>% 
  mutate(remunerated=fct_relevel(remunerated,
                                 "% MPs with non-remunerated executive positions", after = 1L) %>% fct_rev) %>%
  ggplot()+
  labs(title="Austrian Nationalrat: Percentage of MPs with different executive positions per party",
       subtitle=str_wrap("Percentages are shares of each party's total number of members. Absolute numbers in brackets. Lower chamber only (Nationalrat). Data for 2019.", 105),
       caption=my_caption)+
  geom_bar(aes(x=party,
               y=perc,
               fill=party),
           position=position_dodge(),
           stat="identity")+
  geom_text(aes(x=party,
                y=perc+.05,
                label=paste0(scales::percent(perc,
                                             accuracy = 0.1),
                             "\n",
                            "(", number,")"),
                group=remunerated),
            lineheight=0.7,
            position=position_dodge2(width = 1),
            stat="identity")+
  scale_y_continuous(labels=scales::label_percent(),
                     expand=expansion(mult=c(0, .1)))+
  scale_x_discrete(labels=abbreviate)+
  scale_fill_manual(values=party_colors)+
  theme_post()+
  theme(legend.position="none",
        axis.title.x = element_blank())+
  facet_wrap(vars(remunerated),
             labeller=label_wrap_gen(width=30))

```
</details>

```{r echo=FALSE}
pl_executive
```

Note that there is only one MP from the Greens with a paid executive position. Also, there are only three Socialdemocrats in this category. Again, the ÖVP features most MPs with a paid executive position in absolute terms, and Neos in relative terms.

## Parties in comparison (Boxplot)
The boxplot below might be a bit of an overkill, but it gives us a more detailed overview of MPs' additional positions per party. Each dot indicates the number of positions which one specific MP holds in one category of professional positions. In addition, the median, mean, max and minimum values for each party/position combination are provided. Note the comparably high median value of honorary positions among SPÖ MPs.

<details closed>
<summary>Code: boxplots</summary>
```{r echo=TRUE, fig.height=7}
df_position_labels <- df_position_long %>% 
  filter(!position %in% c("executive_non_remunerated_n",
                        "total_n")) %>% 
  group_by(party, position) %>% 
  mutate(party=fct_infreq(party)) %>% 
  summarise(label_median=median(number),
            label_mean=mean(number),
            label_max=max(number),
            label_min=min(number))

pl_box <- df_position_long %>% 
  filter(!position %in% c("executive_non_remunerated_n",
                        "total_n")) %>% 
  mutate(party=fct_infreq(party)) %>% 
  ggplot()+
  labs(caption=my_caption,
       title="Austrian Nationalrat: Number of MPs' other professional positions per party",
       subtitle="Members of lower chamber only (Nationalrat); Data for 2019.",
       x="MP's number of positions")+
  geom_boxplot(aes(y=party,
                  x=number,
                  group=party,
                  color=party),
               orientation="y",
               outlier.shape=NA,
               width=0.1
               )+
  geom_jitter(aes(y=party,
                  x=number,
                  group=party,
                  colour=party),
              alpha=.2,
              orientation="y",
              height=0.1
              )+
  geom_label(data=df_position_labels,
            aes(y=1.2,
                x=0,
                group=party,
                label=paste("median:", round(label_median, 2),
                            # "\n",
                            "mean:", round(label_mean, 2),
                            # "\n",
                            "min:", round(label_min, 2),
                            # "\n",
                            "max:", round(label_max, 2)
                            )),
            fill="white",
            label.padding = unit(0.2,"line"),
            label.r=unit(0.2,"line"),
            label.size = 0,
            family="Roboto Condensed",
            size=3,
            vjust=0,
            hjust=0,
            orientation="y",
            lineheight=0.5)+
  scale_color_manual(values=party_colors)+
  scale_x_continuous(expand=expansion(mult=c(0, .2)))+
  scale_y_discrete(expand=expansion(mult=c(0, .3)),
                   label=abbreviate)+
  theme_post()+
  theme(legend.position = "none",
        panel.grid.major.x = element_line(color="gray30"),
        panel.grid.major.y = element_blank(),
        strip.text.y = element_blank())+
  lemon::facet_rep_grid(party~position,
                        repeat.tick.labels = TRUE,
                        scales = "free_y",
                        labeller = labeller(position=c("executive_n"="executive functions","other_n"="other activities",
"honorary_n"="honorary positions")))

```
</details>
```{r echo=FALSE}
pl_box
```

# Wrap up
So this has become a rather length blog entry. I hope there's been something useful in it for you as well. Again, the emphasis has been largely on various steps in R to extract and subsequently analyze the data of interest. If there's anything you feel like following up, feel free to let me know.

```{r}
df_nr<- df_count %>% 
  left_join(., df_income %>% 
              select(-ends_with("6"), -first_name),
            by=c("name"="name")) %>% 
  select(name, name_clean, party, gender, income_category_19, contains("_n"))

#readr::write_excel_csv2(df_nr, path=here::here("_blog_data", "MP_activities", "NR.csv"))

```

### 2021-01-08 _post/2020-04-28-google-s-mobility-reportsmd/google-s-mobility-reportsmd.Rmd
add google mobility post

---
title: "Google's Mobility reports: changes to mobility patterns during Covid-19 lock-down in Austria"
description: |
  Analysis of Google's mobility reports following introductino of Covid lockdown in March 2020.
date: 04-28-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Corona virus
  - Austria

#preview: submissionsepidemiclaw_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

```{r message=FALSE, include=FALSE}

library(tidyverse)
library(ggplot2)
library(here)
library(hrbrthemes)
library(extrafont)
loadfonts(device = "win", quiet = T)
extrafont::fonts()
library(scales)
library(grid)
library(knitr)
library(geofacet)
library(paletteer)
library(ggtext)

```

```{r setup, echo = F}
# knit_hooks$set(wrap = function(before, options, envir){
#   if (before){
#     paste0('<', options$wrap, '>')
#   } else {
#     paste0('</', options$wrap, '>')
#   }
# })
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dpi = 180, fig.height = 5, fig.width = 7, fig.align = "center")
options(width=180, dplyr.width = 150)

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()

```

# CONTEXT

Here’s a quick post related to Covid-19, but don’t be afraid, it won’t be another infection et al visualization (see [here](https://werk.statt.codes/post/covid-19-corona-virus-quick-look-at-some-data/){target="_blank"}, but also [here](https://www.williamrchase.com/post/why-i-m-not-making-covid19-visualizations-and-why-you-probably-shouldn-t-either/){target="_blank"}). Google and Apple recently released mobility reports which provide some insights on how the lockdown to curtail the epidemic’s spread affected individuals' mobility patterns. While [Apple’s reports](https://www.apple.com/covid19/mobility){target="_blank"} draw on to routing requests to its Apple Map service, Google’s mobility reports highlight how people’s (= mobile device owner who activated their geo-location) presence on specific type of places changed relatively to a baseline period preceding the outbreak of the crisis and the introduction of the lockdown. You can find more details about the dataset [here]( https://www.google.com/covid19/mobility/){target=”_blank” }.

The dataset has been quickly picked up by numerous media outlets and data analysts, I nevertheless thought it might be interesting to have a specific look at the data for Austria. Again, this is predominately an exercise in number crunching and not meant to suggest any specific insights of an epidemiological nature etc. I leave this to the actual experts.

On a more technical level, with this post I also want to highlight the wonderful `geofacet` package by Ryan Hafen  ([link]( https://hafen.github.io/geofacet/){target=”_blank”}). The package, which I used already in a [previous post](https://werk.statt.codes/post/vorzugsstimmen/#candidates-preference-votes-and-electoral-number){target="_blank"}, is an extension to ggplot2, and allows you to arrange a plot’s facets to reflect the position of geographical units on a map. While the facets’ location is a simplification and will hardly reflect the precise and actual size of e.g. a federal state, the approach provides a lucid way of presenting data for different geographical areas.

As always, the entire code for this analysis is available on my github account [here](https://github.com/werkstattcodes){target="_blank"}. If you see any glaring error etc, or if something is unclear, feel free to let me know, best via a direct message at [twitter](https://twitter.com/zoowalk){target="_blank"}.

# IMPORT DATA
But first, let’s import the data. Initially, Google had published only summary graphs (in pdfs) and not the underlying data (what prompted to some very elegant data extraction [exercise](https://joachim-gassen.github.io/2020/04/scrape-google-covid19-cmr-data/){target="_blank"}). But gladly, by now, a csv-file with all data is available on the website.

<details open> 
<summary>Click to hide code</summary>

A few things to highlight: When importing data, the `readr` package generally tries to 'guess' the data type (character, double etc.) by inferring it from the first 1000 lines of the imported data. In some instances, this might not be sufficient or misleading. The **col_types argument** allows to manually specify each column, with **first characters as convenient abbreviations** (c = character, D = date, d = double etc). For details see [here](https://readr.tidyverse.org/articles/readr.html#column-specification){target="_blank"}.
Dplyr 1.0 will provide an option to define the location of a newly created column with the .after/.before argument.

```{r}
# import csv --------------------------------------------------------------
# file_link <- "https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv"
# 
# df_global <- readr::read_csv(file=file_link, col_types = "ccccDdddddd")

df_AT_mobility <- readr::read_csv2(file=here::here("_blog_data", "update_mobility_reports", "AT_mobility_report.csv"))

df_AT <- df_AT_mobility %>% 
  filter(country_region_code=="AT") %>% 
  mutate(date=as.Date(date)) %>% 
  filter(date<as.Date("2020-05-01")) %>% 
  mutate(sub_region_1=case_when(is.na(sub_region_1) ~ "State level",
                   TRUE ~ as.character(sub_region_1)))

df_AT_long <- df_AT %>% 
  pivot_longer(cols=contains("baseline"),
                      names_to="type",
                      values_to="value") %>% 
  mutate(week.day=lubridate::wday(date, label=T), .after=date) %>% 
  mutate(sub_region_1=forcats::as_factor(sub_region_1))

data_as_of <- max(df_AT_long$date, na.rm = T)

#ad description of place types from Google's documentation

df_AT_long <- df_AT_long %>% 
  mutate(place.description=case_when(str_detect(type, "retail") ~ "‘places like restaurants, cafes, shopping centers, theme parks, museums, libraries, and movie theaters.’",
                                     str_detect(type, "grocery") ~ "’places like grocery markets, food warehouses, farmers markets, specialty food shops, drug stores, and pharmacies.’",
                                     str_detect(type, "park") ~ "‘places like local parks, national parks, public beaches, marinas, dog parks, plazas, and public gardens.’",
                                     str_detect(type, "transit") ~ "‘places like public transport hubs such as subway, bus, and train stations.’",
                                     str_detect(type, "work") ~ "‘places of work’",
                                     str_detect(type, "residential") ~ "‘places of residence’",
                                     TRUE ~ as.character("missing"))) %>% 
  mutate(date=as.Date(date))

```
</details>

The mobility reports contain data on the following types of places:

```{r echo=FALSE, message=FALSE, warning=FALSE}
unique(df_AT_long$type)
```

<details open> 
To distinguish mobility patterns prior and during the lockdown we need to define the relevant dates.
<summary>Click to hide code</summary>
```{r}
# set dates ---------------------------------------------------------------
date.start.lockdown <- as.Date("2020-03-16")
date.opening.shops <- as.Date("2020-04-14")

df_dates <- tibble(date.start.lockdown=date.start.lockdown,
                   date.opening.shops=date.opening.shops) %>% 
  pivot_longer(cols=everything(),
               names_to = "event",
               values_to="date")
```
</details>

# GEOFACET GRID

For `geofacet` to be able to position ggplot's facets in (approximate) accordance with units' actual geographical location, it requires a grid which specifies each facet's relative location. The package comes already with a number of predefined grids, and users can upload newly made grids to the package's github. For the present case, I created a new one defining the location of Austria's nine federal states. As suggested by the package's maintainer, the code are ISO-3166-2 codes. The new grid is now uploaded to geofacet's github and can be called by other users by the name AUT_states_grid.

<details open> 
<summary>Click to see code</summary>
```{r}
AUT_states_grid <- data.frame(
  code = c("4", "3", "9", "8", "7", "5", "6", "1", "2"),
  name_en=c("Upper Austria", "Lower Austria", "Vienna", "Vorarlberg", "Tyrol", "Salzburg", "Styria", "Burgenland", "Carinthia"),
  row = c(1, 1, 1, 2, 2, 2, 2, 2, 3),
  col = c(3, 4, 5, 1, 2, 3, 4, 5, 4))

```
</details>

```{r message=FALSE, warning=FALSE}
AUT_states_grid
```

# MAP OF MOBILITY PATTERS IN PARKS

As mentioned above, the mobility reports contain data on specific types of places. Let's plot the dynamics for one of them: Parks.

<details closed> 
<summary>Click to see code</summary> 
```{r}
txt_subtitle <- glue::glue("Changes for each day are compared to a baseline value for that day of the week: The baseline is the median value, for the corresponding day of the week, during the 5-week period Jan 3–Feb 6, 2020. For details see https://www.google.com/covid19/mobility/. Data as of {data_as_of}.")

date_shutdown <- as.Date("2020-03-16")

df_AT_long_type <- df_AT_long %>% 
  filter(str_detect(type, "parks")) %>% 
  filter(!str_detect(sub_region_1, "State"))

pl_map <- df_AT_long_type %>% 
  ggplot()+
  labs(title=paste("Changes in Mobility during Covid-19 crisis:",                  str_to_upper(str_remove_all(unique(df_AT_long_type$type), regex("_percent.*$")))),
       subtitle=str_wrap(txt_subtitle, 130),
       caption=c("data: Google Mobility Report", "graph: Roland Schmidt | @zoowalk| https://werk.statt.codes/"))+
  geom_hline(yintercept=0,
             color="grey20")+
  geom_line(data=df_AT_long %>% 
              filter(str_detect(type, "parks")) %>% 
              filter(!str_detect(sub_region_1, "State")) %>% 
              rename(all_regions=sub_region_1),
            aes(x=date,
                y=value,
                group=all_regions),
            color="grey70",
            show.legend=F,
            stat="identity") +
  geom_line(aes(x=date,
                y=value,
                color=sub_region_1),
            show.legend=F,
            size=1)+
  geom_vline(data=df_dates,
             aes(xintercept=date,
                 linetype=event),
             color="red")+
  scale_linetype_manual(values=c(date.opening.shops="dotted",
                                 date.start.lockdown="solid"),
                        labels=c(date.opening.shops="shops partly open",
                                 date.start.lockdown="start lockdown"))+
  theme_ipsum_rc()+
  scale_x_date(labels = scales::label_date_short(),
               expand=expansion(mult=c(0, 0.15)),
               breaks = c(seq.Date(min(df_AT_long$date), 
                                   max(df_AT_long$date), 
                                   by="2 week"))) +
  scale_y_continuous(labels=scales::label_percent(accuracy = 1,
                                                  scale=1),
                     limits=c(-100, 100))+
  scale_color_paletteer_d("ggsci::default_uchicago")+
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 9),
    axis.text.y = element_text(size = 9),
    axis.title.y=element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.text.y = element_text(angle = 0, vjust = 1),
    legend.position = "bottom",
    legend.justification = "right",
    legend.key.size=unit(0.5, "cm"),
    legend.key.height = unit(0.5, "cm"),
    legend.title = element_blank(),
    legend.text = element_text(size=10),
    # plot.title = element_text(size = 12),
    panel.spacing.x = unit(0.2, "cm"),
    panel.spacing.y = unit(0.1, "cm"),
    plot.margin=margin(l=0, r=0, unit="cm"),
    plot.title.position = "plot",
    plot.subtitle = element_text(size=9, color="grey30"),
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = c(0, 1), color = "grey30"),
    plot.background = element_rect(fill=plot_bg_color, color=plot_bg_color)
  )+
  facet_geo(~sub_region_1, grid = AUT_states_grid)+
  guides(linetype = guide_legend(override.aes = list(colour = "red",
                                                     direction="vertical"),
                                 reverse = T),
         colour = "none")
```
</details>

```{r echo=FALSE, fig.height=6, layout="l-body-outset"}
pl_map

```

Before diving into the interpretation let's recall what the displayed values actually mean. In Google's [own words](https://www.google.com/covid19/mobility/data_documentation.html?hl=en){target="_blank"} *'the data show how visits and lengths of stay at different places change compared to a baseline. Changes for each day are compared to a baseline value for that day of the week. The baseline is the median value, for the corresponding day of the week, during the 5-week period Jan 3–Feb 6, 2020. The changes are based on data from users who have opted-in to Location History for their Google Account, so the data represents a sample of users.'*

So with this in mind, let's look at the graph which depicts the presence in parks. The colored line is the values for the state, grey lines are the values for all other states for the sake of contrast.

I think it's pretty clear that either slightly before or from the introduction of the lockdown onward, individuals' presence in parks decreased considerably. At a closer look, a few variations appear.

When it comes to the *period before the lockdown*, the values for Vorarlberg, Tyrol, Salzburg and Carinthia started decreasing from a relative high level already prior to the lockdown. I don't have any definite answer for that, but I could imagine that mid-February the skiing season had a final peak with schools (e.g. in Germany) having a term break and tourists flocking in strong numbers to the slopes (compared to the baseline). Subsequently, with the skiing season coming to a gradual end, the high numbers started to decrease. But again, whether this dynamic had an impact on parks is purely speculative. In contrast, in most of the other states, the presence level in parks was rather stable prior to the lockdown.

By and large, the lockdown on 16 March lead to a clearly visible reduction in parks. Oddly, there seems to have been some delay in lockdown's impact in Lower Austria and Burgenland.

There is also some variation when it comes to the *dynamics during the lockdown*. While the presence in parks stayed in most states on a comparably low level, Lower Austria, Burgenland and Upper Austria gradually 'rebounded' and the individuals' presence in parks increased again to pre-lockdown levels. Also in Vienna, the presence in parks grew, albeit slower and so far not beyond the baseline level. I don't have any explanation for that, but could imagine that it has something to do with individuals' options to leave their flat to get some fresh air. While in more urban states, parks are the probably the destination of choice, more rural states offer a variety of other 'escapes'. But that's only me speculating here.

# BOXPLOT ON ALL TYPES OF PLACES

Finally, to better distill the changes between before and during the lockdown as well as between the federal states, the plots below present (half) a boxplot summarizing the relative changes of mobility patterns for **all types of places** provided in Google's mobility reports. To supplement the boxplot, I also plot the actual values for each day (= each dot) which represent the relative change of each day to the corresponding weekday in the baseline period. The package `gghalves` allows for combining 'half' plots of two types. Iterating the ggplot command over all types of places covered by Google's mobility reports gives us one map per place type.

To see whether weekends are in any way different, I marked them in red.

<details closed> 
<summary>Click to see code</summary>
```{r}
# boxplot before after ----------------------------------------------------

df_AT_long <- df_AT_long %>% 
  mutate(lockdown.indicator=case_when(date>=date.start.lockdown ~ "lockdown",
                                      date<date.start.lockdown ~ "before lockdown")) %>% 
  mutate(weekend.indicator=case_when(week.day %in% c("Sat", "Sun") ~ "weekend",
                                     TRUE ~ as.character("not weekend"))) #%>%

fn_map <- function(df, my_type) {

  df %>% 
  filter(str_detect(type, my_type)) %>% 
  filter(!str_detect(sub_region_1, "State")) %>% {
    ggplot(., aes(x=lockdown.indicator,
                  y=value))+
      labs(title=paste("Changes in Mobility during Covid-19 crisis:", 
                       str_to_upper(str_remove_all(unique(.$type), regex("_percent.*$")) %>% str_replace_all(., "_", " "))),
           # subtitle=str_wrap(txt_subtitle, 100),
           # subtitle=str_wrap(paste(txt_subtitle, "\n", unique(.$tpye), "are defined as", unique(.$place.description)), 100),

                      subtitle=str_c(str_wrap(txt_subtitle, 130), 
                        str_wrap(paste(str_to_sentence(str_remove_all(unique(.$type), regex("_percent.*$"))) %>% str_replace_all(., "_", " "),  
                                   "are defined as", 
                                   unique(.$place.description)), 130)),

           caption=c("data: Google Mobility Report", 
                          "graph: Roland Schmidt | @zoowalk | http://werk.statt.codes"))+
      gghalves::geom_half_boxplot(side="l",
                                  aes(color=NULL),
                                  outlier.shape = NA,
                                  # nudge=0.2,
                                  show.legend = F)+
      gghalves::geom_half_point_panel(aes(color=weekend.indicator),
                                side="r",
                                range_scale = 0.75,
                                size=0.4,
                                show.legend = T,
                                # nudge=0.2,
                                transformation=position_jitter(width=0.8))+
      scale_x_discrete(labels=c("before lockdown"="before", "lockdown"="during"))+
      scale_color_paletteer_d("ggsci::default_jama")+
      scale_y_continuous(limits=c(-100, 100),
                         labels = scales::label_percent(scale = 1),
                         minor_breaks = NULL)+
      facet_geo(~sub_region_1, grid = AUT_states_grid)+
      theme_ipsum_rc()+
  scale_color_paletteer_d("ggsci::default_uchicago")+
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 9),
    axis.text.y = element_text(size=9),
    axis.title.y=element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.text.y = element_text(angle = 0, vjust = 1),
    legend.position = "top",
    legend.justification = "left",
    legend.key.size=unit(0.5, "cm"),
    legend.key.height = unit(0.5, "cm"),
    legend.title = element_blank(),
    legend.text = element_text(size=10),
    panel.spacing.x = unit(0.2, "cm"),
    panel.spacing.y = unit(0.1, "cm"),
    plot.margin=margin(l=0, r=0, unit="cm"),
    plot.title.position = "plot",
    plot.caption.position = "plot",
    plot.background = element_rect(fill=plot_bg_color, color=plot_bg_color),
    plot.subtitle=element_text(size=9, color="grey30"),
    plot.caption = element_text(hjust = c(0, 1), color = "grey30"))+
  facet_geo(~sub_region_1, grid = AUT_states_grid,
            labeller=as_labeller(function(x) str_remove_all(x, regex("_percent.*$")) %>%
                                   str_replace_all(., "_", " ")))+
      guides(color=guide_legend(override.aes=list(size=3)))}
}

pl_list <- unique(df_AT_long$type) %>% 
  map(., ~fn_map(df=df_AT_long, my_type=.x))

```
</details>

## Retail and Recreation

```{r echo=FALSE, fig.height=6, message=FALSE, warning=FALSE, layout="l-body-outset"}
pl_list[[1]]
```

## Grocery and Pharmacy

```{r echo=FALSE, fig.height=6, message=FALSE, warning=FALSE, layout="l-body-outset"}
pl_list[[2]]
```

## Parks

```{r echo=FALSE, fig.height=6, message=FALSE, warning=FALSE, layout="l-body-outset"}
pl_list[[3]]
```

## Transit stations

```{r echo=FALSE, fig.height=6, message=FALSE, warning=FALSE, layout="l-body-outset"}
pl_list[[4]]
```

## Workplaces

```{r echo=FALSE, fig.height=6, message=FALSE, warning=FALSE, layout="l-body-outset"}
pl_list[[5]]
```

## Residential

```{r echo=FALSE, fig.height=6, message=FALSE, warning=FALSE, layout="l-body-outset"}
pl_list[[6]]
```

Again, I'll refrain from what might become far-fetched interpretations, but what is noteworthy is the overwhelmingly narrow spread of values during the lockdown with the exception of parks. I was a bit puzzled by the relatively weak increase of stays in residential areas compared to the baseline period. Not sure what to make of this.

### 2021-01-08 _post/2020-06-08-update-on-mobility-patterns/update-on-mobility-patterns.Rmd
post update mobility reports; plot bkg color linked to css bkgd

@@
 library(gt)
+
 ```
@@
 options(width=180, dplyr.width = 150)
+
+plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
+  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()
+
 ```
@@
         legend.position="bottom",
-        strip.text.y = element_text(hjust=1)
+        strip.text.y = element_text(hjust=1),
+        plot.background = element_rect(fill=plot_bg_color, color=plot_bg_color)
   )+
@@
         legend.direction = "horizontal",
+                plot.background = element_rect(fill=plot_bg_color, color=plot_bg_color),
         strip.text.y = element_text(hjust=1))+

### 2021-01-08 _post/2020-04-23-covid-19-corona-virus-quick-look-at-some-data/covid-19-corona-virus-quick-look-at-some-data.Rmd
add Ãpost covi data

---
title: "Corona Virus in Austria: Updates on data"
description: |
  Preliminary look at the COVID data on Austria.
date: 04-23-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - Corona Virus

# preview: welcome_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

**Graphs updated on `r format(Sys.Date(), '%A, %d %b %Y')`**

With the COVID epidemic taking a pretty firm grip on our lives, I thought I’ll use this post as a kind of analytic repository regarding the developments in Austria. There are certainly plenty of other, more advanced analyses out there, so see this as a kind of personal attempt to get a grip on the virus and what is happening around us. Hence, as a caveat, this is purely about crunching numbers. I am no epidemiologist, let alone physician who would have any substantive understanding of the matter. Yet, I think looking at ‘the data’ and bearing the above caveat in mind, can be a worthwhile exercise.

I will try to run the code generating this site once a day, so graphs and text will automatically update. If time permits, new analyses will be added. In terms of interpretation, I’ll be rather brief due to lack of time (kids are at home, partner on home office duty…), but I guess most of the graphs are self-explanatory.

As always, and since this is also a blog about R, you can find the code in the folded sections. If you think there is a glaring error etc., don’t hesitate and let me know (best via Twitter [DM](https://twitter.com/zoowalk){target="_blank"}).

```{r message=FALSE, include=FALSE}
library(gtrendsR)
library(tidytext)
library(tidyverse)
library(ggtext)
library(hrbrthemes)
library(extrafont)
library(paletteer)
loadfonts(device = "win", quiet = T)
library(patchwork)
library(magrittr)
library(ggforce)
library(scales)
library(glue)
library(here)
library(knitr)
library(wbstats)
library(rvest)
library(gganimate)

```

```{r setup, echo = F}
# knit_hooks$set(wrap = function(before, options, envir){
#   if (before){
#     paste0('<', options$wrap, '>')
#   } else {
#     paste0('</', options$wrap, '>')
#   }
# })
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dpi = 180, fig.height = 5, fig.width = 7, fig.align = "center")
options(width=180, dplyr.width = 150)

country_color=c("Italy"="lightgreen", "Austria"="yellow", "China"="red")
my_caption <- c("Graph: Roland Schmidt | @zoowalk | <span style='color:black'>**https://werk.statt.codes**</span>")

# bgr_color="white"

bgr_color<- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()

theme_rs <- function(){
  theme_ipsum_pub() %+replace%
  #hrbrthemes::theme_ipsum_rc() %+replace%
  #hrbrthemes::theme_ipsum_tw()+

    theme(panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_line(linetype="dotted"),
          panel.grid.minor.x = element_line(linetype="dotted"),
          axis.title.x=element_blank(),
          legend.position = "top",
          legend.justification = "left",
          legend.title = element_blank(),
          strip.text = element_text(face="bold", 
                                    hjust=0),
          plot.margin=margin(l=0, t=0.5, b=0.5, unit="cm"),
          plot.background=element_rect(fill=bgr_color, color="transparent"),
          plot.title = element_text(size = 14, 
                                    color="grey20",
                                    hjust=0,
                                    #face="bold", 
                                    margin=margin(b=0.1, unit="cm")),
          plot.title.position = "plot",
          plot.subtitle = element_text(size = 10, 
                                       hjust=0,
                                       margin=margin(b=0.5, unit="cm"),
                                       color = "grey30"),
          plot.caption.position = "plot",
          plot.caption = element_markdown(color = "grey30", 
                                          margin=margin(t=0.5, unit="cm"),
                                          hjust = 1))
}

```

# Getting data on the COVID-19

The data for this analysis is taken from Johns Hopkins COVID-19 data repository, which seems to be one of, if not the most frequently used data source on Covid-19. Have a look at their [site](https://coronavirus.jhu.edu/){target="_blank"} for more details.

<details closed> 
<summary>Click to see code to get data</summary>
Get the raw data from the repository with the `readr` package, clean the names with `janitor`; then flip the data into a long format, and summarize to get country totals (as opposed to amounts for subregions). The `datatable` function from the `DT` package quickly produces a searchable html table.

```{r}
# hopkins_link <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv"

hopkins_link <-"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"

df_corona <- readr::read_csv(hopkins_link) %>% 
  janitor::clean_names()

df_corona_long <- df_corona %>% 
  select(-lat, -long) %>% 
  pivot_longer(cols=-c("province_state", "country_region"),
               names_to = "date",
               values_to = "n_obs") %>% 
  mutate(date=str_remove(date, "x") %>% 
           str_replace_all(., "_", "-") %>% 
           lubridate::mdy(.))

df_corona_long <- df_corona_long %>% 
  filter(date < as.Date("2020-04-23"))

df_corona_long <- df_corona_long %>% 
  group_by(country_region, date) %>% 
  summarise(n_obs=sum(n_obs, na.rm=T)) %>% 
  ungroup()
```
</details>

```{r echo=FALSE}
DT::datatable(df_corona_long)
```

# Austria, Italy, and China

To put the Austrian case somewhat into perspective I contrast its numbers with those of Italy and China. The former is the worrying neighbor, representing a scenario which is within the realm of the possible for Austria. Since the absolute number of confirmed cases are quite different, I scaled the numbers on a log scale (note the labels on the y axis).

<details closed> 
<summary>Click to see code  plot</summary>
```{r}
plots_3countries <- df_corona_long %>% 
  filter(str_detect(country_region, "Austria|Italy|China")) %>% 
  ggplot()+
  labs(title="Convid-19: Confirmed cases - Austria, Italy, China",
       subtitle=glue("Case count as of: ", "<span style='color:white'>
                     {format(max(df_corona_long$date), '%A, %d %b %Y')}</span>", 
                       "; ",
                       "Note that <span style='color:white'>y axis is log-transformed </span>.","\n"),
       caption=c("Data: John's Hopkins  COVID-19 Data Repository (github.com/CSSEGISandData/COVID-19)",
                 "\nGraph: Roland Schmidt | @zoowalk | https:://werk.statt.codes"),
       y="number of confirmed cases")+
  geom_line(aes(x=date,
                y=n_obs,
                color=country_region),
            size=1)+
  geom_text(data=. %>% filter(date==max(df_corona_long$date)),
            aes(x=date,
                y=n_obs*1.7,
                label=paste(country_region, 
                            scales::comma(n_obs)),
                color=country_region,
                hjust=1))+
  scale_x_date(breaks=scales::date_breaks(width="1 month"),
               labels = scales::label_date_short(),
               expand=expansion(mult=c(0, 0.05)))+
  scale_y_log10(labels=scales::label_comma())+
  scale_color_manual(values=country_color)+
  theme_ft_rc()+
  theme(legend.position="none",
        plot.caption = element_text(hjust=c(0, 0)),
        panel.grid.minor.y = element_blank(),
        plot.subtitle = element_markdown(),
        axis.title.x = element_blank())
```
</details>
```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE}
plots_3countries
```

In my view, the graph shows quite clearly the staggering dynamic of the virus' spread in Italy. On 20 February, there were only 3 confirmed cases. 10 days later, the number had already risen to 1694 (check the table above for the exact numbers). The graph also shows the flattening curve in China, an indication for the gradual slowdown of the dynamic. As for Austria, the absolute numbers are as of the time of writing still comparably modest, however, there is a clear, steep upward trajectory.

# Comparing Austria's and Italy's dynamics

```{r message=FALSE, include=FALSE}
threshold_value <- 100
```

To better contrast both countries' dynamics, I'll replace the dates on the x-axis with an index which indicates the number of days which have passed since the virus' outbreak in each country. This raises the question when the disease actually breaks out. What is the adequate starting position to contrast the disease's development? The presence of one single case? Or more? Taking the presence of only one case may easily be a distorting threshold since in one country the presence of one case might have been missed for some time. Consequently, the ensuing dynamic might appear steeper than in a country which did not miss the first case. I assume epidemiologists have clear concepts on this, but for my purpose I start the index from the **day the threshold of `r threshold_value` cases is crossed**. In the cases of Italy and Austria, this threshold was crossed on the following dates:

<details closed> 
<summary>Click to see code to set threshold and create index</summary>
```{r}

df_AT_IT <- df_corona_long %>% 
  filter(str_detect(country_region, "Austria|Italy")) %>% 
  group_by(country_region, date) %>% 
  summarise(n_obs=sum(n_obs, na.rm=T)) %>% 
  group_by(country_region) %>% 
  arrange(date, .by_group=T) %>% 
  mutate(threshold_indicator=case_when(n_obs>threshold_value ~ 1,
                                TRUE ~ 0)) %>% 
  filter(threshold_indicator>0) %>% 
  mutate(day_index=row_number())

```
</details>

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
df_AT_IT %>% 
  select(country_region, date) %>% 
  slice(1)
```

<details closed> 
<summary>Click to see code  plot</summary>
To overlap the lines for Austria and Italy, we have to replace the dates of the x-axis with a common index. As a starting point for this index, I chose the day when the number of `r threshold_value`  confirmed cases was crossed for the first time. That's what the first code block does. The second block produces the graph.

```{r}
pl_AT_IT <- df_AT_IT %>% 
  # filter(threshold_indicator>0) %>% 
  # mutate(day_index=row_number()) %>% 
  ggplot()+
  geom_vline(xintercept=df_AT_IT %>% 
               filter(country_region=="Austria") %>% 
               pull(day_index) %>% 
               max(),
             color="white")+
  geom_text(data=.  %>%
               filter(country_region=="Austria") %>%
              tail(1),
            aes(x=day_index,
                y=20000,
                label = format(date, '%a, %d %b %Y')),
            color ="white",
            angle =90,
            nudge_x =-1,
            size=3,
            hjust = 0)+
  geom_line(aes(x=day_index,
                y=n_obs,
                color=country_region))+
  geom_point(data=. %>% filter(day_index==max(df_AT_IT$day_index[df_AT_IT$country_region=="Austria"])),
              aes(x=day_index,
                  size=2,
                y=n_obs,
                color=country_region))+
  labs(title="Covid-19: Comparing dynamics in Austria and Italy",
       subtitle=glue("Starting <span style='color:white'>from day when threshold of {threshold_value} cases was crossed </span>.<br>Data as of <span style='color:white'>{format(max(df_AT_IT$date), '%A, %d %b %Y')}</span>."),
       caption=c("Data: John's Hopkins  COVID-19 Data Repository (github.com/CSSEGISandData/COVID-19)",
                 "\nGraph: Roland Schmidt| @zoowalk | https:://werk.statt.codes"),
       x=glue("# days since more then {threshold_value} confirmed cases"),
       y="# confirmed cases")+
  scale_x_continuous(limits=c(1, NA),
                     breaks=c(1, seq(0, max(df_AT_IT$day_index), 5)),
                     minor_breaks = NULL)+
  scale_color_manual(values=country_color)+

  geom_text(data=. %>%  filter(day_index==max(df_AT_IT$day_index[df_AT_IT$country_region=="Austria"]) &
                                 country_region=="Austria"),
              aes(x=day_index,
                  y=n_obs,
                  color=country_region,
                  label=paste0("latest # Austria;",
                               day_index, " days after threshold:", "\n",
                               comma(n_obs))),
            nudge_x = 0.5,
            nudge_y= 250,
            check_overlap = T,
            hjust=0,
            show.legend = F)+
  geom_text(data=. %>%  filter(day_index==max(df_AT_IT$day_index[df_AT_IT$country_region=="Austria"])&
                                 country_region=="Italy"),
            aes(x=day_index,
                y=n_obs,
                color=country_region,
                label=paste0("Italy ", format(date, '%d %b'), ";",
                             day_index, " days after threshold:", "\n", comma(n_obs))),
            nudge_x = -0.5,
            nudge_y= 500,
            check_overlap = T,
            hjust=1,
            show.legend = F)+
  geom_text(data=. %>%  filter(date==as.Date(max(df_AT_IT$date)) &
                                 country_region=="Italy"),
            aes(x=day_index,
                y=n_obs,
                color=country_region,
                label=paste0("latest # Italy: ", 
                             comma(n_obs))),
            nudge_x = 0,
            nudge_y= 500,
            check_overlap = T,
            hjust=1,
            show.legend = F)+
  scale_y_continuous(limits=c(1, NA),
                     minor_breaks = NULL,
                     breaks=c(1, seq(1000, 12000, 2500)))+
  theme_ft_rc()+
  theme(legend.position = "none",
        #legend.justification = "left",
        plot.subtitle = element_markdown(),
        plot.caption = element_text(hjust=c(0, 0)),
        legend.title=element_blank())

```
</details>

Overall, the graph below suggests that Austria is on a somewhat slower trajectory than Italy. Contrasting both countries' absolute numbers on the **latest index day (`r scales::comma(df_AT_IT %>% filter(country_region=="Austria") %>% pull(day_index) %>% tail(1))`th day as of `r format(df_AT_IT %>% filter(country_region=="Austria") %>% pull(date) %>% tail(1), "%d %B")`**, Italy featured already **`r scales::comma(df_AT_IT %>% filter(country_region=="Italy" & day_index==max(df_AT_IT$day_index[df_AT_IT$country_region=="Austria"])) %>% pull(n_obs) %>% tail(1))` cases**. In contrast, Austria featured **`r scales::comma(df_AT_IT %>% filter(country_region=="Austria") %>% pull(n_obs) %>% tail(1))`** confirmed cases.

```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE}
pl_AT_IT
```

# Correcting infection numbers for population size

Critically, one may adjust the number of confirmed cases with the population size of the country. The graph below presents the modified result.

<details closed> 
<summary>Click to see code for plot</summary>
I retrieve Italy's and Austria's population data from the World Bank's pertaining database, accessed via the `wbstats` package.
```{r}
pop_data <- wb(indicator = "SP.POP.TOTL", startdate = 2018, enddate = 2018) %>% 
  filter(str_detect(country, "Austria|Italy"))

df_AT_IT <- df_AT_IT %>% 
  left_join(., 
            pop_data %>% select(country, pop.size=value),
            by=c("country_region"="country")) %>% 
  mutate(n_obs_rel=n_obs/(pop.size/100000))

pl_AT_IT_rel <- df_AT_IT %>% 
  ggplot()+
  geom_vline(xintercept=df_AT_IT %>% 
               filter(country_region=="Austria") %>% 
               pull(day_index) %>% 
               max(),
             color="white")+
  geom_line(aes(x=day_index,
                y=n_obs_rel,
                color=country_region))+
  geom_point(data=. %>% filter(day_index==max(df_AT_IT$day_index[df_AT_IT$country_region=="Austria"])),
             aes(x=day_index,
                 size=2,
                 y=n_obs_rel,
                 color=country_region))+
  labs(title="Covid-19: Comparing dynamics in Austria and Italy",
       subtitle=glue("Confirmed cases <span style='color:white'>per 100,000 inhabitants</span>. Starting <span style='color:white'>from day when threshold of {threshold_value} cases was crossed </span>.
       \nData as of <span style='color:white'>{format(max(df_AT_IT$date), '%A, %d %b %Y')}</span>."),
       caption=c("Data: John's Hopkins  COVID-19 Data Repository (github.com/CSSEGISandData/COVID-19)",
                 "\nGraph: Roland Schmidt | @zoowalk | https:://werk.statt.codes"),
       x=glue("# days since more then {threshold_value} confirmed cases"),
       y="# confirmed cases per 100,000 inhabitants")+
  scale_x_continuous(limits=c(1, NA),
                     breaks=c(1, seq(0, max(df_AT_IT$day_index), 5)),
                     minor_breaks = NULL)+
  scale_color_manual(values=country_color)+
  geom_text(data=. %>%  filter(day_index==max(df_AT_IT$day_index[df_AT_IT$country_region=="Austria"]) &
                                 country_region=="Austria"),
            aes(x=day_index,
                y=n_obs_rel,
                color=country_region,
                label=paste0("latest # Austria;",
                             day_index, " days after threshold:", 
                             round(n_obs_rel, 1))),
            nudge_x = 1,
            nudge_y= 2,
            check_overlap = T,
            hjust=1,
            show.legend = F)+
  geom_text(data=. %>%  filter(day_index==max(df_AT_IT$day_index[df_AT_IT$country_region=="Austria"])&
                                 country_region=="Italy"),
            aes(x=day_index,
                y=n_obs_rel,
                color=country_region,
                label=paste0("Italy ", format(date, '%d %b'), ";",
                             day_index, " days after threshold:", 
                             round(n_obs_rel, 1))),
            nudge_x = 0.5,
            nudge_y= 0,
            check_overlap = T,
            hjust=0,
            show.legend = F)+
  geom_text(data=. %>%  filter(date==as.Date(max(df_AT_IT$date)) &
                                 country_region=="Italy"),
            aes(x=day_index,
                y=n_obs_rel,
                color=country_region,
                label=paste0("latest # Italy: ", 
                             round(n_obs_rel, 1))),
            nudge_x = -1,
            check_overlap = T,
            hjust=1,
            show.legend = F)+
  #scale_y_percent(labels = scales::label_percent(accuracy=0.001))+
  scale_y_continuous(minor_breaks = NULL)+
  theme_ft_rc()+
  theme(legend.position = "none",
        #legend.justification = "left",
        plot.subtitle = element_markdown(lineheight = .5),
        plot.caption = element_text(hjust=c(0, 0)),
        legend.title=element_blank())

```
</details>
Correcting for the population size, Austria's trajectory starts to look more worrying. While Italy had **`r scales::comma(df_AT_IT %>% filter(country_region=="Italy" & day_index==max(df_AT_IT$day_index[df_AT_IT$country_region=="Austria"])) %>% pull(n_obs_rel) %>% tail(1))`** persons infected with Covid-19 on the `r scales::comma(df_AT_IT %>% filter(country_region=="Austria") %>% pull(day_index) %>% tail(1))` day of the outbreak (if we take the aforementioned threshold), Austria only a considerably larger figure of **`r scales::comma(df_AT_IT %>% filter(country_region=="Austria") %>% pull(n_obs_rel) %>% tail(1))`.**

```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE}
pl_AT_IT_rel
```

# Absolute number of new infections

Let's contrast the absolute number of newly confirmed infections from day to day.

<details closed> 
<summary>Click to see code</summary> 
```{r}

df_new_infections <- df_AT_IT %>% 
  group_by(country_region) %>% 
  arrange(date, .by_group=T) %>% 
  mutate(new_infections=n_obs-lag(n_obs)) %>% 
  mutate(max_new_infections=max(new_infections, na.rm = T)) %>% 
  filter(threshold_indicator>0) %>% 
  mutate(day_index=row_number()) %>% 
  ggplot()+
  labs(title="Covid-19: Daily number of new confirmed cases in Austria and Italy",
       subtitle=glue("Data as of <span style='color:white'>{format(max(df_AT_IT$date), '%A, %d %b %Y')}</span>."),
       caption=c("Data: John's Hopkins  COVID-19 Data Repository (github.com/CSSEGISandData/COVID-19)",
                 "\nGraph: Roland Schmidt | @zoowalk | https:://werk.statt.codes"),
       y="# of new confirmed cases per day")+
  geom_bar(aes(x=date,
               y=new_infections,
               color=country_region,
               fill=country_region),
           stat="identity")+
  # geom_text(data=. %>% filter(day_index %% 2 != 0 ),
  #           aes(x=date,
  #               y=max_new_infections*1.1,
  #               color=country_region,
  #               label=scales::comma(new_infections,
  #                                   accuracy=1),
  #               fill=country_region),
  #           stat="identity")+
  # geom_text(data=. %>% filter(day_index %% 2 == 0 ),
  #           aes(x=date,
  #              y=max_new_infections*1.3,
  #              color=country_region,
  #               label=scales::comma(new_infections,
  #                                   accuracy=1),
  #              fill=country_region),
  #         stat="identity")+
  facet_wrap(vars(country_region),
             scales="free_y",
             nrow=2)+
  scale_y_continuous(labels=scales::label_comma(accuracy = 1),
                     expand=expansion(mult=c(0, 0.1)))+
  scale_fill_manual(values=country_color)+
  scale_color_manual(values=country_color)+
  scale_x_date(labels=scales::label_date_short(),
               breaks=breaks_width(width = "1 day"))+
  theme_ft_rc()+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.subtitle = element_markdown(lineheight = .5),
        plot.caption = element_text(hjust=c(0, 0)),
        legend.title=element_blank())

```
</details>

```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE}
df_new_infections
```

# Growth rates

What if we look at the increase in relative terms, as growth rates:

<details closed> 
<summary>Click to see code</summary>
```{r}

# growth rates --------------------------------------------------------------

pl_growth <- df_corona_long %>% 
  filter(str_detect(country_region, "Austria|Italy")) %>% 
  mutate(threshold_indicator=case_when(n_obs>threshold_value ~ 1,
                                       TRUE ~ 0)) %>% 
  group_by(country_region) %>% 
  arrange(date, .by_group=T) %>% 
  filter(threshold_indicator>0) %>%
  mutate(day_index=row_number()) %>%
  mutate(n_obs_prior=lag(n_obs)) %>% 
  mutate(change=(n_obs-n_obs_prior)/n_obs_prior) %>% 
  #filter(date>as.Date("2020-03-01")) %>% 
  ungroup() %>% 
  ggplot()+
  labs(title="Covid-19: Daily growth rates in Austria and Italy",
       subtitle=glue("<span style='color:white'>Daily Percentage change of number of confirmed cases</span> compared to previous day.<br>Data as of <span style='color:white'>{format(max(df_AT_IT$date), '%A, %d %b %Y')}</span>."),
       caption=c("Data: John's Hopkins  COVID-19 Data Repository (github.com/CSSEGISandData/COVID-19)",
                 "\nGraph: Roland Schmidt | @zoowalk | https:://werk.statt.codes"),
       x="# days since more then 10 confirmed cases",
       x="date",
       y="% change to previous day")+
  geom_bar(aes(x=date,
               y=change,
               fill=country_region,
               group=country_region,
               color=country_region),
           width = 0.8,
           stat ="identity",
           position=position_dodge(preserve="single"))+
  # geom_text(aes(x=date,
  #               y=change+0.13,
  #               group=country_region,
  #               label=scales::number(change, 
  #                                    scale=100,
  #                                    accuracy=0.1),
  #               color=country_region,
  #               fill=country_region),
  #           position=position_dodge(width=1),
  #           size=3,
  #           hjust=1,
  #           angle= 45)+

    # geom_text(data=. %>%  filter(country_region=="Italy"),
    #           aes(x=date,
    #             y=2.5,
    #             group=country_region,
    #             label=scales::number(change, 
    #                                  scale=100,
    #                                  accuracy=0.1),
    #             color=country_region,
    #             fill=country_region),
    #         position=position_dodge(width=1),
    #         size=3,
    #         hjust=.5)+
    #   geom_text(data=. %>%  filter(country_region=="Austria"),
    #           aes(x=date,
    #             y=2.7,
    #             group=country_region,
    #             label=scales::number(change, 
    #                                  scale=100,
    #                                  accuracy=0.1),
    #             color=country_region,
    #             fill=country_region),
    #         position=position_dodge(width=1),
    #         size=3,
    #         hjust=.5)+
  scale_y_continuous(labels=scales::label_percent(scale=100))+
  scale_fill_manual(values=country_color)+
  scale_color_manual(values=country_color)+
  scale_x_date(labels=scales::label_date_short(),
               minor_breaks = NULL,
               breaks=scales::date_breaks(width="1 day"),
               guide = guide_axis(n.dodge=1))+
  theme_ft_rc()+
  theme(legend.position = "top",
        legend.justification = "left",
        plot.subtitle = element_markdown(lineheight = .5),
        plot.caption = element_text(hjust=c(0, 0)),
        legend.title=element_blank())

```
</details>

```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE}
pl_growth
```

# Growth rates synced

<details closed> 
<summary>Click to see code</summary> 
```{r}

# > growth rates on index axis ----------------------------------------------

pl_growth_index <- df_corona_long %>% 
  filter(str_detect(country_region, "Austria|Italy")) %>% 
  mutate(threshold_indicator=case_when(n_obs>threshold_value ~ 1,
                                       TRUE ~ 0)) %>% 
  group_by(country_region) %>% 
  arrange(date, .by_group=T) %>% 
  filter(threshold_indicator>0) %>%
  mutate(day_index=row_number()) %>%
  mutate(n_obs_prior=lag(n_obs)) %>% 
  mutate(change=(n_obs-n_obs_prior)/n_obs_prior) %>% 
  #filter(date>as.Date("2020-03-01")) %>% 
  ungroup() %>% 
  ggplot()+
  labs(title="Covid-19: Daily growth rates in Austria and Italy",
       subtitle=glue("<span style='color:white'>Daily Percentage change of number of confirmed cases</span> compared to previous day.
       <br>Starting <span style='color:white'>from day when threshold of {threshold_value} cases was crossed </span>.
                     <br>Data as of <span style='color:white'>{format(max(df_AT_IT$date), '%A, %d %b %Y')}</span>."),
       caption=c("Data: John's Hopkins  COVID-19 Data Repository (github.com/CSSEGISandData/COVID-19)",
                 "\nGraph: Roland Schmidt | @zoowalk | https:://werk.statt.codes"),
       x="# days since more then 10 confirmed cases",
       x="date",
       y="% change to previous day")+
  geom_bar(aes(x=day_index,
               y=change,
               fill=country_region,
               group=country_region,
               color=country_region),
           width = 0.8,
           stat ="identity",
           position=position_dodge(preserve="single"))+
  # geom_text(aes(x=day_index,
  #               y=change+0.1,
  #               group=country_region,
  #               label=scales::number(change, 
  #                                    scale=100,
  #                                    accuracy=0.1),
  #               color=country_region,
  #               fill=country_region),
  #           position=position_dodge(width=1),
  #           size=3,
  #           angle= 45)+

    # geom_text(data=. %>%  filter(country_region=="Italy"),
    #           aes(x=day_index,
    #             y=2.5,
    #             group=country_region,
    #             label=scales::number(change, 
    #                                  scale=100,
    #                                  accuracy=0.1),
    #             color=country_region,
    #             fill=country_region),
    #         position=position_dodge(width=1),
    #         size=3,
    #         hjust=0)+
    #   geom_text(data=. %>%  filter(country_region=="Austria"),
    #           aes(x=day_index,
    #             y=2.7,
    #             group=country_region,
    #             label=scales::number(change, 
    #                                  scale=100,
    #                                  accuracy=0.1),
    #             color=country_region,
    #             fill=country_region),
    #         position=position_dodge(width=1),
    #         size=3,
    #         hjust=0)+

  scale_y_continuous(labels=scales::label_percent(scale=100))+
  scale_fill_manual(values=country_color)+
  scale_color_manual(values=country_color)+
  # scale_x_date(labels=scales::label_date_short(),
  #              breaks=scales::date_breaks(width="1 day"),
  #              guide = guide_axis(n.dodge=1))+
  theme_ft_rc()+
  theme(legend.position = "top",
        legend.justification = "left",
        plot.subtitle = element_markdown(lineheight = .5),
        plot.caption = element_text(hjust=c(0, 0)),
        legend.title=element_blank())

```
</details>

```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE}
pl_growth_index
```

### 2021-01-09 _post/2020-03-02-primary-schools-in-vienna/primary-schools-in-vienna.Rmd
update

---
title: "Segregation in Vienna's primary schools?"
description: |
  A short description of the post.
date: 03-02-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693

categories:
  - Austria
  - plotly

preview: static-maps-with-mapbox-background-images-snapbox_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

```{r include=FALSE}

library(tidyverse)
library(ggplot2)
library(here)
library(extrafont)
loadfonts(device = "win", quiet = T)
extrafont::fonts()
library(hrbrthemes)
library(scales)
library(ggiraph)
library(knitr)
library(mapdeck)
library(widgetframe)
library(htmlwidgets)
library(plotly)
library(paletteer)
library(ggtext)
library(janitor)
library(magrittr)
options(digits = 12)
options(pillar.sigfig = 12)
library(geosphere)
library(plotly)
library(crosstalk)
```

```{r setup, echo = F}
# knit_hooks$set(wrap = function(before, options, envir){
# if (before){
# paste0('<', options$wrap, '>')
# } else {
# paste0('</', options$wrap, '>')
# }
# })
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dpi = 180, fig.height = 5, fig.width = 7, fig.align = "center")
options(width = 180, dplyr.width = 150)

bgr_color<- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()

```

# Context

Our daughter will turn six this summer and in Austria this means that she will start attending school in autumn. While in most parts of Austria’s countryside it’s a rather foregone conclusion to which school a new pupil will go (since there is mostly only one school ‘in town’), choosing a school for your kid in a city like Vienna can become a somewhat protracted decision making process with numerous schools in the vicinity to select from. 
While visiting the different schools in our neighbourhood and trying to make an informed decision in the best interest of the child, I was startled by a repeated story of two closely located schools, where one school was said to be for kids with migration-background (‘Migrationshintergrund’), and the other for kids from native German speakers. While I am simplifying here and much of this story remains vague - where does it come from, how is this separation made, who makes this separation - I found this account remarkable to say at least.

The inconvenient elephant in the room here seems to be some parents’ fear that kids who do not speak German as a native language would slow down the educational development of their own, native German speakers (pure xenophobic impulses I even do not want to consider here). While I see to some extent the concern, the idea that kids of different backgrounds are effectively separated into different schools strikes me as a completely unacceptable development and brings up associations with the history of a country at the southern cone of Africa (not to speak of Austria’s darker past).

# Analysis
So, triggered by this ‘hearsay’ (and, guilty as charged, by my affinity with R), I wanted to dig a bit deeper and see whether ‘the data’ has anything pertaining to say. The straightforward approach would be simply to look at the share of kids with first languages other than German in each school and compare them. While this data is certainly collected (when enrolling your kid), as far as I can tell, it is not made publicly available on an individual school level (and maybe that’s actually good). However, during our school search I came across a rather informative map provided by Austria’s statistical office (see [here]( https://www.statistik.at/atlas/schulen/#){target=”_blank”}). The map presents not only the location for every (primary) school in Austria, but also data on the school subsequently attended by its pupils once they finished the fourth grade. See e.g. <br>![Example of table of 4th graders' school destination](/post/2020-02-17-primary-schools-in-vienna_files/vienna_schools_table.PNG){width=50% height=50%}.

In short, with this data we know how many % of all pupils ‘graduating’ from a primary school’s 4th grade subsequently attended either a AHS (or equivalent; a high school/gymnasium) or an NMS (or equivalent; ‘new secondary school’). Critically, an AHS offers a direct trajectory to a school leaving examination after eight years which is required to attend university. In contrast, an NMS lasts only for four years after which kids either go for an apprenticeship, start to work, or if their school record was good enough, switch for another four to five years to an AHS or similar school type which offers a school leaving exam. To (over)simplify, leaving for an AHS is by and large considered to be better for the future educational and professional development. I will subsequently call the % of kids per school leaving for an AHS (or equivalent) the __AHS ratio__.^[To be clear, first: this blog does not intend to convey that a kid attending a middle school rather than a high school is on an individual or school level some sort of a failure. Definitely not. Kids can develop at different pace, have different needs and abilities at different times. An NMS might be just the right fit. While I am not ignorant to the particular challenges NMS entail, they should be an environment in which kids can prosper and life chances be seized. That this is not always the reality is a different story. Second: This post also does not mean to say that schools with higher AHS ratios are ‘better’ schools. Better, in my understanding, is based on schools ability to educationally and socially develop kids from where they stand when they enter school to the point when they leave. And in this regard, schools are hardly served with a clean slate (kids have different starting positions).]

## Differences between districts
In an ‘ideal-typical’ world, without any non-randomly distributed confounding factors, the rates of pupils leaving for an AHS should be rather similar across the board. That this is not the case, however, doesn’t come at a surprise. Socio-economic factors of all kinds influence a school’s as well as a kid’s ability to educationally prosper. These factors are frequently geographically clustered and hence can account for different school-leaving ratios between schools.

To see this in the context in Vienna, I downloaded the data (2018/2019) for every Viennese primary school from the above-mentioned map of Statistik Austria and calculated the ratio of kids leaving for an AHS (or equivalent school). If we separate the result of schools by district and whether they are privately or publicly run, we get the graph below. Hover-over each dot/school to get details.

<details closed> 
<summary>Click here for code on how to get data on each primary school </summary>

```{r message=FALSE, warning=FALSE}
# get link to json file with data embedded in the website
data_link <- "https://www.statistik.at/gs-atlas/ATLAS_SCHULE_WFS/ows?service=WFS&version=1.0.0&request=GetFeature&typeName=ATLAS_SCHULE_WFS:ATLAS_SCHULE&outputFormat=application%2Fjson&srsname=EPSG:3857&"

# download jsonfile and extract field with relevant data
# df_schools <- jsonlite::fromJSON(data_link, flatten=TRUE) %>%
# pluck("features") %>% #extract nested list with data on schools
# janitor::clean_names() #clean/standardize names of data columns

df_schools <- jsonlite::fromJSON(data_link, flatten = TRUE)

df_schools <- df_schools$features %>% # extract nested list with data on schools
  janitor::clean_names() # clean/standardize names of data columns

# take only those primary schools located in Vienna
df_primary_schools_vienna <- df_schools %>%
  filter(properties_karto_typ == "VS") %>% # filter only primary schools
  filter(str_detect(properties_plz, "^1")) %>%
  as_tibble() %>%
  mutate(properties_skz = as.numeric(properties_skz)) %>%
  mutate(school_adress = paste0(properties_str, ", ", properties_plz, " ", properties_ort))

# data on 263 primary schools was obtained
n_distinct(df_primary_schools_vienna$properties_skz) # 263
```
</details>

```{r include=FALSE}
# load version which already includes geolocated data
df_primary_schools_vienna <- readr::read_csv2(here("_blog_data", "Vienna_primary_schools", "df_primary_schools_vienna.csv"))
```

<details closed> 
<summary>Click here for code on how to get data on destination of pupils of each primary school </summary>

```{r warning=FALSE}

# link to on pupils' destinations combined with each primary school's ID
# gives vector for all destination tables
library(glue)
df_destinations_links <- glue("https://www.statistik.at/gs-atlas/ATLAS_SCHULE_WFS/ows?service=WFS&version=1.0.0&request=GetFeature&typeName=ATLAS_SCHULE_WFS:ATLAS_SCHULE_UEBERTRITT_OUT_WFS&maxFeatures=500&outputFormat=application%2Fjson&viewparams=SKZ:{df_primary_schools_vienna$properties_skz}") %>%
  enframe(value = "link", name = NULL)
NROW(df_destinations_links) # 263 primary schools in Vienna

# define function to get pupils' destinatino data for one school
fn_get_destination_data <- function(school_link) {
  jsonlite::fromJSON(school_link, flatten = TRUE) %>%
    purrr::pluck("features") %>%
    janitor::clean_names() %>%
    as_tibble() %>%
    mutate(school_link = school_link)
}

# apply function to all schools
df_destination_data <- map_dfr(
  df_destinations_links$link,
  possibly(fn_get_destination_data,
    otherwise = tibble()
  )
)

# only 257 out of 263 primary schools with destination data;
# 6 schools still without 'graduates'
length(unique(df_destination_data$properties_skz_vj))

df_destination_data
```
</details>

<details closed> 
<summary>Click here for code on how to calcualte AHS ratio of each school and district </summary>

```{r warning=FALSE}
# take the destination data and create school-type indicator for schools which provide
# a direct trajectory to a school leaving exam;
# since the data provides a 0 if only three or less pupils transfered to a particular school
# (I assume for data/privacy reasons), I set this number to 2 as some kind of middle ground; see
# the table on the website of Statitik Austria;
# Note the .drop=F option in the group_by function; it provides us with a summary value (0) even if #there was no puipil leaving for an AHS.

df_destination_analysis <- df_destination_data %>%
  mutate(properties_karto_typ = as_factor(properties_karto_typ)) %>%
  mutate(school_type = case_when(
    properties_karto_typ == "NMSH" ~ "Neue Mittelschule",
    properties_karto_typ == "NMSA" ~ "Gymnasium mit Modelversuch Mittelschule",
    properties_karto_typ == "SS" ~ "Sonderschule",
    properties_karto_typ == "VS" ~ "Volksschule",
    TRUE ~ as.character(properties_karto_typ)
  )) %>%
  mutate(school_type_secondary = case_when(
    properties_karto_typ %in% c("AHS", "BMHSK", "BMHST") ~ "matura",
    TRUE ~ as.character("no matura")
  )) %>%
  mutate(school_type_secondary = as_factor(school_type_secondary)) %>%
  mutate(properties_anzahl = case_when(
    properties_anzahl == 0 ~ 2, # changed to two
    TRUE ~ as.numeric(properties_anzahl)
  )) %>% # assumpation
  group_by(properties_skz_vj, school_type_secondary, .drop = F) %>%
  summarise(n_obs = sum(properties_anzahl)) %>%
  ungroup()

df_share <- df_destination_analysis %>%
  mutate(properties_skz_vj = as.numeric(properties_skz_vj)) %>%
  group_by(properties_skz_vj) %>%
  mutate(total = sum(n_obs)) %>%
  mutate(n_rel = n_obs / sum(n_obs)) %>%
  left_join(., df_primary_schools_vienna %>% # merge is wrong; has to be from primary df
    select(
      properties_skz,
      properties_bezeichnung,
      properties_erhalter,
      properties_str,
      properties_plz,
      properties_ort
    ),
  by = c("properties_skz_vj" = "properties_skz")
  ) %>%
  ungroup() %>%
  mutate(district_code = stringr::str_sub(properties_plz, start = 2, end = 3) %>% as.numeric())
```
</details>

<details closed> 
<summary>Click here for code on how to produce the graph </summary>

```{r}
my_caption <- c("Data: Statistik Austria, statistik.at/atlas/schulen/", "Analysis: Roland Schmidt | @zoowalk | <span style='color:black'>**werk.statt.codes**</span>")

plot_district <- df_share %>%
  filter(school_type_secondary == "matura") %>%
  group_by(district_code) %>%
  mutate(n_median = median(n_rel)) %>%
  ungroup() %>%
  mutate(district_code = fct_reorder(as_factor(district_code), n_median)) %>%
  mutate(tooltip = paste("ratio: ", scales::percent(n_rel),
    "\n",
    paste(properties_bezeichnung, properties_str,
      paste0(district_code, ". ", "district"),
      sep = ", "
    ),
    "\n",
    properties_erhalter,
    sep = ""
  )) %>%
  ggplot() +
  labs(
    title = "% of pupils leaving for secondary schools with school leaving certificate (Matura) per district",
    subtitle = "Data for 2018/2019.",
    y = "% of pupils per school leaving for schools with school leaving certificate (Matura)",
    x = "district",
    caption = my_caption
  ) +
  geom_boxplot(aes(
    x = district_code,
    y = n_rel
  ),
  fill = "transparent",
  color = "grey80",
  outlier.shape = NA
  ) +
  geom_jitter_interactive(aes(
    x = district_code,
    y = n_rel,
    color = properties_erhalter,
    tooltip = tooltip
  ),
  width = .20
  ) +
  coord_flip() +
  theme_ipsum_rc() +
  scale_color_paletteer_d("ggsci::default_jama",
    labels = c("öffentlich" = "public", "privat" = "private")
  ) +
  scale_y_continuous(
    labels = scales::label_percent(),
    limits = c(0, 1)
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted"),
    panel.grid.minor.x = element_line(linetype = "dotted"),
    legend.position = "top",
    legend.justification = "left",
    legend.title = element_blank(),
    plot.margin = margin(l = 0, t = 0.5, b = 0.5, unit = "cm"),
    plot.background = element_rect(fill = bgr_color, color = "transparent"),
    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 0, unit = "cm")),
    plot.title.position = "plot",
    plot.subtitle = element_text(size = 12, color = "grey30"),
    plot.caption.position = "plot",
    plot.caption = element_markdown(color = "grey30", hjust = c(0, 1))
  )

plot_district <- girafe(
  ggobj = plot_district,
  options = list(opts_tooltip(
    css = "background-color:#323E4F; 
color: white;
font-family:Roboto Condensed;",
    delay_mouseout = 5000
  )),
  pointsize = 6,
  width_svg = 10,
  height_svg = 7
) # 10 15
```
</details>

```{r echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
plot_district
```
To me the graph reflects two things: First, contrasting the AHS-ratio, there is a considerable difference between districts. While the 11th, 10th and 20th district had median values of around 30 %, those of the 4th, 18th and 19th are above 70 %. Second, private schools (orange dots) have overwhelmingly higher AHS-ratios than the public schools in the same district. Does this come as a surprise? Not really. Is it ok? – Well, not really.

## Role of pupils’ vernacular

To get an idea on any association between schools’ AHS ratio and pupils’ language skills, I added district level data on pupils’ vernacular. The available data (2017!) presents the % of primary school pupils whose vernacular is not German per district.
[Source](https://kurier.at/politik/inland/schule-in-not-wo-deutsch-eine-fremdsprache-ist/400134818){target="_blank"}.
So to be clear, the data is district level data and not school level data and inferences from any association might suffer from an [ecological fallacy](https://en.wikipedia.org/wiki/Ecological_fallacy){target="_blank"}. But in the absence of any better data, and bearing this caveat in mind, it seems worth looking into it. 
The graph below depicts each primary school’s AHS-ratio (y axis) and the pertaining district’s ratio of primary school kids with non-German vernacular (x axis). What becomes apparent is that there is quite a strong association between the two factors: The higher the ratio of kids with non-German vernacular, the lower the AHS ratio (a correlation coefficient of - .44). While this is largely in line with what is generally suggested, it is still rather sobering and highlights a pretty significant shortcoming of the educational system. Again, since we do not have individual/school level data on pupils’ language one has to be cautious and not jump to far fetched conclusions. But the data, in my view, would strongly suggest that non-native German speaking kids are let down by the school system (in the sense that their needs are not adequately addressed).

<details closed> 
<summary>Click here for code on how to produce the graph </summary>
```{r}
# load the data on districts' german/non-german vernacular ratio (source in text),
# recode name of districts to postal codes

df_language <- readr::read_csv(here(
  "_blog_data",
  "Vienna_primary_schools",
  "umgangssprache_deutsch_bezirk_wien.csv"
)) %>%
  clean_names() %>%
  rename(
    non_german_perc = x16_17,
    district = bezirk,
    pupils_total = volksschuler_innen_gesamt,
    german_abs = deutsch,
    non_german_abs = nicht_deutsch,
    non_german_perc_06 = x06_jul,
    change_16_06 = unterschied_in_percent
  ) %>%
  mutate(district_code = case_when(
    district == "Alsergrund" ~ 9,
    district == "Brigittenau" ~ 20,
    district == "Donaustadt" ~ 22,
    district == "Döbling" ~ 19,
    district == "Favoriten" ~ 10,
    district == "Floridsdorf" ~ 21,
    district == "Hernals" ~ 17,
    district == "Hietzing" ~ 13,
    district == "Innere Stadt" ~ 1,
    district == "Josefstadt" ~ 8,
    district == "Landstraße" ~ 3,
    district == "Leopoldstadt" ~ 2,
    district == "Liesing" ~ 23,
    district == "Margareten" ~ 5,
    district == "Mariahilf" ~ 6,
    district == "Meidling" ~ 12,
    district == "Neubau" ~ 7,
    district == "Ottakring" ~ 16,
    district == "Penzing" ~ 14,
    district == "Rudolfsheim-Fünfhaus" ~ 15,
    district == "Simmering" ~ 11,
    district == "Wieden" ~ 4,
    district == "Währing" ~ 18,
    TRUE ~ NA_real_
  ))

# merge data
df_share <- df_share %>%
  left_join(., df_language %>%
    select(district_code, district, non_german_perc),
  by = c("district_code")
  )

# produce plot
plot_langauge <- df_share %>%
  mutate(tooltip = paste("ratio: ", scales::percent(n_rel),
    "\n",
    paste(properties_bezeichnung, properties_str,
      paste0(district_code, ". ", "district"),
      sep = ", "
    ),
    "\n",
    properties_erhalter,
    sep = ""
  )) %>%
  filter(school_type_secondary == "matura") %>%
  filter(properties_erhalter == "öffentl.") %>%
  ggplot() +
  labs(
    y = "AHS ratio",
    title = "Relation between AHS ratio and primary school pupils' vernacular",
    subtitle = str_wrap(
      "Note that AHS ratio are school level data; 
info on vernacular is district level data. 
Only public schools. Hover over dots to get details.",
      150
    ),
    caption = my_caption,
    x = "% of primary school pupils with non-German vernacular per district"
  ) +
  geom_jitter_interactive(aes(
    x = non_german_perc,
    y = n_rel,
    tooltip = tooltip
  )) +
  geom_smooth(aes(
    x = non_german_perc,
    y = n_rel
  ),
  method = "lm", formula = y ~ x
  ) +
  scale_y_continuous(
    labels = scales::percent,
    minor_breaks = NULL
  ) +
  scale_x_continuous(
    labels = scales::label_percent(scale = 1),
    minor_breaks = NULL,
    limits = c(0, 100)
  ) +
  scale_color_paletteer_d("ggsci::default_jama") +
  # facet_wrap(vars(properties_erhalter))+
  hrbrthemes::theme_ipsum_rc() +
  theme(
    panel.grid.major.y = element_line(linetype = "dotted"),
    panel.grid.major.x = element_line(linetype = "dotted"),
    legend.position = "top",
    legend.justification = "left",
    legend.title = element_blank(),
    plot.margin = margin(l = 0, t = 0.5, b = 0.5, unit = "cm"),
    plot.background = element_rect(fill = bgr_color, color = "transparent"),
    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 0, unit = "cm")),
    plot.title.position = "plot",
    plot.subtitle = element_text(size = 12, color = "grey30"),
    plot.caption.position = "plot",
    plot.caption = element_markdown(color = "grey30", hjust = c(0, 1))
  )

# wrap plot as htmlwidget for interactivity
plot_langauge <- girafe(
  ggobj = plot_langauge,
  options = list(opts_tooltip(
    css = "background-color:#323E4F; 
color: white;
font-family:Roboto Condensed;",
    delay_mouseout = 5000
  )),
  pointsize = 6,
  width_svg = 10,
  height_svg = 7
)

# correlation coefficient
cor(
  df_share$n_rel[df_share$school_type_secondary == "matura"],
  df_share$non_german_perc[df_share$school_type_secondary == "matura"]
)
```
</details>

```{r echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
plot_langauge
```

## Same neighbourhood, different school results?

But to come back to my original point of departure, the two graphs above also show that there is significant variation within districts. While Vienna’s districts are frequently perceived as homogeneous subunits, there is considerable variation within these units which may obscure what we are interested in. To account for these differences and seeking to compare like for like schools as far as their neighbourhood is concerned, I retrieved the _GPS position of every primary school and subsequently calculated the geographical distance between them_. Have a look at the annotated code (that's where most of the work went in).

<details closed> 
<summary>Click here for code on how to geolocate schools' adresses </summary>
```{r, eval=F}

# apply function google_geocode to each address
# use map (not map_chr) since you'll get a bunch of geographic details
df_primary_schools_vienna <- df_primary_schools_vienna %>%
  mutate(coord = map(
    school_adress,
    ~ google_geocode(.x, key = your_google_api_key)
  ))

# extract from the nested list the with the wonderful 'pluck' function the values #your #interested (lat, lng)
df_primary_schools_vienna <- df_primary_schools_vienna %>%
  mutate(lat = map_dbl(
    coord,
    purrr::pluck, "results", "geometry", "location",
    "lat"
  )) %>%
  mutate(long = map_dbl(
    coord,
    purrr::pluck, "results", "geometry", "location",
    "lng"
  ))
```
</details>

<details closed> 
<summary>Click here for code on how to get difference of AHS ratios between every pair of schools </summary>
```{r, eval=T}
# take only AHS-ratios
df_diff <- df_share %>%
  ungroup() %>%
  filter(school_type_secondary == "matura") %>%
  select(properties_skz_vj, n_rel)

# function outer creates an array including all combinations of schools
# and calculates the difference of the AHS ratios
diff_matrix <- outer(df_diff$n_rel, df_diff$n_rel, FUN = "-") %>%
  data.frame()

names(diff_matrix) <- df_diff$properties_skz_vj
diff_matrix$school_1 <- df_diff$properties_skz_vj

# with pivot_longer we bring the array in the long format;
# create a pair indictor with pmin/pmax which is indentical for each
# pair irrespective of the sequence of the school id (A-B covers A-B and B-A);
# and we take the absolute value of the difference of the AHS ratio
df_difference <- diff_matrix %>%
  pivot_longer(
    cols = -school_1,
    names_to = "school_2",
    values_to = "difference"
  ) %>%
  mutate(pair_indicator = map2_chr(
    school_1, school_2,
    function(x, y) paste(pmin(x, y), pmax(x, y), sep = "-")
  )) %>%
  group_by(pair_indicator) %>%
  slice(1) %>% # only take one observation; not twice (only one distance sufficient)
  mutate(diff_abs = abs(difference)) %>%
  filter(school_2 != school_1) %>% # remove difference observation of school with itself
  select(pair_indicator, everything())

df_difference
```

```{r, eval=T}
# add details of schools to dataframe containing all combinations of schools (nested);
# first for 1st school in pair, then for 2nd school in pair;
# extract values of interest for easier handling from nested dataframe;
df_main <- df_difference %>%
  mutate_at(vars("school_1", "school_2"), as.numeric) %>%
  nest_join(., df_primary_schools_vienna,
    by = c("school_1" = "properties_skz")
  ) %>%
  rename(school_1_details = df_primary_schools_vienna) %>%
  mutate(school_1_lat = map_dbl(school_1_details, purrr::pluck, "lat")) %>%
  mutate(school_1_long = map_dbl(school_1_details, purrr::pluck, "long")) %>%
  # nest(school_1_coords=c(school_1_lat, school_1_long)) %>%
  nest_join(., df_primary_schools_vienna,
    by = c("school_2" = "properties_skz")
  ) %>%
  rename(school_2_details = df_primary_schools_vienna) %>%
  mutate(school_2_lat = map_dbl(school_2_details, purrr::pluck, "lat")) %>%
  mutate(school_2_long = map_dbl(school_2_details, purrr::pluck, "long")) %>%
  mutate(address1 = map(school_1_details, ~ .x %>% mutate(school_1_address = paste(properties_bezeichnung,
    properties_str,
    properties_plz,
    sep = ", "
  ))) %>%
    map_chr(., purrr::pluck, "school_1_address")) %>%
  mutate(address2 = map(school_2_details, ~ .x %>% mutate(school_2_address = paste(properties_bezeichnung,
    properties_str,
    properties_plz,
    sep = ", "
  ))) %>%
    map_chr(., pluck, "school_2_address")) %>%
  mutate(school_1_erhalter = map(school_1_details, purrr::pluck, "properties_erhalter")) %>%
  mutate(school_2_erhalter = map(school_2_details, purrr::pluck, "properties_erhalter")) %>%
  mutate(school_erhalter_indicator = map2_chr(
    school_1_erhalter, school_2_erhalter,
    function(x, y) paste(pmin(x, y), pmax(x, y), sep = "-")
  )) %>%
  mutate(school_1_erhalter = map(school_1_details, extract, "properties_erhalter") %>% unlist()) %>%
  mutate(school_2_erhalter = map(school_2_details, extract, "properties_erhalter") %>% unlist()) %>%
  mutate(erhalter_type = paste(school_1_erhalter, school_2_erhalter, sep = " - ")) %>%
  left_join(., df_share %>%
    filter(school_type_secondary == "matura") %>%
    select(school_1_n_rel = n_rel, properties_skz_vj),
  by = c("school_1" = "properties_skz_vj")
  ) %>%
  left_join(., df_share %>%
    filter(school_type_secondary == "matura") %>%
    select(school_2_n_rel = n_rel, properties_skz_vj),
  by = c("school_2" = "properties_skz_vj")
  )
```

```{r, eval=T}
# create a tooltip for the graphic
df_main <- df_main %>%
  mutate(tooltip = paste0(
    paste(address1,
      "Wien",
      school_1_erhalter,
      scales::percent(school_1_n_rel),
      sep = ", "
    ),
    "\n",
    paste(address2,
      "Wien",
      school_2_erhalter,
      scales::percent(school_2_n_rel),
      sep = ", "
    )
  ))

# cacluate the factor by which one school's AHS differeds from another school's AHS ratio.
# the pmin/pmax function ensures that the school with the lower AHS ratio is
# is the denominator; makes comparison between school pairs easier;
df_main <- df_main %>%
  mutate(diff_factor = map2_dbl(
    school_1_n_rel, school_2_n_rel,
    function(x, y) pmax(x, y) / pmin(x, y)
  ))
```

```{r eval=T, paged.print=TRUE}
# Here the full table with the data
df_main
```
</details>

<details closed> 
<summary>Click here for code on how to get distance between every pair of schools </summary>

```{r}

# to map the distm function to the four geo-location data points (lat & lon for school 1 and lat & long
# for school 2 use the pmap function from the purrr package)

df_main <- df_main %>%
  ungroup() %>%
  mutate(distance = pmap_dbl(
    list(
      school_1_lat, school_1_long,
      school_2_lat, school_2_long
    ),
    ~ distm(
      x = c(..1, ..2),
      y = c(..3, ..4),
      fun = distHaversine
    )
  ))

df_main %>%
  select(school_1, school_2, distance)
```
</details>

There are 263 primary schools in the data on Statistik Austira’s map; for 257 schools destination data is provided. Calculating the distance between each pair of these schools leads to a matrix with 56644 cells. Since each pair appears twice (A to B, B to A) I created a pair_indicator and take only one observation from it. Furthermore, I removed distances of type A to A which were also in the matrix. The geolocation was done with via Google’s location API and the [googleway package]( https://github.com/SymbolixAU/googleway){target=”_blank”}. The service can cost a few Euros, but there is a 200 Euro / month free access scheme. The distance between these locations was subsequently calculated with the [geosphere package](https://cran.r-project.org/web/packages/geosphere/geosphere.pdf){target=”_blank”} and the distm function. Since the idea is to account for differences in the neighbourhood (compare only schools which are closely located), I look only at those pairs of schools which are not farther away from each other than 500 m. The result is presented in the two graphs below.

In the graph below on the left side, each dot represents a pair of public (!) primary schools. The x axis presents the distance between them in meters. The y axis presents the (absolute) difference between both schools' AHS ratio. E.g. if one school has a ratio of 10 % and the other has a ratio of 30 %, the position on the y axis is 20.

Again, hover with your mouse over each dot to get details and see e.g. which school has a higher AHS ratio. To make things a bit more accessible, if you click on the dot, the map on the right side will highlight the two locations of both schools (you may have to zoom in if schools are very close to each other).

Complementing the absolute difference of the AHS ratios, I also divided the AHS ratios of boths schools, resulting in what I call here (for the lack of a better idea) AHS factor. If school A has a ratio of 10 % and the school B has a ratio of 40 %, the corresponding AHS factor is 4. In other words, the factor indicates how much more likely it is for a kid from one school to ‘graduate’ to an AHS than it is for a kid from the other school. The factor is calculated in such a way that it always takes the ratio of the school with the higher AHS ratio as the numerator and the school with the lower ratio as the denominator. The factor is stated in the tooltip when clicking on the dot (school pair) in the graph.

```{r include=FALSE}
df_main <- readr::read_rds(here("_blog_data", "Vienna_primary_schools", "df_main.rds"))
Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1Ijoiem9vd2FsayIsImEiOiJjazFjMW52NWIwZDVjM21scmE1Y296MnUxIn0.ZXSu71B2VjyBwlUldNUtLg")
```

<details closed> 
<summary>Click here for code on how to produce the scatter plot and the interaction with the map </summary>

Both graphs are produced with [plotly]( https://github.com/ropensci/plotly){target=”_blank”}. Their interaction is facilitated by the [crosstalk]( https://rstudio.github.io/crosstalk/){target=”_blank”} package. Note that when using crosstalk in blogdown (which produces this blog), corsstalk messes up you css styling because it injects bootstrap css. To take care of this, see the unstrap function which I - afer some considerable googeling - found here. It basically removes all bootstrap css tags from the shared/crosstalk object and keeps your own blog style in order.

```{r, eval=T}

# take only those schools located within 500 m vicincity;
# and only public school dyads

shared_df_main <- df_main %>%
  filter(distance < 500) %>%
  filter(school_erhalter_indicator == "öffentl.-öffentl.") %>%
  select(-contains("details"))

# create a crosstalk object
shared_df_main <- SharedData$new(shared_df_main,
  key = ~pair_indicator,
  group = "pair_indicator"
)

# plot: x axis geographical distance;

plot_distance_diff <-
  plot_ly(
    data = shared_df_main,
    mode = "markers",
    type = "scatter",
    hoverinfo = "text",
    text = ~ paste(
      "abs diff AHS ratio (% points):", round((diff_abs * 100), 2),
      "</br>AHS factor:", round(diff_factor, 2),
      "</br>distance (m):", round(distance, 0),
      "</br>", tooltip
    )
  ) %>%
  add_trace(
    # text=~tooltip,
    marker = list(color = "orange"),
    x = ~distance,
    # y=~diff_factor,
    y = ~ scales::percent(diff_abs),
    showlegend = F
  ) %>%
  layout(
    title = list(
      text = "<b>Schools' different AHS ratios and geographical distance</b>",
      x = 0
    ),
    font = list(family = "Roboto Condensed"),
    xaxis = list(title = "distance between schools in meters"),
    yaxis = list(title = "absolute difference of AHS ratios (% points)"),
    plot_bgcolor = bgr_color,
    paper_bgcolor = bgr_color
  )

# plot_distance_diff

# map
# name package::; otherwise confusiong with leaflet addcircle => error

cross_map2 <- plotly::plot_mapbox(
  data = shared_df_main,
  hoverinfo = "text",
  text = ~ paste(
    "abs diff AHS ratio (% points):", round((diff_abs * 100), 2),
    "</br>AHS factor:", round(diff_factor, 2),
    "</br>distance (m):", round(distance, 0),
    "</br>", tooltip
  )
) %>%
  plotly::add_markers(
    y = ~school_1_lat,
    x = ~school_1_long,
    color = "#FF0000" # ,
    # text = ~address1
  ) %>%
  plotly::add_markers(
    y = ~school_2_lat,
    x = ~school_2_long,
    color = "#FF0000" # ,
    # text = ~address2
  ) %>%
  layout(
    mapbox = list(
      style = "mapbox://styles/zoowalk/ck6kz867d0jdi1io2lxyp9008",
      zoom = 10,
      center = list(
        lat = 48.21,
        lon = 16.3720798
      )
    ),
    showlegend = FALSE,
    paper_bgcolor = bgr_color
  )

# combine graphs

# cross_plot <- crosstalk::bscols(
# plot_distance_diff,
# cross_map2)

library(htmltools)
cross_plot <- browsable(tagList(plot_distance_diff, cross_map2))

# see here for how to remove bootstrap css which otherwise messes up you blogdown's css
# https://github.com/rstudio/crosstalk/issues/26

unstrap <- function(x) {
  attr(x, "html_dependencies") <-
    Filter(
      function(dependency) {
        dependency$name != "bootstrap"
      },
      attr(x, "html_dependencies")
    )
  x
}

cross_plot_clean <- unstrap(cross_plot)
```
</details>

```{r echo=FALSE}

cross_plot_clean
```

As the scatter plot shows, there were 36 (public) school pairs with a distance of less than 500 meters between them. A good part of these schools had AHS ratios different by up to 25 % points. What stands out, however, are some cases with differences of above 30 or even 40 % points. The primary schools in the Löwegasse and Deckergasse are only 52 meters away, yet there AHS ratio differ by 35 percentage points. While 64 % of 4th graders in the Löwegasse leave for an AHS, only 28 % do so from the Deckergsse. The primary schools in the Odakergasse and Meindlgasse are only 137 meters away from each other, yet their AHS ratios differ by 44.17 percentage points. In probabilistic terms (AHS factor), a kid from the Meinlgasse is three times (3.08) more likely to transfer to an AHS than a kid from the Odakergasse. These are quite stark contrasts. Similarly, although somewhat farther away (289 m), is the discrepancy between the primary school in the Kindermangasse and Rötzergasse. While the former has an AHS ratio of 69 %, the latter has a ratio of only 22 %. Also a factor of 3.08. For me a critical point here is – the school I was referring to at the beginning of this blog post is among these three schools.

# Conclusion

To recap, two schools which are closely located to each other, have in principle the same pool of kids to take as pupils. Hence, all other things being equal (!), one would expect similar AHS ratios. However, in at least three cases there are very strong differences between these ratios. We know that the percentage of primary school pupils with non-German vernacular on the district level is negatively associated with schools’ AHS ratios. If we make the (reasonable?) assumption that this association also holds on the school level (for which we don’t have data), the results are in line with the story that one school ‘is for migrant kids’ and the other ‘for German speakers’. While this is by no means a proof of some kind, the data at least corroborates the story.

But before finishing, a few caveats are in order. First, the above presented analysis does not claim to provide by any means a watertight chain of causation. There are several limitations, and I will only flag some of them here: At the core of the above bivariate analysis is the assumption that schools would be equal in terms of their ability to make kids leave for an AHS. While this should be the case, there might be some differences in terms of equipment, teachers’ quality etc. Possible? Definitely. Likely to account for the differences observed? I doubt it, though. Second, to make the case stronger one would need data on pupils’ subsequent destination from more than only one year. The data retrieved from the map of Statistik Austria is only from the school year 2018/2019. Do I personally think the data would look much different? Not really. Overall, irrespective of some critical methodological limitations, I find the highlighted discrepancies between closely located schools rather worrisome.

What the data does not tell us is how this discrepancy actually came about. By the best of my imagination, I find it difficult to believe that there is some pro-active, intentional policy going on from the side of the authorities. What seems more likely to me - and I can only speculate - is that we are witnessing the result of a self-perpetuating dynamic. With more German native speakers attending the school in one school, AHS ratios are higher. Consequenlty, German speaking parents send their kids again to the school with the higher AHS ratio and not to the other, similarly close school. Obviously, this still leaves a lot of other questions unaswered (e.g. (why) do non-German speaking parents exhibit the same enrollment behavior as German speaking parents?). However, what seems clear to me though is that this dynamic would warrant some intervention to lift up the schools with the lower AHS ratios.

### 2021-01-09 _post/2020-03-02-primary-schools-in-vienna/primary-schools-in-vienna.Rmd
removed old, invalid token

@@
 df_main <- readr::read_rds(here("_blog_data", "Vienna_primary_schools", "df_main.rds"))
-Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1Ijoiem9vd2FsayIsImEiOiJjazFjMW52NWIwZDVjM21scmE1Y296MnUxIn0.ZXSu71B2VjyBwlUldNUtLg")
+
 ```

### 2021-01-13 _post/2020-03-02-primary-schools-in-vienna/primary-schools-in-vienna.Rmd
adding revised school plot; crosstalk visible now

@@
 description: |
-  A short description of the post.
+  Drawing on public data from Vienna's school administration I look at how many percentages of pupils in each primary schoos subsequently leave for schools offering eventually a higher education entrance qualification (Matura).
+The main interest is in contrasting these ratios between closely (geo-)located schools.
+
 date: 03-02-2020
+
+theme: theme.css
 
@@
 <details closed> 
-<summary>Click here for code on how to geolocate schools' adresses </summary>
+<summary>Click here for code on how to geolocate schools' addresses </summary>
 ```{r, eval=F}
@@
 # use map (not map_chr) since you'll get a bunch of geographic details
+
 df_primary_schools_vienna <- df_primary_schools_vienna %>%
@@
   )
+
+# readr::write_rds(df_main, file=here::here("_blog_data", "Vienna_primary_schools","df_main2.rds"))
+
 ```
@@
 # is the denominator; makes comparison between school pairs easier;
+
 df_main <- df_main %>%
@@
 ```{r, eval=T}
-
 # take only those schools located within 500 m vicincity;
@@
     x = ~school_1_long,
-    color = "#FF0000" # ,
+    color = "#FF0000" #,
     # text = ~address1
@@
     mapbox = list(
-      style = "mapbox://styles/zoowalk/ck6kz867d0jdi1io2lxyp9008",
+      style = "mapbox://styles/zoowalk/ckjrkv7065tf519oax4nadm2k",
+      margin=list(l=0,r=0,t=0,b=0,p=0),
       zoom = 10,
@@
     paper_bgcolor = bgr_color
-  )
+  ) %>% 
+    config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))
+```
+</details>
 
+```{r, eval=T}
 # combine graphs
 
-# cross_plot <- crosstalk::bscols(
-# plot_distance_diff,
-# cross_map2)
+cross_plot <- crosstalk::bscols(
+plot_distance_diff,
+cross_map2)
+#
 
+```
+
+cross_plot <- crosstalk::bscols(
+plot_distance_diff,
+cross_map2)
+```{r eval=FALSE, include=FALSE}
+# cross_plot
+```
+
+cross_plot <- browsable(tagList(cross_plot))
+cross_plot
+
+```{r, eval=T}
 library(htmltools)
-cross_plot <- browsable(tagList(plot_distance_diff, cross_map2))
+# cross_plot <- browsable(tagList(plot_distance_diff, cross_map2))
+cross_plot <- browsable(tagList(cross_plot))
+# cross_plot
+```
 
+```{r, eval=T}
 # see here for how to remove bootstrap css which otherwise messes up you blogdown's css
@@
 ```
-</details>
 
-```{r echo=FALSE}
+cross_plot_clean
+```{r echo=FALSE, layout="l-page", fig.height=30}
 
@@
 With more German native speakers attending the school in one school, AHS ratios are higher.
-Consequenlty, German speaking parents send their kids again to the school with the higher AHS ratio and not to the other, similarly close school.
-Obviously, this still leaves a lot of other questions unaswered (e.g.
+Consequently, German speaking parents send their kids again to the school with the higher AHS ratio and not to the other, similarly close school.
+Obviously, this still leaves a lot of other questions unanswered (e.g.
 (why) do non-German speaking parents exhibit the same enrollment behavior as German speaking parents?).

### 2021-01-13 _post/2020-03-02-primary-schools-in-vienna/primary-schools-in-vienna.Rmd
modified post on primary schools; theme;

@@
 date: 03-02-2020
-
-theme: theme.css
 
@@
 loadfonts(device = "win", quiet = T)
-extrafont::fonts()
 library(hrbrthemes)
@@
 library(crosstalk)
+library(htmltools)
 ```
@@
 ```{r setup, echo = F}
-# knit_hooks$set(wrap = function(before, options, envir){
-# if (before){
-# paste0('<', options$wrap, '>')
-# } else {
-# paste0('</', options$wrap, '>')
-# }
-# })
+
 knitr::opts_chunk$set(warning = FALSE, message = FALSE, dpi = 180, fig.height = 5, fig.width = 7, fig.align = "center")
@@
 
-```{r, eval=T}
+```{r, eval=F}
 # add details of schools to dataframe containing all combinations of schools (nested);
@@
 
-# readr::write_rds(df_main, file=here::here("_blog_data", "Vienna_primary_schools","df_main2.rds"))
+# readr::write_rds(df_main, file=here::here("_blog_data", "Vienna_primary_schools","df_main_base.rds"))
+
+```
+
+```{r, eval=T, include=F}
+df_main <- readr::read_rds(here::here("_blog_data", "Vienna_primary_schools", "df_main_base.rds"))
 
@@
     config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))
-```
-</details>
 
-```{r, eval=T}
-# combine graphs
-
-cross_plot <- crosstalk::bscols(
-plot_distance_diff,
-cross_map2)
-#
-
-```
-
-cross_plot <- crosstalk::bscols(
-plot_distance_diff,
-cross_map2)
-```{r eval=FALSE, include=FALSE}
-# cross_plot
-```
+#combine both plots vertically/as two columns
+#problem: introduces bootstrap css which conflicts with css of blog
+cross_plot <- crosstalk::bscols(plot_distance_diff, cross_map2)
 
 cross_plot <- browsable(tagList(cross_plot))
-cross_plot
 
-```{r, eval=T}
-library(htmltools)
-# cross_plot <- browsable(tagList(plot_distance_diff, cross_map2))
-cross_plot <- browsable(tagList(cross_plot))
-# cross_plot
-```
-
-```{r, eval=T}
 # see here for how to remove bootstrap css which otherwise messes up you blogdown's css
 # https://github.com/rstudio/crosstalk/issues/26
+# only works when combining plots vertically though ! hence css not as intended;
 
@@
 ```
+</details>
 
-cross_plot_clean
-```{r echo=FALSE, layout="l-page", fig.height=30}
+```{r echo=FALSE, fig.height=30, layout="l-page"}
 

### 2021-01-13 _post/2020-03-02-primary-schools-in-vienna/primary-schools-in-vienna.Rmd
add revised primary school post

@@
   - plotly
+  - crosstalk
 
-preview: static-maps-with-mapbox-background-images-snapbox_files/preview.png
+preview: primary-schools-in-vienna_files/preview.png
 
@@
 See e.g.
-<br>![Example of table of 4th graders' school destination](/post/2020-02-17-primary-schools-in-vienna_files/vienna_schools_table.PNG){width=50% height=50%}.
+the image below.
+<br>
+
+![Example of table of 4th graders' school destination](primary-schools-in-vienna_files/vienna_schools_table.PNG){width=50% height=100%}
+<br>
 

### 2021-01-13 _post/2020-12-19-vienna-elections-2020-analysis-of-parties-ads-on-facebook/welcome.Rmd
add revised primary school post

@@
   - social media
-  - election
+  - elections
 date: 12-19-2020

### 2021-01-15 _post/2020-03-02-primary-schools-in-vienna/primary-schools-in-vienna.Rmd
reformated post bih voting hor

@@
 description: |
-  Drawing on public data from Vienna's school administration I look at how many percentages of pupils in each primary schoos subsequently leave for schools offering eventually a higher education entrance qualification (Matura).
+  Drawing on public data from Vienna's school administration I look at how many percentages of pupils in each primary school subsequently leave for schools offering eventually a higher education entrance qualification (Matura).
 The main interest is in contrasting these ratios between closely (geo-)located schools.

### 2021-01-15 _post/2020-02-07-bih-voting-behavior-of-mps-in-house-of-representatives/bih-voting-behavior-of-mps-in-house-of-representatives.Rmd
bih hor - add folder

---
title: "Bosnia: Voting behavior of MPs in the House of Representatives"
description: |
  A short description of the post.
date: 02-07-2020

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693

categories:
  - Bosnia
  - OCR
  - web scraping

#preview: primary-schools-in-vienna_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)

```

# CONTEXT
A while ago I noticed that the website of Bosnia's parliament also provides the voting records for each separate vote taken in both houses. These records also include the voting behavior for each individual MP. Here an example of one of these voting records:

![](bih-voting-behavior-of-mps-in-house-of-representatives_files/record.PNG){width=100% height=100%}

While the dominant role of the ethnic cleavage in Bosnian politics has been a constant theme and continuously emphasized for the last 25 years (to an extent which frequently blanks out other salient issues), I hadn't seen any empirical record on MP's individual voting behavior which could corroborate the cleavage's prevalence in the legislative arena (one insightful exception is Birgit Bahtić-Kunrath's 2011 [paper](https://www.cambridge.org/core/journals/nationalities-papers/article/of-veto-players-and-entityvoting-institutional-gridlock-in-the-bosnian-reform-process/96BB80A5191E12B40F501841546C151D){target="_blank"}).

This post digs into the now available records for the 2014-2018 parliament and seeks to shed some light on the legislative process. The first part presents the results, the second highlights the most relevant steps in R to produce them. The entire code for the analysis is again available on my [github account](https://github.com/werkstattcodes/Bosnia_MP_voting_behavior){target="_blank"}. Since the post got a bit longer than initially expected, this one will mainly focus on the lower chamber (House of Representatives). A later post will take up the upper chamber in detail (House of Peoples).

As always, if you see any glaring error, feel free to let me know (best via [twitter](https://twitter.com/zoowalk){target="_blank"} DM). If you think the blog is interesting and feel like quoting it, please refer to it as _Roland Schmidt, "Bosnia: Voting behavior of MPs in the House of Representatives (2014-2018)", https://werk.statt.codes/post/bih-voting-behavior-of-mps-in-house-of-representatives/, 8 February 2020_.

For those working in R, the blog post might be particularly interesting when it comes to retrieving the actual data. As so often with empirical research the challenges has been less the analysis of the data itself, but 'liberating' it from the format within which it is provided. In this case, the voting records for both the House of Peoples and the House of Representative are provided in pdfs, such as the one above or this [_one_](http://parlament.ba/oLaw/GetOwisDocument/?documentId=206164&data=10788B4568269048D57D8F644F829519&lang=hr){target="_blank"} or this [_one_](http://parlament.ba/oLaw/GetOwisDocument/?documentId=193121&data=E0B8243EF8B51ED931E698C78D196E84&lang=hr){target="_blank"}. Hence, there's been some hefty OCR-ing involved thanks to the powerful [`pdf_tools`](https://docs.ropensci.org/pdftools/){target="_blank"} package by [Jeroen Ooms](https://twitter.com/opencpu){target="_blank"}.

# RESULTS

```{r include=FALSE}
# setup -------------------------------------------------------------------
library(tidyverse)
library(ggplot2)
library(paletteer)
library(glue)
library(extrafont)
library(paletteer)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
library(paletteer)
library(ggtext)
library(scales)
library(patchwork)
library(padr)
library(here)

wdr <- getwd()

bgr_color<- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()

blog_theme <- function(){
    hrbrthemes::theme_ipsum_rc() +
    theme(
        axis.text.x = element_text(angle=0, size=10),
        axis.title = element_blank(),
        legend.position = "bottom",
        legend.justification = "left",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(size = 18, 
                                  face="bold", 
                                  margin=margin(b=0.3, unit="cm")),
        plot.title.position="plot",
        plot.background= element_rect(fill=bgr_color, color=bgr_color),
        plot.subtitle = element_text(size = 15, color = "grey30"),
        plot.margin=margin(l=0, t=0.5, b=0.5, unit="cm"),
       # plot.caption = element_text(hjust=c(0,1)),
        panel.spacing.x = unit(0, "cm"))}

my_caption <- c("Data: parlament.ba", "Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes/")

df_ocr_results_unnested <- readr::read_csv2(here("_blog_data", "Bosnia_MP_voting_behavior", "df_ocr_results_unnested.csv"))

df_ocr_results_unnested <- df_ocr_results_unnested %>% 
  mutate_if(is.character, as.factor)

#parse_date_time order allows to specify different formats and their priority

df_ocr_results_unnested <- df_ocr_results_unnested %>% 
  mutate(session_date=as.character(session_date)) %>% 
  mutate(session_date=lubridate::parse_date_time(session_date,
                                                 orders=c("dmy","ymd", "mdy")) %>% 
           lubridate::as_date(.))

#order factor of vote
df_ocr_results_unnested <- df_ocr_results_unnested %>% 
  mutate(vote=forcats::fct_infreq(vote)) #%>% 
 # mutate(vote=forcats::fct_collapse(vote, abstained=c("abstained", "no vote", "restrained")) %>% 
 #          forcats::fct_drop(.))

voting_colors <- c("no"="firebrick2",
                   "no vote" = "steelblue",
                   "not present" = "grey",
                   "yes" = "springgreen3",
                   "abstained"="orange")

# _ ----------------------------------------------------------------------

# FILTER ------------------------------------------------------------------

#identify start of new legislative session
# date_new_parliament <- df_ocr_results_unnested %>% 
#   filter(str_detect(delegate_name_2, "^Ahmetovic")) %>% 
#   filter(session_date==min(session_date)) %>% 
#   distinct(session_date) %>% 
#   pull(session_date)

df_ocr_results_unnested_truncated <- df_ocr_results_unnested %>% 
  filter(session_date > lubridate::dmy("12/10/2014")) %>% #date of elections
  mutate(party=forcats::fct_drop(party)) %>% #remove empty party factor levels
  mutate_at(vars(contains("name")), ~fct_drop(.))

```

## Number of bills considered and outcome

Overall, during the 2014 to 2018 parliament session 163 bills were considered in both houses. Out of this total, only 37 % were eventually passed and published to become law. 38 % of all bills were rejected in at least one of both houses.^[Bills' outcome was translated as follows: "Usvojen"="adopted", "Procedura - nije preuzet"="procedure - not started", "Procedura"="procedure", "Donesen i objavljen"="submitted and published",
"Odbijen"="rejected", "Obustavljen zakonodavni postupak"="legislative process suspended", "Povucen"="withdrawn",
"Ceka na pokretanje procedure"="waiting for the procedure to start", "Nije razmatran"="not considered". If you think there are more fitting translations, feel free to let me know.]

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE}
df_details_all_status_wide <- readr::read_csv2(here::here("_blog_data","Bosnia_MP_voting_behavior","df_details_all_status_wide.csv")) %>% 
    mutate_at(vars(contains("status")), as.factor)

#standardize colors across plots
colors_outcome <- paletteer_d("ggsci::default_jco", n = 10) %>% as.character() 
names(colors_outcome) <- unique(c(levels(df_details_all_status_wide$HoP_status), levels(df_details_all_status_wide$HoR_status), levels(df_details_all_status_wide$BiH_Parl_final_status)))

df_laws_outcome <- df_details_all_status_wide %>% 
  group_by(BiH_Parl_final_status) %>% 
  summarise(n_obs=n()) %>% 
  ungroup() %>% 
  mutate(n_rel=n_obs/sum(n_obs))

df_laws_outcome %>% 
  ggplot()+
  labs(title="BiH House of Representatives: Outcome of bills",
       subtitle="Outcome of bills considered during legislative period 2014-2018; Total number: 163.",
       y="Number of bills",
       caption=my_caption)+
  geom_bar(aes(x=reorder(BiH_Parl_final_status, n_rel),
               y=n_obs,
               fill=BiH_Parl_final_status),
           stat="identity",
           position=position_dodge())+
  geom_text(aes(x=reorder(BiH_Parl_final_status, n_rel),
                y=70,
                label=paste0(n_obs, " (",
                  percent(n_rel, accuracy = 1),
                  ")")),
            family = "Roboto Condensed",
            color = "grey30",
            hjust = 1,
            size=5,
           # group = name,
            stat="identity")+
  scale_fill_manual(values=colors_outcome,
                    name="Outcome:")+
  scale_x_discrete(labels=function(x) str_wrap(x, 15))+
  scale_y_continuous(breaks=seq(0,60, 20),
                     expand=expansion(mult=c(0.01,0)))+
  blog_theme()+
  theme(panel.grid.major.y = element_blank(),
        legend.position = "none",
        panel.grid.major.x = element_line(),
        axis.text.x= element_text(size=14),
        axis.text.y = element_text(size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_blank(),
        plot.caption = element_text(hjust=c(0, 1), size=12))+
  coord_flip()

```

### by house

Looking separately at each of the two houses provides a more detailed picture. In the lower chamber, the House of Representatives,  153 bills were considered in total out of which a bit more than 50 % were approved. In the upper chamber, the House of Peoples, which comprises five delegates of each constituent people (Bosniak, Croats, Serbs), 56.7 % of all considered bills were passed.

```{r echo=FALSE, fig.height=5, fig.width=10}
#in which house are bills rejected

df_outcome_chambers <- df_details_all_status_wide %>% 
  select(contains("status"), -contains("final")) %>% 
  pivot_longer(cols=contains("status"), names_to="house", values_to="outcome") %>% 
  filter(outcome!="missing") %>%
  mutate(outcome=fct_drop(outcome, only="missing")) %>%
  mutate(outcome=fct_infreq(outcome) %>% fct_rev) %>% 
  group_by(house, outcome, .drop=F) %>% 
  summarise(n_obs=n()) %>% 
  group_by(house) %>% 
  mutate(n_rel=n_obs/sum(n_obs)) %>% 
  arrange(desc(n_obs), .by_group=T) %>% 
  mutate(cumsum_n_obs=cumsum(n_obs)) %>% 
  ungroup() %>% 
  mutate(outcome_label = case_when(house=="HoR_status" ~ paste0(outcome, "\n(", n_obs, ")"),
                                   house=="HoP_status" ~ paste0(outcome, "(", n_obs, ")")))

plot_df_outcome_chambers <- df_outcome_chambers %>% 
  group_split(house) %>% 
  map(~ggplot(.)+
  geom_bar(aes(x=1,
               y=n_obs,
               fill=outcome,
               group=outcome),
           stat="identity",
           position = position_stack())+
  geom_text(aes(x=1,
                y = n_obs,
                label=ifelse(n_obs>0,
                             n_obs,
                             ""),
                group=outcome),
            family = "Roboto Condensed",
            color = "white",
            size=5,
            position = position_stack(vjust= 0.4))+
  geom_text(aes(x=1,
                y = n_obs,
                label=ifelse(n_obs>10,
                             paste0("\n\n", scales::percent(n_rel, 
                                                            accuracy = .1)),
                             ""),
                group=outcome),
            family = "Roboto Condensed",
            color = "white",
            size=5,
            position = position_stack(vjust= 0.4))+  
  labs(y="Number of bills",
       title=ifelse(unique(.$house)=="HoP_status", 
                    "House of Peoples",
                    "House of Representatives"))+
      #subtitle="Bills considered in 2014 - 2018 legislative period.",
      # caption=my_caption)+
  scale_fill_manual(values=colors_outcome,
                    name="Outcome of bill:")+
  # scale_y_continuous(expand=expansion(mult=c(0,0.1)),
  #                    breaks=df_outcome_chambers %>% 
  #                      filter(house==.$house) %>% 
  #                      pull(cumsum_n_obs))+
  scale_y_continuous(expand=expansion(mult=c(0,0.1)),
                     limits=c(0, max(df_outcome_chambers$cumsum_n_obs)),
                     # breaks=.$cumsum_n_obs,
                     breaks=max(.$cumsum_n_obs),
                     guide=guide_axis(n.dodge=2),
                     # labels=c(unique(ifelse(.$n_obs<10,
                     #               ifelse(.$cumsum_n_obs==max(.$cumsum_n_obs),
                     #                      # "", ""),
                     #                      unique(max(.$cumsum_n_obs)), ""),
                     #                      .$cumsum_n_obs))," ")
                     labels=max(.$cumsum_n_obs)
                     )+
  coord_flip()+
  # lemon::facet_rep_wrap(vars(house),
  #                       repeat.tick.labels = F,
  #                       nrow=2,
  #                       labeller = labeller(house=c("HoP_status"="House of Peoples",
  #                                                "HoR_status"="House of Representatives")))+
  blog_theme()+
  theme(panel.grid=element_blank(),
          axis.text.x=element_text(size=15),
          axis.title.x = element_text(size=15),
          axis.title.y = element_blank(),

          panel.grid.major.y=element_blank(),
          axis.text.y = element_blank()))

(patchwork::wrap_plots(plot_df_outcome_chambers))+
  plot_layout(ncol=1, guides="collect") &
  plot_annotation(caption="Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes",

                  theme = theme_ipsum_rc()) &
  theme(legend.position = "bottom",
        plot.background=element_rect(fill=bgr_color, color=bgr_color),
        plot.margin=margin(b=0, t=0, l=0, r=0, unit="cm"),
        legend.text = element_text(size=12),
        legend.justification = "left") 
```

### intersection of houses

If we intersect the developments for each law in each chamber, we get a clearer picture where bills were actually rejected/not passed. The table/heatmap below intends to show this. The x-axis shows a bill's outcome in the House of Representative, the y-axis the outcome of the same bill in the House of Peoples (if a bill was considered only in one chamber, the respective status in the other chamber is 'missing').

Out of the 67 bills adopted in the House of Peoples and 77 in the House of Representatives, 60 were actually considered in both chambers and eventually passed.  With 37 % of all bills it's the largest category in the table. The second largest category comprises 27 bills (16.7 % of all bills) which were rejected in the House of Representatives and never made it to the House of Peoples. The third largest category contains 12 bills (7.4 % of all bills) which were adopted in the lower house but rejected in the upper chamber.

```{r echo=FALSE, fig.width=10, fig.height=7 , message=FALSE}
df_outcome_chambers_combined <- df_details_all_status_wide %>% 
  mutate_at(vars(contains("status")), as.factor) %>% 
  group_by_at(vars(contains("status"), -contains("final")),
              .drop=F) %>% 
  summarise(n_obs=n()) %>% 
  ungroup() %>% 
  mutate(n_rel=n_obs/sum(n_obs)) %>% 
  group_by(HoR_status) %>% 
  mutate(HoR_outcome_label=sum(n_obs) %>% paste0(HoR_status, " (", ., ")")) %>% 
  group_by(HoP_status) %>% 
  mutate(HoP_outcome_label=sum(n_obs) %>% paste0(HoP_status, " (", ., ")")) %>% 
  ungroup() %>%  #ungroup before converting to factor
  mutate_at(vars(contains("label")), as_factor)

#to duplicate axis labels scale has to be numeric; convert to numeric, but
#us character labels;

plot_df_outcome_chambers_combined <- df_outcome_chambers_combined  %>% 
  ggplot()+
  labs(title="Outcome of bills by house",
       x="final status in House of Representatives",
       y=str_wrap("final status in House of Peoples", 20),
       caption=my_caption)+
  geom_tile(aes(x=as.numeric(HoR_outcome_label),
                y=as.numeric(HoP_outcome_label),
                fill=n_obs))+
  geom_text(data=. %>% filter(n_obs!=0),
              aes(x=as.numeric(HoR_outcome_label),
                y=as.numeric(HoP_outcome_label),
                label=paste0(n_obs, " (",
                             percent(n_rel, accuracy = 0.1),
                             ")")),
            family = "Roboto Condensed",
            color = "grey30",
            hjust = .5,
            # group = name,
            stat="identity")+
  scale_fill_gradient(low="white", high="steelblue",
                      na.value="white")+
  scale_x_continuous(labels=levels(df_outcome_chambers_combined$HoR_outcome_label) %>% map_chr(., label_wrap(15)),
                     breaks=seq(1, length(unique(df_outcome_chambers_combined$HoR_outcome_label))),
                     sec.axis = dup_axis())+
  scale_y_continuous(labels=levels(df_outcome_chambers_combined$HoP_outcome_label) %>% map_chr(., label_wrap(20)),
                     breaks=seq(1, length(unique(df_outcome_chambers_combined$HoP_outcome_label))),
                     sec.axis = dup_axis(),
                     trans="reverse")+
  blog_theme()+
  theme(axis.text.x.top=element_blank(),
        axis.title.x.top = element_text(hjust=0, size=11),
        axis.title.x.bottom = element_blank(),
        axis.text.y.left=element_blank(),
        axis.title.y.left = element_text(hjust=0, size=11),
        axis.title.y.right = element_blank(),
        #axis.text.x.bottom=element_text(angle=0, hjust=.5),
        panel.grid = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none",
        panel.border = element_rect(color="grey30", fill=NA),
        axis.title.y = element_text(angle=0),
        plot.title = element_text(size = 14, face="bold", margin=margin(b=0.3, unit="cm")),
        plot.title.position = "plot",
        plot.background= element_rect(fill=bgr_color, color=bgr_color),
        plot.margin=margin(l=0, r=0, b=0, unit="cm"),
        plot.subtitle = element_text(size = 12, color = "grey30"),
        plot.caption.position = "plot",
        plot.caption = element_text(color = "grey30", hjust = c(0, 1)))

plot_df_outcome_chambers_combined

```

## Legislative votes over time

How was the legislative activity distributed over time? The graph below provides the number of ballots in both houses per month. Note that the graph also includes votes which were held in both houses outside of the 2014-2018 legislative session. The reason is that bills considered in 2014-2018 also include bills which were initiated prior to 2014 and hence also ballots which were taken outside of the 2014-2018 legislative session.

If we focus only on the 2014-2018 session, it is rather clear that both chambers were most active in the first half of 2016 and the end of 2017. I am not sure exactly what caused this bout of activity, but it's quite distinct if contrasted e.g. with 2015 or the early parts of 2017.

```{r echo=FALSE, fig.width=10, fig.height=6}
#number of votes per week
#use floor_date to aggregate witin a week
#use pdr to thicken to weeks, then pad for missing weeks, then plot;

election_dates <- c("07/10/2018", "12/10/2014") %>% 
  enframe(., 
          value="election_date",
          name=NULL) %>% 
  mutate(election_date=lubridate::dmy(election_date)) %>% 
  mutate(election_month=lubridate::floor_date(election_date, unit="month"))

df_ocr_results_unnested %>% 
  mutate(month=lubridate::floor_date(session_date, unit="month")) %>%   group_by(house, month) %>% 
  summarise(n_votes=n_distinct(record_id)) %>%
  pad(., interval="month",
      start_val=min(.$month),
      end_val=max(.$month)
  ) %>% 
  padr::fill_by_value() %>% 
  mutate(year=lubridate::year(month)) %>% 
  ggplot()+
  labs(subtitle="Number of items voted on per month.",
       title="Legislative activity",
       caption=my_caption)+
  geom_bar(aes(y=n_votes,
               x=month),
          orientation="x",
          fill="steelblue",
           stat="identity")+
  geom_vline(data=election_dates,
             aes(xintercept=election_month,
                 color="election"))+
  scale_x_date(date_breaks = "1 month",
               labels=scales::label_date(format="%b"))+  #prints year only when year changes
  scale_color_manual(values=c("election"="orange"))+
  facet_grid(house~year,
             scales="free_x",
             space="free_x",
             shrink=F)+
  guides(color=guide_legend(title=NULL))

#check session number and dates

df_sessions <- df_ocr_results_unnested %>% 
  select(house, session_date, session_no) %>% 
  distinct() %>% 
  arrange(house, session_date) %>% 
  mutate(new_parlament=case_when(session_no<lag(session_no) ~ "start",
                                 TRUE ~ NA_character_))

```

## Voting behavior per MP

So far the analysis presented was limited to bills as the unit of analysis. Now let's dig a bit deeper and see how individual MPs voted. The graph below provides the results for every voting decision (including abstentions, no vote, not present) of every MP, by party affiliation, during the 2014-2018 legislative period. Since a bill undergoes several readings before it can become a law/is forwarded to the other chamber, there are more votes than bills. Overall, MPs in the House of Representatives were asked to cast their votes 451 times in total. These votes pertain to 148 laws (out of the initial 162 laws above eleven have not entailed any vote by the end of the 2014-2018 legislative period and three entailed ballots prior to the 2014 elections).

In the graphs below each vertical line (small rectangle) represents an MP's vote in one ballot (hence the x axis is 451 small rectangles long; ordered chronologically). According to the voting records, there are five possible options: an MP can cast 1) a 'yes' (*za*), or 2) an 'against' or 'no' vote (*protiv*); 3) be not present during the ballot (*nije prisuta*); 4) abstained (*suzdržan*), or 5) do not cast a vote (*nije glasao*). Note that the Rules of Procedure for the House of Representatives (as for the House of Peoples) stipulate that 'in case a representative does not vote “in favour”, “against” or “abstained”, but he/she is present at the session during the voting, he/she is considered “abstained” (see [here](https://www.parlament.ba/Content/Read/60?title=Odlu%C4%8Divanje-u-Predstavni%C4%8Dkom-domu){target="_blank"}).'

Unfortunately, I had to remove names' diacritics which added a major source of complication (at least for me).

```{r echo=FALSE, fig.height=12, fig.width=10, warning=FALSE}
# tile graph --------------------------------------------------------------
library(ggforce)
library(ggiraph)

n_record_ids <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  pull(record_id) %>% 
  n_distinct()

plots_tile <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") #%>% 
  #mutate(vote=forcats::fct_collapse(vote, abstained=c("abstained", "no vote", "restrained")))

plots_tile <- plots_tile %>% 
  ggplot()+
  labs(x="items voted at",
       y="Member of House of Representative",
       subtitle="451 decisions in total; 2014-2018 legislative period.",
       title="House of Representative: Vote choice by individual MPs per decision",
       caption=my_caption)+
  geom_tile(aes(x=reorder(record_id, session_date),
                y=fct_rev(delegate_name_family),
                fill=vote))+
  # scale_y_discrete(label=delegate_name_2)+
  scale_fill_manual(values=voting_colors,
                    na.value="white")+
  theme_ipsum_rc() +
  theme(axis.text.x=element_blank(),
        axis.title.x=element_text(hjust=0),
        strip.text = element_text(face="bold"),
        panel.background = element_rect(fill=bgr_color, color=bgr_color),
        panel.grid.major = element_blank(),
        plot.subtitle = element_text(size = 12, color = "grey30"),
        plot.title = element_text(size = 14, face="bold", margin=margin(b=0.3, unit="cm")),
        plot.margin=margin(l=0, t=0.5, b=0.5, unit="cm"),
        plot.title.position = "plot",
        plot.caption = element_text(hjust=c(0, 1)),
        plot.background= element_rect(fill=bgr_color, color=bgr_color),
        legend.position = "top",
        legend.justification = "left")+
  facet_wrap(vars(party),
             ncol = 3,
             scales="free_y")

plots_tile

```

Even without going into numeric details, I find the graph rather informative. While with most parties/MPs the yes votes (green) dominate, the SNSD is a glaring aberration. Milorad Dodik's SNSD clearly stands out as the party which says 'no' (to paraphrase Ian Paisley) or simply doesn't show up when asked to cast a vote.

Note that the representatives Vjekoslav Bevanda and Halid Genjac were only briefly members of the House of Representatives. If not mistaken, Bevanda subsequently became Chairperson of the Council of Ministers. From the graph it appears that his place was taken by Predrag Kožul. Similarly, SDA's Genjac was only for the early sessions member of the HoR. His place seems to have been taken over by Hazim Rančić.

As a general caveat, note that I took MPs' party affiliation as stated on the parliament's website which may omit instances in which MPs have changed their party affiliation in the course of the 2014-2018 legislative period. I am not aware that this was actually the case, but if so, please let me know.^[One way to improve the graph would be to have equally sized heights per MP.]

## MP ranking per vote category

To get a more detailed view on MPs' individual voting behavior, the next graph presents each MP's voting behavior per voting category as % of an MP's number of total voting decisions (incl. abstentions etc). In other words, the graph reveals how often an MP votes for e.g. yes, no etc. as % of all instances in which she was tasked to cast a vote. Party affiliation is presented next to each MP's name. Furthermore, to highlight discrepancies between representatives originating (in terms of mandate) from the Federation and the RS, the bars are colored differently. The variation is rather remarkable, particularly with regard to MPs' 'no' votes and absence (not present and abstentions).

I leave it up to the reader to go through each category, but just to highlight one case: According to the published voting records, Nikola Špirić, one of SNSD's main figure heads (and previously Chairperson of the Council of Ministers), was absent in 44.8 % of all ballots. Even with the benefit of doubt and acknowledging that an MP's work entails more than participating in ballots (work in committees etc), this is a rather, let's say, 'remarkably' low rate of presence.  His party colleagues Markovic and Milovanovic participated only a few times more often. In those instances in which they were present, they voted predominantly with 'no'. Constructive engagement in the legislative process looks different, at least to me.

```{r echo=FALSE, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}
library(tidytext)
library(lemon)

df_votes_delegates  <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  group_by(entity, party, delegate_name_2, delegate_name_family, vote) %>% 
  summarise(n_votes_cat=n()) %>% 
  group_by(party, delegate_name_2, delegate_name_family) %>% 
  mutate(n_votes_total=sum(n_votes_cat)) %>% 
  mutate(n_votes_cat_rel=n_votes_cat/n_votes_total) %>% 
  ungroup()

df_votes_delegates %>% 
  mutate(delegate_name_3=str_extract(delegate_name_2, "[:alpha:]+,\\s[:alpha:]?")) %>% 
  ggplot()+
  labs(title="HoR members' voting decision by category during 2014-2018 legislative period (in %)",
       subtitle ="In how many % of all her votes did the MP vote/was.... Absolute number in category and total in brackets.",
       caption=my_caption)+
  geom_bar(aes(x=reorder_within(paste0(delegate_name_3, " (", party, ")"), n_votes_cat_rel, vote),
               y=n_votes_cat_rel,
               fill=entity),
           stat="identity",
           position=position_stack())+
  geom_text(aes(x=reorder_within(paste0(delegate_name_3, " (", party, ")"), n_votes_cat_rel, vote),
                y=1.2,
                label=paste0(scales::percent(n_votes_cat_rel, accuracy =0.1), " (", n_votes_cat, "/", n_votes_total, ")")),
            family = "Roboto Condensed",
            color = "grey30",
            hjust = 1,
            stat="identity")+
  scale_fill_paletteer_d("ggsci::default_jama")+
  scale_x_reordered()+
  scale_y_continuous(minor_breaks = NULL,
                     labels=scales::label_percent(accuracy=1),
                     breaks=seq(0,0.8,.1),
                     expand=expansion(mult=c(0, 0)))+
  theme_ipsum_rc()+
  theme(axis.text.x = element_text(angle=90),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.spacing = unit(0, "cm"),
        strip.text = element_text(face="bold"),
        plot.title = element_text(size = 14, face="bold", margin=margin(b=0.3, unit="cm")),
        plot.title.position = "plot",
        plot.caption.position = "plot",
        plot.caption = element_text(hjust=c(0,1)),
        plot.margin=margin(l=0, t=0.5, b=0.5, r=0, unit="cm"),
        plot.subtitle = element_text(size = 12, color = "grey30"),
        plot.background=element_rect(fill=bgr_color, color=bgr_color),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "top",
        legend.justification = "left")+
  lemon::facet_rep_wrap(~vote,
                        repeat.tick.labels = T,
                        ncol=2,
                        scales="free_y")+
  coord_flip()

```

## Party differences
As a next step, let's group MPs' voting behavior by party and the respective (ascribed or self-declared) ethnic affiliation. Admittedly, the latter is not without problems, but I guess the categorization below reflects some kind of middle-ground (SDP could probably be moved to the 'Bosniak' category, but well...). The table provides the total number of votes casted by all representatives of each party and their respective share in each vote category. Note the high percentage rate of SDA in the 'yes' category and SNSD's ratio in the 'no' or 'not present' category.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.html.bsTable = TRUE)

df_party_voting <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  mutate(ethnicity=forcats::fct_relevel(ethnicity, "Bosniak", "Croat", "Serb", "civic", "independent")) %>% 
  group_by(ethnicity, party, vote) %>% 
  summarise(n_obs=n()) %>% 
  group_by(ethnicity, party) %>% 
  mutate(n_total=sum(n_obs)) %>% 
  mutate(n_rel=n_obs/n_total)

df_party_voting %>% 
  mutate(n_rel=scales::percent(n_rel)) %>% 
  pivot_wider(., id_cols=c(ethnicity, party, n_total), names_from=vote, values_from=n_rel) %>% 
  rename(`votes total`=n_total) %>% 
  mutate(`number MPs`=`votes total`/451) %>% 
 # mutate(total_reps=map_dbl(`# representatives`, sum)) %>% 
  select(ethnicity, party, `number MPs`, everything()) %>% 
  kable() %>%   
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                fixed_thead = T, 
                font_size = 12) %>%
  scroll_box(width = "100%", height = "400px")

```

Below the same results, put presented as a graph.

```{r echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}

df_party_voting %>% 
  ggplot()+
  geom_bar(aes(x=tidytext::reorder_within(party, n_rel, within=vote),
               fill=vote,
               y=n_rel),
           stat="identity")+
  geom_text(aes(x=tidytext::reorder_within(party, n_rel, within=vote),
                #y=n_rel+.1,
                y=1.1,
                label=paste0(n_obs, "/", n_total, 
                             " (", 
                             scales::percent(n_rel, accuracy=.1),
                             ")")),
            hjust=1,
            size=4,
            family = "Roboto Condensed",
            color = "grey30",
            stat="identity")+
  scale_x_reordered()+
  facet_rep_wrap(vars(vote),
                 repeat.tick.labels = T,
                 ncol=2,
             scales="free_y")+
  scale_y_continuous(expand=expansion(mult=0),
                     limits=c(0, 1.2),
                     labels=scales::label_percent(accuracy=1),
                     breaks=seq(0, .7, .25))+
  scale_fill_manual(values=voting_colors)+
  coord_flip()+
  theme_ipsum_rc()+
  theme(plot.subtitle = element_text(size = 12, color = "grey30"),
        plot.title = element_text(size = 14, face="bold", margin=margin(b=0.3, unit="cm")),
        plot.title.position="plot",
        plot.margin=margin(l=0, t=0.5, b=0.5, r=0, unit="cm"),
        plot.background=element_rect(fill=bgr_color, color=bgr_color),
        panel.grid.major.y =  element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        strip.text = element_text(face="bold"),
        axis.title.y=element_blank(),
        axis.title.x=element_text(hjust=0),
        plot.caption = element_text(hjust=c(0,1)))+
  labs(title="Parties' voting decisions per category (HoR)",
       subtitle="% of all decisions by members of each party; Legislative period 2014-2018.",
       caption=my_caption,
       y="voting decisions in category as % of all decisions by all MPs of each party")

```

```{r eval=FALSE, fig.height=20, fig.width=10, warning=FALSE, include=FALSE}

colors_ethnicity=c("Bosniak"="darkgreen",
                   "Croat"="blue",
                   "Serb"="red",
                   "civic"="orange",
                   "independent"="grey")

df_delegate_voting <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  mutate(ethnicity=forcats::fct_relevel(ethnicity, "Bosniak", "Croat", "Serb", "civic", "independent")) %>% 
  group_by(ethnicity, party, delegate_name, vote, .drop=T) %>% 
  summarise(nobs=n()) %>%
  mutate(total.votes=sum(nobs, na.rm = T)) %>% 
  mutate(perc=nobs/total.votes) %>% 
  ungroup()

plot_delegates <- df_delegate_voting %>% 
  group_split(vote) %>% 
  map(~ggplot(., aes(x=reorder_within(party, desc(perc), ethnicity),
                     y=perc,
                     group=party,
                     tooltip=paste0(delegate_name, "\n",
                                   nobs, " out of ",
                                   total.votes, "\n",
                                   "ratio:", 
                                   scales::percent(perc)),
                     #tooltip=delegate_name,
                     color=ethnicity))+
        geom_jitter_interactive(width=0.03)+
        geom_boxplot(aes(weight=nobs))+
        labs(title=paste(unique(.$vote)))+
        hrbrthemes::theme_ipsum_rc()+
        theme(legend.position="none",
              plot.title = element_text(size = 14, face="bold", margin=margin(b=0.3, unit="cm")),
              plot.subtitle = element_text(size = 12, color = "grey30"),
              axis.title = element_blank(),
              axis.text.x=element_text(angle=45,
                                       vjust=1,
                                       hjust=1,
                                       size=11),
              strip.text.y = element_text(angle=180, 
                                          vjust=1,
                                          hjust=1),
              plot.margin = margin(t=0, b=0, unit="cm"),
              panel.grid.minor = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.spacing.x = unit(0.3, "cm"),
              plot.background=element_rect(fill=bgr_color, color=bgr_color),
              panel.spacing.y = unit(0.3, "cm"))+
        scale_y_continuous(labels=scales::label_percent(accuracy=1),
                           minor_breaks = NULL,
                           limits=c(0,NA),
                           position = "right",
                           sec.axis = dup_axis())+
        scale_x_reordered()+
        scale_color_manual(values=colors_ethnicity, na.value="grey")+
        #scale_color_paletteer_d("ggsci::default_jama")+
        facet_row(~ethnicity,
                  # switch = "y",
                  scales="free_x",
                  space="free"))

wrap_plot_delegats <- wrap_plots(plot_delegates, ncol = 1)

girafe(ggobj=wrap_plot_delegats,
       options=list(opts_tooltip(css = "background-color:lightgray; font-family:Roboto Condensed;",
                                 delay_mouseout = 5000)),
       pointsize=6,
       width_svg=10, height_svg=15)
```

## Conclusion

Overall, the results are pretty much in line with what I expected, but the details nevertheless are quite intriguing. Delegates of SNSD are either absent or vote with 'no' in the House of Representatives. There is a clear gulf between representatives from the Federation and the RS when it comes to 'no' votes.

Since this has been already rather lengthy (not to speak of the actual coding), I leave it here and will continue the analysis with a new post in a bit. If you spotted anything peculiar or have any other constructive comment etc. feel free to let me know and I will try to follow up on it. In any case, thanks for reading.

----

# STEPS IN R
Below the most important steps to obtain the results from above. The entire code for the  (and more) is available at my github [*here*](https://github.com/werkstattcodes/Bosnia_MP_voting_behavior){target="_blank"}.

As already mentioned above, getting the actual data from the voting records/pdfs has been the main challenge. Essentially, there were three steps required:

* First, get the links to each law 
* Second, from there extract
  + data on the status and details of the law in each chamber
  + the links to each voting record(s)
  + use this links to download each voting results file
* Third, get 
  + the raw text from the voting result files and 
  + extract data of interest (e.g. MP and vote)
* Finally, start with the actual analysis ;-)

## Get the data
### Getting the links to each law
The parliament's website provides a list with all bills considered during the 2014-2018 session. See [*here*](http://parlament.ba/oLaw/GetOLawsByStatus?SearchTerm=&MandateId=4&Status=-1){target="_blank"}. Overall, there are 17 pages which contain short summaries on each law plus the links a site with more details on each law. Iterating over these 17 pages and extracting the pertaining links provides us with a dataframe containing the links to the 162 (proposed) laws which were considered in at least one of both houses. To do this, the `purrr` and `rvest` package are our tools of choice. The `possibly` option provides us with a NULL response should the link no yield any info (instead of causing an error).

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(tidyverse)
library(glue)
library(rvest)

#define pages to iterate over
seq_page<- seq(1:17)
seq_page_links <- glue("http://parlament.ba/oLaw/GetOLawsByStatus?page={seq_page}&MandateId=4&Status=-1") %>% 
  enframe(name=NULL)

#define function to extract short info + link of each law 
pb <- progress_estimated(nrow(seq_page_links))

fn_scrap <- function(x){

  pb$tick()$print()

  x %>% 
    read_html() %>% 
    html_nodes("a") %>% 
    html_attr(.,"href") %>% 
    enframe(name=NULL) %>% 
    filter(str_detect(value, "OLawDetails")) %>% 
    mutate(links_to_law_details=paste0("http://parlament.ba", value))

}

#apply function to each of the 17 pages
df_links_to_laws <- seq_page_links$value %>% 
  set_names() %>% 
  map_dfr(.,  possibly(fn_scrap, otherwise=NULL), .id="seq_page_links")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.html.bsTable = TRUE)

df_links_to_laws %>% 
  select(-value) %>%
  kable() %>%   
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T, font_size = 10) %>%
  scroll_box(width = "100%", height = "200px")
```

With these links we are now able to scrap the status and details of each single (proposed) law (see e.g. [*here*](http://parlament.ba/olaw/OLawDetails?lawId=61794){target="_blank"}). In short, the code produces the same result as if we were clicking through our initial list from above and copy-paste the information from each law's details page. The `rvest`'s functions were again applied to each link with the purrr's `map_dfr`.

```{r eval=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

pb <- dplyr::progress_estimated(nrow(df_links_to_laws)) #define length for progress bar

#scrap law details
fn_scrap_law_details <- function(x) {

  pb$tick()$print()

  df.contents <- x %>% 
    read_html() %>% 
    html_nodes("td") %>% 
    html_text(.,trim=T) %>% 
    enframe(name=NULL, value="content") %>% 
    slice(c(1:13))

  df.names<- x %>% 
    read_html() %>% 
    html_nodes("th") %>% 
    html_text(.,trim=T) %>% 
    enframe(name=NULL, value="heading") %>% 
    slice(c(1:13))

  bind_cols(df.names, df.contents) %>% 
    mutate(house=case_when(row_number()>7 ~ "DP",    #House of Peoples
                           TRUE ~ "PS"))           #House of Representatives

}

#Scrap overview of law status 
fn_get_law_status <- function(x) {
  x %>%
    read_html() %>%
    html_nodes("body > section > div.container > div > div.col-md-8 > article > div > div:nth-child(2) > table") %>%
    html_table() %>%
    enframe(name = NULL) %>%
    map_df(., bind_rows) %>%
    rename(
      heading = X1,
      content = X2
    ) %>%
    mutate(heading = case_when(
      lag(heading) %>% str_detect(., "Status u DNPSBiH") ~
        paste("DNPSBiH -", heading),
      lag(heading) %>% str_detect(., "Status u PDPSBiH") ~
        paste("PDPSBiH -", heading),
      TRUE ~ as.character(heading)
    ))
}

#apply function to each link
df_details_all_status<- df_links_to_laws$links_to_law_details %>% 
  set_names() %>% 
  future_map_dfr(.,  fn_get_law_status, 
                 .progress=T,
                 .id="seq_page_links")
```

### Get status of each law
As a first step, I want to get info on each law's status by the end of the 2014-2018 legislative period in each house. We get this info by applying the below scrapping function to each law's link (obtained above). Subsequently I reshape the data to a more convenient wider rectangular form (`pivot_wider`) and extract each field of interest with `stringr` and `regex`.

```{r eval=FALSE}
df_details_all_status_wide <- df_details_all_status %>% 
  mutate(content=str_trim(content, side=c("both"))) %>% 
  mutate(content=iconv(content, from="UTF-8", to="windows-1253")) %>% #removes diacrits; makes it easier to work with
  pivot_wider(id_cols = "seq_page_links",
              names_from = "heading",
              values_from = "content") %>% 
  mutate(law_id=str_extract(seq_page_links, "[0-9]+$")) %>% 
  rename(doc_name ="Naziv dokumenta",
         period = Saziv,
         HoR_draft_law_number_date  = "Broj i datum Prijedloga zakona u PDPSBiH",
         HoR_tabler = `Predlagač u PDPSBiH`,
         HoR_status = "Status u PDPSBiH",
         HoR_order_number_session_date_agenda_item = `PDPSBiH - Red. br. i datum sjednice - tačka dnevnog reda`,
         HoP_draft_law_number_date= `Broj i datum Prijedloga zakona u DNPSBiH`,
         HoP_tabler =`Predlagač u DNPSBiH`,
         HoP_status =`Status u DNPSBiH`,
         HoP_order_number_session_date_agenda_item =`DNPSBiH - Red. br. i datum sjednice - tačka dnevnog reda`,
         BiH_Parl_final_status=`Konačni status u PSBiH`,
         ended_tabler_session =`Utvrđen na sjednici predlagača`) %>% 
  separate(col=HoR_draft_law_number_date, 
           into=c("HoR_draft_law_number", "HoR_draft_law_date"),
           sep="od",
           remove=F) %>% 
  separate(col=HoP_draft_law_number_date, 
           into=c("HoP_draft_law_number", "HoP_draft_law_date"),
           sep="od",
           remove=F) %>% 
  mutate_at(vars(contains("draft_law_date")), ~str_remove(., ".$") %>% 
              lubridate::dmy(.)) %>% 
  mutate_at(vars(ends_with("draft_law_number")), ~str_trim(., side=c("both")) %>% 
            str_remove(., ",$") %>% str_replace(., ",","-")) %>% 
  mutate(HoP_order_number_session_date_agenda_item=ifelse(HoP_order_number_session_date_agenda_item=="",
                                                          NA, HoP_order_number_session_date_agenda_item)) %>% 
  mutate(HoP_order_number=stringr::str_extract_all(HoP_order_number_session_date_agenda_item,
                                                   "^[0-9]+") %>% as.numeric) %>% 
  mutate(HoP_session_date=stringr::str_extract_all(HoP_order_number_session_date_agenda_item,
                                                   "[:digit:]+\\.[:digit:]+\\.[:digit:]{4}") %>% 
           map_chr(., paste, collapse="; ")) %>% 
  mutate(HoP_agenda_item=stringr::str_extract(HoP_order_number_session_date_agenda_item,
                                              "Ad\\. [:digit:]+")) %>% 
#  select(-HoP_order_number_session_date_agenda_item) %>% 
  mutate(HoR_order_number_session_date_agenda_item=ifelse(HoR_order_number_session_date_agenda_item=="",
                                                          NA, HoR_order_number_session_date_agenda_item)) %>% 
  mutate(HoR_order_number=stringr::str_extract_all(HoR_order_number_session_date_agenda_item,
                                                   "^[0-9]+") %>% as.numeric) %>% 
  mutate(HoR_session_date=stringr::str_extract_all(HoR_order_number_session_date_agenda_item,
                                                   "[:digit:]+\\.[:digit:]+\\.[:digit:]{4}") %>% 
           map_chr(., paste, collapse="; ")) %>% 
  mutate(HoR_agenda_item=stringr::str_extract(HoR_order_number_session_date_agenda_item,
                                              "Ad\\. [:digit:]+")) %>% 
#  select(-HoR_order_number_session_date_agenda_item) %>% 
  mutate_at(vars(contains("status")),
            .funs=list(~case_when(
              .=="Usvojen"~"adopted",
              .=="Procedura - nije preuzet" ~ "procedure - not started",  #?
              .=="Procedura" ~ "procedure",
              str_detect(., regex("Povucen")) ~ "withdrawn",  #Povucen encoding issue
              .=="Donesen i objavljen" ~ "submitted and published",
              .=="Odbijen" ~ "rejected",
              .=="Obustavljen zakonodavni postupak" ~ "legislative process suspended",
              str_detect(., "Ceka na pokretanje procedure") ~ "waiting for the procedure to start", #encoding issue
              .=="Nije razmatran" ~ "not considered",
              is.na(.) ~ "missing",
              TRUE ~ .)))

```

Finally, to make things more accessible to clowns like me who haven't managed to properly learn BCS after too many years, I translate each law's title with google translate (checkout the vignette of the powerful [googleLanguageR](https://cran.r-project.org/web/packages/googleLanguageR/index.html){target="_blank"} package and to how get the necessary authorization for the google translate engine).

```{r include=FALSE}
library(googleLanguageR)
gl_auth("C:/Users/Roland/Google Drive/Events - Projects/R-Projects/GoogleAPI/BiH-Laws-49dca6b0c199.json")

```

```{r eval=FALSE, message=FALSE, warning=FALSE}
df_details_all_status_wide<- df_details_all_status_wide %>% 
  mutate(doc_name_eng=googleLanguageR::gl_translate(doc_name, target="en", format=c("text"),
                                                    source="bs")$translatedText) %>%
  select(doc_name, doc_name_eng, everything())

```

So this all yields us with a table containing plenty of information on each bill, including the eventual status in each house, the date of the bills submission and the name of those who tabled the bill (in each house). Not too bad me thinks.

```{r echo=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.html.bsTable = TRUE)

df_details_all_status_wide %>% 
  kable() %>%   
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T, font_size = 10) %>%
  scroll_box(width = "100%", height = "600px")

```

### Get details of each law
In the next step I want to go beyond each bill's eventual status, but get details on each bill's development. The function below, again applied to each link we initially obtained provides us with this. See e.g. [here](http://parlament.ba/olaw/OLawDetails?lawId=61794){target="_blank"}. Note, to speed things up, I use the `furrr` package with the multiprocess option to apply the function in parallel and not in sequence to each link. Hence, rather than scrapping the details of each bill one after the other, they are downloaded in parallel and therefore much quicker.

```{r message=FALSE, warning=FALSE}
library(furrr)
plan(multiprocess)

fn_get_law_status <- function(x) {
  x %>%
    read_html() %>%
    html_nodes("body > section > div.container > div > div.col-md-8 > article > div > div:nth-child(2) > table") %>%
    html_table() %>%
    enframe(name = NULL) %>%
    map_df(., bind_rows) %>%
    rename(
      heading = X1,
      content = X2
    ) %>%
    mutate(heading = case_when(
      lag(heading) %>% str_detect(., "Status u DNPSBiH") ~
        paste("DNPSBiH -", heading),
      lag(heading) %>% str_detect(., "Status u PDPSBiH") ~
        paste("PDPSBiH -", heading),
      TRUE ~ as.character(heading)
    ))
}

df_details_all_status<- df_links_to_laws$links_to_law_details %>% 
  set_names() %>% 
  future_map_dfr(.,  fn_get_law_status, 
                 .progress=F,
                 .id="seq_page_links")

```

### Getting links to each voting record 
Critically, the details page of each bill also contains links to the records for each vote taken pertaining to the bill in question. This can be only one vote in only one house, or multiple votes in each house since a bill goes through several readings and entails also votes on procedural issues (e.g. whether the bill should be fast-tracked). For an example see [here](http://parlament.ba/olaw/OLawDetails?lawId=61565){target="_blank"} and note that the links to the voting records are called 'Listing glasanja o skracenom postupku' or 'Listing glasanja o zahtjevu za hitni postupa'. The critical point here is that links to voting records always contain the word 'listing'. What is hence required at this stage is to extract from all bills' detail pages every link which contains the word 'listing'. Once these links are obtained, the voting records to which these links lead can be downloaded. The function below extracts these links of interest.

```{r message=FALSE, warning=FALSE}

fn_scrap_links_to_voting_records <- function(x) {

    x %>% 
    read_html() %>% 
    html_nodes(xpath="//a[contains(text(), 'Listing')]") %>%  #filters links based on text/name of links
    html_attr('href') %>%  #extracts links
    enframe(name=NULL) %>% 
    mutate(link_to_voting_record=paste0("http://parlament.ba", value))

}

df_voting_results_links <- df_links_to_laws$links_to_law_details %>% 
  set_names() %>% 
  future_map_dfr(.,  possibly(fn_scrap_links_to_voting_records, otherwise=NULL), 
                 .progress = F, .id="seq_page_links") %>% 
  select(-value) %>% 
  mutate(law_id=str_extract(seq_page_links, "[0-9]+$"),
         record_id=str_extract(link_to_voting_record, "(?<=documentId=)[0-9]+")) #pos. look-behind

```

```{r echo=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.html.bsTable = TRUE)

df_voting_results_links %>% 
  kable() %>%   
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T, font_size = 10) %>%
  scroll_box(width = "100%", height = "200px")
```

### Download voting records
Overall we obtain links to 793 voting records which pertain to any of both houses. With these links we can now download each voting record. Note that these voting records pertain to the bills which were considered during the 2014-2018 legislative period. This does not mean the voting took actually place in the 2014-2018 period. A bill might have been initiated prior to 2014 and hence some voting took place prior to the 2014-2018 period. At a later step, once when we have extracted the actual date of each vote, we are able to separate votes which took place during the 2014-2018 session from others.

```{r, eval=FALSE, echo=T}
# download voting records (pdfs) 
pdf_destination <- glue("{wdr}/data/voting_records/law_id_{df_voting_results_links$law_id}-record_id_{df_voting_results_links$record_id}.pdf")

furrr::future_walk2(df_voting_results_links$link_to_voting_record, pdf_destination, download.file, mode = "wb") 
```
## Extracting text from voting records (OCR)
Now, with eventually our objects of desire obtained, we have to retrieve the data we are actually interested in from each pdf, e.g.  eave a look at [*this record*](http://parlament.ba/oLaw/GetOwisDocument/?documentId=201825&data=66A251EE8272F76CD4563213F5B76864&lang=bs){target="_blank"}.

The primary tool which enables us to extract the text from the pdfs is the `pdf_tools` package (see package site [here](https://docs.ropensci.org/pdftools/){target="_blank"}). What hte package basically does is it recognizes the characters in each pdf and extracts them as plain text. Before doing so, one has to download the required language package. Subsequently, apply a function which extracts the entire (!) text from each pdf.

Below the main steps (see the full script at my repo [here.](https://github.com/werkstattcodes/Bosnia_MP_voting_behavior/blob/master/script/ocr_files.R){target="_blank"}).

```{r eval=FALSE}
library(pdftools)

#downlaod required language
tesseract_download("bos")
myengine<- tesseract(language = "bos",
                     datapath =paste0(wdr, "/tesseract_lang/") )

#function to apply to each pdf file/voting record
fn_extract_text <- function(x) {

  x %>% 
    pdf_ocr_text(., language="bos") %>% 
    data.frame(text_raw=.)
}

#apply function in parallel to speed things up
x<- file_list %>% 
  set_names() %>% 
  future_map_dfr(., fn_extract_text, .progress=F, .id="link")

#group extracted text by voting record (unique link); relevant for
#multi-page documents;

df_ocr_results <- x %>% 
  group_by(link) %>% 
  summarise(raw_text=paste(text_raw, collapse=", "),
            pages=n()) %>% 
  mutate(raw_text=iconv(raw_text, from="UTF-8", to="windows-1253"))

```
## Extract data of interest from ocr-ed text
Once this is done, the probably trickiest part of the entire endeavor comes. Extracting our data of interest from the raw text obtained above. Below, in the R code chunk comments which hopefully make a bit clear what's going on.

```{r eval=FALSE}

#check whether the voting record pertains to the upper or lower house and create a pertaining indicator variable
df_ocr_results <- df_ocr_results %>%
  mutate(house_indicator=str_extract(raw_text, "Predstav*|naroda")) %>% 
  mutate(house = case_when(
    str_detect(house_indicator, regex("Predstav*", ignore_case = T)) & !str_detect(house_indicator, regex("naroda", ignore_case = T)) ~ "HoR",
    str_detect(house_indicator, "naroda") ~ "HoP",
    TRUE ~ as.character("missing")
  ))

# Split the long raw text of each bills into multiple lines whenever a line break was in the pdf.

df_ocr_results <- df_ocr_results %>%
  mutate(raw_text2 = stringr::str_split(raw_text, "\n"))

#Extract the session number, the vote number and session date;
#regex has to account for different date formats

df_ocr_results <- df_ocr_results %>%
  mutate(raw_text = stringr::str_squish(raw_text)) %>%
  mutate(session_no = stringr::str_squish(raw_text) %>% stringr::str_extract(., "[0-9]+(?=\\.?\\s?sjednic)")) %>%
  mutate(vote_no = stringr::str_extract(raw_text, "(?<=(Glasanje br:\\s?|Redni broj.{0,20}))[0-9]+")) %>% 
  mutate(session_date = str_extract(raw_text, "[0-9]{1,2}/[0-9]{1,2}/20[0-9]{2}|[0-9]{1,2}.[0-9]{2}.20[0-9]{2}|[0-9]{4}-[0-9]{1,2}-[0-9]{1,2}|[0-9]{1,2}.[:alpha:]+20[1-9]{2}."))

# Extract those lines which contain the votes of each indiviudal member of parliament (yes, no, absent...).
# extract from these lines each MP's name, entity and vote.

df_ocr_results <- df_ocr_results %>%
  mutate(delegate_votes = map(raw_text2, ~ str_subset(., regex(" Za| Nije prisutan| PROTIV| UKUPNO| SUZDRZAN| NIJE GLASAO",
    ignore_case = T
  )) %>%
    str_subset(., " FBiH | RS | Federacija | Republika Srpska ") %>%
    str_remove(., "^[^[:alpha:]]+") %>% # remove every non-alpha before the frist alpha.
    str_squish(.) %>%
    enframe(name = NULL, value = "delegate_raw"))) %>%
  mutate(delegate_votes = map(delegate_votes, ~ .x %>%
    mutate(
      delegate_name = stringr::word(delegate_raw, 1, 2),
      entity = case_when(
        str_detect(delegate_raw, "FBiH|Federacija") ~ "FBiH",
        str_detect(delegate_raw, "RS|Republika Srpska") ~ "RS",
        TRUE ~ as.character("missing")
      ),
      vote = case_when(str_detect(delegate_raw, regex(" za", ignore_case = T)) ~ "yes",  #space before is important; attention suzdrZAn;
                           str_detect(delegate_raw, regex(" protiv", ignore_case = T)) ~ "no",
                           str_detect(delegate_raw, regex(" suzdrzan", ignore_case = T)) ~ "restrained", #reserved
                           str_detect(delegate_raw, regex(" nije glasao", ignore_case = T)) ~ "no vote", 
                           str_detect(delegate_raw, regex(" nije prisutan", ignore_case = T)) ~ "not present",
                       TRUE ~ as.character("missing")
                     ) %>% stringr::str_to_lower()
      ))) %>%
  mutate(delegate_num = map_int(delegate_votes, nrow)) %>% 
  mutate(law_id=str_extract(link, "(?<=law_id_)[0-9]+"),
         record_id=str_extract(link, "(?<=record_id_)[0-9]+")) %>% #pos. look-behind 
  mutate(session_date=as.character(session_date)) %>% 
  mutate(session_date=lubridate::parse_date_time(session_date, #parse_date_time recognizes different date formats
                                                 orders=c("dmy","ymd", "mdy")) %>% 
           lubridate::as_date(.))

#above code provided info on MPs, their votes nested within each voting record; code below unnests and provides one row per MP and vote;
df_ocr_results_unnested <- df_ocr_results %>% 
  unnest(delegate_votes)

#As it turned out one voting record (07.10.2015) caused some ocr erros which need to be accounted for manually; below the necessary modifcations:
df_ocr_results_unnested<- df_ocr_results_unnested %>% 
  mutate(delegate_name = case_when(delegate_name=="Zaim Backovib" ~ "Zaim Backovic",
                                   str_detect(delegate_name, "Aleksandra Pandu") ~ "Aleksandra Pandurevic",
                                   str_detect(delegate_name, "Amir Fazl") ~ "Amir Fazlic",
                                   str_detect(delegate_name, "Asim Sarajlib") ~ "Asim Sarajlic",
                                   str_detect(delegate_name, "Borislav Boji") ~ "Borislav Bojic",
                                   str_detect(delegate_name, "Ljiulja Zovko") ~ "Ljilja Zovko",
                                   str_detect(delegate_name, "D. Nermina") ~ "Nermina Kapetanovic",
                                   str_detect(delegate_name, "D. Safer") ~ "Safer Demirovic",
                                   str_detect(delegate_name, "Magazinovic") ~ "Sasa Magazinovic",
                                   TRUE ~ as.character(delegate_name))) %>% 
  mutate(delegate_name_2=paste(str_remove(delegate_name, "^[:alpha:]+\\s"),
                               word(delegate_name, 1), sep=", ")) %>% 
  mutate(delegate_name_family=stringr::word(delegate_name, -1))
```

The table below presents first 200 out of 24637 rows of the resulting dataframe which is essentially the basis for any further analysis. Each row represents the vote of one MP for a specific ballot. For future research the obtained dataset is made available [here](https://github.com/werkstattcodes/Bosnia_MP_voting_behavior/blob/master/data/df_ocr_results_unnested.csv){target="_blank"}.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.html.bsTable = TRUE)

my_link <- "C:/Users/Roland/Google Drive/Events - Projects/R-Projects/Bosnia_MP_voting_behavior/data/df_ocr_results_unnested.csv"

read_csv2(my_link,
          n_max=200) %>% 
  select(-link) %>% 
  select(law_id, record_id, -raw_text, everything()) %>% 
  kable() %>%   
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T, font_size = 10) %>%
  scroll_box(width = "100%", height = "500px")

```

## Get party affiliation of MPs

Since we are also interested in MP's party affiliation (already included in the table above) we need to get this information. Luckily, the parliament's website also provides us with this detail. The code below extracts the information (for both houses and all legislative sessions).

```{r eval=FALSE}
library(tidyverse)
library(rvest)
library(glue)

wdr <- getwd()

# Members of House of REPRESENTATIVES 
page_index <- seq(1,5,1)
mandate_no <- c(6, 7,8, 9)
df_ids<- crossing(page_index, mandate_no)

MP_links <- glue("https://www.parlament.ba/Representative/List?page={df_ids$page_index}&mandateId={df_ids$mandate_no}")

pb <- dplyr::progress_estimated(length(MP_links))

fn_scrap_MPs <- function(x) {

  pb$tick()$print()

  x  %>% 
    read_html() %>% 
    html_nodes("table") %>% 
    html_table(.,trim=T, fill=T) %>% 
    map_dfr(., as_tibble)
}

df_members_HoRepresentatives <- MP_links %>% 
  set_names() %>% 
  map_dfr(.,  possibly(fn_scrap_MPs, otherwise = NULL), .id = "MP_links")

df_members_HoRepresentatives<- df_members_HoRepresentatives %>% 
  rename(name=X2,
         party2=X3,
         party=X4,
         constituency=X5) %>% 
  #  mutate(name=map(., iconv, from="UTF-8", to="window-1253")) %>% 
  tidyr::separate(name, c("family_name", "first_name"), sep=", ") %>% 
  mutate(party2=stringr::str_remove(party2, "Klup poslanika ")) %>% 
  mutate_at(vars(contains("name"), party), iconv, from="UTF-8", to="windows-1253")

df_members_HoRepresentatives<- df_members_HoRepresentatives %>% 
  select(-X1) %>% 
  mutate(first_name=stringr::str_squish(first_name)) %>% #removes space between double names
  unite(name, c("first_name","family_name"), sep=" " , remove=FALSE) %>% 
  mutate(house="House of Representatives") %>% 
  mutate(entity=case_when(str_detect(constituency, "RS") ~ "RS",
                          str_detect(constituency, "FBiH") ~ "FBiH",
                          TRUE ~ as.character(constituency))) %>% 
  mutate(party.original=party) %>% 
  mutate(party=case_when(party.original=="Samostalni poslanik" ~ "independent",
                         str_detect(party.original, "SBB") ~ "SBB",
                         str_detect(party.original, "BPS") ~ "BPS",
                         str_detect(party.original, "fronta") ~ "DF",
                         str_detect(party.original, "Za evropsku") ~ "A-SDA",
                         str_detect(party.original, "Koalicija HDZ") ~ "HDZ Coalition",
                         TRUE ~ as.character(party.original))) %>% 
  mutate(period=str_extract(MP_links, "(?<=mandateId=)[0-9]+")) %>% 
  mutate(period=case_when(period=="6" ~ "2006-2010",
                          period=="7" ~ "2010-2014",
                          period=="8" ~ "2014-2018",
                          period=="9" ~ "2018-2022",
                     TRUE ~ NA_character_)) %>%
  mutate(house="HoR") %>% 
  mutate_at(vars(contains("name"), party), iconv, from="UTF-8", to="windows-1253") %>% 
  mutate(name=case_when(name=="Nermina Zaimovic - Uzunovic" ~ "Nermina Zaimovic-Uzunovic",
                        name=="Mladen Ivankovic Lijanovic" ~ "Mladen Ivankovic-Lijanovic",
                        name=="Vesna Krstovic – Spremo" ~ "Vesna Krstovic-Spremo",
                        TRUE ~ as.character(name)))

# Members of House of PEOPLES ---------------------------------------------

page_index <- seq(1,5,1)
mandate_no <- c(6, 7, 8, 9)
df_ids<- crossing(page_index, mandate_no)

MP_links <- glue("https://www.parlament.ba/Delegate/List?page={df_ids$page_index}&mandateId={df_ids$mandate_no}")

pb <- dplyr::progress_estimated(length(MP_links))

fn_scrap_MPs <- function(x) {

  pb$tick()$print()

  y <- x  %>% 
    read_html() %>% 
    html_nodes("table") %>% 
    html_table(.,trim=T, fill=T) %>% 
    map_dfr(., as_tibble)
}

df_members_HoPeoples <- MP_links %>% 
  set_names() %>% 
  map_dfr(.,  possibly(fn_scrap_MPs, otherwise = NULL), .id = "MP_links")

df_members_HoPeoples<- df_members_HoPeoples %>% 
  rename(name=X2,
         party2=X3,
         party=X4,
         constituency=X5) %>% 
  tidyr::separate(name, c("family_name", "first_name"), sep=", ") %>% 
  mutate(party2=stringr::str_remove(party2, "Klup poslanika ")) %>% 
  mutate_at(vars(contains("name"), party), iconv, from="UTF-8", to="windows-1253")

df_members_HoPeoples<- df_members_HoPeoples %>%
  select(-X1) %>% 
  mutate(first_name=stringr::str_squish(first_name)) %>% #removes space between double names
  unite(name, c("first_name","family_name"), sep=" " , remove=FALSE) %>% 
  mutate(house="House of Peoples") %>% 
  rename(entity=constituency) %>% 
  mutate(party.original=party) %>% 
  mutate(party=stringr::word(party, 1)) %>% 
  mutate(period=str_extract(MP_links, "(?<=mandateId=)[0-9]+")) %>% 
  mutate(period=case_when(period=="6" ~ "2006-2010",
                          period=="7" ~ "2010-2014",
                          period=="8" ~ "2014-2018",
                          period=="9" ~ "2018-2022",
                          TRUE ~ NA_character_)) %>% 
  mutate(house="HoP")

df_members_parliament <- bind_rows(df_members_HoPeoples, df_members_HoRepresentatives)  
head(df_members_parliament)

```
With this source on MP's party affiliation however comes an important caveat: The source does not provide any details on whether MP's changed party affiliation during (!) a legislative period. I am not aware that this has actually been the case, but if so, it would distort results grouped by party affiliation (please let me know in case you're aware of such a change of party affiliation).

## Analysis
The rest of the R code is more or less 'only' number crunching and producing the graphs above. I won't go into much detail here since i think it's rather straightforward and renders the blog post only unnecessarily longer. In any case, the code is available [*here*](https://github.com/werkstattcodes/Bosnia_MP_voting_behavior/blob/master/script/analysis.R){target="_blank"}.

### 2021-01-22 _post/2019-11-20-vorzugsstimmen/vorzugsstimmen.Rmd
new theme, revised preferential votes post

---
title: "Austrian General Elections 2019: Preference votes"
description: |
  A look at the distribution of preference votes casted in the 2019 general elections.

date: 11-20-2019
draft: false

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - elections
  - stringr
  - tidyr::pivot
  - Austria

#preview: austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
    code_folding: true

creative_commons: CC BY-NC-SA

---

```{r setup, include=FALSE}
library(rvest, quietly = T, verbose = F, warn=F)
library(tidyverse, quietly = T, verbose = F, warn=F)
library(knitr, quietly =T)
library(paletteer)
library(ggtext)
library(ggforce)
library(extrafont)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
library(patchwork)
library(ggiraph)
library(lemon)
library(reactable)

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
  str_trim() %>% 
  str_extract(., regex("^#\\S+"))

knitr::opts_chunk$set(echo = T, background=plot_bg_color)

theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      #panel.border = element_rect(colour = plot_bg_color, fill=NA),
      #plot.border = element_rect(colour = plot_bg_color, fill=NA),
      plot.margin = margin(l = 0, 
                           t = 0.25,
                           unit = "cm"),
      plot.title = element_text(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, b=0.25, unit = "cm"),
        size = 18
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 11
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 11
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(
        angle = 0,
        color = "grey50",
        hjust = 1
      ),
      axis.text.x = element_text(
        size = 12,
        color = "grey50"
      ),
      axis.title.y = element_blank(),
      axis.text.y = element_text(
        size = 9,
        color = "grey50"
      ),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 7
      ),
      legend.text = element_text(
        size = 12,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

my_caption_2 <- c("Data: bmi.gv.at/412/Nationalratswahlen/Nationalratswahl_2019/start.aspx#vorzugsstimmen\nAnalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes")

```

```{r include=F, echo=F}
# preps -------------------------------------------------------------------

party_colors <- c(FPÖ = "#005DA8", NEOS = "#EA5290", ÖVP = "#5DC2CC", SPÖ = "#FC0204", GRÜNE = "#A3C630")

main_parties <- c("SPÖ", "ÖVP", "NEOS", "GRÜNE", "FPÖ")

```

```{r extract_data, include=F, echo=F}
# get state and regional list--------------------------------------

df_state_region <- readxl::read_xlsx(path = here::here("_blog_data", "2019_nrw_vorzugsstimmen", "NRW19_Vorzugsstimmen_Landes_Regionalparteiliste_16102019.xlsx"))

names(df_state_region) <- names(df_state_region) %>%
  stringr::str_remove(., "\r") %>%
  stringr::str_remove(., "\n") %>%
  stringr::str_remove(., "-")

df_state_region <- df_state_region %>%
  janitor::clean_names()

df_state_region_long <- df_state_region %>%
  mutate(party = map_chr(reihung_1, ~ str_extract(., regex("^List.*")))) %>%
  mutate(state_list = map_chr(reihung_1, ~ str_extract(., regex("^Landespartei.*")))) %>%
  mutate(regional_list = map_chr(familien_vorname_2, ~ str_extract(., regex("^Regionalpartei.*")))) %>% # mess in bmi file; heading in different columns
  mutate(regional_list_2 = map_chr(reihung_1, ~ str_extract(., regex("^Regionalpartei.*")))) %>%
  mutate(district = coalesce(state_list, regional_list, regional_list_2)) %>%
  janitor::remove_empty("rows") %>%
  fill(., district, .direction = c("down")) %>%
  fill(., party, .direction = c("down")) %>%
  select(-c("x6", "state_list", "regional_list", "regional_list_2")) %>%
  filter(str_detect(reihung_1, "^[0-9]")) %>%
  select(district, party, everything()) %>%
  rename_at(
    vars(reihung_7:vorz_stimm_11),
    function(x) str_replace_all(x, "[:digit:]+", "b")
  ) %>%
  rename_at(
    vars(reihung_1:vorz_stimm_5),
    function(x) str_replace_all(x, "[:digit:]+", "a")
  ) %>%
  mutate_at(vars(reihung_a:vorz_stimm_b), as.character) %>%
  pivot_longer(
    cols = reihung_a:vorz_stimm_b, names_to = c(".value", "result"),
    names_pattern = "(.*)_([a|b]$)"
  ) %>%
  mutate_at(vars(reihung, geb_jahr, vorz_stimm), as.numeric) %>%
  group_by(district, party) %>%
  arrange(reihung, .by_group = T) %>%
  mutate(district_type = case_when(
    str_detect(district, regex("Landes.*")) ~ "Landeswahlkreis",
    str_detect(district, regex("^Regional.*")) ~ "Regionalwahlkreis",
    TRUE ~ NA_character_
  )) %>%
  select(district_type, district, party, everything(), -result) %>%
  rename(
    name = familien_vorname,
    pref_votes_abs = vorz_stimm,
    # district_type = district_type,
    # district = district,
    rank_list = reihung,
    birth_year = geb_jahr,
  ) %>%
  ungroup()

# get federal list --------------------------------------

df_federal <- readxl::read_xlsx(
  path=here::here("_blog_data", "2019_nrw_vorzugsstimmen", "NRW19_Vorzugsstimmen_bundesweit_16102019.xlsx"),
  col_name = F,
  trim_ws = T
)

names(df_federal) <- c(
  "Nr", "Familienname, Vorname", "Geb.-Jahr", "Beruf",
  "Gesamt", "B", "K", "NÖ", "OÖ", "S", "Stmk", "Tirol",
  "Vlbg", "Wien"
)

df_federal_clean <- df_federal %>%
  janitor::remove_empty(., which = c("rows")) %>%
  filter(!Nr == "Nr.") %>%
  mutate(party = str_extract(Nr, regex("[:alpha:].*"))) %>%
  fill(., party, .direction = c("down")) %>%
  filter(!str_detect(Nr, regex("[:alpha:]"))) %>%
  mutate(district_type = "Bundeswahlkreis") %>%
  pivot_longer(cols = Gesamt:Wien, names_to = "district", values_to = "pref_votes_abs")

# rename to be consistent with regional/state level
df_federal_clean <- df_federal_clean %>%
  select(district_type, district, party,
    rank_list = "Nr", name = "Familienname, Vorname", birth_year = "Geb.-Jahr",
    beruf = "Beruf", everything()
  ) %>%
  mutate_at(vars("rank_list", "birth_year", "pref_votes_abs"), ~ stringr::str_trim(., side = c("both"))) %>%
  mutate_at(vars("rank_list", "birth_year", "pref_votes_abs"), as.numeric)

# bind both bund, laender and regio -------------------------------------------------------------------

df_merge <- bind_rows(
  df_state_region_long,
  df_federal_clean
) %>%
  ungroup() %>% 
  filter(!is.na(name))

# write.csv2(df_merge,
#            file = paste0(wdr, "/data/AUT_NRW19_preference_votes.csv"),
#            row.names = F,
#            fileEncoding = "latin1"
# )

fn_remove_titles <- function(x) {
  titles <- c(
    "Mag", "MMag", "Prof", "Dr", "O\\.Univ\\.Prof", "Dipl-Ing", "Dipl.-Ing.", "Dipl.-Rev.", "Ing",
    "Dipl", "jun", "sen", "Ba", "\\(FH\\)", "MSc",
    "med", "univ", "Bsc", "phil", "iur", "DDr", "BEd", "DI", "Mba", "Mas", "MA", "MIM",
    "vet", "M.A.", "PMM", "MTD", "Päd", "Bakk"
  ) %>%
    map(., paste0, c("\\.", "\\s", "\\,")) %>%
    unlist()

  name_wo_titles <- stringr::str_remove_all(x, regex("[[:alpha:]\\-]+\\.")) %>%
    str_remove_all(., regex("\\([:alpha:]+\\)")) %>%
    str_remove_all(., regex(paste0(titles, collapse = "|"), ignore_case = T)) %>%
    str_squish()
}

fn_short_name <- function(x) {
  family_name <- stringr::str_extract(x, "^[[:alpha:]\\-|\\s]+(?=,)") %>%
    str_to_title() %>%
    str_squish() %>%
    str_trim()

  first_name <- str_extract(x, regex("(?<=\\,)[[:alpha:]\\-\\s]*$")) %>%
    str_trim() %>%
    str_squish() %>%
    str_to_title()

  initals <- str_split(first_name, "\\s") %>%
    map(., str_sub, start = 1L, end = 1L) %>%
    map_chr(., str_c, collapse = ".") %>%
    paste0(., ".")

  name_short <- paste(family_name, initals, sep = ", ")
}

df <- df_merge %>%
  mutate(district = str_remove(district, "Landesparteiliste") %>%
    str_remove(., "Regionalparteiliste") %>%
    str_trim(., side = c("both"))) %>%
  mutate(district = case_when(
    district == "B" ~ "Burgenland",
    district == "S" ~ "Salzburg",
    district == "K" ~ "Kärnten",
    district == "NÖ" ~ "Niederösterreich",
    district == "OÖ" ~ "Oberösterreich",
    district == "Stmk" ~ "Steiermark",
    district == "Vlbg" ~ "Vorarlberg",
    TRUE ~ district
  )) %>%
  mutate(district = as_factor(district)) %>%
  mutate(party2 = str_extract(party, regex("\\([:alpha:]*\\)")) %>% # take parties' short abbrevs.
    str_remove_all(., "\\(|\\)")) %>%
  mutate(name_orig = name) %>%
  mutate(name = map_chr(name_orig, fn_remove_titles)) %>%
  mutate(name_short = map_chr(name, fn_short_name)) %>%
  mutate(state = case_when(
    str_detect(district, "^[0-9].*") ~ str_sub(district, start = 1L, end = 1L),
    TRUE ~ as.character(district)
  )) %>%
  mutate(state = case_when(
    state == 1 ~ "Burgenland",
    state == 2 ~ "Kärnten",
    state == 3 ~ "Niederösterreich",
    state == 4 ~ "Oberösterreich",
    state == 5 ~ "Salzburg",
    state == 6 ~ "Steiermark",
    state == 7 ~ "Tirol",
    state == 8 ~ "Vorarlberg",
    state == 9 ~ "Wien",
    # TRUE ~ NA_character_
    TRUE ~ state,
  )) %>%
  mutate(district_type = fct_relevel(district_type, "Bundeswahlkreis", "Landeswahlkreis", "Regionalwahlkreis")) %>% 
  mutate(name = case_when(
    str_detect(name, "Pistracher") ~ "Pistracher, Gerald",
    TRUE ~ as.character(name)))

readr::write_csv2(df, file=here::here("_blog_data", 
                                      "2019_nrw_vorzugsstimmen",
                                      "df.csv"))
```

```{r include=FALSE}
df <- readr::read_csv2(file=here::here("_blog_data", 
                                       "2019_nrw_vorzugsstimmen",
                                       "df.csv"))

df <- df %>%
  mutate(party3 = fct_other(party2, keep = main_parties) %>%
    fct_relevel(., "ÖVP", "SPÖ", "FPÖ", "GRÜNE", "NEOS", "Other"))
```

```{r include=FALSE}
# INCLUDED top 5 candidates per party ----------------------------------------------

top_candidates <- df %>%
  # filter(party2 %in% main_parties) %>%
  filter(!str_detect(district, "Gesamt")) %>%
  group_by(party3, name) %>%
  summarise(preferences_candidate = sum(pref_votes_abs, na.rm = T)) %>%
  arrange(desc(preferences_candidate), .by_group = T) %>%
  slice(1:5)

df_pl <- df %>%
  filter(name %in% top_candidates$name) %>%
  filter(!str_detect(district, "Gesamt")) %>%
  group_by(party3, name, district_type) %>%
  summarise(sum_candiate = sum(pref_votes_abs, na.rm = T)) %>% # rename sum_candidate
  group_by(party3) %>%
  arrange(desc(sum_candiate), .by_group = T) %>%
  ungroup() %>%
  group_by(name) %>%
  mutate(total_candidate = sum(sum_candiate)) %>%
  ungroup() %>%
  mutate(name = forcats::fct_reorder(name, total_candidate, .desc = F))
```

```{r include=FALSE}
pl_top_5_candidates <- df_pl %>%
  ggplot() +
  geom_bar(aes(
    x = name,
    y = sum_candiate,
    fill = district_type
  ),
  width = 0.8,
  stat = "identity"
  ) +
  geom_text(
    data = top_candidates,
    aes(
      x = name,
      y = max(top_candidates$preferences_candidate)+70000,
      label = scales::comma(preferences_candidate, accuracy = 1)
    ),
    stat = "identity",
    family = "Roboto Condensed",
    color = "grey50",
    hjust = 1
  ) +
  labs(
    caption = my_caption_2,
    title = "Candidates with most preference votes (Top 5)",
    y = "Number of preference votes in 1,000"
  ) +
  scale_x_discrete(label = fn_short_name) +
  scale_y_continuous(
    label = scales::comma_format(scale = 0.001),
    expand = expansion(mult = c(0, 0.1)),
    breaks = seq(0, max(top_candidates$preferences_candidate), 25000),
    minor_breaks = NULL
  ) +
  scale_fill_paletteer_d("ggsci::default_jama",
    labels = c(
      "Regionalwahlkreis" = "Regional lists",
      "Landeswahlkreis" = "State lists",
      "Bundeswahlkreis" = "Federal list"
    )
  ) +
  facet_wrap(vars(party3),
    ncol = 2,
    scale = "free_y"
  ) +
  coord_flip() +
  theme_post()+
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(),
    panel.spacing.x = unit(0, units = "cm"),
    panel.spacing.y = unit(0.25, units = "cm"),
    axis.title.y = element_blank(),
    axis.title.x = element_text(hjust=0, size=13),
    # axis.text.y = element_text(hjust=0),
    axis.text.x = element_text(size=13),
    axis.text.y = element_text(size=13),
    plot.caption=element_text(hjust=0),
    legend.title = element_blank(),
    legend.direction = "horizontal",
    legend.position = "top",
    legend.justification = "left",
    legend.key.size = unit(0.4, units = c("cm")),
    plot.title.position = "panel",
    plot.background = element_rect(fill=plot_bg_color),
    strip.text = element_text(face = "bold",
                              size=14)
  ) +
  guides(fill = guide_legend(rev = T, ))

```

```{r, include=F}
# INCLUDED top per district_type (not included) -----------------------------------------------------------

fn_top_plot <- function(x) {
  x %>%
    ggplot() +
    geom_bar(aes(
      x = name,
      y = pref_votes_abs
    ),
    stat = "identity"
    ) +
    labs(title = paste(unique(x$district_type))) +
    scale_x_discrete(labels = fn_short_name) +
    hrbrthemes::theme_ipsum_ps() +
    facet_wrap(vars(district),
      scales = "free_y"
    ) +
    coord_flip()
}

list_plots <- df %>%
  filter(!district == "Gesamt") %>%
  group_by(district_type, district) %>%
  arrange(desc(pref_votes_abs), .by_group = T) %>%
  slice(1:3) %>%
  ungroup() %>%
  group_split(district_type) %>%
  map(~ fn_top_plot(.))

```

```{r, include=F}
# difference btw ranking and number of preferential votes -----------------

df_rank_diff <- df %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  group_by(party3, district_type, district) %>%
  arrange(desc(pref_votes_abs), .by_group = T) %>%
  mutate(rank_votes = row_number()) %>%
  mutate(rank_diff = rank_list - rank_votes) %>%
  select(district_type, district, party3, name, pref_votes_abs, rank_list, rank_votes, rank_diff) %>%
  arrange(desc(rank_diff))

df_rank_diff_long <- df_rank_diff %>%
  filter(party3 %in% main_parties) %>%
  pivot_longer(
    cols = c(rank_list, rank_votes),
    values_to = "rank_no",
    names_to = "ranking_type"
  )

# df_rank_diff_long %>%
#   ggplot() +
#   geom_point(aes(
#     x = fct_rev(ranking_type),
#     y = rank_no,
#     group = name,
#     color = party3
#   )) +
#   geom_line(aes(
#     x = fct_rev(ranking_type),
#     y = rank_no,
#     group = name,
#     color = party3
#   )) +
#   ggrepel::geom_text_repel(aes(
#     x = fct_rev(ranking_type),
#     y = rank_no,
#     group = name,
#     label = name,
#     color = party3
#   ),
#   size = 2
#   ) +
#   facet_grid(rows = vars(party3), cols = vars(district_type))

# INCLUDED Composition of preferential votes ---------------------------------------

df_list <- df %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  group_by(district_type) %>%
  summarise(district_type_votes = sum(pref_votes_abs, na.rm = T)) %>%
  #  arrange(desc(district_type_votes)) %>%
  mutate(district_type = forcats::fct_relevel(district_type, "Regionalwahlkreis", "Landeswahlkreis", "Bundeswahlkreis")) %>%
  mutate(district_type_votes_cum = cumsum(district_type_votes)) %>%
  mutate(district_type_votes_perc = district_type_votes / sum(district_type_votes)) %>%
  mutate(district_type = forcats::fct_reorder(district_type, district_type_votes_cum, .desc = F))

pl_by_district <- df_list %>%
  ggplot() +
  geom_bar(aes(
    x = 1,
    y = district_type_votes,
    fill = district_type,
    group = district_type
  ),
  stat = "identity"
  ) +
  geom_text(aes(
    x = 1,
    y = district_type_votes,
    group = district_type,
    label = paste(scales::comma(district_type_votes),
                  scales::percent(district_type_votes_perc),
                  sep = "\n"
    )
  ),
  colour = "white",
  position = position_stack(vjust = .5, reverse = F)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    breaks = max(df_list$district_type_votes_cum),
    label = scales::comma(max(df_list$district_type_votes_cum))
  ) +
  scale_fill_paletteer_d("ggsci::default_jama",
                         labels = c(
                           "Regionalwahlkreis" = "Regional lists",
                           "Landeswahlkreis" = "State lists",
                           "Bundeswahlkreis" = "Federal list"
                         )
  ) +
  labs(
    title = "Total number of preference votes per electoral district",
    caption = my_caption_2,
    y = "number of preference votes"
  ) +
  coord_flip() +
  guides(fill = guide_legend(reverse = T))+
  theme_post()+
  theme(legend.position="top",
        legend.justification = "left",
        legend.direction  = "horizontal",
        plot.caption =element_text(hjust=0),
        legend.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank()

  )
```

```{r, include=F}
# INCLUDED by constituency list and party  -------------------------------------------------------------

x <- df %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  group_by(party3, district_type) %>%
  summarise(district_type_votes = sum(pref_votes_abs, na.rm = T)) %>%
  arrange(desc(district_type_votes)) %>%
  mutate(district_type_votes_cum = cumsum(district_type_votes)) %>%
  mutate(district_type_votes_perc = district_type_votes / sum(district_type_votes)) %>%
  ungroup()

pl_constituency_list_party <- x %>%
  filter(party3 != "Other") %>%
  ggplot() +
  geom_bar(aes(
    x = party3,
    y = district_type_votes,
    group = district_type,
    fill = district_type
  ),
  color = NA,
  stat = "identity",
  position = "fill"
  ) +
  geom_text(aes(
    x = party3,
    group = district_type,
    y = district_type_votes_perc,
    label = scales::percent(district_type_votes_perc, accuracy = 0.1)
  ),
  color = "white",
  position = position_stack(vjust = 0.5, reverse = F)
  ) +
  labs(
    title = str_wrap("Contribution of different electoral constituencies to total of preferential votes", 70),
    caption = my_caption_2
  ) +
  scale_fill_paletteer_d("ggsci::default_jama",
    labels = c(
      "Regionalwahlkreis" = "Regional lists",
      "Landeswahlkreis" = "State lists",
      "Bundeswahlkreis" = "Federal list"
    )
  ) +
  scale_y_continuous(
    labels = scales::percent,
    expand = expansion(mult = c(0, 0))
  ) +
  scale_x_discrete(expand = expansion(mult = c(0, 0.2)),
                   position = "top") +
  theme_post()+
  theme(axis.title.x=element_blank(),
        legend.title = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title=element_text(),
        legend.position="top",
        legend.justification = "left",
        legend.direction ="horizontal",
        plot.caption = element_text(hjust=c(0)))+
  guides(fill = guide_legend(reverse = F))

```

```{r, include=F, eval=T}
# save --------------------------------------------------------------------

# x <- girafe_options(
#   x = my_plot,
#   sizingPolicy(padding = "0px")
# )
# 
# saveWidget(x,
#   file = "Party_PVPv_state.html",
#   selfcontained = T,
#   background = "white"
# )
#

# name <- "Party_PVPv_state"
# format <- ".png"
# time <- Sys.Date()
# folder <- paste0(wdr, "/graphs/")
# height <- 20
# width <- 30
# ggsave(
#   filename = paste0(folder, time, name, format),
#   device = "png",
#   type = "cairo",
#   height = height,
#   width = width,
#   scale = 1,
#   unit = c("cm")
# )
#

# intra-party movements due to preference votes --------------------------

# > import election results -----------------------------------------------

nrw19_imp <- readr::read_csv2(
  file = here::here("_blog_data", "2019_nrw_vorzugsstimmen", "wahl_20191003_214253.csv"),
  locale = locale(encoding = "latin1")
) %>%
  janitor::remove_empty(., which = c("cols")) %>%
  janitor::clean_names() %>%
  rename(gkz = x1)

nrw19 <- nrw19_imp %>%
  mutate(state = str_sub(gkz, 2, 2)) %>%
  mutate(district_pol = str_sub(gkz, 3, 4)) %>% # Bezirk
  mutate(municipality = str_sub(gkz, 5, 6)) %>%
  mutate(level = case_when(
    str_sub(gkz, 2, 2) == 0 ~ "gesamt",
    str_detect(gkz, regex("^[A-Z][1-9]00[0,9][0,9]")) ~ "Bundesland",
    str_detect(gkz, regex("^G.{1}[A-Z].+")) ~ "Regional-Wahlkreis",
    str_detect(gkz, regex("^G[1-9][0-9][1-9]00")) ~ "Bezirk",
    str_detect(gkz, regex("^G[1-9][0-9][1-9]99")) ~ "Bezirk",
    str_detect(gkz, regex("^G[1-9][0-9][1-9][0-8][0-9]")) ~ "Gemeinde",
    TRUE ~ as.character("u")
  )) %>%
  mutate(wahlkarten = case_when(
    str_detect(gkz, regex("99$")) ~ "ja",
    TRUE ~ as.character("nein")
  )) %>%
  select(level,
    district_name = gebietsname, gkz, electorate_size = wahlberechtigte,
    votes_mail = wahlkarten, state, votes_casted = abgegebene, district_pol,
    municipality, everything()
  )

nrw19_bezirk <- nrw19 %>%
  filter(votes_mail == "nein") %>%
  filter(level == "Bezirk")

nrw19_bezirk <- nrw19_bezirk %>%
  mutate(turnout = votes_casted / electorate_size * 100) %>%
  mutate_at(vars(spo, ovp, fpo, grune, neos), .funs = list(perc = ~ .x / gultige))

nrw19_long <- nrw19_bezirk %>%
  select(contains("perc"), district_name, gkz, turnout) %>%
  pivot_longer(., cols = contains("perc"), names_to = "party", values_to = "party_perc")

# > federal level ---------------------------------------------------------
results_federal_list <- nrw19 %>%
  filter(level == "gesamt")

results_federal_list_long <- results_federal_list %>%
  pivot_longer(cols = ovp:slp, names_to = "party", values_to = "votes_abs") %>%
  mutate(district_type = "Bundeswahlkreis") %>%
  mutate(district = "00") %>%
  select(district_type, district_name, district, party, votes_abs)

# > state level -----------------------------------------------------------

results_state_list <- nrw19 %>%
  filter(level == "Bundesland") %>%
  group_by(state) %>%
  summarise_if(is.numeric, sum, na.rm = T) %>%
  mutate(district_name = case_when(
    state == 1 ~ "Burgenland",
    state == 2 ~ "Kärnten",
    state == 3 ~ "Niederösterreich",
    state == 4 ~ "Oberösterreich",
    state == 5 ~ "Salzburg",
    state == 6 ~ "Steiermark",
    state == 7 ~ "Tirol",
    state == 8 ~ "Vorarlberg",
    state == 9 ~ "Wien",
    TRUE ~ NA_character_
  )) %>%
  mutate(district = state)

results_state_list_long <- results_state_list %>%
  pivot_longer(cols = ovp:slp, names_to = "party", values_to = "votes_abs") %>%
  mutate(district_type = "Landeswahlkreis") %>%
  select(district_type, district_name, district, party, votes_abs)

# >> details on electoral districts/constituencies -----------------------

library(rvest)

MoI_url <- "https://www.bmi.gv.at/412/Nationalratswahlen/Wahlkreiseinteilung.aspx"

n_mandates_regional_districts <- MoI_url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table() %>%
  map_df(., bind_rows)

# from utils package; allows Umlaute when exporting
write.csv2(n_mandates_regional_districts,
  file = here::here("_blog_data", "2019_nrw_vorzugsstimmen", "n_mandates_regional_districts.csv"),
  row.names = F,
  fileEncoding = "latin1"
)

n_mandates_regional_districts <- read_csv2(
  file = here::here("_blog_data", "2019_nrw_vorzugsstimmen", "n_mandates_regional_districts.csv"),
  locale = locale(encoding = "latin1")
)

n_mandates_landeswahlkreis <- n_mandates_regional_districts %>%
  mutate(state = stringr::str_sub(Regionalwahlkreis, 1, 1)) %>%
  group_by(state) %>%
  summarise(n_mandates = sum(Mandate))

info_constituencies <- dplyr::bind_rows(
  n_mandates_regional_districts %>%
    select(-Stimmbezirke) %>%
    rename(
      district = Regionalwahlkreis,
      district_name = Bezeichnung,
      n_mandates = Mandate
    ),
  n_mandates_landeswahlkreis %>%
    mutate(district = case_when(
      state == "1" ~ "Burgenland",
      state == "2" ~ "Kärnten",
      state == "3" ~ "Niederösterreich",
      state == "4" ~ "Oberösterreich",
      state == "5" ~ "Salzburg",
      state == "6" ~ "Steiermark",
      state == "7" ~ "Tirol",
      state == "8" ~ "Vorarlberg",
      state == "9" ~ "Wien",
      TRUE ~ NA_character_
    )) %>%
    mutate(district_name = district)
) %>%
  mutate(district_name_short = str_replace(district_name, "Burgenland", "Bgld") %>%
    str_replace(., "Kärnten", "Ktn") %>%
    str_replace(., "Niederösterreich", "NÖ") %>%
    str_replace(., "Salzburg", "Sbg") %>%
    str_replace(., "steiermark", "Stmk") %>%
    str_replace(., "Innsbruck", "Ibk") %>%
    str_replace(., "Vorarlberg", "Vbg") %>%
    str_replace(., "\\sund\\s", " u ")) %>%
  select(., -state) %>%
  bind_rows(., data.frame(district = "Gesamt", district_name = "Gesamt", n_mandates = 183, district_name_short = "Gesamt"))

# >> get 'electoral number' --------------------------------------------------

electoral_number <- left_join(results_state_list %>%
  select(state, district_name, district, gultige),
n_mandates_landeswahlkreis,
by = c("district" = "state")
) %>%
  mutate(electoral_number = ceiling(gultige / n_mandates)) # rounded to next higher integer

results_state_list_long <- left_join(results_state_list_long,
  electoral_number %>% select(district, electoral_number),
  by = c("district" = "district")
)

# > regional level -------------------------------------------------------

results_regional_list <- nrw19 %>%
  mutate(district_name = str_remove_all(district_name, "Wahlkarten") %>%
    str_remove_all(., "-") %>%
    str_trim()) %>%
  filter(level == "Regional-Wahlkreis" & str_detect(gkz, "00$")) %>%
  mutate(district_type = "Regionalwahlkreis") %>%
  mutate(district = paste0(state, district_pol) %>% stringr::str_sub(., 1, 2)) %>%
  mutate(state = case_when(
    state == 1 ~ "Burgenland",
    state == 2 ~ "Kärnten",
    state == 3 ~ "Niederösterreich",
    state == 4 ~ "Oberösterreich",
    state == 5 ~ "Salzburg",
    state == 6 ~ "Steiermark",
    state == 7 ~ "Tirol",
    state == 8 ~ "Vorarlberg",
    state == 9 ~ "Wien",
    TRUE ~ NA_character_
  ))

results_regional_list_long <- results_regional_list %>%
  select(district_type, district_name, district, ovp:slp) %>%
  pivot_longer(cols = ovp:slp, names_to = "party", values_to = "votes_abs")

results_regional_list_long <- results_regional_list_long %>%
  left_join(., n_mandates_regional_districts %>%
    select(-Stimmbezirke, -Bezeichnung) %>%
    rename(
      district = Regionalwahlkreis,
      n_mandates = Mandate
    ),
  by = c("district")
  )

# bind results
df_results <- bind_rows(results_federal_list_long, results_state_list_long, results_regional_list_long) %>%
  filter(party %in% c("ovp", "spo", "fpo", "grune", "neos")) %>%
  mutate(party = case_when(
    party == "ovp" ~ "ÖVP",
    party == "spo" ~ "SPÖ",
    party == "grune" ~ "GRÜNE",
    party == "fpo" ~ "FPÖ",
    party == "neos" ~ "NEOS",
    TRUE ~ as.character(party)
  )) %>%
  mutate(district = case_when(
    district == 1 ~ "Burgenland",
    district == 2 ~ "Kärnten",
    district == 3 ~ "Niederösterreich",
    district == 4 ~ "Oberösterreich",
    district == 5 ~ "Salzburg",
    district == 6 ~ "Steiermark",
    district == 7 ~ "Tirol",
    district == 8 ~ "Vorarlberg",
    district == 9 ~ "Wien",
    district == "00" ~ "Gesamt",
    TRUE ~ as.character(district)
  ))

# merge results of preference votes with parties' overall results.
df_comb <- left_join(df %>% select(district_type, district, party3, name, name_short, name_orig,
  preference_votes = pref_votes_abs
),

df_results %>% rename(party_votes = votes_abs),
by = c("district_type" = "district_type", "district" = "district", "party3" = "party")
)

# preference votes vs votes for party -------------------------------------

preference_votes_constituency <- df %>%
  group_by(district_type, district, party3, state) %>%
  summarise(sum_preference_votes = sum(pref_votes_abs, na.rm = T)) %>%
  ungroup()

# federal level would need disaggregation
u <- left_join(preference_votes_constituency,
  df_results %>% select(district_type, district, district_name, party, votes_abs),
  by = c(
    "district_type" = "district_type",
    "district" = "district",
    "party3" = "party"
  )
) %>%
  mutate(p_pvv_ratio = sum_preference_votes / votes_abs) %>%  #p_pvv_ratio=party preference vote vote ratio
  # filter(!(kreis=="Bundeswahlkreis" & district!="Gesamt")) %>%
  filter(district_type == "Regionalwahlkreis") %>%
  filter(party3 != "Other")

labeller_bld <- c(
  "Burgenland" = "Bgld",
  "Kärnten" = "Ktn",
  "Niederösterreich" = "NÖ",
  "Oberösterreich" = "OÖ",
  "Salzburg" = "Sbg",
  "Steiermark" = "Stmk",
  "Tirol" = "Tirol",
  "Vorarlberg" = "Vbg",
  "Wien" = "Wien"
)

u <- dplyr::left_join(u,
  n_mandates_regional_districts %>%
    rename(
      district = Regionalwahlkreis,
      n_mandates = Mandate
    ) %>%
    select(-Bezeichnung, -Stimmbezirke),
  by = c("district")
)

# not included: Parties' preference vote share (facet grid)  -----------------------------------------------------

x <- u %>%
  ggplot() +
  geom_bar(aes(
    x = paste0(district_name, "(", district, ")"),
    y = p_pvv_ratio,
    fill = party3
  ),
  stat = "identity"
  ) +
  labs(
    y = "preference votes in relation to party votes",
    x = "regional constituency list/Regionalwahlkreis",
    title = ""
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.1)),
    position = "right"
  ) +
  scale_fill_manual(values = party_colors) +
  facet_grid(
    rows = vars(party3), cols = vars(state),
    scales = "free_x",
    space = "free",
    switch = "y"
  ) +
  hrbrthemes::theme_ipsum_rc() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0),
    legend.position = "none",
    strip.text = element_text(hjust = 1),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.spacing = unit(0.05, "cm")
  )

# name <- "Parties_preference_vote_share_NOT_INCLUDED_IN_BLOG"
# format <- ".png"
# time <- Sys.Date()
# folder <- paste0(wdr, "/graphs/")
# height <- 30
# width <- 40
# ggsave(
#   filename = paste0(folder, time, name, format),
#   device = "png",
#   type = "cairo",
#   height = height,
#   width = width,
#   scale = 1,
#   unit = c("cm")
# )
#

# not incldued: Parties' preference vote share (boxplot) ---------------------------------------------------------------

# pos <- position_jitter(width = 0.2, height = 0, seed = 1) # define seed for positioning
# 
# u %>%
#   group_by(party3) %>%
#   arrange(desc(p_pvv_ratio)) %>%
#   mutate(index = row_number()) %>%
#   mutate(label = case_when(
#     index < 4 ~ district_name,
#     TRUE ~ NA_character_
#   )) %>%
#   ungroup() %>%
#   ggplot() +
#   labs(
#     title = "Preference votes as share of total votes casted (Parties' preference votes share)",
#     caption = my_caption_2,
#     subtitle = "Only regional constitlevel",
#     y = "Preference votes in relation to total votes"
#   ) +
#   geom_boxplot(aes(
#     x = party3,
#     y = p_pvv_ratio
#   ),
#   outlier.shape = NA,
#   fill = "transparent"
#   ) +
#   geom_point(aes(
#     x = party3,
#     y = p_pvv_ratio
#   ),
#   position = pos
#   ) +
#   ggrepel::geom_text_repel(aes(
#     x = party3,
#     y = p_pvv_ratio,
#     label = label
#   ),
#   position = pos
#   ) +
#   scale_y_continuous(
#     labels = scales::percent_format(accuracy = 1),
#     limits = c(0, .75),
#     breaks = seq(0, .75, .25)
#   ) +
#   hrbrthemes::theme_ipsum_rc() +
#   theme(
#     panel.grid.major.x = element_blank(),
#     plot.caption = element_markdown(
#       color = "grey30",
#       hjust = c(0, 1)
#     ),
#     axis.title.x = element_blank()
#   )
# 
# 
# name <- "Parties_preference_vote_share_boxplot_NOT_INCLUDED_IN_BLOG"
# format <- ".png"
# time <- Sys.Date()
# folder <- paste0(wdr, "/graphs/")
# height <- 30
# width <- 40
# ggsave(
#   filename = paste0(folder, time, name, format),
#   device = "png",
#   type = "cairo",
#   height = height,
#   width = width,
#   scale = 1,
#   unit = c("cm")
# )

# INCLUDED  Parties' Preference Vote Votre (P-PVV) ratio by state  ------------------------------------------------------

pos <- position_jitter(width = 0.2, height = 0, seed = 1) # define seed for positioning

plot_region <- u %>%
  group_by(state) %>%
  arrange(desc(p_pvv_ratio), .by_group = T) %>%
  mutate(index = row_number()) %>%
  mutate(label = case_when(
    index < 4 ~ district_name,
    TRUE ~ NA_character_
  )) %>%
  ungroup() %>%
  group_by(state) %>%
  mutate(median_bundesland = median(p_pvv_ratio)) %>%
  ungroup() %>%
  mutate(state = fct_reorder(state, median_bundesland, .desc = T)) %>%
  ggplot() +
  labs(
    title = "Parties' Preference vote - vote ratio (P-PVV) by state",
    caption = my_caption_2,
    subtitle = "Only regional constituency level (Regionalwahlliste)",
    y = "# of party's preference votes as ratio of party's total votes in the constituency (P-PVV)"
  ) +
  geom_boxplot(aes(
    x = state,
    y = p_pvv_ratio
  ),
  outlier.shape = NA,
  fill = "transparent"
  ) +
  geom_point_interactive(aes(
    x = state,
    color = party3,
    y = p_pvv_ratio,
    data_id = district_name,
    tooltip = paste(
      party3,
      district_name,
      round(p_pvv_ratio * 100, 2),
      "%"
    )
  ),
  position = pos
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, .75),
    breaks = seq(0, .75, .25)
  ) +
  scale_color_manual(values = party_colors) +
  theme_post()+
  theme(
    panel.grid.major.x = element_blank(),
    legend.title = element_blank(),
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    # plot.title = element_text(size = 12),
    # plot.subtitle = element_text(size = 12, color = "grey30"),
    plot.title.position = "plot",
    # plot.caption = element_markdown(
    #   color = "grey30",
    #   hjust = c(0, 1)
    # ),
    plot.caption=element_text(hjust=0),
    # plot.margin = margin(0, unit = "cm"),
    axis.title.x = element_blank()
  )
```

```{r, include=F, eval=F}
plot_region
```

```{r, include=F, eval=T}
plot_region_interactive <- girafe(
  code = print(plot_region),
  width_svg = 10,
  height_svg = 6,
  options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_sizing(rescale = T),
    opts_tooltip(css = "background-color:lightgray;
                 font-family:Roboto Condensed;")
  ))
```

```{r, include=F, eval=T}
# INCLUDED: Party's preference vote vote ratio (P-PVPV ratio) by district magnitude -----------------------------------------------------

pos <- position_jitter(width = 0.2, height = 0, seed = 1) # define seed for positioning

constituencies_annotation <- info_constituencies %>%
  select(district, n_mandates, district_name_short) %>%
  filter(str_detect(district, "^[0-9]")) %>%
  group_by(n_mandates) %>%
  summarise(gebiete = paste0(district_name_short, collapse = ", ")) %>%
  mutate(n_mandates = as_factor(n_mandates))

plot_n_mandates <- u %>%
  mutate(n_mandates = as_factor(n_mandates) %>% fct_rev()) %>%
  group_by(n_mandates) %>%
  # arrange(desc(p_pvpv_ratio), .by_group = T) %>%
  arrange(desc(p_pvv_ratio), .by_group = T) %>%
  mutate(index = row_number()) %>%
  mutate(label = case_when(
    index < 4 ~ district_name,
    TRUE ~ NA_character_
  )) %>%
  ungroup() %>%
  ggplot() +
  labs(
    title = "Parties' preference Votes - party votes ratio (P-PVPV) and district magnitude",
    caption = my_caption_2,
    subtitle = "Only regional constituency level (Regionalwahlliste).",
    y = "# of party's preference votes as ratio of party's total votes in the constituency (P-PVPV)",
    x = "District magnitude (number of available mandates in constituency)"
  ) +
  geom_boxplot(aes(
    x = n_mandates,
    y = p_pvv_ratio
  ),
  outlier.shape = NA,
  fill = "transparent"
  ) +
  geom_point_interactive(aes(
    x = n_mandates,
    color = party3,
    y = p_pvv_ratio, #p_pvv_ratio
    data_id = party3,
    tooltip = paste0(district_name, "\n", party3, ": ", round(p_pvv_ratio * 100, 2), " %") #chagend from p_pvpv_ratio
  ),
  position = pos
  ) +
  geom_rect(
    ymin = .8, ymax = Inf,
    xmin = 0, xmax = Inf,
    fill = plot_bg_color
  ) +
  geom_text(
    data = constituencies_annotation,
    aes(
      y = .75,
      x = n_mandates,
      label = str_wrap(gebiete, width = 40)
    ),
    hjust = 0,
    family = "Roboto Condensed",
    nudge_x = 0.5,
    lineheight = 0.8,
    size = 3.5
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1),
    # expand=expansion(mult=c(0.1,0)),
    breaks = seq(0, .75, .25)
  ) +
  scale_color_manual(values = party_colors) +
  theme_post()+
  theme(
    axis.text.y = element_text(
      size = 12, 
      vjust = 0,
      hjust = 1
    ),
    axis.title.y = element_text(size=12, 
                                color="grey50",
                                hjust=1,
                                angle=90),
    axis.title.x = element_text(hjust=0,
                                color="grey50",
                                size=12),
    panel.grid.major.x = element_line(),
    panel.grid.major.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    # plot.title = element_text(size = 12, face="bold.italic", margin=margin(b=0, unit="cm")),
    plot.subtitle = element_text(size = 12, color = "grey30"),
    plot.caption=element_text(hjust=0),
    plot.margin = margin(0, unit = "cm"),
    plot.title.position = "plot") +
  coord_flip()
```

```{r, include=F, eval=T}
plot_n_mandates_interactive <- girafe(
  code = print(plot_n_mandates),
  width_svg = 10,
  height_svg = 7,
  options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_sizing(rescale = T),
    opts_tooltip(css = "background-color:plot_bg_color;	
                 font-family:Roboto Condensed;")
  ))

```

```{r, include=F}
# INCLUDED Candidate's preference vote vote ratio (C-PVPV) by list  ------------------------------------------

share <- df %>%
  filter(!party3 == "Other") %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  filter(!is.na(pref_votes_abs)) %>%
  group_by(district_type, state, district, party3) %>%
  arrange(desc(pref_votes_abs), .by_group = T) %>%
  mutate(index = row_number()) %>%
  mutate(sum_preference = sum(pref_votes_abs, na.rm = T)) %>%
  ungroup() %>%
  mutate(share_pref = pref_votes_abs / sum_preference) %>%
  # filter(share_pref>0.05) %>%  #change later
  left_join(., info_constituencies,
    by = c("district")
  )

df_thresholds <- tibble::tribble(
  ~threshold, ~district_type,
  .07, "Bundeswahlkreis",
  .10, "Landeswahlkreis",
  .14, "Regionalwahlkreis"
)

df_thresholds_2 <- crossing(df_thresholds, unique(share$party3)) %>%
  rename(party3 = `unique(share$party3)`)

df_share <- share %>%
  left_join(., df_thresholds_2,
    by = c("party3", "district_type")
  ) %>%
  mutate(share_pref = share_pref * 100) %>%
  left_join(., electoral_number %>%
    select(district_name, electoral_number),
  by = c("district" = "district_name")
  ) %>%
  mutate(electoral_number_indicator = case_when(
    pref_votes_abs > electoral_number ~ "yes",
    TRUE ~ as.character("no")
  )) %>% 
  mutate(district_name=case_when(district_name=="Gesamt" ~ "Österreich",
                                 TRUE ~ as.character(district_name))) %>% 
  left_join(., df_results %>% select(district_type, -district_name, district, party, votes_abs) %>% 
              mutate(party=as_factor(party)),
              by=c("district_type"="district_type",
               #  "district_name"="district_name",
                 "district"="district",
                 "party3"="party")) %>% 
  mutate(c_pvpv_ratio=round(pref_votes_abs/votes_abs, 4)) %>%  #c_pvpv candidates' preference vote vote ratio
  mutate(party3=as_factor(party3)) %>% 
  ungroup() %>% 
  group_by(district) %>% 
  mutate(sorting_index=cur_group_id()) #replaced group_indices

x <- df_share %>% 
  select(district, district_name_short, sorting_index) %>% 
  distinct()

plot_c_pvpv_ratio <- df_share %>%
  filter(party3 != "Other") %>%
  ungroup() %>% 
  group_split(district_type) %>%
  # group_by(district_type) %>% 
  # split() %>% 
  map(~ ggplot(.) +
    labs(
      title = paste(unique(.$district_type)),
      subtitle = paste("Candidates securing preference votes more than
                    <span style='color:orange'>", unique(.$threshold) * 100, "</span> % of their party's overall vote in their respective constituency are re-ranked."),
      x = "Candidate's preference votes in % of party votes (C-PVPV)"
    ) +
    geom_point_interactive(aes(
      y = reorder(district_name_short, sorting_index),
      x = c_pvpv_ratio*100,
      # shape = as_factor(electoral_number_indicator),
      color = party3,
      data_id = name_short,
      tooltip = paste(name_short, round(c_pvpv_ratio*100, 2), sep = "; ")
    ),
    position = position_jitter(width = 0, height = 0.2, seed = 2),
    stat = "identity"
    ) +
    geom_vline(aes(xintercept = threshold * 100),
      color = "firebrick"
    ) +
    scale_y_discrete(expand = expansion(add = 0.2)) +
    scale_x_continuous(
      trans = "log10",
      limits = c(NA, 40),
      labels = scales::label_percent(accuracy = 1, scale = 1),
      minor_breaks = NULL,
      breaks = c(NA, 0, 1, 10, 40)
    ) +
    scale_color_manual(values = party_colors) +
    ggforce::facet_row(vars(party3),
      space = "free",
      scale = "free_x",
      shrink = T
    ) +
theme_post()+
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5, size = 9),
      axis.title.y = element_blank(),
      panel.spacing.x = unit(0.2, "cm"),
      panel.spacing.y = unit(0.2, "cm"),
      strip.text.x = element_text(
        angle = 0,
        vjust = 1,
        face = "bold"
      ),
      strip.text.y = element_text(
        angle = 180,
        vjust = 1,
        face = "bold"
      ),
      strip.placement = "outside",
      # plot.title = element_text(size = 11, face = "bold.italic", 
      #                           margin=margin(b=0, unit="cm")),
      plot.subtitle = element_markdown(size = 11, 
                                       color = "grey30"),
      plot.title.position = "plot",
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.justification = "right",
      panel.grid.major.y = element_blank(),
      plot.margin = margin(0, unit = "cm")
    ) +
    guides(color = "none"))

plot_c_pvpv_ratio <- (patchwork::wrap_plots(plot_c_pvpv_ratio, ncol = 1) &
  theme(plot.margin = margin(unit = "cm", 0))) +
  plot_annotation(
    title = "Candidates' preference votes v. party votes ratio (C-PVPV)",
    subtitle = "Candidates above vertical line move up the electoral list.",
    caption = my_caption_2,
    theme = theme(
      plot.title = element_text(family = "Roboto Condensed", 
                                size = 20,
                                face = "bold"),
      plot.subtitle = element_text(family = "Roboto Condensed", 
                                   size = 12,
                                   face = "plain"),
      plot.caption=element_text(hjust=0,
                                color="grey30"),
      plot.background = element_rect(fill=plot_bg_color)))+
      plot_layout(heights = c(1, 9, 39))
```

```{r, include=F}
# > save ------------------------------------------------------------------

# name <- "C_PVPV_by_list"
# format <- ".png"
# time <- Sys.Date()
# folder <- paste0(wdr, "/graphs/")
# height <- 30
# width <- 30
# ggsave(
#   filename = paste0(folder, time, name, format),
#   device = "png",
#   type = "cairo",
#   height = height,
#   width = width,
#   scale = 1,
#   unit = c("cm")
# )

plot_c_pvpv_ratio_interactive <- girafe(
  code = print(plot_c_pvpv_ratio),
  width_svg = 10,
  height_svg = 15,
  options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_sizing(rescale = T),
    opts_tooltip(css = "background-color:plot_bg_color;	
                 font-family:Roboto Condensed;")
  )
)

```

```{r, include=F}
# x <- girafe_options(
#   x = my_plot,
#   sizingPolicy(
#     padding = "0px"
#   )
# )
# 
# saveWidget(x,
#   file = "C_PVPV_by_list.html",
#   background = "white"
# )
#
```

```{r, include=F}
# INCLUDED. Candidates crossing preference vote threshold -------------------------------------------

df_n_candidates_crossing_threshold <- df_share %>%
  mutate(leader=case_when(rank_list==1 ~ "list leader",
                          TRUE ~ as.character("not list leader"))) %>% 
  mutate(threshold_share = case_when(
    c_pvpv_ratio * 100> threshold * 100 ~ "yes",
    TRUE ~ as.character("no")
  )) %>%
  filter(threshold_share == "yes")
 # select(district_type, district_name_short, party3, c_pvpv_ratio, threshold) %>%
  # group_by(district_type, party3, threshold_share) %>%
  # summarise(n = n()) %>%

plot_n_candidates_crossing_threshold <- df_n_candidates_crossing_threshold %>% 
  ggplot() +
  labs(
    title = "Number of candidates crossing preference vote threshold",
    caption = my_caption_2
  ) +
  geom_bar(aes(
    x = party3,
    fill = leader,
    group=leader,
    ),
  stat = "count",
  position=position_stack(vjust=0.5)
  ) +
  geom_text(aes(x=party3,
                group=leader,
                label=..count..),
            color="white",
            stat="count",
            position = position_stack(vjust=0.5))+
  geom_text(aes(x=party3,
                group=party3,
                label=..count..),
            color="black",
            stat="count",
            nudge_y = 0.5,
            #position = position_stack(vjust=1)
            )+
  scale_y_continuous(breaks=seq(0, 20, 2))+
  scale_fill_paletteer_d("ggsci::default_jama")+
  facet_wrap(vars(district_type)) +
  theme_post()+
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(0, unit = "cm"),
    # plot.title = element_text(size = 12, face="bold.italic", margin=margin(b=0, unit="cm")),
    plot.caption = element_text(hjust = c(0))
  )

# name <- "n_candidates_crossing_threshold"
# format <- ".png"
# time <- Sys.Date()
# folder <- paste0(wdr, "/graphs/")
# width <- 26.49
# height <- 11.17
# ggsave(
#   plot = plot_n_candidates_crossing_threshold,
#   filename = paste0(folder, time, name, format),
#   device = "png",
#   dpi = 96,
#   type = "cairo",
#   height = height,
#   width = width,
#   scale = 1,
#   unit = c("cm")
# )

# INCLUDED Candidates crossing electoral number with preference votes (geofacet) -----------------------------------
library(geofacet)

aut_grid <- data.frame(
  row = c(1, 1, 1, 2, 2, 2, 2, 2, 3),
  col = c(3, 4, 5, 1, 2, 3, 4, 5, 4),
  code = c("Oberösterreich", "Niederösterreich", "Wien", "Vorarlberg", "Tirol", "Salzburg", "Steiermark", "Burgenland", "Kärnten"),
  name = c("Oberösterreich", "Niederösterreich", "Wien", "Vorarlberg", "Tirol", "Salzburg", "Steiermark", "Burgenland", "Kärnten")
)

state_electoral_number <- df %>%
  filter(party3 != "Other") %>%
  filter(district_type == "Landeswahlkreis") %>%
  left_join(.,
    electoral_number %>% select(district_name, electoral_number),
    by = c("district" = "district_name")
  ) %>%
  select(district_type, district, name_short, party3, electoral_number, pref_votes_abs)

state_electoral_number_plot <- state_electoral_number %>%
  ggplot() +
  labs(
    title = "Candidates' preference votes and electoral number",
    caption = my_caption_2,
    y = "Preference votes"
  ) +
  geom_point_interactive(aes(
    x = party3,
    color = party3,
    y = pref_votes_abs,
    tooltip = paste(
      party3, "\n",
      name_short, pref_votes_abs
    )
  ),
  position = position_jitter(width = 0.2, height = NULL, seed = 2)
  ) +
  geom_hline(aes(yintercept = electoral_number),
    color = "firebrick"
  ) +
  geom_text(aes(
    x = Inf, y = electoral_number,
    label = scales::comma(electoral_number)
  ),
  color = "firebrick",
  nudge_y = -3000,
  family = "Roboto Condensed",
  hjust = 1,
  check_overlap = T
  ) +
  scale_color_manual(values = party_colors) +
  scale_x_discrete(expand = expansion(mult = 0)) +
  scale_y_continuous(
    labels = scales::label_comma(),
    expand = expansion(mult = 0),
    limits = c(0, 35000),
    minor_breaks = NULL
  ) +
  theme_post()+
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid.major.x = element_blank(),
    strip.text.y = element_text(angle = 0, vjust = 1),
    strip.text.x = element_text(size=14),
    legend.position = "none",
    # plot.title = element_text(size = 12, face="bold.italic", margin=margin(b=0, unit="cm")),
    # plot.title.position = "plot",
    plot.caption = element_text(hjust=0),
    plot.margin = margin(0, r = 0.2, unit = "cm"),
    strip.background.x = element_rect(fill = plot_bg_color, color=NA),
    panel.border = element_rect(color = plot_bg_color, fill = "transparent")
  ) +
  facet_geo(~district, grid = aut_grid) +
  guides(color = "none")

# > save ------------------------------------------------------------------

state_electoral_number_plot_interactive <- girafe(
  code = print(state_electoral_number_plot),
  width_svg = 10,
  height_svg = 6,
  options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_sizing(rescale = T),
    opts_tooltip(css = "background-color:plot_bg_color;	
                 font-family:Roboto Condensed;")
  )
)
```

```{r, include=F}
# x <- girafe_options(
#   x = my_plot,
#   sizingPolicy(padding = "0px")
# )
# 
# saveWidget(x,
#   file = "state_electoral_number_plot.html",
#   background = "white"
# )
```

```{r, include=F}
# INCLUDED Concentration of preference votes (gini) ------------------------------------------------------------------

library(reldist)

# > plot --------------------------------------------------------------------

plot_gini <- df %>%
  filter(party3 != "Other") %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  left_join(., info_constituencies, by = c("district")) %>%
  filter(!is.na(pref_votes_abs)) %>%
  group_by(district_type, state, district_name_short, party3) %>%
  summarise(gini_pref = gini(pref_votes_abs)) %>%
  ggplot() +
  labs(
    title = "Concentration of preference votes",
    caption = my_caption_2,
    y = "Gini coefficient"
  ) +
  geom_point_interactive(aes(
    x = party3,
    y = gini_pref,
    color = party3,
    tooltip = paste(
      district_name_short, "\n",
      party3, round(gini_pref, 2)
    )
  ),
  pos = position_jitter(width = 0.2, height = NULL, seed = 2)
  ) +
  geom_boxplot(aes(
    x = party3,
    y = gini_pref
  ),
  size=0.5,
  color="grey30",
  fill = "transparent",
  outlier.shape = NA
  ) +
  scale_color_manual(values = party_colors) +
  scale_x_discrete(expand = expansion(mult = 0)) +
  scale_y_continuous(
    labels = scales::label_comma(),
    limits = c(0, 1),
    expand = expansion(mult = 0),
    minor_breaks = NULL
  ) +
  facet_wrap(vars(district_type)) +
  theme_post()+
  theme(
    axis.title.x = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.text.y = element_text(angle = 0, 
                                vjust = 1),
    strip.text.x = element_text(size=12),
    legend.position = "none",
    # plot.title = element_text(size = 12, face="bold.italic", margin=margin(b=0, unit="cm")),
    plot.caption = element_text(hjust=0)#,
    # plot.margin = margin(0, unit = "cm")
  ) +
  guides(color = "none")
```

```{r, include=F}
# > save --------------------------------------------------------------------

plot_gini_interactive <- girafe(
  code = print(plot_gini),
  width_svg = 10,
  height_svg = 6,
  options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_sizing(rescale = T),
    opts_tooltip(css = glue::glue("background-color:{plot_bg_color}; font-family:Roboto Condensed")))
)

```

```{r, include=F}
# INCLUDED Competitiveness  ----------------------------------------------------------

# > NOT INCLUDED: Difference as share of total preference votes -------------------------

# df %>%
#   filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
#   # filter(district=="Gesamt") %>%
#   filter(!is.na(pref_votes_abs)) %>%
#   group_by(district_type, state, district, party3) %>%
#   arrange(desc(pref_votes_abs), .by_group = T) %>%
#   mutate(index = row_number()) %>%
#   mutate(sum_preference = sum(pref_votes_abs, na.rm = T)) %>%
#   filter(index < 3) %>%
#   mutate(pref_diff = pref_votes_abs - lead(pref_votes_abs)) %>%
#   mutate(diff_perc = pref_diff / sum_preference) %>%
#   filter(!is.na(diff_perc)) %>%
#   mutate(indicator = paste0(district, "-", name_short)) %>%
#   mutate(indicator = district) %>%
#   ggplot() +
#   labs(
#     title = "Lead",
#     subtitle = "Lead in % of party's total number of preference votes"
#   ) +
#   geom_bar(aes(
#     y = indicator,
#     x = diff_perc
#   ),
#   stat = "identity"
#   ) +
#   lemon::facet_rep_grid(
#     cols = vars(party3),
#     rows = vars(district_type),
#     # repeat.tick.labels = T,
#     scales = "free",
#     drop = T,
#     space = "free"
#   ) +
#   hrbrthemes::theme_ipsum_rc() +
#   theme(axis.text.x = element_text(angle = 90))
```

```{r, include=F}
# > INCLUDED: Difference % leader vs runers-up ---------------------------------------------------

d <- df %>%
  left_join(., info_constituencies, by = c("district")) %>%
  filter(!party3 == "Other") %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  filter(!is.na(pref_votes_abs)) %>%
  group_by(district_type, state, district, party3) %>%
  arrange(desc(pref_votes_abs), .by_group = T) %>%
  mutate(index = row_number()) %>%
  mutate(sum_preference = sum(pref_votes_abs, na.rm = T)) %>%
  filter(index < 3) %>%
  mutate(share_pref = pref_votes_abs / sum_preference) %>%
  mutate(indicator = district) # %>%

d <- d %>%
  ggplot() +
  labs(
    title = "Lead",
    subtitle = "Lead in % of party's total number of preference votes",
    x = "Share of intra-party preference votes; leader and runners-up."
  ) +
  # geom_point(aes(
  #   y = district_name_short,
  #   x = share_pref,
  #   shape = as_factor(index),
  #   color=party3
  # ),
  # stat = "identity"
  # ) +
  geom_point_interactive(aes(
    y = district_name_short,
    x = share_pref,
    shape = as_factor(index),
    color = party3,
    data_id = name_short,
    tooltip = name_short,
  ),
  stat = "identity"
  ) +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1),
    minor_breaks = NULL,
    breaks = seq(0, 1, .5)
  ) +
  scale_shape_manual(
    labels = c("1" = "leader", "2" = "runners-up"),
    values = c(1, 2)
  ) +
  scale_color_manual(values = party_colors) +
  lemon::facet_rep_grid(
    cols = vars(party3),
    rows = vars(district_type),
    repeat.tick.labels = F,
    scales = "free",
    space = "free",
    drop = T
  ) +
  hrbrthemes::theme_ipsum_rc() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5),
    axis.title.y = element_blank(),
    panel.spacing.x = unit(0.1, "cm"),
    panel.spacing.y = unit(0.2, "cm"),
    strip.text.y = element_text(angle = 0, vjust = 1),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.justification = "right"
  ) +
  guides(color = "none")

# >> save ------------------------------------------------------------------

d_interactive <- girafe(
  code = print(d),
  width_svg = 8,
  height_svg = 12,
  options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_sizing(rescale = T)
  )
)

# x <- girafe_options(
#   x = my_plot,
#   sizingPolicy(
#     padding = "0px"
#   )
# )
# 
# saveWidget(x,
#   file = "pi_Difference_leading_runnersup_prefernece_votes.html",
#   background = "gray"
# )
# 
# 
# name <- "Difference_leading_runnersup_prefernece_votes"
# format <- ".png"
# time <- Sys.Date()
# folder <- paste0(wdr, "/graphs/")
# height <- 30
# width <- 30
# ggsave(
#   filename = paste0(folder, time, name, format),
#   device = "png",
#   type = "cairo",
#   height = height,
#   width = width,
#   scale = 1,
#   unit = c("cm")
# )
```

```{r, include=F}
# INCLUDED difference leader - runnersup vs n mandates -----------------------------

d1 <- df %>%
  left_join(., info_constituencies, by = c("district")) %>%
  filter(!party3 == "Other") %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  filter(!is.na(pref_votes_abs)) %>%
  group_by(district_type, state, district, party3) %>%
  arrange(desc(pref_votes_abs), .by_group = T) %>%
  mutate(index = row_number()) %>%
  mutate(sum_preference = sum(pref_votes_abs, na.rm = T)) %>%
  filter(index < 3) %>%
  mutate(share_pref = pref_votes_abs / sum_preference) %>%
  mutate(indicator = district) %>%
  select(
    district_type, state, district, party3, name_short, district_name, n_mandates, district_name_short,
    indicator, share_pref, index
  ) %>%
  group_by(district_type, state, district, party3, district_name, district_name_short, indicator) %>%
  mutate(
    runners_up = lead(name_short),
    runners_up_share = lead(share_pref)
  ) %>%
  rename(
    leader = name_short,
    leader_share = share_pref
  ) %>%
  filter(index == 1) %>%
  mutate(share_diff = leader_share - runners_up_share) %>%
  ungroup() %>%
  mutate(n_mandates = fct_reorder(as_factor(n_mandates), as.numeric(n_mandates)))

plot_difference <- d1 %>%
  ggplot() +
  labs(
    title = "Difference between top-2 candidates' preference vote share by constituency list",
    y = "Differnce between top-2 candidates' preference vote share",
    caption = my_caption_2
  ) +
  geom_point_interactive(aes(
    x = party3,
    y = share_diff,
    color = party3,
    tooltip = paste(
      district_name_short, "\n",
      leader, round(leader_share * 100, 2), "%", "\n",
      runners_up, round(runners_up_share * 100, 2), "%", "\n"
    )
  ),
  pos = position_jitter(width = 0.2, height = NULL, seed = 2)
  ) +
  geom_boxplot(aes(
    x = party3,
    y = share_diff
  ),
  fill = "transparent",
  outlier.shape = NA
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  scale_color_manual(values = party_colors) +
  facet_wrap(vars(district_type)) +
  theme_post()+
  theme(
    # plot.title = element_text(size = 12, face="bold.italic", margin=margin(b=0, unit="cm")),
    plot.title.position = "plot",
    legend.position = "none",
    plot.caption = element_text(hjust = c(0)),
    axis.title.x = element_blank(),
    plot.margin = margin(0, unit = "cm"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

# >> save ------------------------------------------------------------------

plot_difference_interactive <- girafe(
  code = print(plot_difference),
  width_svg = 10,
  height_svg = 6,
  options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_sizing(rescale = T),
    opts_tooltip(css = "background-color:plot_bg_color;	font-family:Roboto Condensed;")
  )
)

```

# Context

On 29 September, Austria held its general elections to the national assembly (*Nationalrat*). By now a flurry of analyses and comments has been published and the caravan of political commentary has largely moved on, at least when it comes to the analysis of the results itself (the [government formation](https://www.nytimes.com/reuters/2019/11/10/world/europe/10reuters-austria-politics.html){target="_blank"} train  is finally also getting up to speed). One aspect, I personally have never looked into, and which seems to fall a bit under the radar, is the use of preference votes by Austria's electorate.  This post is essentially my first 'exploratory' go at preference votes. No specific question, no specific theoretical framework in mind, just poking. I'll first present the results, then detail some of the steps implemented in R to obtain, analyze and visualize the relevant data. The entire code for the analysis is available on my github [account](https://github.com/zoowalk/NRW19_preference_votes/){target="_blank"}. As always, if you spot any glaring error etc, don't hesitate and let me know.

Preference vote (*Vorzugsstimme*) refer to voters' option (but not obligation) to indicate their support/preference for specific candidates running for the party of their choice. Since Austrian candidates run on a nation wide federal district list (*Bundeswahlkreis*), one of nine state district lists (*Landeswahlkreis*), or/and one of 39 regional district lists (*Regionalwahlkreis*), voters can cast up to three preference votes [(here](https://www.help.gv.at/Portal.Node/hlpd/public/resources/documents/NRW2017_Vorzugsstimmen.jpg){target="_blank"} a sample ballot slip). Eventually, a candidate's number of preference votes can be consequential when it comes to distributing the seats which a party won to the individual candidates. Without preference votes, seats are distributed in accordance to candidates' position on the electoral list. The lower a candidate is on a list, the lower the likelihood that she can actually get a seat. However, with a sufficient large number of preference votes, the pre-election order on the electoral list can change and initially lower ranked candidates can move up and eventually gain a seat. For details see [here](https://www.oesterreich.gv.at/themen/leben_in_oesterreich/wahlen/1/Seite.320260.html){target="_blank"}.

# Results

## by district list

Overall, there were 1,8 million preference votes cast. The largest part (more than 60 %) were cast for candidates running on regional district lists. Probably this could be read as voters been particularly keen to support 'their' local/regional candidates to make it into the new parliament. Ties between voters and regional candidates are likely to be different than those between voters and candidates running on the more 'removed' federal or state constituency list.

```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE, code_folding=NULL}
pl_by_district
```

## by constituency list and party

If we further disaggregate these numbers and differentiate between parties, we see that preference votes obtained on the regional level accounted for the majority of preference votes for all parties. However, there is a noticeable difference between the comparable low 55.9 % of the Freedom Party (FPÖ) and e.g. the 65.7 % of the Social Democrats (SPÖ) or the 64.1 % of the NEOS. Or put differently, with 34.1 % the share of preference votes obtained on the federal level, the Freedom Party stands out. As for the underlying reasons, I can only speculate. One reading could be that voters of the FPÖ have weaker ties to their candidates on the regional or state level and hence felt less inclined to cast a preference vote for them. Another one could be that the relation among FPÖ's candidates on the federal list were particularly competitive and hence a disproportionate high number of voters were mobilized to cast their preference votes on the federal list. Considering the exit of former party leader HC Strache ([Ibiza](https://www.theguardian.com/world/2019/may/20/austria-ibiza-scandal-sting-operation-what-happened-why-does-it-matter){target="_blank"}...) and the subsequent (not yet consolidated?) change in the leadership, this seems plausible (see below). It however does not explain why FPÖ voters did not cast preference votes for candidates on state or regional lists.

```{r echo=FALSE, message=FALSE, warning=FALSE, code_folding=NULL, background=plot_bg_color}
pl_constituency_list_party
```

## Top 5 candidates
If we are interested in candidates, which candidates were particularly successful in securing preference votes?

```{r echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE, code_folding=NULL}
pl_top_5_candidates

```

The graph above clearly demonstrates the predominate role of former chancellor Sebastian Kurz when it comes to preference votes. With more than 150,000 votes his result clearly outclasses those of any other candidate. With the notable exception of the FPÖ, the leading candidates/party leaders of all parties succeeded in securing most of the preference votes. In the case of the FPÖ, former Minister of Interior Herbert Kickl overtook his party's leader, Norbert Hofer. What is also noticeable is that party leaders almost exclusively secured their preference votes from the federal constituency list, while their colleagues largely relied on votes from the regional constituency list.^[Reportedly, Hofer's 'defeat' to Kickl might be due to an administrative issue though. There was another FPÖ candidate with the family name Hofer running on the federal constituency list. When voters put only the candidate's family name on the ballot to indicate their preference, it was not clear which Hofer was actually meant. As a consequence, these preference votes were reportedly ruled as invalid.]

Looking at the rankings within parties, I have to admit that I was a bit surprised by the strong performance of a some candidates which I at least had never heard of. The result of T. Mete for the SPÖ, and J Hechenberger for the ÖVP were particularly surprising to me.

## Regional dimension

### by district magnitude

The plot below displays a party's total number of preference votes as percentage of a party's total vote in each regional electoral constituency.  Let's call this for the sake of easier reference **parties' preference votes - party votes ratio**. Or to predend a serious color, **P-PVPV** ratio.^[I assume that the literature on preference votes has already coined a term for this indicator, but for the limited analysis here I leave it with this.] This PVPV ratio are grouped by constituencies' magnitude (number of mandates). The motivation behind is the assumption that the fewer mandates are available the higher will be also intra-party competition. With a smaller cake to distribute (fewer mandates), candidates will try particularly hard to win preference votes and move up the electoral list. Here the analysis is limited to the 39 regional electoral lists. To avoid the indivudal dots overlapping, some random vertical variation was introduced.

The plot seems to provide at least some backing to this proposition. The constituency with only 1 mandate to compete for (Osttirol) features the highest median of parties' PVPV ratio. While the picture is not entirely clear cut, there seems to be indeed a decreasing median share as constituencies' magnitude increases from to two to seven mandates. The constituencies with eight and nine mandates (Graz-Umgebung, Hausruckviertel, OberStmk), however, do deviate. 
Hover over the individual dots to get details.^[There seems to be an issue with the hoover over effect on mobile devices. If anyone knows how to iron this out, please let me know.]

```{r, echo=F, code_folding=NULL}
plot_n_mandates_interactive
```

### by state

If we are interested in whether there are any regional differences when it comes to parties' PVPV ratios, grouping them by states can be of some help. As the boxplots below show, on average parties' PVPV ratio was the largest in Burgenland followed by Vorarlberg. These two states are Austria's two smallest in terms of population. On the other end, Upper Autria (Oberösterreich), Lower Austria (Niedrösterreich), Styria (Steiermark), and Vienna featured on average clearly lower PVPV ratios. These states are Austria's  largest. This difference would be in line with the proposition that candidates in smaller states are better in mobilizing personal support, i.e. preference votes. Closer candidate-voter relations due to smaller population size sounds rather probable to me. But this would certainly need further analysis to corroborate the claim.

```{r, code_folding=NULL, echo=F}
plot_region_interactive
```

## Intra-party dynamics
As already outlined above, preference votes are first and foremost interesting from an intra-party perspective. With sufficient votes, candidates which embarked on the electoral campaign from a lower ranked position can move up the intra-party ladder and secure a seat.

So when do candidate's move up the ballot list?^[See [here](https://www.oesterreich.gv.at/themen/leben_in_oesterreich/wahlen/1/Seite.320260.html){target="_blank"} for details]

+ On the federal level: when a candidate got a preference votes of at least 7 % of those who voted for her party.
+ On the state level: when a candidate got preference votes of of at least 10 % of those who voted for her party, *or* at least as many preference votes as the 'electoral number' (*Wahlzahl*)^[By dividing the number of valid votes in a state level constituency by the number of the constituency's available mandates one gets the 'electoral number'].
+ On the regional level: when a candidate got preference votes by at least 14 % of those who voted for her party.

Did this happen often?

### Candidates' preference vote party vote ratio
The graph below presents **candidates' preference votes - party votes ratio**, or, to stay ~~serious~~ clumsy, **C-PVPV**. This ratio is candidates' number of preference votes expressed as % of the total number of votes of her party (in the respective constituency/list). The vertical red lines indicate the stipulated thresholds which candidates have to pass in order to move up on the party's electoral list. Hover over the individual points to get the data pertaining to each candidate. Note that the x-axis is log-transformed to ensure candidates with low C-PVPV ratios remain (somewhat) visible. Note also that only those candidates are included which actually obtained at least one preference vote.

```{r, echo=F, code_folding=NULL}
plot_c_pvpv_ratio_interactive
```

While the graph provides plenty of details, a few general things standout. First, on the federal level, only two candidates crossed the threshold. Former chancellor Kurz (ÖVP) and former Minister of Interior Kickl (FPÖ). Second, on the state level, not a single candidate secured sufficient preference votes. However, there are several candidates running on regional lists who were able to cross the pertaining threshold. Note that there is no district in which more than one candidate from one party crossed the threshold.

### Candidates' preference votes and electoral number
Aside from candidates' preference-vote party-vote ratio, candidates have one further avenue to be promoted. See point two of the conditions listed above. On the state constituency level, candidates can move up the ladder not only if they crossed the 10 % threshold, but also if they would have reached more preference votes than the electoral number. The graph below presents the pertaining figures. In short, none of the candidates succeeded in doing so in these elections. In fact, candidates are pretty far from this threshold.

```{r, code_folding=NULL, echo=F, layout="l-body-outset", fig.width=12}
state_electoral_number_plot_interactive
```

With the above, we know by now who was able to muster sufficient preference votes to cross the relevant thresholds. However, when aggregating, it makes sense to differentiate between list leaders and those of lower ranks.

```{r, code_folding=NULL, echo=F}
plot_n_candidates_crossing_threshold
```

The graph below provides the number of candidates per party who were able to cross the required preference vote threshold. On the regional list level, ÖVP, SPÖ, and FPÖ had the same number of candidates, followed by the Greens and Neos. What the graph also reveals is that in the predominant part of cases, the successful candidates were in fact parties' federal/regional list leaders. In total, only seven candidates which were not list leaders succeeded in securing the necessary preference votes. Probably noteworthy, 4 of these 7 cases originate from the FPÖ. Does this mean anything? I am tempted to suspect that parties with clear hiearchial structures are less likely to see lower rank candidates overtake their party's list leaders. In cases, though, in which e.g. the leadership question is not entirely settled, or leaders are confronted with very strong 'colleagues' (Feind, Todfeind... Parteifreund), successful preference vote campaigns are more likely to occur. At least that's an interpretation which came to my mind when seeing the numbers pertaining to the FPÖ (did I hear 'selecting on the dependent variable'?). On the other hand, this proposition is immediately relativized when checking who these candidates were. With Kunasek (former Minister of Defense, leader of Styria's FPÖ) and Haimbuchner (leader of Upper Austria's FPÖ), we have two FPÖ representatives from the party's upper echelons, who put their names on the ballot list, but - as far as I am aware - see their future as for now (again) in their home states rather than in the national parlament. I am only speculating here, but I found it at least noteworthy that the FPÖ somewhat deviates.

### Difference Leader and Runners-up

Another angle to look at preference votes could be to examine the difference between the two candidates leading in terms of the number of preference votes. This gives us some idea on how dominant a candidate has been, i.e. whether the intra-party contestation at the top was particularly high. The graph above seeks to display the difference in candidates' **preference vote share** per party and electoral district. Preference vote share is a candidate's share of her party's total preference votes in the respective list/constituency. Here, the higher the difference between those, the lower the contestation for the top spot. Hover over the dots to see the two top candidate and their respective share of preference votes. Not the small gap between SPÖ's leader Rendi-Wagner and Mete.

```{r, code_folding=NULL, echo=F}
plot_difference_interactive
```

### Concentration of preference votes

Last but not least, to round off the picture, let's see how different parties are when it comes to the concentration of preference votes. Are a party's preference votes going almost exclusively to one or only a few candidates (high concentration)? Or, are preference votes relatively evenly shared among different candidates (low concentration)? A measure of concentration (Gini coefficient) can provide us with another angle on intra-party dynamics. Below the results:

```{r, code_folding=NULL, echo=F}
plot_gini_interactive
```

# Steps in R
Below some of the most important steps in R to obtain the results above. The full code is published on my github account [here](https://github.com/werkstattcodes/NRW19_preference_votes){target="_blank"}

## Obtaining the preference vote results
The first potential hurdle is to virtually 'liberate' the results pertaining to the preference vote. Just have a look a the format of the data [here](https://www.bmi.gv.at/412/Nationalratswahlen/Nationalratswahl_2019/files/NRW19_Vorzugsstimmen_Landes_Regionalparteiliste_16102019.pdf){target="_blank"} and [here](https://www.bmi.gv.at/412/Nationalratswahlen/Nationalratswahl_2019/files/NRW19_Vorzugsstimmen_bundesweit_16102019.pdf){target="_blank"}. Particularly the results for the state and regional electoral district lists are not what a follower of '[tidy data](https://vita.had.co.nz/papers/tidy-data.pdf){target="_blank"}' would hope for. So let's start wrangling:

```{r, ref.label="extract_data", eval=F, code_folding="Code: Extract data", include=T}
```

Et voila (here only the first entries). If needed, you can downlaod the cleaned file [here](https://github.com/werkstattcodes/NRW19_preference_votes/master/data/AUT_NRW19_preference_votes.csv){target="_blank"} github repo.

```{r echo=FALSE, message=FALSE, warning=TRUE}
df_merge %>% 
  reactable(.,
          filterable = T,
          pagination = T,
          height = 250,
          outlined = T,
          bordered = F,
          compact=T,
          theme=reactableTheme(backgroundColor = plot_bg_color,
                           style=list(fontFamily="Karla",
                                      fontSize="12px"),
                           filterInputStyle = list(
                             color="black",
                             backgroundColor = plot_bg_color)))

```

If needed, you can download the cleaned file [here](https://github.com/zoowalk/NRW19_preference_votes/master/data/AUT_NRW19_preference_votes.csv){target="_blank"} from the github repo.

## Abbreviating names, removing titles
An issue which has as been a frequent source of annoyance is the verbose use of academic and other (obviously even more important) titles in names. It only unnecessarily takes space from graphs and has something, in my personal view, deeply inegalitarian to it (the very purpose of using titles is probably to signal inegalitrianism). So, let's remove all titles and only use the first letter of the first name. [`Regex`](https://stringr.tidyverse.org/articles/regular-expressions.html){target="_blank"} and [`stringr`](https://stringr.tidyverse.org/index.html){target="_blank"} are your friends here. Since this is a task which comes up time and again, it actually screams for a r package...

```{r, eval=F}
fn_remove_titles <- function(x) {
  titles <- c(
    "Mag", "MMag", "Prof", "Dr", "O\\.Univ\\.Prof", "Dipl-Ing", "Dipl.-Ing.", "Dipl.-Rev.", "Ing",
    "Dipl", "jun", "sen", "Ba", "\\(FH\\)", "MSc",
    "med", "univ", "Bsc", "phil", "iur", "DDr", "BEd", "DI", "Mba", "Mas", "MA", "MIM",
    "vet", "M.A.", "PMM", "MTD", "Päd", "Bakk"
  ) %>%
    map(., paste0, c("\\.", "\\s", "\\,")) %>%
    unlist()

  name_wo_titles <- stringr::str_remove_all(x, regex("[[:alpha:]\\-]+\\.")) %>%
    str_remove_all(., regex("\\([:alpha:]+\\)")) %>%
    str_remove_all(., regex(paste0(titles, collapse = "|"), ignore_case = T)) %>%
    str_squish()
}

fn_short_name <- function(x) {
  family_name <- stringr::str_extract(x, "^[[:alpha:]\\-|\\s]+(?=,)") %>%
    str_to_title() %>%
    str_squish() %>%
    str_trim()

  first_name <- str_extract(x, regex("(?<=\\,)[[:alpha:]\\-\\s]*$")) %>%
    str_trim() %>%
    str_squish() %>%
    str_to_title()

  initals <- str_split(first_name, "\\s") %>%
    map(., str_sub, start = 1L, end = 1L) %>%
    map_chr(., str_c, collapse = ".") %>%
    paste0(., ".")

  name_short <- paste(family_name, initals, sep = ", ")
}

df <- df_merge %>%
  mutate(district = str_remove(district, "Landesparteiliste") %>%
    str_remove(., "Regionalparteiliste") %>%
    str_trim(., side = c("both"))) %>%
  mutate(district = case_when(
    district == "B" ~ "Burgenland",
    district == "S" ~ "Salzburg",
    district == "K" ~ "Kärnten",
    district == "NÖ" ~ "Niederösterreich",
    district == "OÖ" ~ "Oberösterreich",
    district == "Stmk" ~ "Steiermark",
    district == "Vlbg" ~ "Vorarlberg",
    TRUE ~ district
  )) %>%
  mutate(district = as_factor(district)) %>%
  mutate(party2 = str_extract(party, regex("\\([:alpha:]*\\)")) %>% # take parties' short abbrevs.
    str_remove_all(., "\\(|\\)")) %>%
  mutate(name_orig = name) %>%
  mutate(name = map_chr(name_orig, fn_remove_titles)) %>%
  mutate(name_short = map_chr(name, fn_short_name)) %>%
  mutate(state = case_when(
    str_detect(district, "^[0-9].*") ~ str_sub(district, start = 1L, end = 1L),
    TRUE ~ as.character(district)
  )) %>%
  mutate(state = case_when(
    state == 1 ~ "Burgenland",
    state == 2 ~ "Kärnten",
    state == 3 ~ "Niederösterreich",
    state == 4 ~ "Oberösterreich",
    state == 5 ~ "Salzburg",
    state == 6 ~ "Steiermark",
    state == 7 ~ "Tirol",
    state == 8 ~ "Vorarlberg",
    state == 9 ~ "Wien",
    TRUE ~ NA_character_
  )) %>%
  mutate(district_type = fct_relevel(district_type, "Bundeswahlkreis", "Landeswahlkreis", "Regionalwahlkreis"))

```
With the above we are basically ready to roll and start analyzing the data in earnest. The code below gives you the first plot (preference votes by district lists).

## Plot - Total number of preference votes per electoral district

```{r, eval=F, code_folding="Code: Plot total number of preference votes per electoral district"}

df_list <- df %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  group_by(district_type) %>%
  summarise(district_type_votes = sum(pref_votes_abs, na.rm = T)) %>%
  #  arrange(desc(district_type_votes)) %>%
  mutate(district_type = forcats::fct_relevel(district_type, "Regionalwahlkreis", "Landeswahlkreis", "Bundeswahlkreis")) %>%
  mutate(district_type_votes_cum = cumsum(district_type_votes)) %>%
  mutate(district_type_votes_perc = district_type_votes / sum(district_type_votes)) %>%
  mutate(district_type = forcats::fct_reorder(district_type, district_type_votes_cum, .desc = F))

pl_by_district <- df_list %>%
  ggplot() +
  geom_bar(aes(
    x = 1,
    y = district_type_votes,
    fill = district_type,
    group = district_type
  ),
  stat = "identity"
  ) +
  geom_text(aes(
    x = 1,
    y = district_type_votes,
    group = district_type,
    label = paste(scales::comma(district_type_votes),
      scales::percent(district_type_votes_perc),
      sep = "\n"
    )
  ),
  colour = "white",
  position = position_stack(vjust = .5, reverse = F)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    breaks = max(df_list$district_type_votes_cum),
    label = scales::comma(max(df_list$district_type_votes_cum))
  ) +
  scale_fill_paletteer_d("ggsci::default_jama",
    labels = c(
      "Regionalwahlkreis" = "Regional lists",
      "Landeswahlkreis" = "State lists",
      "Bundeswahlkreis" = "Federal list"
    )
  ) +
  labs(
    title = "Total number of preference votes per electoral district",
    caption = my_caption_2,
    y = "number of preference votes"
  ) +

  # hrbrthemes::theme_ipsum_tw() +
  # theme(
  #   plot.background = element_rect(fill=)
  #   legend.position = "top",
  #   legend.justification = "left",
  #   plot.title.position = "plot",
  #   legend.title = element_blank(),
  #   axis.text.y = element_blank(),
  #   axis.title.y = element_blank(),
  #   plot.title = element_text(size = 13),
  #   panel.grid = element_blank(),
  #   plot.caption = element_markdown(hjust = c(0, 1), color = "grey30"),
  #   plot.margin = margin(t = 0, l = 0, b = 0, r = 1, unit = "cm")
  # ) +
  coord_flip() +
  guides(fill = guide_legend(reverse = T))+
  theme_post()+
  theme(legend.position="top",
        legend.justification = "left",
        legend.direction  = "horizontal",
        plot.caption =element_text(hjust=0),
        # plot.background=element_rect(fill = "plot_bg_color", color="plot_bg_color"),
        legend.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank()

        )

```

```{r code_folding=NULL, echo=F, fig.height=3, fig.width=10, dev="png"}
pl_by_district
```

## Group_split, map, and ggplot

Another part of the code which I'd like to highlight is the one producing the [*plot on candidates' preference votes share*](http://werk.statt.codes/post/vorzugsstimmen/#candidates-preference-votes-share). Here, rather than using a `facet_grid`, I split the dataframe by the electoral district lists, and mapped the plot function to each of the three parts, and subsequently combined them with `patchwork`. The advantage was that graph is more efficient in using the available space.

```{r, eval=F, code_folding="plot on candidates’ preference votes share"}

share <- df %>%
  filter(!party2 == "Other") %>%
  filter(!(district_type == "Bundeswahlkreis" & district != "Gesamt")) %>%
  filter(!is.na(pref_votes_abs)) %>%
  group_by(district_type, state, district, party2) %>%
  arrange(desc(pref_votes_abs), .by_group = T) %>%
  mutate(index = row_number()) %>%
  mutate(sum_preference = sum(pref_votes_abs, na.rm = T)) %>%
  ungroup() %>%
  mutate(share_pref = pref_votes_abs / sum_preference) %>%
  # filter(share_pref>0.05) %>%  #change later
  left_join(., info_constituencies,
    by = c("district")
  )

df_thresholds <- tibble::tribble(
  ~threshold, ~district_type,
  .07, "Bundeswahlkreis",
  .10, "Landeswahlkreis",
  .14, "Regionalwahlkreis"
)

df_thresholds_2 <- crossing(df_thresholds, unique(share$party3)) %>%
  rename(party3 = `unique(share$party3)`)

df_share <- share %>%
  left_join(., df_thresholds_2,
    by = c("party3", "district_type")
  ) %>%
  mutate(share_pref = share_pref * 100) %>%
  left_join(., electoral_number %>%
    select(district_name, electoral_number),
  by = c("district" = "district_name")
  ) %>%
  mutate(electoral_number_indicator = case_when(
    pref_votes_abs > electoral_number ~ "yes",
    TRUE ~ as.character("no")
  )) %>% 
  mutate(district_name=case_when(district_name=="Gesamt" ~ "Österreich",
                                 TRUE ~ as.character(district_name))) %>% 
  left_join(., df_results %>% select(district_type, -district_name, district, party, votes_abs) %>% 
              mutate(party=as_factor(party)),
              by=c("district_type"="district_type",
               #  "district_name"="district_name",
                 "district"="district",
                 "party3"="party")) %>% 
  mutate(c_pvpv_ratio=round(pref_votes_abs/votes_abs, 4)) %>%  #c_pvpv candidates' preference vote vote ratio
  mutate(party3=as_factor(party3)) %>% 
  ungroup() %>% 
  mutate(sorting_index=group_indices(., district))

x <- df_share %>% 
  select(district, district_name_short, sorting_index) %>% 
  distinct()

plot_c_pvpv_ratio <- df_share %>%
  filter(party3 != "Other") %>%
  group_split(district_type) %>%
  map(~ ggplot(.) +
    labs(
      title = paste(unique(.$district_type)),
      subtitle = paste("Candidates securing preference votes more than
                    <span style='color:orange'>", unique(.$threshold) * 100, "</span> % of their party's overall vote in their respective constituency are re-ranked."),
      x = "Candidate's preference votes in % of party votes (C-PVPV)"
    ) +
    geom_point_interactive(aes(
      y = reorder(district_name_short, sorting_index),
      x = c_pvpv_ratio*100,
      # shape = as_factor(electoral_number_indicator),
      color = party3,
      data_id = name_short,
      tooltip = paste(name_short, round(c_pvpv_ratio*100, 2), sep = "; ")
    ),
    position = position_jitter(width = 0, height = 0.2, seed = 2),
    stat = "identity"
    ) +
    geom_vline(aes(xintercept = threshold * 100),
      color = "firebrick"
    ) +
    scale_y_discrete(expand = expansion(add = 0.2)) +
    scale_x_continuous(
      trans = "log10",
      limits = c(NA, 40),
      labels = scales::label_percent(accuracy = 1, scale = 1),
      minor_breaks = NULL,
      breaks = c(NA, 0, 1, 10, 40)
    ) +
    scale_color_manual(values = party_colors) +
    ggforce::facet_row(vars(party3),
      space = "free",
      scale = "free_x",
      shrink = T
    ) +
theme_post()+
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5, size = 9),
      axis.title.y = element_blank(),
      panel.spacing.x = unit(0.2, "cm"),
      panel.spacing.y = unit(0.2, "cm"),
      strip.text.x = element_text(
        angle = 0,
        vjust = 1,
        face = "bold"
      ),
      strip.text.y = element_text(
        angle = 180,
        vjust = 1,
        face = "bold"
      ),
      strip.placement = "outside",
      plot.title = element_text(size = 11, face = "bold.italic", 
                                margin=margin(b=0, unit="cm")),
      plot.subtitle = element_markdown(size = 11, color = "grey30",
                                       face="italic"),
      plot.title.position = "plot",
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.justification = "right",
      panel.grid.major.y = element_blank(),
      plot.margin = margin(0, unit = "cm")
    ) +
    guides(color = "none"))

plot_c_pvpv_ratio <- (patchwork::wrap_plots(plot_c_pvpv_ratio, ncol = 1) &
  theme(plot.margin = margin(unit = "cm", 0))) +
  plot_annotation(
    title = "Candidates' preference votes v. party votes ratio (C-PVPV)",
    subtitle = "Candidates above vertical line move up the electoral list.",
    caption = my_caption,
    theme = theme(
      plot.title = element_text(family = "Roboto Condensed", size = 12,face = "bold"),
      plot.subtitle = element_text(family = "Roboto Condensed", size = 12,face = "plain")))+
      plot_layout(heights = c(1, 9, 39))

```
## Obtaining district magnitude

A convenient way to get data on the districts' magnitude (number of mandates) is to scrap the pertainnig data from the website of the Ministry of Interior. The `rvest`package´ is our reliable tool here.

```{r code_folding="Scrap data from site of MoI", eval=F}
library(rvest, quietly = T, verbose = F, warn=F)

MoI_url <- "https://www.bmi.gv.at/412/Nationalratswahlen/Wahlkreiseinteilung.aspx"

n_mandates_regional_districts <- MoI_url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table() %>%
  map_df(., bind_rows)
```

```{r message=FALSE, warning=FALSE, paged.print=TRUE, code_folding=NULL, echo=F}
library(reactable)
n_mandates_regional_districts %>% 
  reactable(.,
            filterable = T,
          pagination = T,
          height = 250,
          outlined = T,
          bordered = F,
          compact=T,
          theme=reactableTheme(backgroundColor = plot_bg_color,
                           style=list(fontFamily="Karla",
                                      fontSize="12px"),
                           filterInputStyle = list(
                             color="black",
                             backgroundColor = plot_bg_color)))

```

## Geo_facet
Finally, as the last item, I'd like to highlight the wonderful [`geo_fact` package](https://hafen.github.io/geofacet){target="_blank"}. Rather than plotting data to a map by means of different color (shades), the `geo_facet` package allows facets of a plot to be arranged (more or less) in accordance with the map of e.g. a country. Here, I first created a grid which approximates the position the Austria's states (*Bundesländer*), i.e. state electoral districts. This grid is subsequently passed on to the faceting function used in an otherwise ordinary ggplot command. The result you saw above [here](http://werk.statt.codes/post/vorzugsstimmen/#candidates-preference-votes-and-electoral-number). Sweet, isn't ;-)

Gladly, the author of the `geo_facet` package offers users the possibility to [submit](https://hafen.github.io/grid-designer/){target="_blank"} new grids to be integrated into the package. Once I have polished the grid (e.g. add iso codes), I'll submit it and you can use it by simply calling the grid's name in the geo_facet function (update: the grid [was submitted](https://github.com/hafen/geofacet/issues/226){target="_blank"} and now part of the package).

```{r, eval=F, code_folding="Code geo_facet grid Austria"}
library(geofacet)

aut_grid <- data.frame(
  row = c(1, 1, 1, 2, 2, 2, 2, 2, 3),
  col = c(3, 4, 5, 1, 2, 3, 4, 5, 4),
  code = c("Oberösterreich", "Niederösterreich", "Wien", "Vorarlberg", "Tirol", "Salzburg", "Steiermark", "Burgenland", "Kärnten"),
  name = c("Oberösterreich", "Niederösterreich", "Wien", "Vorarlberg", "Tirol", "Salzburg", "Steiermark", "Burgenland", "Kärnten")
)

state_electoral_number <- df %>%
  filter(party3 != "Other") %>%
  filter(district_type == "Landeswahlkreis") %>%
  left_join(.,
    electoral_number %>% select(district_name, electoral_number),
    by = c("district" = "district_name")
  ) %>%
  select(district_type, district, name_short, party3, electoral_number, pref_votes_abs)

state_electoral_number_plot <- state_electoral_number %>%
  ggplot() +
  labs(
    title = "Candidates' preference votes and electoral number",
    caption = my_caption_2,
    y = "Preference votes"
  ) +
  geom_point_interactive(aes(
    x = party3,
    color = party3,
    y = pref_votes_abs,
    tooltip = paste(
      party3, "\n",
      name_short, pref_votes_abs
    )
  ),
  position = position_jitter(width = 0.2, height = NULL, seed = 2)
  ) +
  geom_hline(aes(yintercept = electoral_number),
    color = "firebrick"
  ) +
  geom_text(aes(
    x = Inf, y = electoral_number,
    label = scales::comma(electoral_number)
  ),
  color = "firebrick",
  nudge_y = -3000,
  family = "Roboto Condensed",
  hjust = 1,
  check_overlap = T
  ) +
  scale_color_manual(values = party_colors) +
  scale_x_discrete(expand = expansion(mult = 0)) +
  scale_y_continuous(
    labels = scales::label_comma(),
    expand = expansion(mult = 0),
    limits = c(0, 35000),
    minor_breaks = NULL
  ) +
  hrbrthemes::theme_ipsum_rc() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid.major.x = element_blank(),
    strip.text.y = element_text(angle = 0, vjust = 1),
    legend.position = "none",
    plot.title = element_text(size = 12, face="bold.italic", margin=margin(b=0, unit="cm")),
    plot.title.position = "plot",
    plot.caption = element_markdown(hjust = c(0, 1), color = "grey30"),
    plot.margin = margin(0, r = 0.2, unit = "cm"),
    strip.background.x = element_rect(fill = "lightgrey"),
    panel.border = element_rect(color = "lightgrey", fill = "transparent")
  ) +
  facet_geo(~district, grid = aut_grid) +
  guides(color = "none")
```

### 2021-01-22 _post/2019-10-24-women-in-parliament/women-in-parliament.Rmd
revised blog women in parliament

---
title: "Women in parliament"
description: |
    How well are women represented in Austria's new parliament. A comparative look at the female share of MPs.

date: 10-24-2019

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693

categories:
  - Austria
  - women

#preview: austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
    code_folding: true

creative_commons: CC BY-NC-SA

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(tidyverse)
library(ggrepel)
```

# Context
So, Austria has a new parliament. Following the elections in September, the new lower house got constituted yesterday and as a few noted it is more diverse than previous ones. [Here](https://www.derstandard.at/story/2000110261600/fuehlen-sie-sich-im-parlament-vertreten?ref=rec)'s a nice visualization if you want to get a quick overview. Among others, it was the share of women in parliament which was particularly often noted. None of Austria's previous lower houses had an as high share of women at the start of a new legislative period. This made me a bit curious how Austria's lower house actually compares to other countries. First I'll show the results, then how to get there in R. The entire code is available at my github [account](https://github.com/zoowalk/women_in_parliaments).

# Result

```{r fig.height=20, echo=F, code_folding=NULL}
knitr::include_graphics("women-in-parliament_files/2019-10-24Women_in_Parliament_combined.png")
```

So, how is Austria doing? Well, not bad if compared with most other countries. However, given that there are only four (!) countries in which women are at least represented on par with men (min 50 %) it's overall quite a sobering affair. Don't get me started on Papua New Guinea...

# How to get there

## Getting the data
Data comes from the Inter-Parliamentary Union, based in Geneva, which seeks to strengthen international cooperation between parliaments and support their work.

```{r, eval=F}
df <- readr::read_csv(
  file = "https://data.ipu.org/api/women-ranking.csv?load-entity-refs=taxonomy_term%2Cfield_collection_item&max-depth=2&langcode=en&month=9&year=2019",
  skip = 5
)

names(df) <- c(
  "rank", "country",
  "lower_elections", "lower_seats", "lower_women_abs", "lower_women_rel",
  "upper_elections", "upper_seats", "upper_women_abs", "upper_women_rel"
)

df <- df %>%
  mutate_at(vars(contains("lower"), contains("upper")), .funs = list(num = ~ as.numeric(.))) %>%
  mutate(country = stringr::str_remove_all(country, regex("\\(.*\\)") %>% str_trim()))
```

## Dotplot: Connecting dots with labels

As for the dotplot, when it comes to annotating data `ggrepel` is your friend.

Creating the dot plot essentially entailed two critical points. First, the plotting of the dots. To avoid over-plotting, the position argument of `geom_point` introduces a random variation. Note that the seed of the `position` argument is set in advance with the  `position_jitter` command. This is critical since the same seed has to be used when subsequently placing the labels and connecting lines. If you don't use the same seeds, dots and labels would be positioned with different random values and consequently not match.

The second part, which took me a while to figure out, is the creation of an additional variable/label (`label2`). This label is empty for those countries which we do not want to label. While 'normally' one could use a different dataframe for the annotation, this avenue seems not to be possible since the dependency on the same (random) position would get messed up. Hence, this approach does not create filtered additional dataset, but uses the entire set but only provides labels for those countries which we want to highlight.

More details can be found in the `ggrepel` vignette and at this pertaining [issue](https://github.com/slowkow/ggrepel/issues/123) filed on github.

```{r, eval=F}
pos <- position_jitter(width = 0.2, seed = 1) #define seed for positioning

df %>%
  mutate(label2 = paste(country, scales::percent(lower_women_rel_num / 100,
    accuracy = .1))) %>%
  mutate(label2 = case_when(     #create label only for thos countries you want to highlight
    indicator == "other" ~ "",
    TRUE ~ as.character(label2))) %>%
  ggplot(., aes(
    x = .5,
    y = lower_women_rel_num,
    color = indicator,
    label = label2,
    group = country)) +
  geom_point(
    position = pos  #same pos
  ) +
  geom_rect(
    xmin = 0.75,
    xmax = Inf,
    ymin = -Inf,
    ymax = Inf,
    color = "white",
    fill = "white"
  ) +
  geom_text_repel(
    xlim = c(.75, 1.2),
    position = pos,  #same pos
    fontface = "italic",
    color = "grey30",
    segment.color = "grey30",
    hjust = 0
  ) +
  labs(
    title = "Women in Parliaments",
    subtitle = "Share of women in lower houses; top and bottom 3 \ncountries plus Austria, Bosnia, and USA.",
    caption = c("Data: Inter-Parliamentary Union (data.ipu.org) as of end of 2018, Austria as of <br>23 Oct 2019.", "<br><br><br>Roland Schmidt | @zoowalk | <span style='color:black'>**werk.statt.codes**</span>")
  ) +
  scale_y_continuous(
    limits = c(-5, 100),
    breaks = seq(0, 100, 25),
    minor_breaks = NULL,
    expand = expansion(mult = c(0.1, 0)),
    labels = scales::percent_format(scale = 1, accuracy = 1)
  ) +
  scale_x_continuous(
    expand = expansion(mult = 0),
    limits = c(0.15, 1.2)
  ) +
  scale_color_manual(values = c(
    "Austria" = "red",
    "other" = "grey30",
    "top" = "green",
    "bottom" = "steelblue",
    "select" = "orange"
  )) +
  hrbrthemes::theme_ipsum_rc() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.subtitle = element_text(color = "grey30"),
    legend.position = "none",
    plot.title.position = "plot",
    plot.caption.position = "plot",
    plot.caption = element_markdown(color = "grey30", hjust = c(0, 1)),
    axis.title = element_blank(),
    axis.text.x = element_blank())
```

## Barplot

As for the barplot, I'll only light to highlight the `cut` function. It's simple as powerful when it comes to cutting continuous variables into factors. Note the` right=F` argument to open intervals on the left side.

```{r, eval=F}
df_interval <- df %>% 
  mutate(women_interval=cut(x=lower_women_rel_num, 
                            breaks=seq(0,100,10),
                            right=F)) %>% 
  mutate(women_interval=forcats::fct_explicit_na(women_interval, na_level=">70")) %>% 
  mutate(indicator = case_when(
    str_detect(country, "Austria") ~ "Austria",
    TRUE ~ "other"
  ))

plot_distribution <- df_interval %>% 
  ggplot()+
  geom_bar(aes(x=women_interval,
               fill=indicator),
           stat="count") +
  geom_text(aes(x=women_interval, 
                label=..count..),   
            stat="count",
            nudge_y=5)+
  labs(title="Number of parliaments per share of female MPs",
       subtitle=paste("Only lower houses.","Position of <span style='color:red'>Austria</span> indicated in red."),
       caption = c("Data: Inter-Parliamentary Union (data.ipu.org) as of end of 2018, Austria as of <br>23 Oct 2019.", "<br><br><br>Roland Schmidt | @zoowalk | <span style='color:black'>**werk.statt.codes**</span>"),
       x="% of female MPs",
       y="Number of lower houses")+
  scale_y_continuous(expand=expansion(mult=c(0.01,0.1)))+
  scale_fill_manual(values=c("Austria"="red", "other"="steelblue"))+
  hrbrthemes::theme_ipsum_rc()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.caption.position = "plot",
        plot.subtitle = element_markdown(color= "grey30"),
        plot.caption = element_markdown(color = "grey30", hjust = c(0, 1)),
        )

```

### 2021-02-14 _post/2020-02-07-bih-voting-behavior-of-mps-in-house-of-representatives/bih-voting-behavior-of-mps-in-house-of-representatives.Rmd
main update

@@
 
-bgr_color<- readr::read_file(file=here::here("theme.css")) %>% 
-  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()
+plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
+  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
+  str_trim() %>% 
+  str_extract(., regex("^#\\S+"))
 
@@
         plot.title.position="plot",
-        plot.background= element_rect(fill=bgr_color, color=bgr_color),
+        plot.background= element_rect(fill=plot_bg_color, color=plot_bg_color),
         plot.subtitle = element_text(size = 15, color = "grey30"),
@@
   theme(legend.position = "bottom",
-        plot.background=element_rect(fill=bgr_color, color=bgr_color),
+        plot.background=element_rect(fill=plot_bg_color, color=plot_bg_color),
         plot.margin=margin(b=0, t=0, l=0, r=0, unit="cm"),
@@
         plot.title.position = "plot",
-        plot.background= element_rect(fill=bgr_color, color=bgr_color),
+        plot.background= element_rect(fill=plot_bg_color, color=plot_bg_color),
         plot.margin=margin(l=0, r=0, b=0, unit="cm"),
@@
         strip.text = element_text(face="bold"),
-        panel.background = element_rect(fill=bgr_color, color=bgr_color),
+        panel.background = element_rect(fill=plot_bg_color, color=plot_bg_color),
         panel.grid.major = element_blank(),
@@
         plot.caption = element_text(hjust=c(0, 1)),
-        plot.background= element_rect(fill=bgr_color, color=bgr_color),
+        plot.background= element_rect(fill=plot_bg_color, color=plot_bg_color),
         legend.position = "top",
@@
         plot.subtitle = element_text(size = 12, color = "grey30"),
-        plot.background=element_rect(fill=bgr_color, color=bgr_color),
+        plot.background=element_rect(fill=plot_bg_color, color=plot_bg_color),
         axis.title.y = element_blank(),
@@
         plot.margin=margin(l=0, t=0.5, b=0.5, r=0, unit="cm"),
-        plot.background=element_rect(fill=bgr_color, color=bgr_color),
+        plot.background=element_rect(fill=plot_bg_color, color=plot_bg_color),
         panel.grid.major.y =  element_blank(),
@@
               panel.spacing.x = unit(0.3, "cm"),
-              plot.background=element_rect(fill=bgr_color, color=bgr_color),
+              plot.background=element_rect(fill=plot_bg_color, color=plot_bg_color),
               panel.spacing.y = unit(0.3, "cm"))+

### 2021-02-14 _post/2020-03-02-primary-schools-in-vienna/primary-schools-in-vienna.Rmd
main update

@@
     plot.margin = margin(l = 0, t = 0.5, b = 0.5, unit = "cm"),
-    plot.background = element_rect(fill = bgr_color, color = "transparent"),
+    plot.background = element_rect(fill = plot_bg_color, color = "transparent"),
     plot.title = element_text(size = 14, face = "bold", margin = margin(b = 0, unit = "cm")),
@@
     yaxis = list(title = "absolute difference of AHS ratios (% points)"),
-    plot_bgcolor = bgr_color,
-    paper_bgcolor = bgr_color
+    plot_bgcolor = plot_bg_color,
+    paper_bgcolor = plot_bg_color
   )
@@
     showlegend = FALSE,
-    paper_bgcolor = bgr_color
+    paper_bgcolor = plot_bg_color
   ) %>% 

### 2021-02-14 _post/2020-04-28-google-s-mobility-reportsmd/google-s-mobility-reportsmd.Rmd
main update

@@
 plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
-  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()
+  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
+  str_trim() %>% 
+  str_extract(., regex("^#\\S+"))
 

### 2021-02-14 _post/2020-10-10-vienna-elections-2020-age-profile/vienna-elections-2020-age-profile.Rmd
correct intro segment to vienna elections/age post
main update

@@
 ```
+</details>
 

### 2021-02-14 _post/2021-01-22-covidconsultationduration/covidconsultationduration.Rmd
main update

---
title: "Austria's Covid legislation: Duration of parliamentarian consultation period in comparision."
description: |
  A (relative) deep dive into the length of the public consultation period for draft bills in Austria, and whether three days for three pages are actually 'normal'.
date: 01-22-2021
draft: false
author:

  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - Corona
  - pdftools
#preview: 2020-10-10-vienna-elections-2020-age-profile/preview.png
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
    code_folding: TRUE
creative_commons: CC BY-NC-SA
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(xml2)
library(rvest)
library(ggtext)
library(paletteer)
library(ggiraph)
library(reactable)
library(svglite)
knitr::opts_chunk$set(echo = T)

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
  str_trim() %>% 
  str_extract(., regex("^#\\S+"))

caption <- "Data: https://www.parlament.gv.at; XIV (4.11.1975) to XXVII (as of 24.01.2021) legislative period.\nAnalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes"

theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      #panel.border = element_rect(colour = plot_bg_color, fill=NA),
      #plot.border = element_rect(colour = plot_bg_color, fill=NA),
      plot.margin = margin(l = 0, 
                           t = 0.25,
                           unit = "cm"),
      plot.title = element_markdown(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 13
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 11
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(
        angle = 0,
        color = "grey50",
        hjust = 1
      ),
      axis.text.x = element_text(
        size = 9,
        color = "grey50"
      ),
      axis.title.y = element_blank(),
      axis.text.y = element_text(
        size = 9,
        color = "grey50"
      ),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 9
      ),
      legend.text = element_text(
        size = 9,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

rt_theme <- reactableTheme(backgroundColor = plot_bg_color,
                           style=list(fontFamily="Roboto Condensed",
                                      fontSize="12px"),
                           tableBodyStyle=list(lineHeight="100%"),
                           filterInputStyle = list(
                             color="black",
                             backgroundColor = plot_bg_color))

```

# Context
This is an addition to my previous posts on COVID related legislation in Austria.

Over New Year (!), the Austrian government introduced another bill with the aim to get a grip on the COVID pandemic. With the speed with which things have been developping and the flurry of hard/soft/de facto lockdowns, I have to confess that I would have to look into the text to recall the changes introduced by bill 88/ME (in full: *'Epidemiegesetz, COVID-19-Maßnahmengesetz, Änderung'*). However, what made the bill remarkable even during these times was its legislative genesis, with the government limiting the public consultative process to only three days. As mentioned in previous posts, Austria's legislative procedure includes a public consultation process which allows citizens, NGOs, churches etc. to file submissions in which they can raise concerns and provide feedback on the draft law. At least in theory, it's a feedback loop which allows the government to solicit input and revise its bills.

Unsurprisingly, at least to me, the government was harshly criticized for this short consultation period and accused of rendering any meaningful consideration impossible. I am no expert on the legislative process, but a quick Google search led me to a [circular](https://www.bundeskanzleramt.gv.at/dam/jcr:32bd35f3-2815-42ab-b276-be8aaba18124/RS_600614Begutachtungsfristen.doc){target="_blank"} from 2008 in which the Chancellery's own constitutional service urged ministries to provide a consultative period of at least six weeks. In contrast, some pointed out that the bill had only three pages and that - in light of the urgency due to a pandemic - three days should suffice to read and comment on the bill.

This debate made me wonder how long other bills were open for the consultative process and how the bill 88/ME would compare, also when considering their documents' lengths. To answer these questions, I a) extracted from the parliament's website the start and end dates of each bill's consultative period to calculate the length of the consultative process; b) downloaded the text of all pertaining bills and retrieved their number of pages. Putting the length of the consultation process and the length of the bill together should allow to better evaluate the three day notice for the amendment to the epidemic bill.

I'll first present the results and other stuff I found noteworthy, and then focus on some of the required steps in R to obtain them. The raw dataset containing the consolidated results is [here](#datatable).

```{r chunk_get_bills_overview, eval=F, code_folding=NULL, include=F}
fn_get_bill_overview <- function(rss_feed){

  df_rss <- xml2::read_xml(rss_feed)
  data <- xml_child(df_rss, 1)

  df_bill_pages_link <- data %>% 
    xml2::xml_find_all("//link") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="link_single_bill_page") %>% 
    mutate(link_single_bill_page=str_squish(link_single_bill_page)) %>% 
    filter(id>2) #removes first two rows which don't include data on bills

  #get title
  df_bill_pages_name <- data %>% 
    xml2::xml_find_all("//title") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="title") %>% 
    # mutate(name=str_extract(title, regex("(?<=\\>).*(?=\\<)"))) %>% 
    mutate(title=str_trim(title)) %>% 
    filter(id>2)

  #get publication date
  df_bill_pages_pub_date <- data %>% 
    xml2::xml_find_all("//pubDate") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="date") %>% 
    mutate(date=date %>% str_squish() %>% lubridate::dmy_hms(., tz="Europe/Vienna"))

  #combine to one dataframe
  df_bill <- bind_cols(
    df_bill_pages_name,
    df_bill_pages_pub_date,
    df_bill_pages_link,
  ) %>% 
    select(-contains("id"))

  df_bill %>% 
    mutate(indicator_me_numeric=str_extract(link_single_bill_page, regex("(?<=_)\\d*"))) %>% 
    mutate(indicator_me=glue::glue("{as.numeric(indicator_me_numeric)}/ME")) %>% 
    mutate(rss_link_submissions=glue::glue("https://www.parlament.gv.at/PAKT/VHG/XXVII/ME/ME_{indicator_me_numeric}/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP=XXVII&ITYP=ME&INR={as.numeric(indicator_me_numeric)}&SUCH=&listeId=142&FBEZ=FP_142
"))

}

# create vec wtih rss links -----------------------------------------------

periods <- as.roman(seq(14, 27, 1)) #freaking nice
rss_links <- glue::glue("https://www.parlament.gv.at/PAKT/MESN/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP={periods}&MESN=ME&R_MESN=ME&MIN=ALLE&SUCH=&listeId=102&FBEZ=FP_002")

df_bills_all <- rss_links %>% 
  purrr::set_names() %>% 
  map_dfr(., fn_get_bill_overview, .id="rss_links")

# readr::write_csv2(df_bills_all, file=here::here("_blog_data",
#                                           "2021_covid_consultation_duration",
#                                           "df_bills_all.csv"))

```

```{r chunk_fn_consult, eval=F, code_folding=NULL, include=F}
#function to extract details (duration and pdf links) from bills' overview
fn_consult <- function(link_bill)  {

  df_consultation_dates <- link_bill %>% 
    read_html() %>% 
    html_nodes("#content > div.contentBlock.tabs-responsive__contentBlock > div:nth-child(2) > table") %>% 
    html_table() %>% 
    map_df(., bind_rows) %>% 
    janitor::clean_names() %>% 
    filter(str_detect(stand_des_parlamentarischen_verfahrens, regex("Begutachtungsfrist"))) %>% 
    rename(date_start=datum) %>% 
    mutate(date_end=str_extract(stand_des_parlamentarischen_verfahrens, regex("\\d{2}\\.\\d{2}\\.\\d{4}"))) %>% 
    select(-protokoll)

  # get link to text of bill - define function ---------------------------------------------------------

df_link_text <-   link_bill %>% 
    read_html() %>% 
    html_nodes(xpath="//a[contains(text(), 'Ministerialentwurf') 
                and contains(text(), 'PDF') 
                or contains(text(), 'Gesetzestext')
                or contains(text(), 'Entwurf')]") %>%  #works! case sensetive
    html_attr("href")  %>% 
   tibble::enframe(name=NULL, value="link_bill_pdf") %>% 
   mutate(link_bill_pdf=paste0("https://www.parlament.gv.at/", link_bill_pdf))

  bind_cols(df_consultation_dates, df_link_text) %>% 
    select(text_raw="stand_des_parlamentarischen_verfahrens", date_start, date_end,
           link_bill_pdf)

}

#APPLY FUNCTION
tbl_missing <- tibble(link_single_bill_page=NA_character_, text_raw=NA_character_, date_start=NA, date_end=NA, link_bill_pdf=NA_character_)

library(furrr)
plan(multisession, workers = 2)

df_res <- df_bills_all$link_single_bill_page %>% 
  purrr::set_names() %>% 
  future_map_dfr(., possibly(fn_consult, otherwise=tbl_missing), .id="link_single_bill_page",
                 .progress=T) %>% 
  select(link_single_bill_page, text_raw, date_start, date_end, link_bill_pdf)

df_res <- df_res %>% 
  mutate(across(.cols=contains("date"), lubridate::dmy)) %>% 
  mutate(legis_period=str_extract(link_single_bill_page, regex("[^/]*?(?=/ME/)"))) %>% 
  mutate(bill_id=str_extract(link_single_bill_page, regex("ME_.*?(?=/)")))

df_res <- df_res %>% 
  left_join(., df_bills_all %>% 
  select(link_single_bill_page, title))

# readr::write_csv2(df_res, file=here::here("_blog_data",
#                                           "2021_covid_consultation_duration",
#                                           "df_res.csv"))

```

```{r chunk_duration, eval=T, code_folding=NULL, include=F}
df_res <-  readr::read_csv2(file=here::here("_blog_data", "2021_covid_consultation_duration",
                                           "df_res.csv")) %>% 
    mutate(across(.cols=contains("date"), lubridate::ymd))

#remove dupes due to multiple pdf documents
df_duration <- df_res %>% 
  distinct(link_single_bill_page, 
           text_raw, 
           date_start, date_end, 
           legis_period, bill_id, title)

df_duration_mult <- df_duration %>% 
  group_by(link_single_bill_page, legis_period, bill_id, title) %>% 
  mutate(n_obs_duration=n()) %>% 
  ungroup()

df_dates_missing <- df_duration_mult %>% 
  filter(if_any(starts_with("date"), is.na)) #new w dplyr 1.04
nrow(df_dates_missing) #23

df_duration_collected <- df_duration_mult %>% 
  filter(!is.na(date_start) & !is.na(date_end)) %>%  #remove those with missing dates
  group_by(link_single_bill_page, legis_period, bill_id, title) %>% 
  summarize(date_start_min=min(date_start, na.rm = T),
         date_end_max=max(date_end, na.rm = T),
         n_obs_duration=mean(n_obs_duration),
         dates_collected=paste(paste(as.character(date_start), "to", 
                              as.character(date_end)),
         collapse="; ")) %>% 
         # dates_collected=list(paste(as.character(date_start), "to", 
         #                      as.character(date_end)))) %>% 
  mutate(duration_collected=date_end_max-date_start_min+1) %>% 
  ungroup()

table(df_duration_collected$duration_collected<0)
#109 with neg duration
table(df_duration_collected %>% 
        filter(duration_collected<0) %>% 
        pull(n_obs_duration)) #all negatives have only one table row on duration

#check those negative duration
df_negative <- df_duration_collected %>% 
  filter(duration_collected<0)
#randomly checked source; appears to be errors in data collection/source;
#remove these from further analysis

table(df_duration_collected$n_obs_duration)
#3 obs - check:OK
# https://www.parlament.gv.at/PAKT/VHG/XXIV/ME/ME_00294/index.shtml

#2 obs - check
#https://www.parlament.gv.at/PAKT/VHG/XXIII/ME/ME_00139/index.shtml - ok
#https://www.parlament.gv.at/PAKT/VHG/XX/ME/ME_00361/index.shtml - ok
#https://www.parlament.gv.at/PAKT/VHG/XX/ME/ME_00153/index.shtml - ok

```

```{r BILL_LENGTH, eval=F, code_folding=NULL, include=F}
df_pages <- df_res %>% 
  select(link_single_bill_page, link_bill_pdf, 
         legis_period, bill_id, title) %>% 
  distinct() %>% #remove duplicates
  filter(link_single_bill_page %in% df_duration_collected$link_single_bill_page) %>% #only for those where we have valid durations
  mutate(file_name=str_extract(link_bill_pdf, regex("[^\\/]*pdf$"))) %>% 
  mutate(file_name=glue::glue("{legis_period}_{bill_id}_{file_name}")) %>% 
  mutate(download_destination=glue::glue("{here::here('_blog_data',
                                         '2021_covid_consultation_duration',
                                         'bills_pdf')}/{file_name}") %>%
           as.character() %>% 
           str_trim(., side=c("both")))

#download laws
safe_download <- safely(~ download.file(.x , .y, mode = "wb", quiet=T))

library(magrittr)
df_pages  %$%
  purrr::walk2(
    link_bill_pdf,
    download_destination,
    safe_download, 
    mode = "wb")

```

```{r get_pages, eval=F, code_folding=NULL, include=F}
# get page number ---------------------------------------------------------

library(pdftools)

df_pages <- df_pages %>% 
  filter(link_single_bill_page %in% df_duration_collected$link_single_bill_page) %>% 
  mutate(pdf_details=map(download_destination, possibly(pdftools::pdf_info, otherwise = NA_integer_))) %>% 
  mutate(pdf_pages=map_dbl(pdf_details, purrr::pluck, "pages", .default=NA)) %>% #note default=NA, not NULL
  select(-pdf_details)

# readr::write_excel_csv2(df_pages, file=here::here("_blog_data", "2021_covid_consultation_duration", "df_pages.csv"))

df_pages <- readr::read_csv2(here::here("_blog_data", "2021_covid_consultation_duration", "df_pages.csv"))

df_pages_collected <- df_pages %>% 
  distinct() %>% #double check to remove duplicates
  group_by(link_single_bill_page) %>% 
  summarise(pdf_pages_sum=sum(pdf_pages, na.rm = T),
            n_obs_docs=n())

```

```{r COMBINE_DURATION_AND_BILL_LENGTH, eval=F, code_folding=NULL, include=F}
df_consolidated <- df_duration_collected %>% 
  left_join(., df_pages_collected)

# nrow(df_duration_collected)
# nrow(df_pages_collected)
# nrow(df_consolidated)

# readr::write_excel_csv2(df_consolidated, 
#                         file=here::here("_blog_data", "2021_covid_consultation_duration", "df_consolidated.csv"))

```

```{r message=FALSE, warning=FALSE, include=FALSE, code_folding=NULL}
df_consolidated <- readr::read_csv2(file=here::here("_blog_data", "2021_covid_consultation_duration", "df_consolidated.csv"))

#create link to rss submission feeds for all bills
df_consolidated <- df_consolidated %>% 
  mutate(rss_submissions=glue::glue('https://www.parlament.gv.at/PAKT/VHG/{legis_period}/ME/{bill_id}/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP={legis_period}&ITYP=ME&INR={str_extract(bill_id, regex("[:digit:]*$")) %>% as.numeric()}&SUCH=&listeId=142&FBEZ=FP_142'))

```

```{r GET_SUBMISSIONS, include=FALSE, eval=FALSE, code_folding=NULL}
# get dates of all submissions
# check whether they fall within the calculated dates of the consultation period as  provided on the bill's overview page

#function to extract details of submissions
fn_submission_details <- function(link) {

submission_rss_feed <- xml2::read_xml(link)

#get link to subpages with link to submissions
df_submission_link <- submission_rss_feed %>% 
  xml2::xml_find_all("//link") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="link_single_submission_page") %>% 
  mutate(link_single_submission_page=str_squish(link_single_submission_page)) %>% 
  filter(id>2)

#get title
df_submission_title <- submission_rss_feed %>% 
  xml2::xml_find_all("//title") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="title") %>% 
  mutate(name=str_extract(title, regex("(?<=\\>).*(?=\\<)"))) %>% 
  filter(id>2) %>% 
  select(-title)

#get publication date
df_submission_date <- submission_rss_feed %>% 
  xml2::xml_find_all("//pubDate") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="date") %>% 
  mutate(date=date %>% str_squish() %>% lubridate::dmy_hms(., tz="Europe/Vienna"))

#combine to one dataframe
df_submissions <- bind_cols(
  df_submission_link,
  df_submission_title,
  df_submission_date,
  ) %>% 
  select(-contains("id"))

df_submissions

}

df_missing <- tibble(link_single_submission_page=NA_character_,
                     name=NA_character_,
                     date=NA)

df_submissions_pages <- df_consolidated$rss_submissions %>% 
  purrr::set_names() %>% 
  map_dfr(., possibly(fn_submission_details, otherwise=df_missing), .id="rss_submissions")

#fills with details to all (!) submissions over entire period
# readr::write_excel_csv2(df_submissions_pages, file=here::here("_blog_data", "2021_covid_consultation_duration", "df_submissions_pages.csv"))
```

```{r GET_SUBMISSIONS_2, include=FALSE, eval=T, code_folding=NULL}
df_submissions_pages <- readr::read_csv2(file=here::here("_blog_data",
                                                         "2021_covid_consultation_duration", "df_submissions_pages.csv"))

#number of submissions per bill
df_submissions_per_bill <- df_submissions_pages %>% 
  group_by(rss_submissions) %>% 
  summarise(n_obs_submissions=n())

df_consolidated <- df_consolidated %>% 
  left_join(.,
            df_submissions_per_bill) %>% 
  mutate(covid_bills=case_when(legis_period=="XXVII" &
                                 str_detect(title, "Epidemie") ~ "covid-19",
                               TRUE ~ as.character("other"))) %>% 
  mutate(covid_measures_bill=case_when(bill_id=="ME_00088" & legis_period=="XXVII" ~ "COVID-19-Maßnahmengesetz",
TRUE ~ as.character("other")))

```

```{r chunk_pl_submissions_per_bill, include=FALSE, code_folding=NULL}
pl_submissions_per_bill<- df_consolidated %>% 
  mutate(bill_indicator=case_when(str_detect(title, "Epidemie") & legis_period=="XXVII" ~ "yes",
                                  TRUE ~ as.character("no"))) %>% 
  ungroup() %>% 
  ggplot()+
  labs(title="LEGISLATIVE CONSULTATION PROCESS:<br>Number of filed submissions per bill.",
       subtitle="Note log scale of y-axis.",
       x="Date",
       y="Number of submissions (log scale)",
       caption=caption)+
  geom_segment(aes(x=as.Date("2017-01-01"),
               xend=as.Date("2017-01-01"),
               y=0,
               yend=4000),
               color="grey50")+
  geom_text(label=str_wrap("Introduction of possibility to file electronic submissions via parliament's website",30),
            aes(x=as.Date("2016-06-01"),
            y=6000),
            color="grey50",
            lineheight=0.7,
            check_overlap = T,
            hjust=1,
            vjust=1,
            size=3,
            family="Roboto condensed")+
  geom_jitter_interactive(aes(x=date_end_max,
               y=n_obs_submissions,
               tooltip=glue::glue("{str_wrap(title, 40)}
                                  Number of submissions: {n_obs_submissions}
                                  Bill ID: {bill_id}/{legis_period},

                                  Click to open submission page"),
               onclick=paste0('window.open("', link_single_bill_page, '#tab-Stellungnahmen', '")'),
               color=bill_indicator))+
    geom_text(label=str_wrap("Epidemiegesetz, COVID-19-Maßnahmengesetz, Änderung (88/ME)
", 40),
aes(x=as.Date("2021-01-01"),
    y=19000),
color="firebrick",
lineheight=0.7,
check_overlap = T,
hjust=1,
size=3,
family="Roboto Condensed")+
  scale_x_date(breaks=c(seq.Date(as.Date("1990-01-01"), as.Date("2020-01-01"), by="10 year"), as.Date("2017-01-01"),
                        as.Date(min(df_consolidated$date_end_max))),
               date_labels = "%Y" )+
  scale_y_log10(labels=scales::label_comma(accuracy=1),
                limits=c(1, 20000))+
  scale_color_manual(values=c("yes"="firebrick",
                              "no"="grey50"),
                     labels=c("yes"="Covid-19 related bills",
                              "no"="other"))+
  guides(color=guide_legend(reverse=T))+
  theme_post()+
  theme(plot.title.position = "panel",
        legend.position = "top",
        legend.direction = "horizontal",
        legend.justification = "left",
        legend.title = element_blank(),
        axis.title.x = element_text(hjust=0,
                                    color="grey30"),
        axis.title.y=element_text(hjust=0, 
                                  color="grey30",
                                  angle=90))

pl_submissions_per_bill <- girafe(ggobj = pl_submissions_per_bill,
       height_svg = 5,
       options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_tooltip(css = glue::glue("background-color:{plot_bg_color}; 
                 line-height:100%;
                 color:black;
                 font-size:80%;
                 font-family:'Roboto Condensed';)"))
  ))

```

```{r chunk_reactable, code_folding=NULL, include=F}
tb_consolidated <- df_consolidated %>% 
  filter(duration_collected>=0) %>% 
  mutate(link_submissions=paste0(link_single_bill_page, "#tab-Stellungnahmen")) %>% 
  select(legis_period, 
         title, 
         bill_id, 
         pdf_pages_sum, 
         date_start_min, 
         date_end_max, 
         duration_collected, 
         n_obs_submissions, 
         link_single_bill_page,
         link_submissions) %>%   
  reactable(.,
            # columnGroups = list(
            #   colGroup(name="consultation",
            #             columns=c("date_start_min",
            #                       "date_end_max",
            #                       "duration_collected"))
            # ),
            columns=list(
              legis_period=colDef(name="legilsative period",
                                  width=70),
              title=colDef(name="bill's name",
                           minWidth=160),
              bill_id=colDef(name="bill id", 
                             align="right",
                             minWidth=50),
              pdf_pages_sum=colDef(name="# pages total",
                                   minWidth=50),
              date_start_min=colDef(show=F,
                                    name="start",
                                    align="right",
                                    ),
              date_end_max=colDef(show=F, 
                                  name="end",
                                  align="right"),
              duration_collected=colDef(name="duration",
                                        align="right",
                                        minWidth=50),
              n_obs_submissions=colDef(name="# submissions",
                                       minWidth=70),
              link_single_bill_page=colDef(
                name="link bill",
                show=T,
                filterable = F,
                width=50, 
                align="right",
                cell=function(value){
                  htmltools::tags$a(href=value,
                                    target="_blank",
                                    "Link")}),
              link_submissions=colDef(name="link submissions",
                                     show=T,
                                     filterable = F,
                                     width=100,
                                     align="right",
                                     cell = function(value) {
                                       htmltools::tags$a(href=value, target = "_blank", paste0("Link"))})),
            defaultSorted = list(n_obs_submissions = "desc"),
            sortable = T,
            compact = T,
            filterable = T,
            outlined = TRUE,
            theme=rt_theme,
            details=function(index) {
              df_consolidated_pos <- df_consolidated %>% filter(duration_collected>=0)
              df_duration_details <- df_duration[df_duration$link_single_bill_page == df_consolidated_pos$link_single_bill_page[index],]
                df_duration_details %>% 
                select(starts_with("date")) %>% 
                reactable(., 
                          columns = list(
                            date_start=colDef(show=T, 
                                              name="consultation start"),
                            date_end=colDef(show=T,
                                            name="consultation end")),
                          fullWidth = F,
                          theme=reactableTheme(style=list(
                            margin="0px 0px 0px 50px",
                            backgroundColor=plot_bg_color)))
              })
```

```{r chunk_duration_legis_period, echo=F, code_folding=NULL, include=F}
pl_duration_legis_period<- df_consolidated %>%
  filter(duration_collected>=0) %>% 
  mutate(legis_period_fac=as.roman(legis_period) %>% as.integer %>% as.character() %>%  fct_inseq()) %>% 
  ungroup() %>% 
  ggplot()+
  labs(title="Distribution of length of consultative periods.",
       subtitle="",
       x="legislative period",
       y="duration in days",
       caption=caption) +
  geom_violin(aes(x=legis_period_fac,
                  y=duration_collected),
              fill="grey50",
              color="grey50"
              )+
  geom_hline(yintercept=6*7,
             linetype="dotted",
             color="grey50")+
  geom_text(label="requested six\nweek duration\n(42 days)",
            y=6*7+5,
            x=length(unique(df_consolidated$legis_period))+1,
            color="grey50",
            vjust=0,
            hjust=0,
            size=3,
            lineheight=0.7,
            family="Roboto Condensed",
            check_overlap = T)+
  geom_point_interactive(data=. %>% 
                           group_by(legis_period_fac) %>% summarise(duration_collected_median=median(duration_collected, na.rm = T)),
                         aes(
                           x=legis_period_fac,
                           y=duration_collected_median,
                           group=legis_period_fac,
                           tooltip=glue::glue("Median duration of consultative period: {duration_collected_median}")
                         ),
                         color="firebrick")+
  scale_color_manual(label=c("firebrick"="median length of consultative period"),
                     values=c("firebrick"="firebrick"))+
  scale_x_discrete(labels = function(x) as.roman(x),
                   expand=expansion(mult=c(0.05,.3)))+
  scale_y_continuous(breaks=c(seq(0, 400, 100), 42))+
  coord_cartesian(ylim=c(0, 400))+  #set limits here and not in scale_y to zoom in
  theme_post()+
  theme(plot.title.position = "panel",
        legend.position = "top",
        legend.direction = "horizontal",
        legend.justification = "left",
        legend.title = element_blank(),
        axis.title.x = element_text(hjust=0,
                                    color="grey50"),
        axis.title.y=element_text(hjust=0, 
                                  color="grey50",
                                  angle=90))

pl_duration_legis_period <- girafe(ggobj = pl_duration_legis_period,
       height_svg = 4,
       options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_tooltip(css = glue::glue("background-color:{plot_bg_color}; 
                 line-height:100%;
                 color:black;
                 font-size:80%;
                 font-family:'Roboto Condensed';)"))  
    ))

```

```{r submissions_max_min, eval=F, code_folding=NULL, include=F}
#check whether all submissions were indeed filed within the 
#extracted consultation period;

df_submissoins_max_min <- df_submissions_pages %>% 
  group_by(rss_submissions) %>% 
  summarise(sub_date_min=as.Date(min(date)),
            sub_date_max=as.Date(max(date)))

df_consolidated2 <- df_consolidated %>% 
  left_join(., df_submissoins_max_min) %>% 
  mutate(period_check=case_when(
    sub_date_min >= date_start & sub_date_max <= date_end ~ "within",
    TRUE ~ as.character("outside"))) %>% 
  select(link_single_bill_page, contains("date"), period_check)

table(df_consolidated2$period_check)

```

```{r chunk_pl_length_duration, echo=F, include=F, code_folding=NULL}
caption_note <- "Note: Each dot represents a specific bill of x pages and y days available for filing a submission as part of the public consultation process. Dots within the same rectangle have the same page - duration combination. Random variation was added to avoid overplotting of dots. Only bills with max 10 pages and max 20 days consultation period are shown. See blog post for details."

pl_length_duration <- df_consolidated %>% 
  filter(duration_collected<21) %>% 
  filter(duration_collected>-1) %>% 
  filter(pdf_pages_sum<11) %>% 
  mutate(pdf_pages_sum_fct=forcats::fct_inseq(as.character(pdf_pages_sum))) %>% 
  mutate(duration_collected_fct=fct_inseq(as.character(duration_collected)) %>% fct_rev()) %>% 
  ggplot()+
  labs(title="Duration of consultation process and length of considered bills:",
       subtitle = "",
       caption=glue::glue("{str_wrap(caption_note, 125)}

                          {caption}"),
    x="Number of bill's pages",
       y="Number of days to file a submission")+
  geom_jitter_interactive(aes
                          (x=pdf_pages_sum_fct,
                            y=duration_collected_fct,
                            color=covid_measures_bill,
                            tooltip=glue::glue("Name of bill: {stringr::str_trunc(title, width=25, side=c('right'))},
                              Legislation period: {legis_period}, Bill ID: {bill_id}
                              pages: {pdf_pages_sum}
                              consultation duration: {duration_collected}
                              link: <a href={link_single_bill_page}>Click to open</a>"),
                              onclick=paste0('window.open("', link_single_bill_page , '")'))) +
  scale_color_manual(values=c("COVID-19-Maßnahmengesetz"="orange", "other"="grey50"),
                     labels=c("COVID-19-Maßnahmengesetz"="Epidemiegesetz ME88/XXVII",
                              "other"="other"))+
  facet_grid(duration_collected_fct ~ pdf_pages_sum_fct,
             drop=F,
             switch="both",
             scales="free")+
  theme_post()+
  theme(panel.spacing.x = unit(0, "cm"),
        panel.spacing.y = unit(0, "cm"),
        panel.grid.major.x =  element_blank(),
        panel.grid.major.y =  element_blank(),
        strip.text.y.left = element_text(angle=0,
                                         vjust=0.5,
                                         hjust=1,
                                         color="grey30",
                                         face="plain"), #.left needed
        strip.text.x.bottom = element_text(angle=0,
                                           hjust=0.5,
                                           color="grey30",
                                           face="plain"), #.left needed
        legend.position = "top",
        legend.direction = "horizontal",
        legend.justification = "left",
        legend.title = element_blank(),
        axis.title.x = element_text(hjust=0,
                                    color="grey30"),
        axis.title.y=element_text(hjust=0, 
                                  color="grey30",
                                  angle=90),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, size = .2))

pl_length_duration <- girafe(ggobj = pl_length_duration,
       height_svg = 5,
       options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_tooltip(css = glue::glue("background-color:{plot_bg_color}; 
                 line-height:100%;
                 color:black;
                 font-size:80%;
                 font-family:'Roboto Condensed';)"))
  ))

```

```{r chunk_3days, code_folding=NULL, include=F}
pl_consolidated <- df_consolidated %>% 
  filter(duration_collected>=0) %>% #remove those with negative duration
  count(duration_collected<=3) %>% 
  janitor::clean_names() %>%
  mutate(duration_3=as_factor(duration_collected_3)) %>% 
  ggplot()+
  labs(title="Number of bills with three or fewer days of consultation period",
       subtitle="Bill 88/ME XXVII amending the 'law on epidemics' was open for consultation for three days. \nHow many other bills had a similar or shorter consultation period?",
       y="Number of bills",
       x="Length of consultation period",
       caption=caption)+
  geom_bar(aes(x=duration_3,
               y=n,
               fill=duration_3),
           stat="identity",
           position=position_dodge(width=0.9))+
  geom_text(aes(x=as.numeric(duration_3)-(.5*.9),
                y=n,
                label=n),
            # position=position_dodge(width=0.9),
            size=10,
            color="grey30",
            vjust=0,
            hjust=0)+
  scale_x_discrete(labels=c("FALSE"="more than 3 days",
                            "TRUE"="3 days or fewer"),
                   expand = expansion(mult=c(0,0.1)))+
  scale_y_continuous(labels=scales::label_comma(),
                     limits=c(0,NA),
                     expand=expansion(mult=c(0,0.1)))+
  scale_fill_manual(values=c("FALSE"="grey30",
                    "TRUE"="orange"))+
  theme_post()+
  theme(legend.position = "none",
        axis.text.x = element_text(color="grey10"),
        axis.title.y = element_text(color="grey30",
                                    angle=90,
                                    vjust=1,
                                    hjust=0),
        axis.title.x = element_blank())
```

# Results

## Number of submissions

```{r ref.label=c("chunk_pl_submissions_per_bill"), code_folding="Code plot", eval=F}

```

To start with, let's have a look at the number of submissions filled during the consultation process per bill. The plot below  shows quite clearly that bills related to Covid triggered record breaking numbers of submissions. Apart from the three latest Covid bills, only the two drafts related to security issues (*'Sicherheitspolizeigesetz'*) with around 9,000 submissions were within a similar level. To get details, hover over the dots. Note that the y-axis is log-scaled to keep the variation among bills with fewer submissions visible.

```{r fig.height=5, echo=F, code_folding=NULL}
pl_submissions_per_bill
```
<aside>
Unrelated but still interesting, bills related to the educational sector triggered also quite high numbers of submissions. Prior to the Covid-related bills, these bills, and police bills, came out on top. 
</aside>

## Duration of consultation process and length of bills
But what about the initial question? How unusual is it that a bill of three pages length, i.e. bill 88/ME, is open for consultation for only three days? The plot below essentially addresses the question put forward in the introduction. As the orange dot in the graph below highlights, there has never been another bill with three pages and as little as three days of consultation. The next bills in the same page-length category had eight days (among them a bill amending the law on gambling/*Glückspielgesetz*...). So at least from this perspective, our bill in question has been quite an aberration.

However, if we broaden our view a bit, and also consider bills with fewer or more than three pages, we see that there were several other bills for which consultation periods shorter than three days were granted.

```{r ref.label="chunk_pl_length_duration", code_folding="Code plot", eval=F}
```

```{r fig.height=5, code_folding=NULL, echo=F}
pl_length_duration
```

Nevertheless, out of the `r nrow(df_consolidated)` bills open for consultation, `r df_consolidated %>% filter(duration_collected>3) %>% nrow()` featured longer consultation periods than three days. There were only `r df_consolidated %>% filter(duration_collected>=0) %>% filter(duration_collected<=3) %>% nrow()` bills with three or fewer days (excluding results with negative duration or incomplete start/end dates, see [here](#wrongdata)).

```{r ref.label=c("chunk_3days"), code_folding="Code plot", eval=F}

```

```{r, fig.height=5, echo=F, code_folding=NULL}
pl_consolidated
```

## Overview table {#datatable}
If you are interested in the details pertaining to each bill, the table below provides them.

```{r ref.label="chunk_reactable", code_folding=NULL, eval=F, echo=F}

```

```{r, layout="l-body-outset", code_folding=NULL, echo=F}
tb_consolidated
```

## Duration of consultation process per leigslative  period
With the data available I started to wonder whether there are actually any considerable differences between different legislative periods/governments and the lengths of their bills' consultation processes. To further contrast the lengths, I added the requested minimum duration of six weeks as raised in the 2007 circular of the Chancellery's constituitonal service.

The plot provides below shows the distribution of consultation periods' length (in days) per legislative period. The horizontal line indicates the recommended length of six weeks.

```{r ref.label="chunk_duration_legis_period", code_folding="Code plot", eval=F}

```

```{r, code_folding=NULL, echo=F}
pl_duration_legis_period
```

# Analysis in R

In the following section I'll highlight the most important steps how to retrieve the results from above.

## Getting the duration of the consultation period
### Get links to all bills
To get the required data I start with retrieving data on each bill's title, its date and a link eventually leading us to each bill's details on the different steps in the legislative process. To get these data, I make use of the rss-feed provided on the parliament's website (see [here](https://www.parlament.gv.at/PAKT/MESN/index.shtml){target="_blank"} to extract the data.

```{r code_folding=NULL, echo=F}
knitr::include_graphics("covidconsultationduration_files/images/rss_bills_overview.png")
```

First, I define the function `fn_get_bill_overview`.

```{r chunk_get_bills_overview_copy, eval=F, code_folding=NULL, include=T}
fn_get_bill_overview <- function(rss_feed){

  df_rss <- xml2::read_xml(rss_feed)
  data <- xml_child(df_rss, 1)

  df_bill_pages_link <- data %>% 
    xml2::xml_find_all("//link") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="link_single_bill_page") %>% 
    mutate(link_single_bill_page=str_squish(link_single_bill_page)) %>% 
    filter(id>2) #removes first two rows which don't include data on bills

  #get title
  df_bill_pages_name <- data %>% 
    xml2::xml_find_all("//title") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="title") %>% 
    mutate(title=str_trim(title)) %>% 
    filter(id>2)

  #get publication date
  df_bill_pages_pub_date <- data %>% 
    xml2::xml_find_all("//pubDate") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="date") %>% 
    mutate(date=date %>% str_squish() %>% lubridate::dmy_hms(., tz="Europe/Vienna"))

  #combine to one dataframe
  df_bill <- bind_cols(
    df_bill_pages_name,
    df_bill_pages_pub_date,
    df_bill_pages_link,
  ) %>% 
    select(-contains("id"))

  df_bill %>% 
    mutate(indicator_me_numeric=str_extract(link_single_bill_page, regex("(?<=_)\\d*"))) %>% 
    mutate(indicator_me=glue::glue("{as.numeric(indicator_me_numeric)}/ME")) %>% 
    mutate(rss_link_submissions=glue::glue("https://www.parlament.gv.at/PAKT/VHG/XXVII/ME/ME_{indicator_me_numeric}/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP=XXVII&ITYP=ME&INR={as.numeric(indicator_me_numeric)}&SUCH=&listeId=142&FBEZ=FP_142
"))

}
```

Then I apply this function to all the links of the rss feeds for the different legislative periods. As you can see in the image above, the parliament's website provides data for the XIV to the XXVII legislative period. Since the roman number of the period forms part of the rss-link, I make use of the `as.roman` function which let's us easily iterate over the sequence from 14 to 27.

```{r chunk_get_bills_overview_copy_2, eval=F, code_folding=NULL, include=T}

#link to rss feeds of all available legislative period
periods <- as.roman(seq(14, 27, 1)) #freaking nice
rss_links <- glue::glue("https://www.parlament.gv.at/PAKT/MESN/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP={periods}&MESN=ME&R_MESN=ME&MIN=ALLE&SUCH=&listeId=102&FBEZ=FP_002")

#apply function
df_bills_all <- rss_links %>% 
  purrr::set_names() %>% 
  map_dfr(., fn_get_bill_overview, .id="rss_links")
```

The result is a data frame providing us with some basic information on each bill, but more importantly, a link to each bill's overview page (`link_single_bill_page`) and page with submissions filed during the consultiatve period (`rss_link_submisison`).

```{r include=T, echo=F, code_folding=NULL}
readr::read_csv2(file=here::here("_blog_data","2021_covid_consultation_duration", "df_bills_all.csv")) %>% 
  select(title, date, contains("indicator"), contains("link")) %>% 
  glimpse()
```

## Get details on bills' duration
The overview page of each bill contains a table which details the bill's different stages in the legislative process as well as links to the text of the  bill open for consultation. It is here where we can retrieve the start and the end date of the consultation process, and subsequently calculate the process' duration. We can also get the links to download the bill's file(s) and subsequently extract their page numbers.

```{r fig.height=5, echo=F, code_folding=NULL, include=T}
knitr::include_graphics("covidconsultationduration_files/images/table_law_stages.png")
```

To do so I constructed the function `fn_consult` which reads the content of the overview table with the `rvest` package. I'll explain the function in two parts:

First, to get the start and end date, I keep (filter) only those rows which include the word 'Begutachtungsfrist' (consultation deadline). From this row we can further extract the start and end date.

Second, to eventually get the text of the bills, the function extracts those weblinks which include 'Ministerialentwurf' (ministrial draft) and the file suffix 'pdf', or 'Gesetzestext' (law text) or 'Entwurf' (draft) in their text. The `xpath` argument  of the `rvest::html_node` function allows for `regex` expressions what is just... beautiful. These links will be later used to download the bills and extract their number of pages.

Third,

```{r chunk_fn_consult_copy, eval=F, code_folding=NULL, include=T}
#function to extract details (duration and pdf links) from bills' overview
fn_consult <- function(link_bill)  {

#1) GET ROWS WITH START AND END DATES OF CONSULTATION PERIOD  
  df_consultation_dates <- link_bill %>% 
    read_html() %>% 
    html_nodes("#content > div.contentBlock.tabs-responsive__contentBlock > div:nth-child(2) > table") %>% 
    html_table() %>% 
    map_df(., bind_rows) %>% 
    janitor::clean_names() %>% 
    filter(str_detect(stand_des_parlamentarischen_verfahrens, regex("Begutachtungsfrist"))) %>% #Identifier to select rows with dates
    rename(date_start=datum) %>% 
    mutate(date_end=str_extract(stand_des_parlamentarischen_verfahrens, regex("\\d{2}\\.\\d{2}\\.\\d{4}"))) %>% #extract number for dates
    select(-protokoll)

#2) GET LINKS TO PDFS OF DRAFT BILLS
df_link_text <-   link_bill %>% 
    read_html() %>% 
    html_nodes(xpath="//a[contains(text(), 'Ministerialentwurf') 
                and contains(text(), 'PDF') 
                or contains(text(), 'Gesetzestext')
                or contains(text(), 'Entwurf')]") %>%  #works! case sensetive
    html_attr("href")  %>% 
   tibble::enframe(name=NULL, value="link_bill_pdf") %>% 
   mutate(link_bill_pdf=paste0("https://www.parlament.gv.at/", link_bill_pdf))

  bind_cols(df_consultation_dates, df_link_text) %>% 
    select(text_raw="stand_des_parlamentarischen_verfahrens", date_start, date_end,
           link_bill_pdf)

}

#3 APPLY FUNCTION AND CREATE A TIBBLE
tbl_missing <- tibble(link_single_bill_page=NA_character_, text_raw=NA_character_, date_start=NA, date_end=NA, link_bill_pdf=NA_character_)

library(furrr)
plan(multisession, workers = 2) #to speed up process, use multicore

#wrap function with 'possibly' to be on the safe side;
df_res <- df_bills_all$link_single_bill_page %>% 
  purrr::set_names() %>% 
  future_map_dfr(., possibly(fn_consult, otherwise=tbl_missing), .id="link_single_bill_page",
                 .progress=T) %>% 
  select(link_single_bill_page, text_raw, date_start, date_end, link_bill_pdf)

#house keeping; add legislative period and bill id;
df_res <- df_res %>% 
  mutate(across(.cols=contains("date"), lubridate::dmy)) %>% 
  mutate(legis_period=str_extract(link_single_bill_page, regex("[^/]*?(?=/ME/)"))) %>% 
  mutate(bill_id=str_extract(link_single_bill_page, regex("ME_.*?(?=/)")))

df_res <- df_res %>% 
  left_join(., df_bills_all %>% 
  select(link_single_bill_page, title))
```

The resulting dataframe is already an important building block, but we're not yet there. A few potential pitfalls have to be controlled for: When a bill has more than one link to the bill's text, our function provides us with multiple observations (rows). This has to be controlled for when aggregating the consultation period's overall duration (i.e. duplicates should not be summed up). On the other hand, there are bills in which the consultation period was extended. As a consequence we obtained multiple rows with start and end dates for one distinct bill.

```{r chunk_duration_copy, eval=T, code_folding=NULL, include=F}
df_res <-  readr::read_csv2(file=here::here("_blog_data", "2021_covid_consultation_duration",
                                           "df_res.csv")) %>% 
    mutate(across(.cols=contains("date"), lubridate::ymd))
```

```{r chunk_duration_copy_collect, eval=F, code_folding=NULL, include=T}
#remove dupes due to multiple pdf documents
df_duration <- df_res %>% 
  distinct(link_single_bill_page, 
           text_raw, 
           date_start, date_end, 
           legis_period, bill_id, title)

df_duration_mult <- df_duration %>% 
  group_by(link_single_bill_page, legis_period, bill_id, title) %>% 
  mutate(n_obs_duration=n()) %>% 
  ungroup()

#take each bill's earliest (minimum) start date and and latest (maximum) end date to calculate the overall duration of the consultation period;
df_duration_collected <- df_duration_mult %>% 
  filter(!is.na(date_start) & !is.na(date_end)) %>%  #remove those with missing dates
  group_by(link_single_bill_page, legis_period, bill_id, title) %>% 
  summarize(date_start_min=min(date_start, na.rm = T),
         date_end_max=max(date_end, na.rm = T),
         n_obs_duration=mean(n_obs_duration),
         dates_collected=paste(paste(as.character(date_start), "to", 
                              as.character(date_end)),
         collapse="; ")) %>% 
  mutate(duration_collected=date_end_max-date_start_min+1) %>% 
  ungroup()
```

The example below illustrates the issue of multiple consultation periods.

```{r fig.height=5, echo=F, code_folding=NULL, include=T}
knitr::include_graphics("covidconsultationduration_files/images/example_multiple_periods.png")
```

Here's how the case is included in the data table above:

```{r code_folding=NULL, include=T, echo=F, layout="l-body-outset"}
df_consolidated %>% 
  filter(duration_collected>=0) %>% 
  filter(bill_id=="ME_00294") %>% 
  filter(legis_period=="XXIV") %>% 
  mutate(link_submissions=paste0(link_single_bill_page, "#tab-Stellungnahmen")) %>% 
  select(legis_period, 
         title, 
         bill_id, 
         pdf_pages_sum, 
         date_start_min, 
         date_end_max, 
         duration_collected, 
         n_obs_submissions, 
         link_single_bill_page,
         link_submissions) %>%   
  reactable(.,
              columns=list(
              legis_period=colDef(name="legilsative period",
                                  width=70),
              title=colDef(name="bill's name",
                           minWidth=160),
              bill_id=colDef(name="bill id", 
                             align="right",
                             minWidth=50),
              pdf_pages_sum=colDef(name="# pages total",
                                   minWidth=50),
              date_start_min=colDef(show=F,
                                    name="start",
                                    align="right",
                                    ),
              date_end_max=colDef(show=F, 
                                  name="end",
                                  align="right"),
              duration_collected=colDef(name="duration",
                                        align="right",
                                        minWidth=50),
              n_obs_submissions=colDef(name="# submissions",
                                       minWidth=70),
              link_single_bill_page=colDef(
                name="link bill",
                show=T,
                filterable = F,
                width=50, 
                align="right",
                cell=function(value){
                  htmltools::tags$a(href=value,
                                    target="_blank",
                                    "Link")}),
              link_submissions=colDef(name="link submissions",
                                     show=T,
                                     filterable = F,
                                     width=100,
                                     align="right",
                                     cell = function(value) {
                                       htmltools::tags$a(href=value, target = "_blank", paste0("Link"))})),
            defaultSorted = list(n_obs_submissions = "desc"),
            sortable = T,
            compact = T,
            filterable = T,
            outlined = TRUE,
            theme=rt_theme,
            details=function(index) {
              df_consolidated_pos <- df_consolidated %>%
                filter(duration_collected>=0) %>%
                filter(bill_id=="ME_00294") %>% 
                filter(legis_period=="XXIV") 
              df_duration_details <- df_duration[df_duration$link_single_bill_page == df_consolidated_pos$link_single_bill_page[index],]
                df_duration_details %>% 
                select(starts_with("date")) %>% 
                reactable(., 
                          columns = list(
                            date_start=colDef(show=T, 
                                              name="consultation start"),
                            date_end=colDef(show=T,
                                            name="consultation end")),
                          fullWidth = F,
                          theme=reactableTheme(style=list(
                            margin="0px 0px 0px 50px",
                            backgroundColor=plot_bg_color)))
              })
```

## Missing or wrong data {#wrongdata}
The data obtained by now is essentially the outcome presented in the table above. There were however some results which were incomplete or plainly wrong. I think it's important to highlight that these observations were excluded from the results presented in the first section.

There were `r nrow(df_duration_mult %>% filter(if_any(starts_with("date"), is.na)))` bills where no start or end date was provided:

```{r layout="l-body-outset", code_folding=NULL, echo=F}
#there are 21 bills with missing dates
df_dates_missing <- df_duration_mult %>% 
  filter(if_any(starts_with("date"), is.na)) #new w dplyr 1.04
#nrow(df_dates_missing) #21

df_dates_missing %>% 
  select(legis_period, title, bill_id, contains("date"), link_single_bill_page) %>% 
  reactable(.,
            columns=list(
               legis_period=colDef(name="legilsative period",
                                  width=70),
              title=colDef(name="bill's name",
                           minWidth=160),
              bill_id=colDef(name="bill id", 
                             align="right",
                             minWidth=50),
              link_single_bill_page=colDef(
                name="link bill",
                show=T,
                filterable = F,
                width=50, 
                align="right",
                cell=function(value){
                  htmltools::tags$a(href=value,
                                    target="_blank",
                                    "Link")})),
            theme=rt_theme)

```

There were also `r nrow(df_duration_collected %>% filter(duration_collected<0))`  bills where start and end dates were provided, but they effectively meant a negative duration. As far as I can tell, the reason behind these erroneous results are either due to obviously wrong data entries or an inconsistency in the data representation (and subsequently wrongly extracted data).

```{r, eval=T, code_folding=NULL, include=T, layout="l-body-outset"}
df_duration_collected %>% 
  filter(duration_collected<0) %>% 
  mutate(duration_collected=as.numeric(duration_collected)) %>% 
  select(legis_period, title, bill_id, contains("date_"), duration_collected, link_single_bill_page) %>% 
  reactable(.,
            columns=list(
              legis_period=colDef(name="legilsative period",
                                  width=70),
              title=colDef(name="bill's name",
                           minWidth=160),
              bill_id=colDef(name="bill id", 
                             align="right",
                             minWidth=50),
              duration_collected=colDef(name="duration",
                                        format = colFormat(digits = 0)),
              link_single_bill_page=colDef(
                name="link bill",
                show=T,
                filterable = F,
                width=50, 
                align="right",
                cell=function(value){
                  htmltools::tags$a(href=value,
                                    target="_blank",
                                    "Link")})),
            theme=rt_theme)
```

Overall, there were `r nrow(df_duration_collected %>% filter(duration_collected<0)) + nrow(df_duration_mult %>% filter(if_any(starts_with("date"), is.na)))` cases in which the obtained duration was not helpful and eventually omitted from the further analysis. With a bit more time, one could go through these cases and correct the errors.

## Getting the length of the  bills
The second part of the analysis required the extraction of the number of pages per bill. Embarking from our initial dataframe `df_res` which included already the links to all submission pdfs, we can construct file names (`file_name`) and a download destination.

```{r BILL_LENGTH_copy, eval=F, code_folding=NULL, include=T}
df_pages <- df_res %>% 
  select(link_single_bill_page, link_bill_pdf, 
         legis_period, bill_id, title) %>% 
  distinct() %>% #remove duplicates
  filter(link_single_bill_page %in% df_duration_collected$link_single_bill_page) %>% #only for those where we have valid durations
  mutate(file_name=str_extract(link_bill_pdf, regex("[^\\/]*pdf$"))) %>% 
  mutate(file_name=glue::glue("{legis_period}_{bill_id}_{file_name}")) %>% 
  mutate(download_destination=glue::glue("{here::here('_blog_data',
                                         '2021_covid_consultation_duration',
                                         'bills_pdf')}/{file_name}") %>%
           as.character() %>% 
           str_trim(., side=c("both"))) 
```

With the `purrr:walk2` function we can take 1) the links to the files (`link_bill_pdf`) and the 2) download destination (`download_destination`) as inputs and feed them to the download function which I additionally wrapped with the `safely` adverb. Since there is a considerable number of documents to download, this may take a while (a few hours).

```{r eval=F, code_folding=NULL, include=T}
safe_download <- safely(~ download.file(.x , .y, mode = "wb", quiet=T))

library(magrittr)
df_pages  %$%  #%$% exposes df_pages's names
  purrr::walk2(
    link_bill_pdf,
    download_destination,
    safe_download, 
    mode = "wb")

```

To eventually extract the number of pages from each document, I make once again use of the wonderful `pdftools` package. Mapping each downloaded file to pdftools' `pdf_info` function gives us a list including the number of each document's pages. Similar to the calculation of a consultation period's length, we also have to account for bills which comprise multiple documents and, in these cases, have to aggregate them to a total number of a bill's page.

```{r eval=F, code_folding=NULL, include=T}
library(pdftools)

df_pages <- df_pages %>% 
  filter(link_single_bill_page %in% df_duration_collected$link_single_bill_page) %>% 
  mutate(pdf_details=map(download_destination, possibly(pdftools::pdf_info, otherwise = NA_integer_))) %>% 
  mutate(pdf_pages=map_dbl(pdf_details, purrr::pluck, "pages", .default=NA)) %>% #note default=NA, not NULL
  select(-pdf_details)

df_pages_collected <- df_pages %>% 
  distinct() %>% #double check to remove duplicates
  group_by(link_single_bill_page) %>% 
  summarise(pdf_pages_sum=sum(pdf_pages, na.rm = T),
            n_obs_docs=n())

```

# Possible next steps

As so often, the above presented results - as insightful as they may (or may not) be - are only stepping stones and further improvements are definitely possible. One obvious step for 'fine-tuning' could be to correct the duration of the consultation period for the number of actual working days. As of now, there is no differentiation between working and non-working days. Another, possible avenue could be not to take the number of pages of a bill, but the number of characters as an indicator for the document's length. This step would require to OCR the entire text of all bills. While this 'technically' not a big deal (again thanks to the `pdftools` package), it is rather time consuming endeavor (at least on my laptop; after 36 hours I interrupted the process). Having said this, though, I am not sure how much different the results would actually look like and whether its really worth the effort.

### 2021-02-14 _post/2021-02-12-thankyoublogdownhellodistill/thankyoublogdownhellodistill.Rmd
main update

---
title: "Thank you blogdown, hello distill"
description: |
  A short description of the post.
author:
  - name: Roland Schmidt
    url: https://werk.statt.codes
date: 02-12-2021

draft: T

output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

So, I moved my collection of posts ('blog' sounds a bit to serious) from blogdown to distill. 
In principle, setting up the new blog (damn...) was quite straighforward, particularly thanks to the
comprehensive documentation on the distill site and posts by Thomas, XXXX and XXXX.

Nevertheless, a few quick remarks which might be helpful to those who are also considerng to move:

Search function:
  Not that the search bar will only show up once you've deployed your site (e.g. via githut to netlify). 
  I initially was trying to find an error until I mustered the courage to post a pertaining question on github, and
  Jeff was kind enough to quickly respond. The mentioned 'sevr' package gives you a handy tool to test-drive your
  site before deploying.

Where to put static files:
_blog_data

theming:
css

performance:
tick slower than blogdown / based on hugo

### 2021-02-16 _post/2020-11-18-recreating-nyt-election-map/recreating-nyt-election-map.Rmd
update nyt map a facebook vienna

@@
 description: |
-  How to re-create election maps by the NYT with ggplot
+  How to re-create election maps by the NYT with ggplot.
 date: 11-18-2020
 
-draft: TRUE
+draft: FALSE
 
@@
 creative_commons: CC BY-NC-SA
-
 ---
@@
 
-![](index_files/Capture.PNG)
+![](recreating-nyt-election-map_files/images/Capture.PNG)
 
@@
 library(tidyverse)
+library(opencage)
 
 plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
-  str_extract(., regex("(?<=background-color:).*?(?=;)")) %>% str_trim()
+  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
+  str_trim() %>% 
+  str_extract(., regex("^#\\S+"))
+
 ```
 
-<details closed>
+</details>
 
@@
 
-```{r setup, echo = T}
-knit_hooks$set(wrap = function(before, options, envir) {
-  if (before) {
-    paste0("<", options$wrap, ">")
-  } else {
-    paste0("</", options$wrap, ">")
-  }
-})
-
+```{r setup, echo = F}
 knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
@@
 <details open>
+
 <summary>Code: Calculate centroids</summary>
@@
 ```
+
 </details>
 
-Use the `googleway` package to get the long/lat details for Philly and Pittsburgh (see [here](https://werk.statt.codes/post/primary-schools-in-vienna/){target="_blank"} for another blog where I used the package quite heavily.)
+Use the ~~`googleway`~~ `opencage` package to get the long/lat details for Philly and Pittsburgh.[^1]
+
+[^1]: In a previous version I used the googleway package. The opencage package allows to get geolocation without needing to access the google API.
 
 <details open>
+
 <summary>Code: Geocode cities</summary>
-```{r message=FALSE, warning=FALSE}
+
+```{r message=FALSE, warning=FALSE, eval=F, include=F}
 # data for pittsburgh and philadelphia ------------------------------------
 
-library(googleway)
+# library(googleway)
+# cities <- c("Philadelphia", "Pittsburgh")
+# df_cities <- map_dfr(cities, ~google_geocode(address=.x, 
+#                                              key=Sys.getenv("googleway"),
+#                                              simplify = T)) 
+# 
+# df_cities_geom <- df_cities$results$geometry$location
+# 
+# df_cities_add <- df_cities$results$formatted_address %>% enframe(., name=NULL, value="city")
+# 
+# df_cities <- bind_cols(df_cities_geom, df_cities_add)
+# 
+# sf_cities <- st_as_sf(df_cities, coords = c("lng", "lat"), remove = FALSE, 
+#          crs = 4326, agr = "constant") %>% 
+#   st_transform(32617)
+```
+
+```{r}
 cities <- c("Philadelphia", "Pittsburgh")
-df_cities <- map_dfr(cities, ~google_geocode(address=.x, 
-                                             key=Sys.getenv("googleway"),
-                                             simplify = T))
+df_cities <- map_dfr(cities, ~oc_forward_df(placename=.x))
 
-df_cities_geom <- df_cities$results$geometry$location
-
-df_cities_add <- df_cities$results$formatted_address %>% enframe(., name=NULL, value="city")
-
-df_cities <- bind_cols(df_cities_geom, df_cities_add)
-
-sf_cities <- st_as_sf(df_cities, coords = c("lng", "lat"), remove = FALSE, 
+sf_cities <- st_as_sf(df_cities, coords = c("oc_lng", "oc_lat"), remove = FALSE, 
          crs = 4326, agr = "constant") %>% 
   st_transform(32617)
+
 ```
+
 </details>
@@
 <details open>
+
 <summary>Code: map</summary>
+
 ```{r message=FALSE, warning=FALSE}
@@
   geom_sf_text(data = sf_cities,
-                aes(label=city %>% 
+                aes(label=placename %>% 
                       str_extract(., regex("[:alpha:]*"))),
@@
 ```
+
 </details>
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, dev.args=list(bg=plot_bg_color)}
+```{r echo=FALSE, message=FALSE, warning=FALSE, dev.args=list(bg=glue::glue("{plot_bg_color}")), dev="ragg_png"}
 pl_map
@@
 <details open>
+
 <summary>Code: legend inset</summary>
+
 ```{r message=FALSE, warning=FALSE}
@@
 ```
+
 </details>
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, dev.args=list(bg=plot_bg_color)}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, dev.args=list(bg=glue::glue("{plot_bg_color}")), dev="ragg_png"}
 pl_inset
@@
 <details open>
+
 <summary>Code: Combine plots</summary>
+
 ```{r message=FALSE, warning=FALSE}
@@
 ```
+
 </details>
@@
 
-```{r echo=FALSE, fig.asp=.6, fig.width=9, message=FALSE, warning=FALSE, dev.args=list(bg=plot_bg_color), out.height="110%"}
+```{r echo=FALSE, fig.asp=.6, fig.width=9, message=FALSE, warning=FALSE, dev="ragg_png", dev.args=list(bg=glue::glue("{plot_bg_color}")), out.height="110%", crop=T}
 pl_combined

### 2021-02-16 _post/2020-12-19-vienna-elections-2020-analysis-of-parties-ads-on-facebook/welcome.Rmd
update nyt map a facebook vienna

@@
 
-Do I see this text
-
 ```{r}
@@
 
-glimpse(df_ads)
-
 ```
@@
 #readr::write_csv2(df_impression_check,file=here::here("blog_data","vienna_elections_2020_facebook","df_impression_check.csv"))
-```
-
-```{r echo=FALSE}
-glimpse(df_ads_wide)
 ```
@@
 df_pl_pages <- df_ads_wide %>% 
+  ungroup() %>% 
   group_by(party) %>% 
@@
                                   other_level="all other")) %>% 
+  ungroup() %>% 
   group_by(party, page_name_lum) %>% 
   summarise(ads_n=n(),
-            across(.cols=contains("spend_") & where(is.numeric),
-                   .fns=sum)) %>% 
-  select(party, page_name_lum, ads_n, spend_lower_bound_rev, spend_mid, spend_upper_bound) %>%   ungroup() %>% 
+            across(.cols=c(contains("spend_") & where(is.numeric)),
+                   .fns=sum
+                   )) %>% 
+  ungroup() %>% 
+  select(party, page_name_lum, ads_n, spend_lower_bound_rev, spend_mid, spend_upper_bound) %>%   
   arrange(-ads_n)

### 2021-02-16 _post/2021-01-22-covidconsultationduration/covidconsultationduration.Rmd
update format post similarity
update nyt map a facebook vienna

@@
 Over New Year (!), the Austrian government introduced another bill with the aim to get a grip on the COVID pandemic.
-With the speed with which things have been developping and the flurry of hard/soft/de facto lockdowns, I have to confess that I would have to look into the text to recall the changes introduced by bill 88/ME (in full: *'Epidemiegesetz, COVID-19-Maßnahmengesetz, Änderung'*).
-However, what made the bill remarkable even during these times was its legislative genesis, with the government limiting the public consultative process to only three days.
+With the speed with which things have been developping and the flurry of hard/soft/de facto lockdowns, I have to confess that I would have to look into the text to recall the changes introduced by bill 88/ME (XXVII) (in full: *'Epidemiegesetz, COVID-19-Maßnahmengesetz, Änderung'*).
+However, what made the bill memorable even during these times was its legislative genesis, with the government limiting the public consultative process to only three days.
 As mentioned in previous posts, Austria's legislative procedure includes a public consultation process which allows citizens, NGOs, churches etc.
@@
 
-This debate made me wonder how long other bills were open for the consultative process and how the bill 88/ME would compare, also when considering their documents' lengths.
-To answer these questions, I a) extracted from the parliament's website the start and end dates of each bill's consultative period to calculate the length of the consultative process; b) downloaded the text of all pertaining bills and retrieved their number of pages.
+This debate made me wonder how long other bills were open for the consultative process and how bill 88/ME (XXVII) would compare, also when considering their documents' lengths.
+To answer these questions, I a) extracted from the parliament's website the start and end dates of all bill's consultative period since 1975, *as published on the parliament's website*.
+Based on these dates I calculated the length of the consultative process; b) downloaded the text of all pertaining bills and retrieved their number of pages.
 Putting the length of the consultation process and the length of the bill together should allow to better evaluate the three day notice for the amendment to the epidemic bill.
@@
        x="Length of consultation period",
-       caption=caption)+
+       caption=glue::glue("{caption}\nAnalysis excluded {nrow(df_duration_mult %>% filter(if_any(starts_with('date'), is.na)))+nrow(df_duration_collected %>% 
+  filter(duration_collected<0))} bills with errors in the data source. See blog for details.")
+       )+
   geom_bar(aes(x=duration_3,
@@
                                     hjust=0),
-        axis.title.x = element_blank())
+        axis.title.x = element_text(color="grey30",
+                                    # angle=90,
+                                    # vjust=1,
+                                    hjust=0))
+
 ```
@@
 
-To start with, let's have a look at the number of submissions filled during the consultation process per bill.
+To start with, let's have a look at the number of submissions filed during the consultation process per bill.
 The plot below  shows quite clearly that bills related to Covid triggered record breaking numbers of submissions.
@@
 How unusual is it that a bill of three pages length, i.e.
-bill 88/ME, is open for consultation for only three days?
-The plot below essentially addresses the question put forward in the introduction.
+bill 88/ME (XXVII), is open for consultation for only three days?
+The plot below addresses this question.
 As the orange dot in the graph below highlights, there has never been another bill with three pages and as little as three days of consultation.
 The next bills in the same page-length category had eight days (among them a bill amending the law on gambling/*Glückspielgesetz*...).
-So at least from this perspective, our bill in question has been quite an aberration.
+So at least from this perspective, the consultation period for bill 88/ME was quite an aberration.
 
@@
 Nevertheless, out of the `r nrow(df_consolidated)` bills open for consultation, `r df_consolidated %>% filter(duration_collected>3) %>% nrow()` featured longer consultation periods than three days.
-There were only `r df_consolidated %>% filter(duration_collected>=0) %>% filter(duration_collected<=3) %>% nrow()` bills with three or fewer days (excluding results with negative duration or incomplete start/end dates, see [here](#wrongdata)).
+There were only `r df_consolidated %>% filter(duration_collected>=0) %>% filter(duration_collected<=3) %>% nrow()` bills with three or fewer days (excluding results with negative duration or incomplete start/end dates due to errors in the data source, see [here](#wrongdata)).
 
@@
 
-```{r, fig.height=5, echo=F, code_folding=NULL}
+```{r, fig.height=5, echo=F, code_folding=NULL, dev="ragg_png"}
 pl_consolidated
@@
 If you are interested in the details pertaining to each bill, the table below provides them.
+Note that for the sake of completeness this table also includes those bills for which I retrieved impossible/wrong duration data (see [here](#wrongdata) for details.) However, the latter were excluded from the analysis.
 

### 2021-02-17 _post/2020-07-21-austrian-parliament-mps-transparency-and-conflict-of-interest-report/austrian-parliament-mps-transparency-and-conflict-of-interest-report.Rmd
css remove line
without theme

@@
     orcid_id: 0000-0002-6897-1693
+
 categories:
@@
   message = FALSE,
-  warning = FALSE,
-  dpi = 300
+  warning = FALSE
 )
@@
 <summary>Code: Define plot theme</summary> 
-```{r echo=TRUE}
-theme_post <- function() {
+```{r echo=F}
+plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
+  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
+  str_trim() %>% 
+  str_extract(., regex("^#\\S+"))
+
+theme_post2 <- function() {
   hrbrthemes::theme_ipsum_rc() +
     theme(
+      plot.background = element_rect(fill = plot_bg_color, color=NA),
+      panel.background = element_rect(fill = plot_bg_color, color=NA),
       plot.margin = margin(l = 0, unit = "cm"),
@@
 
+theme_post <- function() {
+  hrbrthemes::theme_ipsum_rc() +
+    theme(
+      plot.background = element_rect(fill = plot_bg_color, color=NA),
+      panel.background = element_rect(fill = plot_bg_color, color=NA),
+      #panel.border = element_rect(colour = plot_bg_color, fill=NA),
+      #plot.border = element_rect(colour = plot_bg_color, fill=NA),
+      plot.margin = margin(l = 0, 
+                           t = 0.25,
+                           unit = "cm"),
+      plot.title = element_markdown(
+        color = "grey20",
+        face = "bold",
+        margin = margin(l = 0, unit = "cm"),
+        size = 13
+      ),
+      plot.title.position = "plot",
+      plot.subtitle = element_text(
+        color = "grey50",
+        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
+        size = 11
+      ),
+      plot.caption = element_text(
+        color = "grey50",
+        size = 8,
+        hjust = c(0)
+      ),
+      plot.caption.position = "panel",
+      axis.title.x = element_text(
+        angle = 0,
+        color = "grey50",
+        hjust = 1
+      ),
+      axis.text.x = element_text(
+        size = 9,
+        color = "grey50"
+      ),
+      axis.title.y = element_blank(),
+      axis.text.y = element_text(
+        size = 9,
+        color = "grey50"
+      ),
+      panel.grid.minor.x = element_blank(),
+      panel.grid.major.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      panel.spacing = unit(0.25, "cm"),
+      panel.spacing.y = unit(0.25, "cm"),
+      strip.text = element_text(
+        angle = 0,
+        size = 9,
+        vjust = 1,
+        face = "bold"
+      ),
+      legend.title = element_text(
+        color = "grey30",
+        face = "bold",
+        vjust = 1,
+        size = 9
+      ),
+      legend.text = element_text(
+        size = 9,
+        color = "grey30"
+      ),
+      legend.justification = "left",
+      legend.box = "horizontal", # arrangement of multiple legends
+      legend.direction = "vertical",
+      legend.margin = margin(l = 0, t = 0, unit = "cm"),
+      legend.spacing.y = unit(0.07, units = "cm"),
+      legend.text.align = 0,
+      legend.box.just = "top",
+      legend.key.height = unit(0.2, "line"),
+      legend.key.width = unit(0.5, "line"),
+      text = element_text(size = 5)
+    )
+}
+
 my_caption=c("data: Austrian Parliament, Liste gemäß § 9 Bezügebegrenzungs-BVG, 2020. \nanalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes")
@@
 
-![](/post/2020-07-03-austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/list_example - Copy.PNG){width=100%}
+![](austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/list_example - Copy.PNG){width=100%}
 
@@
 
-```
-</details>
-
-```{r echo=FALSE}
+pl_gender_category
+
+```
+</details>
+
+```{r echo=FALSE, fig.height=6, fig.width=10, include=T}
 pl_gender_category

### 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
add ggiraph test

---
title: "ggiraph text test"
description: |
  A short description of the post.
author:
  - name: Roland Schmidt
    url: https://werk.statt.codes
date: 02-17-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

```{r}
library(ggplot2)
library(ggiraph)
library(svglite)
library(ggtext)
data <- mtcars
data$carname <- row.names(data)

```

# Normal, theme_minmal

```{r}
gg_point = ggplot(data = data) +
  labs(title="long title long title long title")+
    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
    tooltip = carname, data_id = carname)) +
  theme_minimal()

girafe(ggobj = gg_point)
```

# Normal, theme_minmal, plot_title_elmenent_markdown

```{r}
gg_point = ggplot(data = data) +
  labs(title="long title long title long title")+
    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
    tooltip = carname, data_id = carname)) +
  theme_minimal()+
  theme(plot.title=element_markdown())

girafe(ggobj = gg_point)
```

# Normal, theme_minmal, plot_title_elmenent_markdown, device svglite

```{r, dev="svglite"}
gg_point = ggplot(data = data) +
  labs(title="long title long title long title")+
    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
    tooltip = carname, data_id = carname)) +
  theme_minimal()+
  theme(plot.title=element_markdown())

girafe(ggobj = gg_point)
```

### 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
add theme_post to ggiraph test

@@
 knitr::opts_chunk$set(echo = T)
+```
+
+```{r}
+theme_post <- function() {
+  hrbrthemes::theme_ipsum_rc() +
+    theme(
+      # plot.background = element_rect(fill = plot_bg_color, color=NA),
+      # panel.background = element_rect(fill = plot_bg_color, color=NA),
+      # #panel.border = element_rect(colour = plot_bg_color, fill=NA),
+      #plot.border = element_rect(colour = plot_bg_color, fill=NA),
+      plot.margin = margin(l = 0, 
+                           t = 0.25,
+                           unit = "cm"),
+      plot.title = element_markdown(
+        color = "grey20",
+        face = "bold",
+        margin = margin(l = 0, unit = "cm"),
+        size = 13
+      ),
+      plot.title.position = "plot",
+      plot.subtitle = element_text(
+        color = "grey50",
+        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
+        size = 11
+      ),
+      plot.caption = element_text(
+        color = "grey50",
+        size = 8,
+        hjust = c(0)
+      ),
+      plot.caption.position = "panel",
+      axis.title.x = element_text(
+        angle = 0,
+        color = "grey50",
+        hjust = 1
+      ),
+      axis.text.x = element_text(
+        size = 9,
+        color = "grey50"
+      ),
+      axis.title.y = element_blank(),
+      axis.text.y = element_text(
+        size = 9,
+        color = "grey50"
+      ),
+      panel.grid.minor.x = element_blank(),
+      panel.grid.major.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      panel.spacing = unit(0.25, "cm"),
+      panel.spacing.y = unit(0.25, "cm"),
+      strip.text = element_text(
+        angle = 0,
+        size = 9,
+        vjust = 1,
+        face = "bold"
+      ),
+      legend.title = element_text(
+        color = "grey30",
+        face = "bold",
+        vjust = 1,
+        size = 9
+      ),
+      legend.text = element_text(
+        size = 9,
+        color = "grey30"
+      ),
+      legend.justification = "left",
+      legend.box = "horizontal", # arrangement of multiple legends
+      legend.direction = "vertical",
+      legend.margin = margin(l = 0, t = 0, unit = "cm"),
+      legend.spacing.y = unit(0.07, units = "cm"),
+      legend.text.align = 0,
+      legend.box.just = "top",
+      legend.key.height = unit(0.2, "line"),
+      legend.key.width = unit(0.5, "line"),
+      text = element_text(size = 5)
+    )
+}
 ```
@@
 ```
+
+# Normal, theme_post, plot_title_elmenent_markdown, device svglite
+
+```{r, dev="svglite"}
+gg_point = ggplot(data = data) +
+  labs(title="long title long title long title")+
+    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
+    tooltip = carname, data_id = carname)) +
+  theme_post()+
+  theme(plot.title=element_markdown())
+
+girafe(ggobj = gg_point)
+```

### 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
use only theme_ipsum_rc

--- 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
@@
 library(ggtext)
+library(hrbrthemes)
+library(extrafont)
 data <- mtcars
@@
 
+```
+
+```{r}
+theme_post <- function() {hrbrthemes::theme_ipsum_rc()}
+```
+
+```{r, eval=F}
+theme_post <- function() {
+  hrbrthemes::theme_ipsum_rc() +
+    theme(
+      # plot.background = element_rect(fill = plot_bg_color, color=NA),
+      # panel.background = element_rect(fill = plot_bg_color, color=NA),
+      # #panel.border = element_rect(colour = plot_bg_color, fill=NA),
+      #plot.border = element_rect(colour = plot_bg_color, fill=NA),
+      plot.margin = margin(l = 0, 
+                           t = 0.25,
+                           unit = "cm"),
+      plot.title = element_markdown(
+        color = "grey20",
+        face = "bold",
+        margin = margin(l = 0, unit = "cm"),
+        size = 13
+      ),
+      plot.title.position = "plot",
+      plot.subtitle = element_text(
+        color = "grey50",
+        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
+        size = 11
+      ),
+      plot.caption = element_text(
+        color = "grey50",
+        size = 8,
+        hjust = c(0)
+      ),
+      plot.caption.position = "panel",
+      axis.title.x = element_text(
+        angle = 0,
+        color = "grey50",
+        hjust = 1
+      ),
+      axis.text.x = element_text(
+        size = 9,
+        color = "grey50"
+      ),
+      axis.title.y = element_blank(),
+      axis.text.y = element_text(
+        size = 9,
+        color = "grey50"
+      ),
+      panel.grid.minor.x = element_blank(),
+      panel.grid.major.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      panel.spacing = unit(0.25, "cm"),
+      panel.spacing.y = unit(0.25, "cm"),
+      strip.text = element_text(
+        angle = 0,
+        size = 9,
+        vjust = 1,
+        face = "bold"
+      ),
+      legend.title = element_text(
+        color = "grey30",
+        face = "bold",
+        vjust = 1,
+        size = 9
+      ),
+      legend.text = element_text(
+        size = 9,
+        color = "grey30"
+      ),
+      legend.justification = "left",
+      legend.box = "horizontal", # arrangement of multiple legends
+      legend.direction = "vertical",
+      legend.margin = margin(l = 0, t = 0, unit = "cm"),
+      legend.spacing.y = unit(0.07, units = "cm"),
+      legend.text.align = 0,
+      legend.box.just = "top",
+      legend.key.height = unit(0.2, "line"),
+      legend.key.width = unit(0.5, "line"),
+      text = element_text(size = 5)
+    )
+}
 ```

### 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
only theme_ipsum

--- 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
@@
 
+```
+
+```{r eval=T}
+theme_post <- function() {
+  hrbrthemes::theme_ipsum_rc() +
+    theme(
+      # plot.background = element_rect(fill = plot_bg_color, color=NA),
+      # panel.background = element_rect(fill = plot_bg_color, color=NA),
+      # panel.border = element_rect(colour = plot_bg_color, fill=NA),
+      # plot.border = element_rect(colour = plot_bg_color, fill=NA),
+      plot.margin = margin(l = 0, 
+                           t = 0.25,
+                           unit = "cm"),
+      plot.title = element_markdown(
+        color = "grey20",
+        face = "bold",
+        margin = margin(l = 0, unit = "cm"),
+        size = 13
+      ),
+      plot.title.position = "plot",
+      plot.subtitle = element_text(
+        color = "grey50",
+        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
+        size = 11
+      ),
+      plot.caption = element_text(
+        color = "grey50",
+        size = 8,
+        hjust = c(0)
+      ),
+      plot.caption.position = "panel",
+      axis.title.x = element_text(
+        angle = 0,
+        color = "grey50",
+        hjust = 1
+      ),
+      axis.text.x = element_text(
+        size = 9,
+        color = "grey50"
+      ),
+      axis.title.y = element_blank(),
+      axis.text.y = element_text(
+        size = 9,
+        color = "grey50"
+      ),
+      panel.grid.minor.x = element_blank(),
+      panel.grid.major.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      panel.spacing = unit(0.25, "cm"),
+      panel.spacing.y = unit(0.25, "cm"),
+      strip.text = element_text(
+        angle = 0,
+        size = 9,
+        vjust = 1,
+        face = "bold"
+      ),
+      legend.title = element_text(
+        color = "grey30",
+        face = "bold",
+        vjust = 1,
+        size = 9
+      ),
+      legend.text = element_text(
+        size = 9,
+        color = "grey30"
+      ),
+      legend.justification = "left",
+      legend.box = "horizontal", # arrangement of multiple legends
+      legend.direction = "vertical",
+      legend.margin = margin(l = 0, t = 0, unit = "cm"),
+      legend.spacing.y = unit(0.07, units = "cm"),
+      legend.text.align = 0,
+      legend.box.just = "top",
+      legend.key.height = unit(0.2, "line"),
+      legend.key.width = unit(0.5, "line"),
+      text = element_text(size = 5)
+    )
+}
 ```
@@
 ```
+
+# Normal, theme_post, plot_title_elmenent_markdown, device svglite
+
+```{r, dev="svglite"}
+gg_point = ggplot(data = data) +
+  labs(title="long title long title long title")+
+    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
+    tooltip = carname, data_id = carname)) +
+  theme_post()+
+  theme(plot.title=element_markdown())
+
+girafe(ggobj = gg_point)
+```
+
+# Normal, theme_ipsum_rc, plot_title_elmenent_markdown, device svglite
+
+```{r dev="svglite", eval=F}
+gg_point = ggplot(data = data) +
+  labs(title="long title long title long title")+
+    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
+    tooltip = carname, data_id = carname)) +
+  theme_ipsum_rc()+
+  theme(plot.title=element_markdown())
+
+girafe(ggobj = gg_point)
+```

### 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
only theme_ipsum

@@
 
-```{r dev="svglite", eval=F}
+```{r dev="svglite", eval=T}
 gg_point = ggplot(data = data) +

### 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
without svglite
wo element_markdown

@@
 ```
+
+# Normal, theme_ipsum_rc, plot_title_elmenent_markdown, dno evice svglite
+
+```{r dev="svglite", eval=T}
+gg_point = ggplot(data = data) +
+  labs(title="long title long title long title")+
+    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
+    tooltip = carname, data_id = carname)) +
+  theme_ipsum_rc()+
+  theme(plot.title=element_markdown())
+
+girafe(ggobj = gg_point)
+```
+
+# Normal, theme_ipsum_rc, plot_title_elmenent_markdown, dno evice svglite, o element_markdown
+
+```{r dev="svglite", eval=T}
+gg_point = ggplot(data = data) +
+  labs(title="long title long title long title")+
+    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
+    tooltip = carname, data_id = carname)) +
+  theme_ipsum_rc()
+
+girafe(ggobj = gg_point)
+```

### 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
modify theme_post element_text

@@
                            unit = "cm"),
-      plot.title = element_markdown(
+      plot.title = element_text(
         color = "grey20",
@@
     tooltip = carname, data_id = carname)) +
-  theme_post()+
-  theme(plot.title=element_markdown())
+  theme_post()#+
+  # theme(plot.title=element_text())
 

### 2021-02-19 _post/2020-07-21-austrian-parliament-mps-transparency-and-conflict-of-interest-report/austrian-parliament-mps-transparency-and-conflict-of-interest-report.Rmd
without css for perfomanc testing

--- 2021-01-08 _post/2020-07-21-austrian-parliament-mps-transparency-and-conflict-of-interest-report/austrian-parliament-mps-transparency-and-conflict-of-interest-report.Rmd
@@
 
-preview: austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/preview.png
+# preview: austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/preview.png
 
@@
     self_contained: false
+    code_folding: TRUE
 
 creative_commons: CC BY-NC-SA
-
 ---
@@
 ```{r include=FALSE}
-knitr::opts_chunk$set(echo = T)
+knitr::opts_chunk$set(echo = T,
+                      dev = "png")
 ```
@@
 The publication is mandatory by law, and basically seeks to curtail potential conflicts of interest and to increase transparency as to MPs' activities outside of parliament.
-While the list comes, unsurprisingly, with a number of caveats ^[For details see [here.](https://www.parlament.gv.at//POOL/SWBRETT/25020/0010/BezBegrBVGPar9Erkl-NR.pdf#page=1){target="_blank"}], and by itself can hardly constitute a watertight safeguard against the various 'undemocratic temptations', the report nevertheless seems to provide interesting data on MPs' financial statuses and professional machinations.
+While the list comes, unsurprisingly, with a number of caveats [^1], and by itself can hardly constitute a watertight safeguard against the various 'undemocratic temptations', the report nevertheless seems to provide interesting data on MPs' financial statuses and professional machinations.
 All in all enough reason to give it a closer look, i.e.
-a blog post.^[Halfway through the blog, I realized that the transparency website meineabgeordneten.at had alreay published a similar, although non-technial blog on this matter.
+a blog post.[^2] Note that I will focus only on Austria's lower chamber (*Nationalrat*) in this post.
+If time permits, I'll follow up with one on the upper chamber (*Bundesrat*) at a later point.
+
+[^1]: For details see [here.](https://www.parlament.gv.at//POOL/SWBRETT/25020/0010/BezBegrBVGPar9Erkl-NR.pdf#page=1){target="_blank"}
+
+[^2]: Halfway through the blog, I realized that the transparency website meineabgeordneten.at had alreay published a similar, although non-technial blog on this matter.
 If you speak German, have a look [here](https://www.meineabgeordneten.at/News/detail/Nationalrat_Nebeneinkommen_Offenlegung_2019){target="_blank"}.
 In a previous version I also had wrongly identified the gender of one MP.
-Kindly, Marion Breitschopf from meineabgeordneten.at helped clarifying this.] Note that I will focus only on Austria's lower chamber (*Nationalrat*) in this post.
-If time permits, I'll follow up with one on the upper chamber (*Bundesrat*) at a later point.
+Kindly, Marion Breitschopf from meineabgeordneten.at helped clarifying this.
 
@@
 # Setup
-But step by step:
-Let's first take care of the necessary R packages and blog/rmarkdown setup.
+
+But step by step: Let's first take care of the necessary R packages and blog/rmarkdown setup.
 
 <details closed>
+
 <summary>Code: Load packages</summary>
+
 ```{r echo=TRUE, message=FALSE, warning=FALSE}
@@
 library(hrbrthemes)
-hrbrthemes::update_geom_font_defaults(family = "Roboto Condensed",
-                                      size=3)
+# hrbrthemes::update_geom_font_defaults(family = "Roboto Condensed",
+                                      # size=3)
 library(scales)
@@
 <details closed>
+
 <summary>Code: Define rmarkdown options</summary>
+
 ```{r setup, echo = T}
@@
 
-knitr::opts_chunk$set(
-  echo = TRUE,
-  fig.align = "left",
-  # fig.height = 5,
-  # fig.width = 7,
-  out.width = "100%",
-  message = FALSE,
-  warning = FALSE,
-  dpi = 300
-)
-
-options(width = 180, dplyr.width = 150)
-```
+# knitr::opts_chunk$set(
+#   echo = TRUE,
+#   fig.align = "left",
+#   # fig.height = 5,
+#   # fig.width = 7,
+#   out.width = "100%",
+#   message = FALSE,
+#   warning = FALSE
+# )
+# 
+# options(width = 180, dplyr.width = 150)
+```
+
 </details>
@@
 <details closed>
+
 <summary>Code: Define plot theme</summary>
-```{r echo=TRUE}
-theme_post <- function() {
+
+```{r echo=F}
+plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
+  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
+  str_trim() %>% 
+  str_extract(., regex("^#\\S+"))
+
+# theme_post <- function() {
   hrbrthemes::theme_ipsum_rc() +
     theme(
+      plot.background = element_rect(fill = plot_bg_color, color=NA),
+      panel.background = element_rect(fill = plot_bg_color, color=NA),
       plot.margin = margin(l = 0, unit = "cm"),
-      plot.title = element_markdown(
+      plot.title = element_text(
         color = "grey20",
@@
     )
+#}
+
+theme_post <- function() {
+  hrbrthemes::theme_ipsum_rc() 
 }
@@
 # Data extraction
+
 As so often, one of the main challenges to actually analyze available data is to 'liberate' the information of interest from the format in which they are provided.
@@
 
-![](/post/2020-07-03-austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/list_example - Copy.PNG){width=100%}
+![](austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/list_example - Copy.PNG){width="100%"}
 
@@
 The income categories are:
+
 ```{r echo=FALSE}
@@
 <details open>
+
 <summary>Code: Extract text from pdfs </summary>
+
 ```{r}
@@
 The expression `str_replace_all(text_raw, "\\h+", " ")` replaces all horizontal white space with a single space, but keeps line breaks (\\n and \\r).
-The latter will be needed at a later stage when it comes to identifying the location of data of interest and extract them.^[I am grateful to the respondents to my pertaining question on [stackoverflow](https://stackoverflow.com/questions/62678640/stringr-remove-multiple-spaces-but-keep-linebreaks-n-r/62679016#62679016){target="_blank"}]
+The latter will be needed at a later stage when it comes to identifying the location of data of interest and extract them.[^3]
+
+[^3]: I am grateful to the respondents to my pertaining question on [stackoverflow](https://stackoverflow.com/questions/62678640/stringr-remove-multiple-spaces-but-keep-linebreaks-n-r/62679016#62679016){target="_blank"}
 
@@
 ## Get income category
+
 As a first step, let's extract the name of the MPs and the income category from the raw strings obtained above.
@@
 Note the `(?s)` term in the regex expressions.
-While `.` would match any pattern but not line breaks, the `(?s)` alters this limitation and forces `.` to match also `\n`  and `\r`.^[See [here](https://stackoverflow.com/questions/159118/how-do-i-match-any-character-across-multiple-lines-in-a-regular-expression/45981809#45981809){target="_blank"} for a very helpful answer on SO.]
+While `.` would match any pattern but not line breaks, the `(?s)` alters this limitation and forces `.` to match also `\n` and `\r`.[^4]
+
+[^4]: See [here](https://stackoverflow.com/questions/159118/how-do-i-match-any-character-across-multiple-lines-in-a-regular-expression/45981809#45981809){target="_blank"} for a very helpful answer on SO.
 
 <details open>
+
 <summary>Code: Extract MPs' names and income category </summary>
+
 ```{r}
@@
 <details open>
+
 <summary>Code: Account for multi-page declarations </summary>
+
 ```{r}
@@
 ## Missing income data
+
 Let's have a first, quick look at the information obtained so far:
@@
 <details closed>
+
 <summary>Code: Number of MPs per income bracket </summary>
+
 ```{r}
@@
 <details closed>
+
 <summary>Code: MPs with with missing income data</summary>
+
 ```{r}
@@
 ```
-</details>
+
+</details>
+
 ```{r echo=FALSE}
@@
 <details closed>
+
 <summary>Code: Get entry dates for MPs with missing income data</summary>
+
 ```{r eval=FALSE}
@@
 <details closed>
+
 <summary>Code: Remove MPs joining outside of reporting period </summary>
+
 ```{r echo=TRUE}
@@
 <details open>
+
 <summary>Code: Get MPs' gender</summary>
+
 ```{r}
@@
 ## Female share per income group
+
 What the subsequent analysis reveals is that the share of women per income category decreases the further we move up the income ladder.
@@
 <details closed>
+
 <summary>Code: Calculate gender ratios and create plots</summary>
+
 ```{r}
@@
                    position = "left")+
-  theme_post()+
-  theme(legend.position="none",
-        plot.title=element_markdown(),
-        axis.title.x=element_blank(),
-        axis.title.y.left = element_text(angle = 90,
-                                         color="grey50",
-                                         hjust=1),
-        axis.text.y.left = element_text(),
-        axis.text.y.right = element_blank(),
-        axis.title.y.right = element_blank(),
-        panel.grid.major.y  = element_blank())
-
-```
-</details>
-
-```{r echo=FALSE}
+  theme_post()#+
+  # theme(legend.position="none",
+  #       plot.title=element_markdown(),
+  #       axis.title.x=element_blank(),
+  #       axis.title.y.left = element_text(angle = 90,
+  #                                        color="grey50",
+  #                                        hjust=1),
+  #       axis.text.y.left = element_text(),
+  #       axis.text.y.right = element_blank(),
+  #       axis.title.y.right = element_blank(),
+  #       panel.grid.major.y  = element_blank())
+
 pl_gender_category
@@
 
+</details>
+
+```{r echo=FALSE, fig.height=6, fig.width=10, include=T, device="ragg_png"}
+pl_gender_category
+
+```
+
 ## Share of income groups per gender group
+
 Another gender sensitive perspective is to look at the distribution of the MPs across income categories within each gender group (in contrast to within each income group as in the previous graph).
@@
 <details closed>
+
 <summary>Code: graph intra-gender distribution </summary>
+
 ```{r}
@@
   facet_wrap(vars(gender))+
-  theme_post()+
+  # theme_post()+
   theme(legend.position = "none",
@@
 
+```{r}
+mtcars %>% 
+  ggplot()+
+  geom_point(aes(
+    x=mpg,
+    y=disp
+  ))
+```
+
 ## Get MPs' party membership
+
 As the next step to facilitate the analysis of the data, let's add MPs' party affiliation.
@@
 <details closed>
+
 <summary>Code: Get party affiliation </summary>
+
 ```{r}
@@
 <details closed>
+
 <summary>Code: plots on income categories</summary>
+
 ```{r}
@@
 ## Extracting data
+
 Extracting MPs' additional professional positions from the pdfs has been somewhat challenging.
@@
 <details open>
+
 <summary>Code: identifying new positions</summary>
+
 ```{r}
@@
 <details open>
+
 <summary>Code: Table with one row per position</summary>
+
 ```{r paged.print=TRUE}
@@
 <details closed>
+
 <summary>Code: row numbers of sub-headings</summary>
+
 ```{r}
@@
 ## Putting it all into a function
+
 I put all the above outlined (and subsequent) steps into a function which provides me with the number of positions per category for each MP.
@@
 <details open>
+
 <summary>Code: Function to extract positions </summary>
+
 ```{r}
@@
 <details open>
+
 <summary>Code: Apply function </summary>
+
 ```{r}
@@
 ## Paid positions and income category
+
 Let's have a look at MPs' number of paid positions and income category.
@@
 <details closed>
+
 <summary>Code: Number of paid positions and income category</summary>
+
 ```{r}
@@
 <details closed>
+
 <summary>Code: MPs with no income but paid positions</summary>
+
 ```{r}
@@
 But irrespective of this, there is nothing in the additional text what would suggest - at leat to me - that Himmelbauer did not receive at least for one month a salary during the reporting period.
-Similarly, Markus Koza and Meri Dioski had both paid positions for their own parities up until 31.10.2019. Nevertheless, this income was not reported as indicated by the income category 0.
+Similarly, Markus Koza and Meri Dioski had both paid positions for their own parities up until 31.10.2019.
+Nevertheless, this income was not reported as indicated by the income category 0.
 
@@
 <details open>
+
 <summary>Code: MP share w at least one add. position</summary>
+
 ```{r}
@@
 ## Executive positions {#unpaid_executive}
+
 Below I further disaggregate executive positions into those which entail a salary and those which don't.
-In the parliament’s report, executive positions without salary were marked with an asterisk (\*).
+In the parliament's report, executive positions without salary were marked with an asterisk (\*).
 When retrieving the data above, I already accounted for this difference (`str_detect(., "\\*")`).
@@
 <details open>
+
 <summary>Code: different types of executive positions</summary>
+
 ```{r}
@@
 ## Parties in comparison (Boxplot)
+
 The boxplot below might be a bit of an overkill, but it gives us a more detailed overview of MPs' additional positions per party.
@@
 <details closed>
+
 <summary>Code: boxplots</summary>
+
 ```{r echo=TRUE, fig.height=7}
@@
 # Wrap up
+
 So this has become a rather length blog entry.

### 2021-02-19 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
without css for perfomanc testing

@@
     self_contained: false
+    highlight: rstudio
+    highlight_downlit: false
+
+draft: true    
 ---

### 2021-02-19 _post/2020-07-21-attransperency/attransperency.Rmd
without css

--- 2021-02-19 _post/2020-07-21-austrian-parliament-mps-transparency-and-conflict-of-interest-report/austrian-parliament-mps-transparency-and-conflict-of-interest-report.Rmd
@@
 description: |
-  Analysis of MP's transparency and conflict of interest report.
+  A short description of the post.
+
 date: 07-21-2020
-
 author:
@@
 creative_commons: CC BY-NC-SA
+
 ---
-
-```{r include=FALSE}
-knitr::opts_chunk$set(echo = T,
-                      dev = "png")
-```
 
@@
 library(hrbrthemes)
-# hrbrthemes::update_geom_font_defaults(family = "Roboto Condensed",
-                                      # size=3)
+hrbrthemes::update_geom_font_defaults(family = "Roboto Condensed",
+                                      size=3)
 library(scales)
@@
 
-# theme_post <- function() {
+theme_post <- function() {
   hrbrthemes::theme_ipsum_rc() +
@@
       legend.key.width = unit(0.5, "line"),
-      text = element_text(size = 5)
+      text = element_text(size = 3)
     )
-#}
-
-theme_post <- function() {
-  hrbrthemes::theme_ipsum_rc() 
 }
@@
 
-![](austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/list_example - Copy.PNG){width="100%"}
+![](attransperency_files/list_example - Copy.PNG){width="100%"}
 
@@
 
-<details open>
-
-<summary>Code: Extract text from pdfs </summary>
-
-```{r}
+```{r code_folding="Code: Extract text from pdfs"}
 my_path <- here::here("_blog_data", "MP_activities", "BezBegrBVGPar9-NR.pdf")
@@
 
-</details>
-
 The code above provides us with a tibble containing as many rows as the report has pages (186).
@@
 
-```{r echo=FALSE, tidy=FALSE, R.options=list(width=10)}
+```{r echo=FALSE, tidy=FALSE, R.options=list(width=10), code_folding=NULL}
 stringr::str_wrap(df_all$text_raw[1], 10)
@@
 
-<details open>
-
-<summary>Code: Extract MPs' names and income category </summary>
-
-```{r}
+```{r, code_folding=FALSE}
 df_all <- df_all %>%
@@
 
-</details>
-
 As already highlighted, since the declarations of some MPs run over multiple pages and are hence included several times in the above output, we have to limit the result to one row per MP.
@@
 
-<details open>
-
-<summary>Code: Account for multi-page declarations </summary>
-
-```{r}
+```{r code_folding=FALSE}
 df_income <- df_all %>%
@@
   distinct()
-```
-
-```{r}
+
 nrow(df_all)
@@
 
-</details>
-
 ## Missing income data
@@
 
-<details closed>
-
-<summary>Code: Number of MPs per income bracket </summary>
-
-```{r}
+```{r, code_folding="Code: Number of MPs per income bracket"}
 tbl_income <- df_income %>% 
@@
 
-</details>
-
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=NULL}
 tbl_income
@@
 
-<details closed>
-
-<summary>Code: MPs with with missing income data</summary>
-
-```{r}
+```{r, code_folding="Code: MPs with with missing income data"}
 df_income_missing <- df_income %>% 
@@
 
-</details>
-
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=NULL}
 tbl_missing
@@
 
-<details closed>
-
-<summary>Code: Get entry dates for MPs with missing income data</summary>
+```{r include=FALSE}
+# readr::write_rds(df_mp_dates, path=here::here("_blog_data", "MP_activities", "MP_dates.rds"))
+
+df_mp_dates <- readr::read_rds(file=here::here("_blog_data", "MP_activities", "MP_dates.rds"))
+```
 
@@
 #apply funciton to all links
-# df_mp_dates <- df %>% 
-#   mutate(mp_dates=map(link, ~fn_get_dates(.)))
+df_mp_dates <- df %>%
+  mutate(mp_dates=map(link, ~fn_get_dates(.)))
 
@@
 ```{r include=FALSE}
-# readr::write_rds(df_mp_dates, path=here::here("_blog_data", "MP_activities", "MP_dates.rds"))
-
-df_mp_dates <- readr::read_rds(file=here::here("_blog_data", "MP_activities", "MP_dates.rds"))
-
-```
-
-```{r}
 #extract family name for subsequent merger with income data
@@
 
-</details>
-
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=NULL}
 df_income_missing %>% 
@@
 
-<details closed>
-
-<summary>Code: Remove MPs joining outside of reporting period </summary>
-
-```{r echo=TRUE}
+```{r echo=TRUE, code_folding="Code: Remove MPs joining outside of reporting period"}
 nrow(df_income)
@@
 ```
-
-</details>
 
@@
 
-<details open>
-
-<summary>Code: Get MPs' gender</summary>
-
-```{r}
+```{r, code_folding=FALSE}
 #get a clean name, without any academic or prefessional titles
@@
 
-</details>
-
 ## Female share per income group
@@
 
-<details closed>
-
-<summary>Code: Calculate gender ratios and create plots</summary>
-
-```{r}
+```{r, code_folding=T}
 df_gender <- df_income %>% 
@@
 
-```
-
-```{r}
 pl_gender_category <- df_gender %>% 
@@
                    position = "left")+
-  theme_post()#+
-  # theme(legend.position="none",
-  #       plot.title=element_markdown(),
-  #       axis.title.x=element_blank(),
-  #       axis.title.y.left = element_text(angle = 90,
-  #                                        color="grey50",
-  #                                        hjust=1),
-  #       axis.text.y.left = element_text(),
-  #       axis.text.y.right = element_blank(),
-  #       axis.title.y.right = element_blank(),
-  #       panel.grid.major.y  = element_blank())
-
-pl_gender_category
-
-```
-
-</details>
-
-```{r echo=FALSE, fig.height=6, fig.width=10, include=T, device="ragg_png"}
+  theme_post()+
+  theme(legend.position="none",
+        plot.title=element_markdown(),
+        axis.title.x=element_blank(),
+        axis.title.y.left = element_text(angle = 90,
+                                         color="grey50",
+                                         hjust=1),
+        axis.text.y.left = element_text(),
+        axis.text.y.right = element_blank(),
+        axis.title.y.right = element_blank(),
+        panel.grid.major.y  = element_blank())
+
+```
+
+```{r echo=FALSE, fig.height=6, fig.width=10, code_folding=NULL}
 pl_gender_category
@@
 
-```{r}
+```{r, code_folding=F}
 pl_gender_intra_dist <- df_gender %>% 
@@
   facet_wrap(vars(gender))+
-  # theme_post()+
+  theme_post()+
   theme(legend.position = "none",
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=NULL}
 pl_gender_intra_dist
@@
 # Party affiliation
-
-```{r}
-mtcars %>% 
-  ggplot()+
-  geom_point(aes(
-    x=mpg,
-    y=disp
-  ))
-```
-
 ## Get MPs' party membership
@@
 
-<details closed>
-
-<summary>Code: Get party affiliation </summary>
-
-```{r}
+```{r, code_folding="Code: Get party affiliation"}
 wiki_link <- "https://de.wikipedia.org/wiki/Liste_der_Abgeordneten_zum_%C3%96sterreichischen_Nationalrat_(XXVII._Gesetzgebungsperiode)"
@@
 
-</details>
-
 Below the result with all MPs, their party affiliation, and income data.
 
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=F}
 datatable(df_income %>% select(name, party, gender, income_category_19),
@@
 
-<details closed>
-
-<summary>Code: plots on income categories</summary>
-
-```{r}
+```{r, code_folding="Code: plots on income categories"}
 df_agg <- df_income %>% 
@@
                 ),
-            fill="white",
+            fill=plot_bg_color,
             label.size=0,
@@
 
-</details>
-
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=NULL}
 pl_rel
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=NULL}
 pdf_file <- pdftools::pdf_data(pdf = my_path) 
@@
 
-<details open>
-
-<summary>Code: identifying new positions</summary>
-
-```{r}
+```{r, code_folding="Code: identifying new positions"}
 df_linebreaks <- pdf_file[[8]] %>%
@@
 
-</details>
-
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=NULL}
 datatable(df_linebreaks,
@@
 
-<details open>
-
-<summary>Code: Table with one row per position</summary>
-
-```{r paged.print=TRUE}
+```{r paged.print=TRUE, code_folding="Code: Table with one row per position"}
 df_rows <- df_linebreaks %>% 
@@
 
-</details>
-
-```{r}
+```{r, ech=F, code_folding=NULL}
 df_rows
@@
 
-<details closed>
-
-<summary>Code: row numbers of sub-headings</summary>
-
-```{r}
+```{r, code_folding="Code: row numbers of sub-headings"}
 # headings
@@
 
-</details>
-
 ## Putting it all into a function
@@
 
-<details open>
-
-<summary>Code: Function to extract positions </summary>
-
-```{r}
+```{r, code_folding="Code: Function to extract positions"}
 fn_get_positions <- function(x) {
@@
 
-</details>
-
 This function is applied to our previous results of the `pdf_data` function.
@@
 
-<details open>
-
-<summary>Code: Apply function </summary>
-
-```{r}
+```{r, code_folding="Code: Apply function"}
 df_data <- pdf_file %>%
@@
 
-</details>
-
-```{r echo=FALSE}
+```{r echo=FALSE, code_folding=NULL}
 datatable(df_count,
@@
 
-<details closed>
-
-<summary>Code: Number of paid positions and income category</summary>
-
-```{r}
+```{r, code_folding="Code: Number of paid positions and income category"}
 df_income_positions <- df_count2 %>% 
@@
 
-</details>
-
-```{r echo=FALSE, fig.height=15}
+```{r echo=FALSE, fig.height=15, code_folding=NULL}
 girafe(ggobj = pl_income_positions)
@@
 
-<details closed>
-
-<summary>Code: MPs with no income but paid positions</summary>
-
-```{r}
+```{r, code_folding="Code: MPs with no income but paid positions"}
 df_mp5 <- df_income_positions %>% 
@@
 
-</details>
-
 ```{r echo=FALSE}
@@
 
-<details open>
-
-<summary>Code: MP share w at least one add. position</summary>
-
-```{r}
+```{r, code_folding=">Code: MP share w at least one add. position"}
 df_bar_position <- df_position_long %>% 
@@
 
-<details open>
-
-<summary>Code: different types of executive positions</summary>
-
-```{r}
+```{r, code_folding="Code: different types of executive positions"}
 df_executive <- df_count2 %>% 
@@
                             )),
-            fill="white",
+            fill=plot_bg_color,
             label.padding = unit(0.2,"line"),
@@
 
-```{r}
+```{r, include=F, eval=F}
 df_nr<- df_count %>% 

### 2021-02-20 _post/2019-11-20-vorzugsstimmen/vorzugsstimmen.Rmd
css backgroud white

@@
 
-```{r echo=FALSE, message=FALSE, warning=TRUE}
+```{r echo=FALSE, message=FALSE, warning=TRUE, eval=F}
 df_merge %>% 
@@
 
-```{r message=FALSE, warning=FALSE, paged.print=TRUE, code_folding=NULL, echo=F}
+```{r message=FALSE, warning=FALSE, paged.print=TRUE, code_folding=NULL, echo=F, eval=F}
 library(reactable)

### 2021-02-20 _post/2020-07-21-attransperency/attransperency.Rmd
css backgroud white

@@
       legend.key.width = unit(0.5, "line"),
-      text = element_text(size = 3)
+      text = element_text(size = 5)
     )
@@
 
-```{r echo=FALSE, code_folding=F}
+```{r echo=FALSE, code_folding=NULL}
 datatable(df_income %>% select(name, party, gender, income_category_19),
@@
 
-```{r echo=FALSE, code_folding=NULL}
+```{r echo=FALSE, fig.height=7, fig.width=10, code_folding=NULL, dev="svglite"}
 pl_rel

### 2021-02-20 _post/2020-12-22-submissionsepidemiclaw/submissionsepidemiclaw.Rmd
update format post similarity
css backgroud white

@@
           theme = reactableTheme(
-            backgroundColor = "#f0eff0",
+            backgroundColor = plot_bg_color,
                 filterInputStyle = list(
                   color="green",
-                  backgroundColor = "#f0eff0")))
+                  backgroundColor = plot_bg_color)))
 ```
@@
 ```{r include=FALSE}
-df_submission <- readr::read_rds(file="submissionsepidemiclaw_files/df_submission_150.rds")
+df_submission <- readr::read_rds(file=here::here("_blog_data", "epidemic_law_submissions",
+                                                 "df_submission_150.rds"))
 
@@
           theme = reactableTheme(
-            backgroundColor = "#f0eff0",
+            backgroundColor = plot_bg_color,
                 filterInputStyle = list(
-                  color="green",backgroundColor = "#f0eff0")
+                  color="green",backgroundColor = plot_bg_color)
           )
@@
           theme = reactableTheme(
-            backgroundColor = "#f0eff0",
+            backgroundColor = plot_bg_color,
                 filterInputStyle = list(
                   color="green",
-                  backgroundColor = "#f0eff0")))
+                  backgroundColor = plot_bg_color)))
 

### 2021-02-20 _post/2021-01-22-covidconsultationduration/covidconsultationduration.Rmd
css backgroud white

@@
   A (relative) deep dive into the length of the public consultation period for draft bills in Austria, and whether three days for three pages are actually 'normal'.
-date: 01-22-2021
+date: 02-20 -2021
 draft: false
@@
   ggplot()+
-  labs(title="LEGISLATIVE CONSULTATION PROCESS:<br>Number of filed submissions per bill.",
+  labs(title="LEGISLATIVE CONSULTATION PROCESS:\nNumber of filed submissions per bill.",
        subtitle="Note log scale of y-axis.",
@@
 
-```{r, fig.height=5, echo=F, code_folding=NULL, dev="ragg_png"}
+```{r, fig.height=5, echo=F, code_folding=NULL, dev="svglite"}
 pl_consolidated

### 2021-02-22 _post/2019-11-20-vorzugsstimmen/vorzugsstimmen.Rmd
modified post.json

--- 2021-01-22 _post/2019-11-20-vorzugsstimmen/vorzugsstimmen.Rmd
@@
 
-```{r echo=FALSE, message=FALSE, warning=TRUE}
+```{r echo=FALSE, message=FALSE, warning=TRUE, eval=F}
+#table.start
 df_merge %>% 
@@
                              backgroundColor = plot_bg_color)))
-
+#table.end
 ```
@@
 
-```{r message=FALSE, warning=FALSE, paged.print=TRUE, code_folding=NULL, echo=F}
+```{r message=FALSE, warning=FALSE, paged.print=TRUE, code_folding=NULL, echo=F, eval=F}
+#table.start
 library(reactable)
@@
 
+#table.end
 ```

### 2021-02-22 _post/2020-06-08-update-on-mobility-patterns/update-on-mobility-patterns.Rmd
modified post.json

@@
 #get all stations in Austria
-df_AT <- ncdc_stations(locationid='FIPS:AU')$data
+df_AT <- ncdc_stations(locationid='FIPS:AU', token=Sys.getenv("noaakey"))$data
 
@@
 #get temperature data
-df_res <- ncdc(datasetid = "GHCND", stationid = df_stations$id, datatypeid = "TAVG", startdate = "2020-01-01", enddate = "2020-06-01",
-     add_units=T, limit=1000)
+df_res <- ncdc(datasetid = "GHCND", 
+               stationid = df_stations$id, 
+               datatypeid = "TAVG", 
+               startdate = "2020-01-01", 
+               enddate = "2020-06-01", 
+               token=Sys.getenv("noaakey"),
+               add_units=T, 
+               limit=1000)
 

### 2021-02-22 _post/2020-07-21-attransperency/attransperency.Rmd
modified post.json

@@
 
-```{r echo=TRUE, fig.height=7}
+```{r echo=TRUE, fig.height=7, code_folding=F}
 df_position_labels <- df_position_long %>% 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, fig.height=7, fig.width=10, code_folding=NULL}
 pl_box

### 2021-02-22 _post/2020-10-10-vienna-elections-2020-age-profile/vienna-elections-2020-age-profile.Rmd
modified post.json

@@
 description: |
-  A short description of the post.
+  An empirical look at candidates' age.
 date: 10-10-2020
@@
 
-![](/post/2020-10-10-vienna-elections-2020-age-profile_files/Inkedtext_flow_LI_Moment.jpg){width=50% height=50%}
+![](vienna-elections-2020-age-profile_files/Inkedtext_flow_LI_Moment.jpg){width=50% height=50%}
 
@@
 ```{r echo=FALSE}
+#table.start
 html_tb_all
+#table.end
 ```
@@
 ```{r echo=FALSE}
+#table.start
 html_tb_young
+#table.end
 ```
@@
 ```{r echo=FALSE}
+#table.start
 html_tb_old
-
+#table.end
 ```
@@
 <br>
+
 Hover with the mouse over the box plot and dots to obtain the pertaining data.
+
 <details closed> 
@@
 ```{r echo=FALSE}
+#table.start
 html_age_list_district
+#table.end
 ```
@@
 ```{r echo=FALSE}
+#table.start
 tb_residence_party
+#table.end
 ```
@@
 ```{r echo=FALSE}
+#table.start
 tb_residence_district
+#table.end
 ```

### 2021-02-22 _post/2020-12-19-vienna-elections-2020-analysis-of-parties-ads-on-facebook/welcome.Rmd
css backgroud white
modified post.json

@@
 greens <- "#A3C630"
-strache <- "White"
+strache <- "grey50"
 bier <- "yellow"
@@
 
+#table.start
 fn_reactable_headings(title="WIEN-WAHL 2020: Facebook pages within scope and affiliated party", subtitle="", body = rt_pages, caption=my_caption)
+#table.end
 
@@
 
+#table.start
 fn_reactable_headings(title="WIEN-WAHL 2020: All Facebook ads",
@@
                       caption=my_caption)
+#table.end
 ```
@@
 ```{r echo=FALSE}
+#table.start
 fn_reactable_headings(title="WIEN-WAHL 2020: Number and total amount spent by party",
@@
                       caption=my_caption)
-
+#table.end
 ```
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
+#table.start
 fn_reactable_headings(title="Impressions",
@@
                       caption = my_caption)
+#table.end
 ```
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
+#table.start
 fn_reactable_headings(title="WIEN-WAHL 2020: Accounts which placed ads with a median female audience share of >50%", subtitle="", body = rt_female, caption=my_caption)
-
+#table.end
 ```

### 2021-02-22 _post/2020-12-22-submissionsepidemiclaw/submissionsepidemiclaw.Rmd
modified post.json

@@
 ```{r echo=FALSE, paged.print=TRUE}
+#table.start
 reactable(df_submission, 
@@
                   backgroundColor = plot_bg_color)))
+#table.end
 ```
@@
 ```{r echo=FALSE}
+#table.start
 tb_text
+#table.end
 ```
@@
 ```{r echo=FALSE}
+#table.start
 reactable(df_collocs,
@@
 ```{r echo=FALSE}
+#table.start
 reactable(head(df_similarity, n = 10^3),
@@
                   backgroundColor = plot_bg_color)))
+#table.end
 

### 2021-02-22 _post/2021-01-22-covidconsultationduration/covidconsultationduration.Rmd
modified post.json

@@
   labs(title="Distribution of length of consultative periods.",
-       subtitle="",
+       subtitle="Red dot indicates median value.",
        x="legislative period",
@@
 
-```{r ref.label="chunk_reactable", code_folding=NULL, eval=F, echo=F}
-
-```
-
 ```{r, layout="l-body-outset", code_folding=NULL, echo=F}
+#table.start
 tb_consolidated
+#table.end
 ```
@@
 ```{r code_folding=NULL, include=T, echo=F, layout="l-body-outset"}
+#table.start
 df_consolidated %>% 
@@
               })
+#table.end
 ```
@@
 
+#table.start
 df_dates_missing %>% 
@@
             theme=rt_theme)
-
+#table.end
 ```
@@
 ```{r, eval=T, code_folding=NULL, include=T, layout="l-body-outset"}
+
+#table.start
 df_duration_collected %>% 
@@
             theme=rt_theme)
+#table.end
 ```

### 2021-02-23 _post/2020-02-07-bih-voting-behavior-of-mps-in-house-of-representatives/bih-voting-behavior-of-mps-in-house-of-representatives.Rmd
subtitles added

@@
 description: |
-  A short description of the post.
+  Creating and analysing a new dataset on the voting behavior of BiH's members of the House of Representatives.
 date: 02-07-2020

### 2021-02-23 _post/2020-07-21-attransperency/attransperency.Rmd
subtitles added

@@
 description: |
-  A short description of the post.
+  Analysis of Austria MP's transparency reports.
 
@@
 
-# preview: austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/preview.png
+preview: attransperency_files/preview.png #speeling errors
 

### 2021-02-23 _post/2021-01-22-covidconsultationduration/covidconsultationduration.Rmd
subtitles added

@@
   filter(duration_collected<21) %>% 
-  filter(duration_collected>-1) %>% 
+  filter(duration_collected>0) %>% 
   filter(pdf_pages_sum<11) %>% 

### 2021-02-25 _post/2021-01-22-covidconsultationduration/covidconsultationduration.Rmd
misssing para in covid consultation post
typo in heading

@@
 ---
-title: "Austria's Covid legislation: Duration of parliamentarian consultation period in comparision."
+title: "Austria's Covid legislation: Duration of parliamentarian consultation period in comparison."
 description: |
@@
 
-Third,
+Third, apply the function to all links.
 

### 2021-02-25 _post/2021-02-12-thankyoublogdownhellodistill/thankyoublogdownhellodistill.Rmd
typo in heading

@@
   A short description of the post.
-author:
-  - name: Roland Schmidt
-    url: https://werk.statt.codes
+
 date: 02-12-2021
@@
 
+author:
+  - first_name: "Roland"
+    last_name: "Schmidt"
+    url: https://werk.statt.codes
+    orcid_id: 0000-0002-6897-1693
+categories:
+  - distill
+
+#preview: 2020-10-10-vienna-elections-2020-age-profile/preview.png
 output:
   distill::distill_article:
+    toc: true
+    toc_depth: 3
+    toc_float: true
     self_contained: false
+    code_folding: TRUE
+creative_commons: CC BY-NC-SA
+editor_options: 
+  chunk_output_type: console
+
 ---
@@
 
-So, I moved my collection of posts ('blog' sounds a bit to serious) from blogdown to distill. 
-In principle, setting up the new blog (damn...) was quite straighforward, particularly thanks to the
-comprehensive documentation on the distill site and posts by Thomas, XXXX and XXXX.
+So, I moved my collection of posts from blogdown to distill.
+What pulled me to the latter was the perspective of a simpler, yet effective means to put posts on the web.
+Blogdown undoubtedly can be a very powerful tool and the web is full of visually very appealing blog (just see here) and the speed of its underlying Hugo language seems unrivaled.
 
-Nevertheless, a few quick remarks which might be helpful to those who are also considerng to move:
+Yet, after using it for about a year or so, I found the lack of/the required work to get some features running too cumbersome. Hence, the move to distill.
+
+Among the helpful resources which made the transition rather smooth, I'd like to particularly highlight the documents on RStudio's distill [site](https://rstudio.github.io/distill/){target="_blank"}, [Thomas Mock's](https://themockup.blog/posts/2020-08-01-building-a-blog-with-distill/){target="_blank"} and [Lisa Lendway's](https://lisalendway.netlify.app/posts/2020-12-09-buildingdistill/){target="_blank"} related posts.
+I also found it helpful to have a look at the code underlying the blogs of [Miles McBain](https://milesmcbain.xyz/){target="_blank"} or others who have made their repositories publicly available on github.
+And last, but certainly not least, the authors and maintainers of the package who were overwhelmingly quite quick to respond to issues I had posted on [github](https://github.com/rstudio/distill){target="_blank"}.
+So essentially, I don't have much too add, except a few details which I thought might be helpful to those blogdown users considering the switch.
+
+## Search function
+Distill comes with a search function which can be activated by setting the pertaining option to true in the site's yml file. Two points here:
+1) The search field will become visible only *after having the site deployed*.
+I was wondering for a while what was wrong with with my setup since the search field didn't show up.
+An alternative to deploying is to test run the site with the search fied is to 'virtually' run the site with the servr package (ht to )
 

### 2021-02-26 _post/2021-02-24-hellodistill/hellodistill.Rmd
post hello distill

---
title: "Thank you {blogdown}, hello {distill}."
description: |
  A few notes on moving from blogdown to distill.

date: 02-26-2021

draft: false

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693

categories:
  - distill
  - blogdown

preview: hellodistill_files/distill_logo_small.jpg
# preview: hellodistill_files/blue.jfif

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
    code_folding: false

creative_commons: CC BY-NC-SA

editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(tidyverse)
```

So, I moved my collection of posts from blogdown to distill. What pulled me to the latter was the perspective of a simpler, yet effective means to put posts on the web. Blogdown undoubtedly can be a very powerful tool, the web is full of visually very appealing blogs and the speed of its underlying Hugo language seems unrivaled.

<aside>
```{r, echo=F, preview=F}
knitr::include_graphics("hellodistill_files/distill_logo_small.png")
```
</aside>

Yet, after using it for about a year or so, I found the lack of/the required work to get some features too cumbersome or at least to intimidating. Hence, the move to distill.

Among the helpful resources which made the transition rather smooth, I'd like to particularly highlight the documents on RStudio's distill [site](https://rstudio.github.io/distill/){target="_blank"}, [Thomas Mock's](https://themockup.blog/posts/2020-08-01-building-a-blog-with-distill/){target="_blank"} and [Lisa Lendway's](https://lisalendway.netlify.app/posts/2020-12-09-buildingdistill/){target="_blank"} related posts. I also found it helpful to have a look at the code underlying the blogs of [Miles McBain](https://milesmcbain.xyz/){target="_blank"} or others who have made their repositories publicly available on github. And last but certainly not least, the authors and maintainers of the package were overwhelmingly quick to respond to issues I had posted on [github](https://github.com/rstudio/distill){target="_blank"}. So essentially, I don't have much too add, except a few details which I thought might be helpful to those users considering the switch.

# Search function
Distill comes with a search function which can be [activated](https://rstudio.github.io/distill/blog.html#site-search){target="_blank"} by setting the pertaining option to true in the site's yml file. Two points here:
1) The *search field will become visible only after having the site deployed*. I was wondering for a while what was wrong with my setup since the search field didn't show up before pushing the site. An alternative to deploying is to test-run the site with the search field is to 'virtually' run the site with the `servr` package (ty to [jjallaire](https://github.com/rstudio/distill/issues/264){target="_blank"}).

Another issue I encountered pertains to the speed of the search function. As far as I can tell, the search function 'scans' the `post.json` file which comprises the contents of all posts (it's located in the `_site/post/` folder). Most of my posts contain tables, sometimes comprising a few thousand rows. Also, almost all of my posts contain a graph or two produced with the `ggiraph` package. Both, but particularly the latter, can create very very long entries in the post.json file. The consequence is that the *search function becomes very slow, or as in my case, even freezes due to a bloated `post.json`*.

To overcome this obstacle, I first build the blog, producing the large `post.json` file, and afterwards run a script removing everything that is between the `<svg>` and `</svg>` tags. Plots created by `ggiraph` are svg files.

```{r, eval=F}
library(tidyverse)
library(jsonlite)

post <- readr::read_file(file=here::here("_site", "post", "post.json"))
post <- str_remove_all(post, regex("<svg.*?\\\\/svg>", dotall=T))
readr::write_file(post, file=here::here("_site", "post", "post.json"))
```

And only *afterwards* I push the blog to netlify via github. The drawback that this content is missing from the search results is in my opinion negligible, since the code which produces the svg is still there (and can be searched for in the unlikely case users are searching code snippets).

# Speed
An issue which might be not bother everyone is the blog's performance, i.e. how quickly it is loaded in the browser. Moving from blogdown I was curious to see whether there is actually a difference and used Google's [Pagespeed](https://developers.google.com/speed/pagespeed/insights/){target="_blank"} service to get a clearer idea. Unfortunately, the result is imho somewhat sobering.

While the blog featured a value of 96 on blogdown, it now has 'only' 85 on distill. For mobile devices the value is even only 61. I am by no  means an expert on site performances etc, but with longer and data heavy posts I can definitely see the difference. I fear, the speed of blogdown is something I'll miss.

# Location of data sources
When starting with the distill blog, I wasn't sure where to actually put my heavy data sources. Putting them in to the `_post` folder would mean that they are automatically copied into the `_site` folder every time when building the blog. And more than once, when pushing the site to github I ran subsequently into serious trouble because of exceeding the 100 MB file size limit. The solution, I came up with is to put all data sources into a folder called `_blog_data`. The critical thing here is that folders or files starting with an underscore will not be copied when building the blog. So far this approached worked well for me (as long as I haven't figured out the git command to permanently ignore files larger than 50 MBs.)

# Theming via css
Blogdown offers a myriad of options to theme and style your blog. Maybe too many, for my liking. Distill, on the other hand, comes with a uniform style, but nevertheless offers a few options via the underlying css. Since I turned quite a few of those css-screws (mostly in a trial and error manner), here the [link](https://github.com/werkstattcodes/distill_clean/blob/master/theme.css){target="_blank"} to actual file. Maybe it's helpful for some.

### 2021-03-16 _post/2020-06-08-update-on-mobility-patterns/update-on-mobility-patterns.Rmd
post authorship

@@
   - climate
-  - Corona virus
+  - COVID
   - shiny

### 2021-03-16 _post/2020-11-18-recreating-nyt-election-map/recreating-nyt-election-map.Rmd
post authorship

@@
   - Austria
-  - Corona virus
+  - COVID
   - OCR

### 2021-03-16 _post/2021-03-10-covidsubmissionsnoauthor/covidsubmissionsnoauthor.Rmd
trim index

---
title: "Austria's Covid legislation: The remarkable absence of authorship in public submissions to the consultative process"
description: |
  Austria's legislative procedure provides for a consultative process which allows all citizens to make submissions in which they can raise their concerns as to a proposed bill. When making such a submission, submitters can decide whether their authorship and submission are published on the website of the parliament. Remarkably, the share of undisclosed authorships and statements is particularly high with bills introducing various restrictions to curtail the Covid epidemic.

date: 03-16-2021

author:  
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - COVID
preview: covidsubmissionsnoauthor_files/images/preview.png
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
    code_folding: TRUE

creative_commons: CC BY-NC-SA
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, code_collapse=NULL}
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
  str_trim() %>% 
  str_extract(., regex("^#\\S+"))

theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      plot.margin = margin(l = 0, 
                           t = 0.25,
                           unit = "cm"),
      plot.title = element_text(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 13
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 11
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(
        angle = 0,
        color = "grey50",
        hjust = 1
      ),
      axis.text.x = element_text(
        size = 9,
        color = "grey50"
      ),
      axis.title.y = element_blank(),
      axis.text.y = element_text(
        size = 9,
        color = "grey50"
      ),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 9
      ),
      legend.text = element_text(
        size = 9,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

```

# Context

Here's yet another post (see [here](https://werk.statt.codes/post/2021-01-22-covidconsultationduration/){target="_blank"} and [here](https://werk.statt.codes/post/2020-12-22-submissionsepidemiclaw/){target="_blank"}) on public submissions to Covid-related bills as part of the public legislative consultation process in Austria (*Stellungnahmen im Begutachtungsverfahren*).

Earlier this month, the government introduced another Covid related bill (98/ME XXVII), which triggered again a new record number of filed submissions. This numerical peak in the legislative arena was  paralleled by unprecedented hostile and partly violent demonstrations against the government. As in many countries, Covid demonstrations bring a colorful bunch of individuals to the street, including not only (if there is such a thing as) 'ordinary' citizens who may be exhausted and fearful for their e.g. economic future, but also right wing extremists (e.g. our former minister of interior), and other peculiar characters of highly dubious political shades.

All in all, the unprecedented 'submission activism' and the significant turnout at the demonstrations are suggesting that the opposition to the government's Covid measures is far more than an aberration at the fringes or some freak cases. I think it's not an exaggeration to describe the public discourse as increasingly antagonistic and polarized, even if one acknowledges that submissions as part of a legislative consultative process or participating in a demonstration can be completely legitimate expressions of a democratic discourse.

# Submissions and non-disclosed authorship

However, with opposition to Covid measures increasingly visible, I was struck to notice that the text and authorship of many submissions related to 98/ME XXVII were not published on the parliament's website. The records just state 'Keine öffentliche Stellungnahme' (*No public statement*). While it's nothing unusual to have some submitters who prefer not to be named in public and hence choose the pertaining option when filing their submission, the frequency was remarkable when randomly scrolling through the list for the latest bill (see [here](https://www.parlament.gv.at/PAKT/VHG/XXVII/ME/ME_00098/index.shtml#tab-Stellungnahmen){target="_blank"}; to be clear, the parliament's relevant office has the names of the authors, they are just not published on the website).

<aside>
```{r, echo=F, preview=F, code_folding=NULL}
knitr::include_graphics("covidsubmissionsnoauthor_files/images/no_public_record.png")
```
</aside>

This made me wonder whether my observation was just based on an unrepresentative sample or the reflection of a wider pattern: **Are Covid-related bills featuring more submissions with not publicly disclosed authorship?**

This is just a blog post and not rigorously thought through, but I found the question particularly interesting since it reminded me of reports about right wing voters who either submit wrong or no answers at all when surveyed as to their voting intentions. Since their actual voting preference may be e.g. perceived as controversial, they prefer to keep it to themselves and do not answer truthfully. This is only a very crude description (see e.g. 'Shy Trump voters 2016', Bradley effect etc...), and I am not claiming that the same mechanism is actually at play in the Covid submission case, but I found the parallel intriguing enough to crunch the numbers.

# Code

Unfold the snippet below to get the code for the analysis. I inserted comments where I thought they might be helpful.

```{r, eval=F, code_folding="Code: get overview of all bills"}
# packages ----------------------------------------------------------------
library(tidyverse)
library(rvest)
library(xml2)

# define a function which gives gives a list of all bills

fn_get_bill_overview <- function(rss_feed) {
  df_rss <- xml2::read_xml(rss_feed)
  data <- xml_child(df_rss, 1)

  df_bill_pages_link <- data %>%
    xml2::xml_find_all("//link") %>%
    html_text() %>%
    enframe(.,
      name = "id",
      value = "link_single_bill_page"
    ) %>%
    mutate(link_single_bill_page = str_squish(link_single_bill_page)) %>%
    filter(id > 2) # removes first two rows which don't include data on bills

  # get title
  df_bill_pages_name <- data %>%
    xml2::xml_find_all("//title") %>%
    html_text() %>%
    enframe(.,
      name = "id",
      value = "title"
    ) %>%
    # mutate(name=str_extract(title, regex("(?<=\\>).*(?=\\<)"))) %>%
    mutate(title = str_trim(title)) %>%
    filter(id > 2)

  # get publication date
  df_bill_pages_pub_date <- data %>%
    xml2::xml_find_all("//pubDate") %>%
    html_text() %>%
    enframe(.,
      name = "id",
      value = "date"
    ) %>%
    mutate(date = date %>% str_squish() %>% lubridate::dmy_hms(., tz = "Europe/Vienna"))

  # combine to one dataframe
  df_bill <- bind_cols(
    df_bill_pages_name,
    df_bill_pages_pub_date,
    df_bill_pages_link,
  ) %>%
    select(-contains("id"))

  df_bill %>%
    mutate(legis_period=str_extract(link_single_bill_page, regex("(?<=VHG\\/).*(?=\\/ME\\/)")), .before=1) %>% 
    mutate(indicator_me_numeric = str_extract(link_single_bill_page, regex("(?<=_)\\d*"))) %>%
    mutate(indicator_me = glue::glue("{as.numeric(indicator_me_numeric)}/ME")) %>%
    mutate(link_single_bill_submissions=glue::glue("{link_single_bill_page}#tab-Stellungnahmen")) %>% 
    mutate(rss_single_bill_submissions = glue::glue("https://www.parlament.gv.at/PAKT/VHG/{legis_period}/ME/ME_{indicator_me_numeric}/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP={legis_period}&ITYP=ME&INR={as.numeric(indicator_me_numeric)}&SUCH=&listeId=142&FBEZ=FP_142
")) %>% 
    relocate(link_single_bill_page, .before="link_single_bill_submissions")
}

# create a vector with links to the rss feed of the bills of the different legislative periods (in roman numbers)

periods <- as.roman(seq(14, 27, 1)) 
rss_bills_overview <- glue::glue("https://www.parlament.gv.at/PAKT/MESN/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP={periods}&MESN=ME&R_MESN=ME&MIN=ALLE&SUCH=&listeId=102&FBEZ=FP_002")

# apply the function
df_bills_all <- rss_bills_overview %>%
  purrr::set_names() %>%
  map_dfr(., fn_get_bill_overview, .id = "rss_bills_overview") %>% 
  arrange(date) %>% 
  distinct() #some duplicates
```

```{r, eval=F, code_folding="Code: get data and analyze", include=F}
# some tests

#some links are duplicates; remove
test_rss <- df_bills_all %>% 
  group_by(link_single_bill_page) %>% 
  mutate(n_rss_links=n())

df_bills_all %>% 
  group_by(legis_period) %>% 
  summarise(n_obs=n()) %>% 
  mutate(legis_period_num=as.roman(legis_period) %>% as.integer()) %>% 
  arrange(legis_period_num)

# readr::write_csv2(df_bills_all, file = here::here(
#   "_blog_data",
#   "2021_3_covid_authorship",
#   glue::glue("{Sys.Date()}-all_bills.csv")
# ))
```

```{r, eval=F, code_folding="Code: get data on submissions for all bills"}
# define function extracting list of submissions for a single bill

fn_get_submission_details <- function(link) {

  data <- xml2::read_xml(link)

  # get link to subpages with link to submissions
  df_submission_pages_link <- data %>%
    xml2::xml_find_all("//link") %>%
    html_text() %>%
    enframe(.,
      name = "id",
      value = "link_single_submission_page"
    ) %>%
    mutate(link_single_submission_page = str_squish(link_single_submission_page)) %>%
    filter(id > 2) # removes first two rows which don't include data on submissions

  # get title
  df_submission_pages_name <- data %>%
    xml2::xml_find_all("//title") %>%
    html_text() %>%
    enframe(.,
      name = "id",
      value = "title"
    ) %>%
    mutate(name = str_extract(title, regex("(?<=\\>).*(?=\\<)"))) %>%
    filter(id > 2) %>%
    select(-title)

  # get publication date
  df_submission_pages_pub_date <- data %>%
    xml2::xml_find_all("//pubDate") %>%
    html_text() %>%
    enframe(.,
      name = "id",
      value = "date"
    ) %>%
    mutate(date = date %>% str_squish() )#%>% lubridate::dmy_hms(., tz = "Europe/Vienna"))

  # combine to one dataframe
  df_submission <- bind_cols(
    df_submission_pages_name,
    df_submission_pages_pub_date,
    df_submission_pages_link,
  ) %>%
    select(-contains("id"))

  # if(nrow(df_submission)==0) {
  #   
  #   df_submission <- tibble(name=NA_character_, 
  #                           date=NA_character_, 
  #                           link_single_submission_page=NA_character_)
  # }

  df_submission
}

tbl_missing <- tibble(name=NA_character_, 
                      date=NA_character_, 
                      link_single_submission_page=NA_character_)

# apply function
df_all_submissions <- df_bills_all$rss_single_bill_submissions %>% 
  purrr::set_names() %>% 
  map_dfr(., possibly(fn_get_submission_details,
                      otherwise = tbl_missing),
          .id="rss_single_bill_submissions")
```

```{r, eval=F, code_folding="Code: get data and analyze", include=F}
# > test ------------------------------------------------------------------
length(unique(df_all_submissions$rss_single_bill_submissions)) #3769; 31 missing

df_missing_bills_submissions <- df_bills_all %>% 
#  select(rss_single_bill_submissions) %>% 
  filter(!rss_single_bill_submissions %in% unique(df_all_submissions$rss_single_bill_submissions))
nrow(df_missing_bills_submissions)
#function doesn't trigger error in these cases; if to be included
#include if clause on nrow==0

df_test <- df_missing_bills_submissions$rss_single_bill_submissions %>% 
  purrr::set_names() %>% 
  map_dfr(., fn_get_submission_details, .id="rss_single_bill_submissions")

# > save ------------------------------------------------------------------

write_csv2(df_all_submissions, 
           file=here::here("_blog_data",
                           "2021_3_covid_authorship",
                           "df_all_submissions.csv"))

# load data ---------------------------------------------------------------

# df_bills_all <- readr::read_csv2(file=here::here("data", "2021-03-10-all_bills.csv"))
# df_all_submissions <- readr::read_csv2(file=here::here("data", "df_all_submissions.csv"))
# head(df_all_submissions$rss_single_bill_submissions)
```

```{r, eval=F, code_folding="Code: analysis", include=T}
df_bill_submissions <- df_bills_all %>% 
  left_join(., df_all_submissions %>% 
              rename(date_submission=date),
            by="rss_single_bill_submissions") %>% 
  mutate(covid_bill=str_detect(title, regex("Covid|Epidemie", ignore_case = T))) %>% 
  mutate(public_author=!str_detect(name, regex("Keine öffentliche Stellungnahme"))) %>% 
  mutate(bill_id=glue::glue("{legis_period}-{indicator_me}"))

df_submissions_per_bill <- df_bill_submissions%>% 
  group_by(bill_id, date, title, covid_bill, public_author) %>% 
  summarise(submissions_n=n()) %>% 
  pivot_wider(., id_cols=c(bill_id, date, title, covid_bill),
              names_from=public_author,
              names_glue="public_author_{public_author}",
              values_from=submissions_n,
              values_fill = 0) %>% 
  mutate(submissions_total=public_author_TRUE+public_author_FALSE) %>% 
  mutate(non_public_author_share=public_author_FALSE/submissions_total)

# write_csv2(df_submissions_per_bill, file=here::here("_blog_data",
#                                                     "2021_3_covid_authorship",
#                                                     "df_submissions_per_bill.csv"))
```

```{r, eval=T, include=F}
df_submissions_per_bill <- readr::read_csv2(file=here::here("_blog_data",
                                                            "2021_3_covid_authorship",
                                                            "df_submissions_per_bill.csv"))
```

# Result

I guess the plot below speaks (hopefully) for itself: Those bills with Covid relation (in orange) not only feature much more filed submissions (size of the dots), they also feature much more submissions without the author and the text of the submission being publicly disclosed (location on the x-axis). The difference to submissions related to other bills is quite remarkable.

```{r, code_folding="Code: plot"}
library(ggiraph)
library(ggtext)

color_no <- "deepskyblue4"
color_yes <- "orange"

caption <- glue::glue("<span style='font-size:10pt; text-align: justify;
'>Seit September 2017 können alle, auch von einem Gesetzesvorschlag nicht unmittelbar betroffene Bürger<br>und Institutionen Stellungnahmen zu diesem im Rahmen des erweiterten Begutachtungsverfahrens (auch<br>online) einreichen. Es bleibt den Einreichenden überlassen, ob Name sowie Stellungnahme auf der Homepage<br>des Parlaments veröffentlicht wird. Die Graphik oben zeigt, dass der Anteil nicht-veröffentlichter Namen und<br>Stellungnahmen bei Einreichungen zur Begutachtung von Gesetzesvorschlägen mit COVID-Bezug außer-<br>gewöhnlich hoch war.</span><br><br>Data: https://www&#46;parlament&#46;gv&#46;at; {df_submissions_per_bill %>% 
  filter(date>as.Date('2017-09-01')) %>% nrow() %>% map_chr(., scales::label_comma())} Gesetzesvorschläge, {df_submissions_per_bill %>% filter(date>as.Date('2017-09-01')) %>% pull(submissions_total) %>% sum() %>% map_chr(., scales::label_comma())} Stellungnahmen. <br>Analysis: Roland Schmidt | @zoowalk | https://werk&#46;statt&#46;codes")

pl_public<- df_submissions_per_bill %>% 
  filter(date>as.Date("2017-09-01")) %>% #opening of consultation process to general public
  mutate(covid_bill_facet=case_when(covid_bill=="FALSE" ~ glue::glue("Stellungnahmen zu<br>Gesetzesvorschlägen <span style='color:{color_no}; font-size:12pt;'>ohne<br></span> COVID-Maßnahmen Bezug "),
                                    covid_bill=="TRUE" ~ glue::glue("Stellungnahmen zu<br>Gesetzesvorschlägen <span style='color:{color_yes}; font-size:12pt;'>mit<br></span> COVID-Maßnahmen Bezug"))) %>% 
  ggplot()+
  labs(y="Anteil nicht-veröffentlichter Stellungnahmen",
       x="Covid19 related bill",
       title=glue::glue("BEGUTACHTUNGSVERFAHREN: Anteil nicht-veröffentlichter Einreicher-Namen<br>bei Stellungnahmen zu Gesetzesvorschlägen <span style='color:{color_yes}'>mit</span> und <span style='color:{color_no}'>ohne</span> COVID-Maßnahmen<br>Bezug"),
   #    subtitle="",
       caption=caption)+
  geom_jitter_interactive(
    aes(
      x=covid_bill_facet,
      y=non_public_author_share,
      color=covid_bill,
      size=submissions_total,
      tooltip=glue::glue("{date}
                         {bill_id}
                         {title}
                         Submissions total: {submissions_total}
                         Submissions not public: {scales::percent(non_public_author_share)}
                         ")),
    width=0.3)+
  scale_y_continuous(labels = scales::label_percent(),
                     expand=expansion(mult=c(0,0.03)))+
  # scale_x_discrete(expand=expansion(mult=c(0.1,0)))+
  scale_color_manual(guide=NULL,
                       values=c("FALSE"=color_no, 
                                "TRUE"=color_yes))+
  scale_size_continuous(name="Anzahl Stellungnahmen gesamt:",
                        labels=scales::label_comma())+
  coord_flip()+
  theme_post()+
  theme(legend.direction = "horizontal",
        legend.position = c(1,1),
        legend.justification = c(1,0),
        legend.text = element_text(color="grey50"),
        legend.title = element_text(color="grey50",
                                    face = "plain",
                                    vjust=0.5),
        strip.text.y.left = element_markdown(hjust=.5,
                                             vjust=.5,
                                             angle = 0),
        strip.text = element_markdown(),
        axis.text.y = element_markdown(),
        plot.title = element_markdown(margin = margin(b=.5, unit="cm")),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(),
        plot.caption = element_markdown(),
        plot.caption.position = "plot")+
    guides(size=guide_legend(override.aes = list(color="grey50")))

```

```{r, echo=F, code_folding=NULL}
girafe(ggobj = pl_public,
       height_svg = 4.5,
       options = list(
    opts_toolbar(saveaspng = FALSE),
    opts_tooltip(css = glue::glue("background-color:{plot_bg_color}; 
                 line-height:100%;
                 color:black;
                 font-size:80%;
                 font-family:'Roboto Condensed';)"))
  ))
```

### 2021-03-16 _post/2021-03-10-covidsubmissionsnoauthor/covidsubmissionsnoauthor.Rmd
heading graph authorship; line break on mobile

@@
 
-```{r, echo=F, code_folding=NULL}
+```{r, echo=F, code_folding=NULL, dev="svglite"}
 girafe(ggobj = pl_public,

### 2021-03-16 _post/2021-03-10-covidsubmissionsnoauthor/covidsubmissionsnoauthor.Rmd
ggiraph ggtext incomp? svglite?
add font to ggiraph

--- 2021-03-16 _post/2021-03-10-covidsubmissionsnoauthor/covidsubmissionsnoauthor.Rmd
@@
 library(tidyverse)
+library(svglite)
 
@@
 theme_post <- function() {
-  hrbrthemes::theme_ipsum_rc() +
+  hrbrthemes::theme_ipsum_rc(base_family="RobotoCondensed") +
     theme(
@@
 
-```{r, echo=F, code_folding=NULL}
+```{r, echo=F, code_folding=NULL, dev="svglite"}
 girafe(ggobj = pl_public,
        height_svg = 4.5,
+       font=list(sans="RobotoCondensed"),
        options = list(

### 2021-03-16 _post/2021-03-10-covidsubmissionsnoauthor/covidsubmissionsnoauthor.Rmd
font
revert to original font

--- 2021-03-16 _post/2021-03-10-covidsubmissionsnoauthor/covidsubmissionsnoauthor.Rmd
@@
 library(tidyverse)
+library(svglite)
+
+library(hrbrthemes)
 
@@
 
-```{r, echo=F, code_folding=NULL}
+```{r, echo=F, code_folding=NULL, dev="svglite"}
 girafe(ggobj = pl_public,
        height_svg = 4.5,
+  #     font=list(sans="Roboto"),
        options = list(

### 2021-08-19 _post/2021-08-19-reactablefmtr-and-title-size-issue/reactablefmtr-and-title-size-issue.Rmd
adding reactablefrmtr test

---
title: "reactablefmtr and title size issue?"
description: |
  A short description of the post.
author:
  - name: Roland Schmidt
    url: https://werk.statt.codes
date: 08-19-2021
output:
  distill::distill_article:
    self_contained: false
---

This is quick test of how to change the font size in tables created by `reactable` and formated with the `reactablefrmtr` package.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
library(reactable)
library(reactablefmtr)
```

# table without font size defined

```{r}
mtcars %>% 
  reactable() %>% 
  add_title(title="this is a title without font-size defined")
```

# table with font size 20

```{r}
mtcars %>% 
  reactable() %>% 
  add_title(title="this is a title in font 20",
            font_size = 20)
```

# table with font size 5

```{r}
mtcars %>% 
  reactable() %>% 
  add_title(title="this is a title in font 20",
            font_size = 5)
```

### 2021-08-20 _post/2021-08-19-reactablefmtr-and-title-size-issue/reactablefmtr-and-title-size-issue.Rmd
add frmtr post

@@
 author:  
-  - name: Roland Schmidt
+  - first_name: "Roland"
+    last_name: "Schmidt"
     url: https://werk.statt.codes
+    orcid_id: 0000-0002-6897-1693
 date: 08-19-2021
@@
   distill::distill_article:
+    toc: true
+    toc_depth: 3
+    toc_float: true
     self_contained: false
+    code_folding: TRUE
 ---

### 2021-10-07 _post/2021-07-18-the-ibizia-inquiry-in-numbers/the-ibizia-inquiry-in-numbers.Rmd
adding ibizia u post

---
title: "Extracting the transcripts of the 'Ibizia Inquiry'"
description: In mid-June, the so-called 'Ibiza Commission of Inquiry' heard its last witness after almost one year at work. The commission had been established by the Austrian Parliament and set up to investigate the prevalence of corruption during the coalition government of the ÖVP and FPÖ (Dec, 2017- May, 2019). This post diggs into its published transcripts. The emphasis of the post is, first, on demonstrating how to extract the text from the hearings' transcripts in R and, second, on crunching a few numbers to get some substantive insights. The compiled dataset, covering all statements (incl. questions, answers etc.), is available for download.

date: 10-07-2021

author:  
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - Parliament
  - web scraping
preview: the-ibizia-inquiry-in-numbers_files/images/subpage_link_to_transcript.PNG
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
    code_folding: TRUE
draft: FALSE
creative_commons: CC BY-NC-SA
editor_options: 
  chunk_output_type: console
---

# Context

In mid-June, the Austrian Parliament's *'Commission of Inquiry concerning the alleged corruptibility of the turquois-blue Federal Government'*[^1] heard its last respondent (*'Auskunftsperson'*). More informally, the commission is simply called 'Ibiza inquiry', named after the location of a secretly taped video which showed high ranking members of the extreme right FPÖ party speaking freely - liberated by alcohol - with a fake niece of a Russian oligarch about actual or intended corruption in Austria's political system. As a consequence of the video's release, the then ruling coalition government of ÖVP ('turquoise') and FPÖ ('blue') collapsed and the commission was set up. In short, a pretty wild and bewildering story.

[^1]: ['Untersuchungsausschuss betreffend mutmaßlicher Käuflichkeit der türkis-blauen Bundesregierung'](https://www.parlament.gv.at/PAKT/VHG/XXVII/A-USA/A-USA_00002_00906/index.shtml){target="_blank"}

The commission, empowered with a fairly broad mandate and armed with access to plenty of WhatsApp etc messages, offered an unprecedented view into the inner-working of Chancellor Kurz's first government and the mindset of some of its protagonists. While the initial impetus for setting up the commission was first and foremost the Ibiza video featuring the FPÖ leadership, the inquiry's focus gradually shifted (not least due to the opposition's efforts) to the wheeling and dealing of Chancellor Sebastian Kurz's ÖVP and its affiliates.

Having said this, the purpose of this post is not to give a recap of the inquiry or its results, but first and foremost procedural in the sense that it intends to detail the necessary steps to extract statements from the inquiry's transcripts and subsequently obtain some exemplary insights with R. As always, if you spot any error, feel free to contact me, best via [twitter DM](https://twitter.com/zoowalk){target="_blank"}. And if you use any of the work provided here, grateful if you acknowledge my input.

If you are not interested in any of the coding steps generating the data, jump directly to the [Analysis] section which still contains plenty of code, but you're not burdened with how to obtain the data in the first place.

# Getting the data

In this section I will layout the necessary steps to obtain the relevant data. I'll first detail each step with one single, sample session of the inquiry commission. Subsequently, I'll apply the demonstrated steps to all sessions by means of one general function.

But before, let's load the packages we'll need along the way and define some templates.

```{r, code_folding=F}
# libraries ---------------------------------------------------------------
#load the required libraries
library(tidyverse)
library(rvest)
library(xml2)
library(fuzzyjoin)
library(reactable)
library(reactablefmtr)
library(hrbrthemes)

#define party colors
vec_party_col <- c(
  "ÖVP"="#5DC2CC",
  "SPÖ"="#FC0204", 
  "Grüne"="#A3C630",
  "FPÖ"="#005DA8", 
  "NEOS"="#EA5290"
  ) 
```

```{r, include=F}
# theme for reactables
options(reactable.theme=reactableTheme(
  style=list(fontSize="12px",
             lineHeight="14px",
             fontFamily="Roboto condensed",
             compact=T,
             bordered = T,
             highlight = T)))

plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
  str_trim() %>% 
  str_extract(., regex("^#\\S+"))

#ggplot theme
theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      plot.margin = margin(l = 0, 
                           t = 0.25,
                           unit = "cm"),
      plot.title = element_text(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 13
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 11
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(
        angle = 0,
        color = "grey50",
        hjust = 1
      ),
      axis.text.x = element_text(
        size = 9,
        color = "grey50"
      ),
      axis.title.y = element_blank(),
      axis.text.y = element_text(
        size = 9,
        color = "grey50"
      ),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 9
      ),
      legend.text = element_text(
        size = 9,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

```

## Getting links to transcripts

In a first step, let's get the links leading to the transcripts of the sessions. This [link](https://www.parlament.gv.at/PAKT/VHG/XXVII/A-USA/A-USA_00002_00906/index.shtml#tab-VeroeffentlichungenBerichte){target="_blank"} leads us to the commission's overview page including all published reports, including the sessions' transcripts (*Protokolle*).

![](the-ibizia-inquiry-in-numbers_files/images/overview_links.PNG)

The code below extracts links related to the latter. Explanatory comments are inserted directly into the code chunk.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F}
#link to overview page
site_link <- "https://www.parlament.gv.at/PAKT/VHG/XXVII/A-USA/A-USA_00002_00906/index.shtml#tab-VeroeffentlichungenBerichte"

# get links to pages where links to protocols are located
# link has to include words 'Protokolls' in text;
df_links_to_subpages <- site_link %>%
  rvest::read_html() %>%
  #define a filter to get only the links related to transcripts (protocols)
  # filters links based on text/name of links
  rvest::html_elements(xpath = "//a[contains(text(), 'Protokolls')]") %>%
  html_attr("href") %>%
  # extracts links
  enframe(name = NULL, value = "link") %>%
  #links of interest inlude "KOMM"
  filter(str_detect(link, regex("KOMM"))) %>%
  #complete the link
  mutate(link_to_subpages = paste0("https://www.parlament.gv.at/", link)) %>%
  select(link_to_subpages)
```

Here the first ten links:

```{r, code_folding=NULL, echo=F}
head(df_links_to_subpages, n=10) %>% 
  reactable()
```

Each of these links leads to a subpage which provides details on the record in question and the link to the actual file containing the transcribed text. Below one such [subpage](https://www.parlament.gv.at//PAKT/VHG/XXVII/KOMM/KOMM_00271/index.shtml){target="_blank"}.

![](the-ibizia-inquiry-in-numbers_files/images/subpage_link_to_transcript.PNG)

Note the link leading to the HTML version of the transcript. To access the transcript we need the link's target address. The function below extracts the link leading to actual text. Subsequently, the function is applied to all subpage links which we obtained in the previous step.

```{r, code_folding=F}
# function to extract link to protocol from details page

fn_get_link_to_record <- function(link_to_subpage) {
  link_to_subpage %>%
    rvest::read_html() %>%
    rvest::html_elements("a") %>%
    html_attr("href") %>%
    enframe(
      name = NULL,
      value = "link"
    ) %>%
    #link to transcript contains "fnameorig"
    filter(str_detect(link, regex("fnameorig"))) %>%
    #complete link
    mutate(link_to_record = paste0("https://www.parlament.gv.at/", link)) %>%
    select(link_to_record)
}

library(furrr)
plan(multisession, workers = 2)

# apply function to all links
df_links_to_records <- df_links_to_subpages %>%
  pull(link_to_subpages) %>%
  purrr::set_names() %>%
  future_map_dfr(., fn_get_link_to_record, .id = "link_to_subpages")

```

What we obtain is a dataframe including the links which lead to *all* transcripts of the inquiry (only first 5 are shown).

```{r, echo=F, code_folding=NULL}
reactable::reactable(df_links_to_records %>% 
                       slice_head(., n=5))
```

## Extracting text

Now with the links to the transcripts available, let's have a look at one such text, e.g. [here](https://www.parlament.gv.at/PAKT/VHG/XXVII/KOMM/KOMM_00271/fnameorig_990146.html){target="_blank"}.

![](the-ibizia-inquiry-in-numbers_files/images/transcript.PNG){width="2703"}

Importantly, notice that statements given before the inquiry commission are always introduced with the *speaker's name (and position) in bold and underlined letters*. This (almost) consistently applied formatting will eventually allow us to distinguish between the actual statement and its speaker, and the start/end of different statements. I'll first extract these names and subsequently assign these names to their respective statements.

### Extract speakers

```{r, include=F}
    link_to_record <- "https://www.parlament.gv.at//PAKT/VHG/XXVII/KOMM/KOMM_00271/fnameorig_990146.html"

#test link
  # link_to_record <- "https://www.parlament.gv.at/PAKT/VHG/XXVII/KOMM/KOMM_00271/fnameorig_990146.html"

```

To extract the speakers from the text I'll use once again the powerfull `rvest` package. To identify those text parts which are in bold and underlined, its `html_elements` function is used. As for the resort to `xml_contents` I am grateful for the answer to [this](https://stackoverflow.com/questions/38195120/how-can-i-retain-format-in-a-page-webscraped-with-rvest/38195693){target="_blank"} Stackoverflow question.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T}
# get those elements which are bold and underlined
df_speakers <- link_to_record %>%
    rvest::read_html() %>%
    #extract elements which are bold and underlined; note that the sequence has to be that of the html tags 
    rvest::html_elements("b u") %>%
    map(., xml_contents) %>%
    map(., html_text) %>%
    enframe(value = "speaker") %>%
    mutate(speaker = str_trim(speaker, side = c("both"))) %>%
    mutate(speaker = str_squish(speaker)) %>%
    # keep only those elements which end with colon;
    # filter(str_detect(speaker, regex("\\:$"))) %>%
    #remove colon at end; needed to unify names of speakers where some instances end/do not end with colon
    mutate(speaker=str_remove(speaker, regex("\\:$"))) %>% 
    filter(str_count(speaker, regex("\\S+"))>1) %>% 
    #removes heading of transcript which is also bold and underlined
    filter(!str_detect(speaker, regex("^Befragung der"))) %>% 
    distinct(speaker)
```

Here's the result for our one our sample session.

```{r, echo=F, code_folding=NULL}
df_speakers
```

As you can see, the approach worked rather well, but not 100 % perfect. There are a few rows which contain incomplete names of speakers or some related fragments (Sobotka without a). These unwanted results are - as far as I can tell - due to some inconsistent formatting of speakers' names (even if their appearance in the document is identical) or editorial errors (e.g. the position of a speaker is mentioned in one instance, but not in another, e.g. Abgeordneter mag. Klaus Führlinger). These glitches have to be corrected 'manually'. The code chunk below does this for our specific sample case.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=F}
# introduce corrections to errors in transcript for one test link; not published link
  df_speakers <- df_speakers %>%
    #remove fragments; speaker has to comprise at least two words
    filter(str_count(speaker, regex("\\S+"))>1) %>% 
    mutate(speaker = case_when(
      str_detect(speaker, regex("Fürlinger")) ~ "Abgeordneter Mag. Klaus Fürlinger (ÖVP):",
      str_detect(speaker, regex("^Nina Tomaselli")) ~ "Abgeordnete Mag. Nina Tomaselli (Grüne):",
      str_detect(speaker, regex("Hanger")) ~ "Abgeordneter Mag. Andreas Hanger (ÖVP):",
      str_detect(speaker, regex("Gödl")) ~ "Abgeordneter Mag. Erst Gödl (ÖVP):",
      TRUE ~ as.character(speaker))) %>%
    # remove colon at end; some are with, some without
    mutate(speaker = str_remove(speaker, regex("\\:$"))) %>%
    distinct(speaker)

#correction: remove entries where names of deputies were extracted twice since
#party membership was missing in at least one instance
df_speaker <- df_speakers %>% 
  filter(!str_detect(speaker, regex("^Abge.*[^\\)]$")))

```

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=F}
# introduce corrections to errors in transcript
  df_speakers <- df_speakers %>%
    #remove fragments; speaker has to comprise at least two words
    filter(str_count(speaker, regex("\\S+"))>1) %>% 
    mutate(speaker = case_when(
      str_detect(speaker, regex("Fürlinger")) ~ "Abgeordneter Mag. Klaus Fürlinger (ÖVP):",
      str_detect(speaker, regex("^Vorsitzender Mag\\. Wolfgang Sobot")) ~ "Vorsitzender Mag. Wolfgang Sobotka:",
      TRUE ~ as.character(speaker))) %>%
    # remove colon at end; some are with, some without
    mutate(speaker = str_remove(speaker, regex("\\:$"))) %>%
    distinct(speaker)

#correction: remove entries where names of deputies were extracted twice since
#party membership was missing in at least one instance
df_speakers <- df_speakers %>% 
  filter(!str_detect(speaker, regex("^Abge.*[^\\)]$")))

```

After these modifications we get a clean dataframe of those who made statements during the session in question (our sample link).

```{r, echo=F, code_folding=NULL}
df_speakers
```

In a later step, we will search the entire transcript of the session for the presence of these speakers' names (incl. position and title) to identify the start of a statement. Since this pattern matching will require regular expressions (*regex*), the names have to be modified accordingly (e.g. a dot . has to be escaped and becomes `\\.`, for further info see [here](https://stringr.tidyverse.org/articles/regular-expressions.html#escaping-1){target="_blank"}).

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T}
# create regex
  df_speakers <- df_speakers %>%
    mutate(
      speaker_pattern =
        str_replace_all(speaker, "\\.", "\\\\.") %>%
          str_replace_all(., "\\:", "\\\\:") %>%
          # str_replace_all(., "\\,", "\\\\,") %>%
          str_replace_all(., "\\)", "\\\\)") %>%
          str_replace_all(., "\\(", "\\\\(") %>% 
        #match has to be at start of string; avoids mismatches where name appears in middle of text
        paste0("^",.))
```

### Identify and extract statements

Now, let's get the entire text of a transcript. Again, I use the `rvest` package, this time, however, the hmtl-tag `<p>` is targeted.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=T}
## Extract entire text from transcript

df_text <- link_to_record %>%
    rvest::read_html() %>%
    rvest::html_elements("p") %>%
    rvest::html_text() %>%
    enframe(., 
            value = "text",
            name="row") %>%
    mutate(text = str_squish(text) %>%
      str_trim(., side = c("both")))
```

```{r, paged.print=TRUE, echo=F, code_folding=NULL}
df_text %>% reactable::reactable(.,
                                 columns = list(
                                   row=colDef(width=50),
                                   text=colDef(width = 600)))
```

The result is a dataframe with one row for each line in the transcript. The challenge is now to identify those rows where a speaker's statement starts and ends/the next speaker's statement begins. I'll do this by - as already mentioned above - pattern matching the names of the retrieved speakers with the content of each row. In simpler terms, does the row start with the name of a speaker which we previously identified by their bold and underlined format? To do this I'll make use of `fuzzyjoins::regex_left_join`.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=T}
# get speaker name from text
  df_text_2 <- df_text %>%
    #search for match only in opening section of line
  #  mutate(text_start=stringr::str_sub(text, start=1, end=40)) %>% 
    fuzzyjoin::regex_left_join(.,
      df_speakers,
      by = c("text" = "speaker_pattern")
    ) # speaker_pattern
```

```{r, include=F}
dupes <- df_text_2 %>% 
  count(row, sort=T) %>% 
  filter(n>1)

df_dupes <- df_text_2 %>% 
  semi_join(.,
            dupes) %>% 
  reactable()

```

Have a look at the result below. There's now a new column *indicating the start of a statement* with the speaker's name (since transcripts are starting with some introductory text you'll see speakers' names appear only on later pages in the table).

```{r, echo=F, code_folding=NULL}
df_text_2 %>% 
  select(contains("speaker"), everything()) %>% 
  reactable::reactable(.,
                       columns = list(
                                   row=colDef(width=50),
                                   text=colDef(width = 600)),
                       filterable = T)
```

In the next step, I 1) `dplyr::fill` all empty speaker rows after a statement's start with the speaker's name (details [here](https://tidyr.tidyverse.org/reference/fill.html){target="_blank"}), and 2) create a numerical grouping variable which increases each time the speaker is changing (= start of new statement). Remaining rows without a speaker are removed since they contain only additional text and not statements.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=T}
df_text_2 <- df_text_2 %>%
    # fill rows with speaker and pattern as basis for grouping
    fill(speaker, .direction = "down") %>%
    fill(speaker_pattern, .direction = "down") %>%
    filter(!is.na(speaker)) %>%
    # create grouping id; later needed to collapse rows
    mutate(grouping_id = ifelse(speaker == dplyr::lag(speaker, default = "start"),
      0,
      1
    )) %>%
    mutate(grouping_id_cum = cumsum(grouping_id)) %>%
    # remove speaker info from actual spoken text
    mutate(text = str_remove(text, speaker_pattern)) %>%
    # remove colon; this approach keeps annotations which are included between speaker name and colon; e.g "(zur Geschäftsbehandlung)";
    mutate(text = str_remove(text, regex("\\:"))) %>%
    mutate(text = str_trim(text, side = c("both")))
```

```{r, echo=F, code_folding=NULL}
df_text_2 %>% 
   select(contains("grouping"), 
          contains("speaker"), 
          everything()) %>% 
  reactable::reactable(.,
                                 columns = list(
                                   row=colDef(width=50),
                                   text=colDef(width = 600))) 
```

With each row now attributed to a speaker, and each statement assigned with a distinct indicator, we can collapse a statement's multiple lines into one single row with its single speaker.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=T}
df_text_3 <- df_text_2 %>%
    # collapse rows
    group_by(grouping_id_cum) %>%
    summarise(
      text = paste(text, collapse = " "),
      speaker = unique(speaker)
    ) %>%
    relocate(speaker, .before = "text") %>%
    mutate(text = str_trim(text, side = c("both")))
```

```{r, echo=F, code_folding=NULL}
df_text_3 %>% 
   select(contains("grouping"), 
          contains("speaker"), 
          everything()) %>% 
  reactable::reactable(.,
                                 columns = list(
                                   text=colDef(width = 600)))
```

The resulting table already represents most of the required data for the later analysis (for one specific sample respondent). I complement it with additional data on an MP's party affiliation (if applicable); a speaker's position (if stated); and the session's date, duration and number.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=T}
df_text_3 <- df_text_3 %>%
    # extract party of MP; pattern matches only if speaker starts with Abgeordneter
    mutate(speaker_party=str_extract(speaker, regex("(?<=^Abgeordnete[^\\(]{1,40})\\(.*\\)$"))) %>% 
    mutate(speaker_party=str_extract(speaker_party, regex("[:alpha:]+"))) %>% 
  # extract position of speaker
    mutate(speaker_position = case_when(
      str_detect(speaker, regex("^Abgeord")) ~ "Abgeordneter",
      str_detect(speaker, regex("^Verfahrensanwalt")) ~ "Verfahrensanwalt",
      str_detect(speaker, regex("^Vorsitzender-Stellvertreter")) ~ "Vorsitzender-Stellvertreter",
      str_detect(speaker, regex("^Vorsitzende")) ~ "Vorsitzende/r",
      str_detect(speaker, regex("^Verfahrensrichter-Stellvertreter")) ~ "Verfahrensrichter-Stellvertreter",
      str_detect(speaker, regex("^Verfahrensricht")) ~ "Verfahrensrichter",
      str_detect(speaker, regex("^Vertrauensperson")) ~ "Vertrauensperson",
      TRUE ~ as.character("Auskunftsperson")
    ))

```

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=T}
# add session details
# extract date of session
  vec_respondent <- df_text %>%
    filter(str_detect(text, regex("^Befragung der Auskunftsperson"))) %>%
    pull(text) %>%
    str_remove(., "^Befragung der Auskunftsperson ")

  # extract duration of ENTIRE session (not only respondent)
  vec_duration <- df_text %>%
    filter(str_detect(text, regex("^Gesamtdauer der"))) %>%
    pull(text) %>%
    str_extract(., regex("(?<=Sitzung).*"))

  # extract session number
  vec_session_no <- df_text %>%
    filter(str_detect(text, regex("^Gesamtdauer der"))) %>%
    pull(text) %>%
    str_extract(., regex("\\d*(?=\\. Sitzung)"))

  # extract date of session
  vec_date <- df_text %>%
    filter(str_detect(text, regex("^Montag|^Dienstag|^Mittwoch|^Donnerstag|^Freitag|^Samstag|^Sonntag"))) %>%
    pull(text)

df_text_4 <- df_text_3 %>%
    mutate(
      session_date = vec_date,
      session_no = vec_session_no,
      session_duration = vec_duration,
      respondent = vec_respondent
    ) %>%
    select(session_no, session_date, session_duration, respondent, speaker, speaker_position, speaker_party, text, -grouping_id_cum)
```

### Party affiliation of MP asking question(s)

A further detail which we can extract is the party affiliation of an MP who asks a question of the respondent. In other words, each row from a respondent obtains an column indicating the party of the MP who asked the question. To do so, however, implies the assumption that any row/statement by an MP which precedes a row/statement by the respondent is actually a question by the MP. This assumption may not always be warranted. In some cases statements/rows by both an MP and a respondent may e.g. actually refer to an earlier question by the commission's chair and hence not be an interaction between the latter two.

Bearing this qualification in mind, the approach seems promising enough to identify interactions between an MP and a respondent and consider the former's party affiliation.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=F, eval=T, include=T}
# identify party
  df_text_4 <- df_text_4 %>%
    mutate(respondent_questioner_party = case_when(
      str_detect(lag(speaker_position), regex("^Abge")) &
        speaker_position=="Auskunftsperson" ~ lag(speaker_party),
      TRUE ~ NA_character_)) %>%
  # calculate answer length without annotations included in transcripts
    mutate(text_length_old=stringi::stri_count_words(text, locale="de")) %>% 
    mutate(text_length = str_remove_all(text, regex("\\([^\\(\\)]*\\)")) %>% 
             stringi::stri_count_words(., locale="de"))
```

```{r, echo=F, code_folding=NULL}
df_text_4 %>% 
   select(contains("grouping"), 
          contains("speaker"), 
          everything()) %>% 
  mutate(text=str_trunc(text, width = 300, side = "right")) %>% 
  reactable::reactable(., columns = list(
                                   text=colDef(width = 600)))
```

### Function for all sessions

The example above detailed the code and results for only one sample respondent to illustrate each individual step. In the chunk below, I wrap all implemented steps into a function which is subsequently applied to every single link leading to a transcript text. Unfold the code chunk to below to see the entire function.

```{r, cache=F, cache.extra=Sys.Date(), code_folding=T, eval=T}
# Extract data - Function -------------------------------------------------

link_to_record <- "https://www.parlament.gv.at//PAKT/VHG/XXVII/KOMM/KOMM_00271/fnameorig_990146.html"

fn_get_data <- function(link_to_record) {

  # get those elements which are bold and underlined
  df_speakers <- link_to_record %>%
    rvest::read_html() %>%
    rvest::html_elements("b u") %>%
    map(., xml_contents) %>%
    map(., html_text) %>%
    enframe(value = "speaker") %>%
    mutate(speaker = str_trim(speaker, side = c("both"))) %>%
    mutate(speaker = str_squish(speaker)) %>%
    #remove colon at end; needed to unify names of speakers where some instances end/do not end with colon
     mutate(speaker=str_remove(speaker, regex("\\:$"))) %>% 
     #remove fragments shorter than 2 words
     filter(str_count(speaker, regex("\\S+"))>1) %>% 
     #removes heading of transcript which is also bold and underlined
     filter(!str_detect(speaker, regex("^Befragung der"))) %>% 
     distinct(speaker)

  #remove fragmensts of deputy names: starts with Abgeordnete but doesn't end with last
  #parenthesis enclosing party affiliation or colon
  df_speakers <- df_speakers %>%
    filter(!str_detect(speaker, regex("^Abge.*[^\\):]$")))

  # 'manually' introduce corrections to formatting/editorial errors in transcript
  df_speakers <- df_speakers %>%
    mutate(speaker=str_replace(speaker, "Wolfang", "Wolfgang")) %>% 
    mutate(speaker=str_replace(speaker, "Sobotk\\b", "Sobotka")) %>% 
    mutate(speaker = case_when(
      str_detect(speaker, regex("Stellvertreter Mag\\. Friedrich Ofenauer")) ~ "Vorsitzender-Stellvertreter Mag. Friedrich Ofenauer",
      #Ofenauer is not a MP participating in inquiry
      str_detect(speaker, regex("^Mag. Friedrich Ofenauer \\(ÖVP\\)")) ~ "Abgeordneter Mag. Friedrich Ofenauer (ÖVP):",
      str_detect(speaker, regex("^Abgeordneter Mag. Friedrich Ofenauer")) ~ "Abgeordneter Mag. Friedrich Ofenauer (ÖVP):",
      str_detect(speaker, regex("Nina Tomaselli")) ~ "Abgeordnete Mag. Nina Tomaselli (Grüne):",
      str_detect(speaker, regex("Gerstl")) ~ "Abgeordneter Mag. Wolfgang Gerstl (ÖVP)
:",
      str_detect(speaker, regex("Hanger")) &
        !str_detect(speaker, "Vorsitz") ~ "Abgeordneter Mag. Andreas Hanger (ÖVP):",
      str_detect(speaker, regex("Hanger")) &
        str_detect(speaker, "Vorsitzender-Stellvertreter") ~ "Vorsitzender-Stellvertreter Mag. Andreas Hanger:",

      str_detect(speaker, regex("^Andreas Holzer")) ~ "Mag. Andreas Holzer, MA:",
      # double check; Sobotka was Vorsitzender and respondent
      #? controls for typo in transcript (Wolfang, Lisa Wieser respondent)
      str_detect(speaker, regex("^Wolfg?ang Sobotka")) ~ "Vorsitzender Mag. Wolfgang Sobotka:",
      #str_detect(speaker, regex("^Mag\\. Wolfgang Sobotka")) ~ "Mag. Wolfgang Sobotka:",
      # str_detect(speaker, regex("^Vorsitzender Mag\\. Wolfgang Sobot")) ~ "Vorsitzender Mag. Wolfgang Sobotka:",
      str_detect(speaker, regex("^Hans-Peter Weiss")) ~ "Dipl.-Ing. Hans-Peter Weiss",
      str_detect(speaker, regex("^Bernhard Perner")) ~ "Dipl.-Ing. Bernhard Perner",
      str_detect(speaker, regex("^Dr\\. Martin Graf \\(FPÖ\\)")) ~ "Abgeordneter Mag. Dr. Martin Graf (FPÖ):",
      str_detect(speaker, regex("^Dr\\. Daniel Varro, LL\\.M\\.")) ~ "MMag. Dr. Daniel Varro, LL.M.",
      str_detect(speaker, regex("orsitzender-Vertreterin Doris Bures")) ~ "Vorsitzender-Vertreterin Doris Bures",
      str_detect(speaker, regex("Nikolaus Kern, akad.$")) ~ "Nikolaus Kern, akad. M&S (WU)",
      str_detect(speaker, regex("^M?ag\\. Wolfgang Gerstl \\(ÖVP\\)")) ~ "Abgeordneter Mag. Wolfgang Gerstl (ÖVP)",
      str_detect(speaker, regex("Krisper")) ~ "Abgeordnete Dr. Stephanie Krisper (NEOS):",
      str_detect(speaker, regex("Kai Jan")) ~ "Abgeordneter Kai Jan Krainer (SPÖ):",
      str_detect(speaker, regex("Fürlinger")) ~ "Abgeordneter Mag. Klaus Fürlinger (ÖVP):",
      str_detect(speaker, regex("Mag\\. Erich Leopold")) ~ "Mag. Erich Leopold, MSc:",
      str_detect(speaker, regex("Gödl")) ~ "Abgeordneter Mag. Ernst Gödl (ÖVP):",
      str_detect(speaker, regex("Hafenecker")) ~ "Abgeordneter Christian Hafenecker, MA (FPÖ)",
      str_detect(speaker, regex("Mag\\. Christina Jilek")) ~ "Mag. Christina Jilek, LL.M.",
      str_detect(speaker, regex("Verfahrensanwalt-Stellvertreterin Mag\\. Dr\\. Barbara Weiß, LL\\.")) ~ "Verfahrensanwalt-Stellvertreterin Mag. Dr. Barbara Weiß, LL.M.",
      TRUE ~ as.character(speaker)
    )) %>%
    # remove wrongly formatted part of Kern
    filter(!str_detect(speaker, regex("^M\\&S \\(WU\\)"))) %>%
    # remove instances where only position fragments were extracted; keeping them would
    #lead to dupes in the fuzzy joins since the complete 'version' is in any case
    #already included
    filter(!str_detect(speaker, regex("^Vorsitzender Mag\\.$"))) %>% 
    filter(!str_detect(speaker, regex("^Vorsitzender-Stellvertreter Mag\\.$"))) %>% 
    filter(!str_detect(speaker, regex("^Abgeordnete(r)? Mag\\.$"))) %>% 
    filter(!str_detect(speaker, regex("^Abgeordnete(r)? M$"))) %>% 
    # remove colon at end; some are with, some without
    mutate(speaker = str_remove(speaker, regex("\\:$"))) %>%
    distinct(speaker)

  # create regex
  df_speakers <- df_speakers %>%
    mutate(
      speaker_pattern =
        str_replace_all(speaker, "\\.", "\\\\.") %>%
          str_replace_all(., "\\:", "\\\\:") %>%
          # str_replace_all(., "\\,", "\\\\,") %>%
          str_replace_all(., "\\)", "\\\\)") %>%
          str_replace_all(., "\\(", "\\\\(") %>% 
        #match has to be at start of string; avoids mismatches where name appears in middle of text
        paste0("^",.))

  df_speakers <- df_speakers %>%
    # remove titles after comma
    mutate(speaker_name = str_remove(speaker, regex(",.*\\:"))) %>%
    # every word ending with a dot (= academic abbreviations)
    mutate(speaker_name = str_remove_all(speaker_name, regex("\\S*\\."))) %>%
    # remove first word = position
    mutate(speaker_name = str_remove(speaker_name, regex("^\\S*"))) %>%
    # remove colon at end of speaker name
    mutate(speaker_name = str_remove(speaker_name, regex("\\:"))) %>%
    mutate(speaker_name = str_trim(speaker_name, side = c("both")))

## Extract entire text

# extract full text from minutes
  df_text <- link_to_record %>%
    rvest::read_html() %>%
    rvest::html_elements("p") %>%
    rvest::html_text() %>%
    enframe(., 
            name="row",
            value = "text") %>%
    mutate(text = str_squish(text) %>%
      str_trim(., side = c("both")))

  #manually correct for typo in text
  df_text <- df_text %>% 
    mutate(text=str_replace(text, pattern=regex("Grün(?=\\))"), 
                            replacement = "Grüne"))

  # extract date of session
  vec_respondent <- df_text %>%
    filter(str_detect(text, regex("^Befragung der Auskunftsperson"))) %>%
    #inserting corrections
    mutate(text=case_when(
      str_detect(text, regex("Blümel")) ~ "Bundesminister für Finanzen Mag. Gernot Blümel, MBA",
      str_detect(text, regex("Andreas Holzer")) ~ "Ministerialrat General Mag. Andreas Holzer",
                 TRUE ~ as.character(text))) %>% 
    pull(text) %>%
    str_remove(., "^Befragung der Auskunftsperson") %>% 
    str_trim(., side=c("both"))
    #insert correction for spelling error with Blümel

  # extract duration of ENTIRE session (not only respondent)
  vec_duration <- df_text %>%
    filter(str_detect(text, regex("^Gesamtdauer der"))) %>%
    pull(text) %>%
    str_extract(., regex("(?<=Sitzung).*"))

  # extract session number
  vec_session_no <- df_text %>%
    filter(str_detect(text, regex("^Gesamtdauer der"))) %>%
    pull(text) %>%
    str_extract(., regex("\\d*(?=\\. Sitzung)"))

  # extract date of session
  vec_date <- df_text %>%
    filter(str_detect(text, regex("^Montag|^Dienstag|^Mittwoch|^Donnerstag|^Freitag|^Samstag|^Sonntag"))) %>%
    pull(text)

  # get speaker name from text
  df_text_2 <- df_text %>%
    #search for match only in opening section of line
  #  mutate(text_start=stringr::str_sub(text, start=1, end=40)) %>% 
    fuzzyjoin::regex_left_join(.,
      df_speakers,
      by = c("text" = "speaker_pattern")
    ) # speaker_pattern

if (nrow(df_text_2)!=nrow(df_text)) {
  print(link_to_record)
  print(nrow(df_text_2))
  print(nrow(df_text))
  print("fuzzyjoin leads to duplicates")
  print(df_speakers)

  break
}

  df_text_2 <- df_text_2 %>%
    # fill rows with speaker and pattern as basis for grouping
    fill(speaker, .direction = "down") %>%
    fill(speaker_pattern, .direction = "down") %>%
    fill(speaker_name, .direction = "down") %>%
    filter(!is.na(speaker)) %>%
    # create grouping id; later needed to collapse rows
    mutate(grouping_id = ifelse(speaker == dplyr::lag(speaker, default = "start"),
      0,
      1
    )) %>%
    mutate(grouping_id_cum = cumsum(grouping_id)) %>%
    # remove speaker info from actual spoken text
    mutate(text = str_remove(text, speaker_pattern)) %>%
    mutate(text = str_remove(text, regex("\\:"))) %>%
    mutate(text = str_trim(text, side = c("both")))

  df_text_3 <- df_text_2 %>%
    # collapse rows
    group_by(grouping_id_cum) %>%
    summarise(
      text = paste(text, collapse = " "),
      speaker = unique(speaker)
    ) %>%
    relocate(speaker, .before = "text") %>%
    mutate(text = str_trim(text, side = c("both")))

  df_text_3 <- df_text_3 %>%
    # extract party of MP; pattern matches only if speaker starts with Abgeordneter
    mutate(speaker_party=str_extract(speaker, regex("(?<=^Abgeordnete[^\\(]{1,40})\\(.*\\)$"))) %>% 
    mutate(speaker_party=str_extract(speaker_party, regex("[:alpha:]+")))

  # add session details
  df_text_4 <- df_text_3 %>%
    mutate(
      session_date = vec_date,
      session_no = vec_session_no,
      session_duration = vec_duration,
      respondent = vec_respondent#,
      # row_id=paste(vec_respondent,< row, sep="-")
    ) %>%
    select(session_no, session_date, session_duration, respondent, speaker, speaker_party, text, -grouping_id_cum)

#CORRECTIONS SPEAKER
df_text_4 <- df_text_4 %>% 
  mutate(speaker=case_when(
         #Vorsitzender-Vertreterin Doris Bures
    str_detect(speaker, regex("Vorsitzender-Vertreterin Doris Bures")) ~ "Vorsitzender-Stellvertreterin Doris Bures",
    #one instance with out Mag.
    str_detect(speaker, regex("Martin Graf")) ~ "Abgeordneter Mag. Dr. Martin Graf (FPÖ)",
    #
    str_detect(speaker, "Rohrer") ~ "Verfahrensrichter-Stellvertreter Dr. Ronald Rohrer",

    #there is an error in the transcript of the 45 session; Ofenauer is once identified as a deputy and not as Vorsitzender-Stellvertreter although he participates as the latter in the session in question
      session_no==45 & 
      str_detect(speaker, "Ofenauer")==T &
      str_detect(speaker, regex("^Abge"))==T ~ "Vorsitzender-Stellvertreter Mag. Friedrich Ofenauer",
    str_detect(respondent, "Schwarzbartl") &
      str_detect(speaker, "Sobotka") ~  "Vorsitzender Mag. Wolfgang Sobotka",

    str_detect(speaker, regex("Verfahrensrichter-Vertreter Ronald Rohrer")) ~ "Verfahrensrichter-Stellvertreter Ronald Rohrer",

    TRUE ~ as.character(speaker))) %>%

  #correcting for error in  transcript; wrong identification of Ofenauer as deputy with a party affiliation instead as of Vorsitzender-Stell.
  mutate(speaker_party=case_when(
    str_detect(speaker, "Vorsitzender-Stellvertreter Mag. Friedrich Ofenauer") ~ NA_character_,
    TRUE ~ as.character(speaker_party)
  ))

  df_text_4 <- df_text_4 %>% 
  # extract position of speaker
    mutate(speaker_position = case_when(
      str_detect(speaker, regex("^Abgeord")) ~ "Abgeordneter",
      str_detect(speaker, regex("^Verfahrensanwalt\\b")) ~ "Verfahrensanwalt",
      str_detect(speaker, regex("^Vorsitzender-(Stell)?vertreter(in)?", ignore_case = T)) ~ "Vorsitzender-Stellvertreter",
      str_detect(speaker, regex("^Vorsitzende(r)?\\b")) ~ "Vorsitzende/r",
      str_detect(speaker, regex("^Verfahrensrichter(in)?-Stellvertreter(in)?")) ~ "Verfahrensrichter-Stellvertreter",
      str_detect(speaker, regex("^Verfahrensrichter(in)?\\s")) ~ "Verfahrensrichter",
      str_detect(speaker, regex("^Vertrauensperson")) ~ "Vertrauensperson",
      TRUE ~ as.character("Auskunftsperson")
    ))

  #CORRECTING SPEAKER POSITION
  df_text_4 <- df_text_4  %>%

  mutate(speaker_position=case_when(
    str_detect(speaker, "Rohrer") ~ "Verfahrensrichter-Stellvertreter",
    #there is an error in the transcript of the 45 session; Ofenauer is once identified as a     deputy and not as Vorsitzender-Stellvertreter although he particpates as the latter in       the session in question
    session_no==45 &
      str_detect(speaker, "Ofenauer")==T &
      str_detect(speaker_position, regex("^Abge"))==T ~ "Vorsitzender-Stellvertreter",

    str_detect(respondent, regex("Schwarzbartl")) &
      str_detect(speaker, regex("Sobotka")) ~ "Vorsitzende/r",

    TRUE ~ as.character(speaker_position) ))

  # identify party
  df_text_4 <- df_text_4 %>%
    mutate(respondent_questioner_party = case_when(
      str_detect(lag(speaker_position), regex("^Abge")) ~ lag(speaker_party),
      TRUE ~ NA_character_)) %>%
    # calculate answer length without annotations included in transcripts
    mutate(text_length = str_remove_all(text, regex("\\([^\\(\\)]*\\)")) %>% 
             stringi::stri_count_words(., locale="de"))

    #create short name of speaker
df_text_4 <- df_text_4 %>%
  #remove everything after comma
  mutate(speaker_name=str_remove(speaker, regex("\\,.*?$"))) %>% 
  #remove expresssions in brackets
  mutate(speaker_name=str_remove_all(speaker_name, regex("\\(.*?\\)"))) %>% 
  #remove academic titles
  mutate(speaker_name=str_remove_all(speaker_name, regex("\\S*\\."))) %>% 
  mutate(speaker_name=str_trim(speaker_name, side=c("both"))) %>% 
  mutate(speaker_name=str_squish(speaker_name))

df_text_4 <- df_text_4 %>% 
  mutate(procedural=str_detect(text, regex("^\\(zur Geschäftsbehandlung\\)"))) %>% 
  mutate(text=str_remove(text, regex("^\\(zur Geschäftsbehandlung\\)")) %>% 
           str_trim(., side="left"))

  # test - if there more than to respondents in a record something went wrong
  respondents_dupes <- df_text_4 %>%
    filter(speaker_position == "Auskunftsperson") %>%
    distinct(speaker)

  n_respondents <- respondents_dupes %>%
    nrow()

  if (n_respondents > 1) {
    print("more than one respondent; something wrong with extraction")
    print(link_to_record)
    print(respondents_dupes)
    break
  }

#final test to check whether any breaks/start of new statements were missed

df_text_4 <- df_text_4 %>% 
  mutate(id_statement=row_number())

df_speakers_name_family_unique <- df_text_4 %>% 
  distinct(speaker_name) %>% 
  mutate(speaker_name_family=str_extract(speaker_name, 
                                         regex("\\S*$"))) %>% 
  # mutate(speaker_name_family_regex=glue::glue("[^\\(]{20}{speaker_name_family}(?=.{0,20}:)"))
  mutate(speaker_name_family_regex=paste0(speaker_name_family, "(?=.{0,10}:)"))

df_missing <- df_text_4 %>% 
  mutate(text_clean=str_remove_all(text, regex("\\(.*?\\)"))) %>% 
  regex_semi_join(., 
                  df_speakers_name_family_unique,
                  by=c(text_clean="speaker_name_family_regex")) %>% 
  mutate(id_check=T)

df_text_4 <- df_text_4 %>% 
  left_join(.,
            df_missing  %>% 
              select(id_check, id_statement))

  # print(link_to_record)

  df_text_4
}
```

```{r, include=F, eval=F}
#test
# Vorsitzender-Stellvertreter Mag. Andreas Hanger

test_link <- "https://www.parlament.gv.at//PAKT/VHG/XXVII/KOMM/KOMM_00244/fnameorig_986573.html"

o <- fn_get_data(test_link)

o %>% 
  distinct(speaker, speaker_name, speaker_position)
```

Now let's apply the function to all links leading to transcripts of the inquiry's sessions.

```{r, cache=T, cache.extra=Sys.Date(), code_folding=F, echo=T, eval=T}
# Apply function to extract data -------------------------------------------

#To speed up the process I apply the function in parallel by using the furrr package.
library(furrr)
plan(multisession, workers = 2)

tictoc::tic()
df_data <- df_links_to_records %>%
  pull(link_to_record) %>%
  purrr::set_names() %>% 
  future_map_dfr(., fn_get_data, 
                 .progress = T,
                 .id="link")
tictoc::toc()

df_data <- df_data %>% 
  mutate(speaker_party= fct_relevel(speaker_party, names(vec_party_col)))
```

```{r, include=F, eval=F}
#these are all statements where a line break could have been missed
df_check <- df_data %>% 
  filter(id_check==T) %>% 
  mutate(txt_check=str_remove_all(text, regex("\\(.*?\\)")) %>% 
           str_extract_all(., regex(".{0,100}:.{0,20}", dotall = T))) %>% 
  select(session_no, link, id_statement, txt_check) %>% 
  unnest_longer(txt_check)

#check manually; OK
readr::write_excel_csv2(df_check, file=here::here("_blog_data", "2021_7_ibiza_inquiry","df_check.csv"))

```

```{r, eval=F, include=F}
file_name <- paste("df_data", Sys.Date(), ".csv", sep="_")

readr::write_excel_csv2(df_data %>% 
                          select(-contains("id")),
  file = here::here("_blog_data", "2021_7_ibiza_inquiry", file_name)
)

```

```{r, include=F, eval=F}
df_data_l <- readr::read_csv2(file=here::here("_blog_data", "2021_7_ibiza_inquiry",
                                            "df_data_2021-09-01_.csv"))

```

```{r, include=F, eval=T}

#make party a factor; important later when aggregating over groups and accounting for 'missing' levels (.drop=F)
df_data <- df_data %>% 
    mutate(speaker_party= fct_relevel(speaker_party, names(vec_party_col))) %>% 
    mutate(respondent_questioner_party= fct_relevel(respondent_questioner_party, names(vec_party_col))) %>% 
    mutate(respondent_questioner_party=forcats::as_factor(respondent_questioner_party)) %>% 
    mutate(session_no=as.numeric(session_no))

```

The dataframe obtained comprises the transcripts of `r length(unique(df_data$link))` sessions, which cover `r scales::number(nrow(df_data), big.mark=",")` single statements (rows of our dataframe). If someone is interested in the entire dataset, it can be downloaded here. Note that the dataframe only covers statements, and not the accompanying text (introduction etc) of the transcript (if you use the dataset, grateful if you quote this post as its source).

# Analysis

Everything we've done so far in the preceding section was preparatory work to eventually analyze the data. Now let's dig into it.

## Some general discriptives

Let's start with some general descriptives:

*How many sessions were there in which respondents were questioned:*

```{r}
df_data %>% 
  distinct(session_no) %>% 
  nrow()
```

*How many persons testified:*

```{r}
df_data %>% 
  distinct(respondent) %>% 
  nrow()
```

```{r}
df_data %>% 
  # filter(speaker_position=="Auskunftsperson") %>% 
  distinct(respondent) %>% 
  reactable(.,
            columns = list(
              respondent=colDef(name="Auskunftspersonen")
            ),
            filterable = T)
```

*How many testimonies were there:*

```{r}
df_data %>% 
  distinct(link) %>% 
  nrow()
```

*Which individuals testified multiple times:*

```{r}
df_data %>% 
  distinct(session_no, respondent) %>% 
  group_by(respondent) %>% 
  summarise(n_obs=n()) %>% 
  arrange(desc(n_obs)) %>% 
  filter(n_obs>1) %>% 
  reactable(.,
            filterable = T)
```

```{r, include=F}
# *Who spoke most words over the course of the entire inquiry:*
df_data %>% 
  group_by(speaker) %>% 
  summarise(text_length_sum=sum(text_length, na.rm=T)) %>% 
  arrange(desc(text_length_sum))
```

*Which MP spoke most (number of words)?:*

```{r paged.print=TRUE}
df_data %>% 
  filter(speaker_position=="Abgeordneter") %>% 
  group_by(speaker) %>% 
  summarise(text_length_sum=sum(text_length, na.rm=T)) %>% 
  arrange(desc(text_length_sum)) %>% 
  reactable(.,
            columns=list(
              speaker=colDef(filterable = T),
              text_length_sum=colDef(
                name="number of words",
                format = colFormat(separators = T))
              ))
```

*Which party spoke most (sum of words spoken by parties' MPs)?*

```{r}
df_data %>% 
  filter(speaker_position=="Abgeordneter") %>% 
  group_by(speaker_party) %>% 
  summarise(text_length_sum=sum(text_length, na.rm=T)) %>% 
  arrange(desc(text_length_sum)) %>% 
  reactable(.,
            columns = list(
              speaker_party=colDef(
                name="Partei"),
              text_length_sum=colDef(
                name="Wortanzahl insg.",
                format = colFormat(separators = T)
              )
            ))
```

## Respondents' length of statements

Let's dig a bit deeper now. One aspect I was wondering about is the length of answers by respondents to questions by MPs.

As outlined above, identifying an answer of a respondent to an MP is a bit tricky, and implies the assumption that a question by an MP always directly precedes the answer of the respondent. Bear this assumption in mind when reading the later results.

Below a plot which shows the length of every response by a respondent, grouped by the party membership of the MPs to which the respondent answered. The length of an answer is measured by the number of words spoken; the plot shows only a selection of respondents which I considered particularly interesting.

```{r fig.height=8, layout="l-body-outset"}
#create vectors of respondents' names; 
vec_respondents_oevp <- c(
  "Kurz", "Blümel", "Bonelli", "Mei-Pochtler",
  "Spiegelfeld-Quester", "Nehammer", "Purkart", "Sobotka")
vec_respondents_oevp_or <- paste(vec_respondents_oevp, collapse="|")

vec_respondents_fpoe <- c("Strache", "Gudenus", "Hofer", "Kickl")
vec_respondents_fpoe_or <- paste(vec_respondents_fpoe, collapse="|")

vec_respondents_gruene <- c("Zadic")
vec_respondents_gruene_or <- paste(vec_respondents_gruene, collapse="|")

vec_respondents_spoe <- c("Christian Kern")
vec_respondents_spoe_or <- paste(vec_respondents_spoe, collapse="|")

vec_respondents_all_or <- paste(vec_respondents_oevp_or, vec_respondents_fpoe_or, 
                                vec_respondents_gruene_or, vec_respondents_spoe_or,
                                sep="|")

library(cowplot)
library(rlang)
library(ggtext)

element_textbox_highlight <- function(..., 
                                      hi.labels = NULL, hi.fill = NULL,
                                      hi.col = NULL, hi.box.col = NULL,
                                      hi.labels2 = NULL, hi.fill2 = NULL,
                                      hi.col2 = NULL, hi.box.col2 = NULL) {
  structure(
    c(element_textbox(...),
      list(hi.labels = hi.labels, hi.fill = hi.fill, hi.col = hi.col, hi.box.col = hi.box.col,
           hi.labels2 = hi.labels2, hi.fill2 = hi.fill2, hi.col2 = hi.col2, hi.box.col2 = hi.box.col2)
    ),
    class = c("element_textbox_highlight", "element_textbox", "element_text", "element",
              "element_textbox_highlight", "element_textbox", "element_text", "element")
  )
}

element_grob.element_textbox_highlight <- function(element, label = "", ...) {
  if (label %in% element$hi.labels) {
    element$fill <- element$hi.fill %||% element$fill
    element$colour <- element$hi.col %||% element$colour
    element$box.colour <- element$hi.box.col %||% element$box.colour
  }
  if (label %in% element$hi.labels2) {
    element$fill <- element$hi.fill2 %||% element$fill
    element$colour <- element$hi.col2 %||% element$colour
    element$box.colour <- element$hi.box.col2 %||% element$box.colour
  }
  NextMethod()
}

library(lemon)
df_data %>%
  filter(str_detect(respondent, regex(vec_respondents_all_or))) %>%
  filter(speaker_position == "Auskunftsperson") %>% 
  filter(!is.na(respondent_questioner_party)) %>%
  mutate(facet_label=glue::glue("{speaker_name} ({session_no})")) %>% 
  ggplot() +
  labs(title="IBIZA-AUSSCHUSS: Länge der Antworten von Auskunftspersonen \nper Parteimitgliedschaft der fragenden Abgeordneten",
       subtitle="Ausgewählte Auskunftspersonen. Sitzungsnummer in Klammer.",
       caption=c("Jeder Punkt stellt eine Antwort/Wortmeldung der Auskunftsperson dar. Die Position auf der x-Achse gibt an welcher Partei der/die \nAbgeordnete angehört, der/die zuvor eine Frage gestellt hatte bzw. an den/die die Antwort der Auskunftsperson gerichtet war. Die Position auf der y-Achse zeigt die Länge der Antwort der Auskunftsperson als Anzahl an Wörtern an.\n\nRoland Schmidt | @zoowalk | http://werk.statt.codes"),
       # y="Länge der Antwort (Anzahl v. Wörtern)",
       x="Parteizugehörigkeit der/s Abgeordneten")+
  geom_jitter(aes(
    x = respondent_questioner_party,
    y = text_length,
    color=respondent_questioner_party),
    size=0.5
  )+
  scale_color_manual(values=vec_party_col)+
  scale_y_continuous(expand=expansion(mult=c(0.01, 0.1)),
                     breaks=c(0,500, 1000, 1400),
                     labels=c("0", "500", "1000", "1400 Wörter"))+
  lemon::facet_rep_wrap(~facet_label,
             ncol=3,
             repeat.tick.labels = "x",
             labeller=label_wrap_gen(width = 30))+
  theme_post()+
  theme(
    # strip.text = element_text(color="grey20"),
    axis.text.x=element_text(size=7),
    axis.title.x=element_text(hjust=0,
                              size=7),
    axis.text.y=element_text(size=7),
    panel.spacing.x = unit(0, units="cm"),
    legend.position = "none",
    axis.title.y=element_blank(),
    panel.grid.major.x=element_blank(),
    strip.text = element_textbox_highlight(
      size = 8,
      # unnamed set (all facet windows except named sets below)
      color = "black", 
      fill = "white", 
      box.color = "white",
      halign = 0, 
      valign=0.5,
      linetype = 1, 
      # r = unit(5, "pt"), 
      # height=unit(1, "cm"),
      width = unit(1, "npc"), #expands text box to full width of facet
      #padding = margin(2, 0, 1, 0), 
      #margin = margin(3, 3, 3, 3),
      # this is new relative to element_textbox():
      # first named set
      hi.labels = c("Sebastian Kurz (55)", "Herbert Kickl (41)"),
      hi.fill = "black", 
      hi.box.col = "black", #line around box
      hi.col = "white"))

```

The graph pertaining to the testimony of Chancellor Kurz during the commission's 55th session stands out. Chancellor Kurz's answers to the questions by his own party colleagues were strikingly long. There are some instances of long answers also with the testimony of Herbert Kickl and Johann Gudenus, when replying to questions from members of their won party (although I am not sure whether Gudenus is formally still FPÖ member).

However, there's one critical qualification warranted. It could be that a respondent gives few, but some particularly lengthy answers, in contrast to another respondent who gives many, but shorter answers. This could be e.g. due to a different dynamic in the conversation with an MP. Hence the plot above might give a distorted impression. Nevertheless, judging from the number of dots in the graph (i.e. number of replies), the plot suggests that this was not the case.

Another, more accurate avenue to address this issue is to a) calculate the total length of answers (= total number of words) addressed to each party, and subsequently b) contrast these numbers and check whether the largest portion was actually directed at questions of MPs who were from the respondent's own party.

The table below shows how the total number of words answered (excluding opening statements) by a respondent were distributed among the party membership of MPs asking questions to the respondent. Note that the table provides the results for all (!) respondents who testified during the inquiry.

Search for the entry for Sebastian Kurz or sort the ÖVP column. You'll see there was no other respondent who dedicated a larger share of his overall answer to questions raised by ÖVP MPs.

```{r, layout="l-body-outset", eval=T}
df_party_response_length<- df_data %>% 
  #keep only answers from respondents; no opening statement etc.
  filter(!is.na(respondent_questioner_party)) %>% 
  filter(speaker_position=="Auskunftsperson") %>% 
  #calculate sum of length of answer by respondent and party of asking MP;
  #drop=F because we want to have 0 if a respondent never answered to a party; and not a missing observation
  group_by(session_no, speaker_name, respondent, respondent_questioner_party, .drop = F) %>% 
  summarize(text_length_sum=sum(text_length, na.rm = T)) %>% 
  ungroup()

df_respondent_response_party <- df_party_response_length %>%
  group_by(session_no, respondent, .drop = F) %>%
  mutate(text_length_sum_total=sum(text_length_sum, na.rm = T)) %>%
  mutate(text_length_sum_rel=text_length_sum/text_length_sum_total) %>%
  group_by(session_no, respondent) %>%
  arrange(desc(text_length_sum_rel), .by_group = T)

library(reactablefmtr)

df_respondent_response_party %>% 
  select(-speaker_name) %>% 
  pivot_wider(.,
              id_cols=c(session_no, respondent, text_length_sum_total),
              names_from = c(respondent_questioner_party),
              names_sort = T,
              values_from = c(text_length_sum_rel),
              #adds 0 as response length if respondent did not answer a party
              values_fill = 0) %>% 
  arrange(desc(ÖVP)) %>% 
  reactable(.,
            columns = list(
              session_no = colDef(name = "Sitzung",
                                  align = "center",
                                  sticky = "left",
                                  width = 60
                                  ),
              respondent = colDef(name = "Auskunftsperson",
                                  sticky = "left",
                                  filterable = T,
                                  width = 250
                                  ),
              text_length_sum_total = colDef(name = "Antwortlänge insg. (Wörter)",
                                             format = colFormat(separators = T),
                                             sticky = "left",
                                             width=150,
                                             align="center"
                                             ),
              ÖVP = colDef(width = 60,
                           align = "left",
                           cell = data_bars(., 
                                            text_size = 13,
                                            # bar_height = 13,
                                            fill_color = "#5DC2CC", 
                                            text_position = "inside-base",
                                            number_fmt = scales::percent,
                                            max_value = 1
                                        )
                           ),
              SPÖ = colDef(format=colFormat(percent = T, 
                                            digits = 1),
                           width = 60,
                           align = "left",
                           cell = data_bars(., 
                                            fill_color = "#FC0204", 
                                            text_position = "inside-base",
                                            text_size = 13,
                                            #bar_height = 13,
                                            brighten_text=FALSE,
                                            number_fmt = scales::percent,
                                            max_value = 1
                                            )
                           ),
              Grüne = colDef(format=colFormat(percent = T,
                                              digits = 1),
                             width = 60,
                             align = "left",
                             cell = data_bars(., 
                                              fill_color = "#A3C630",
                                              text_size = 13,
                                              # bar_height = 13,
                                              text_position = "inside-base",
                                              number_fmt = scales::percent,
                                              max_value=1
                                              )
                             ),
              FPÖ = colDef(format=colFormat(percent = T, 
                                            digits = 1),
                           width = 60,
                           align = "left",
                           cell = data_bars(., 
                                            text_size = 13,
                                            fill_color = "#005DA8",
                                            # bar_height = 13,
                                            text_position ="inside-base",
                                            number_fmt = scales::percent,
                                            max_value=1)
                           ),
              NEOS = colDef(format=colFormat(percent = T, 
                                             digits = 1),
                            width = 60,
                            cell = data_bars(., 
                                             text_size = 13,
                                             # bar_height = 13,
                                             fill_color = "#EA5290", 
                                             text_position = "inside-base",
                                             number_fmt = scales::percent,
                                             max_value=1)
                            )
              ),
            theme=nytimes(cell_padding=0,
                          font_size=13)) %>% 
  add_title(title="IBIZA-AUSSCHUSS: Wem antworten Auskunftspersonen?") %>% 
  add_subtitle(subtitle="Verteilung der Antworten (Wortanzahl) von Auskunftspersonen nach Parteizugehörigkeit der fragenden Parlamentarier.") %>% 
  add_source(source="Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes · Data: parlament.gv.at",
             font_size=11,
             font_weight = "normal",
             font_color = "#999999",
             font_family = "Helvetica",
             align="right") 
```

```{r fig.height=8, layout="l-body-outset", include=F, eval=F}
#NOT INCLUDED FOR BREVITY REASONS
# The graph below shows the largest response share (= total length of answers to one party divided by total length of answers to all parties) of respondents. Note that the plot shows the top ten of all (!) respondents who had testified during the inquiry. In other words,

f_party_response_length %>% 
  mutate(speaker_session=glue::glue("{speaker_name} ({session_no})")) %>% 
  filter(str_detect(respondent, regex(vec_respondents_all_or))) %>% 
  ggplot()+
  labs(title="")+
  geom_bar(
    aes(
      x=respondent_questioner_party,
      y=text_length_sum,
      fill=respondent_questioner_party),
    stat="identity",
    position=position_dodge())+
  scale_fill_manual(values=vec_party_col)+
  theme_post()+
  theme(
    # strip.text = element_text(color="grey20"),
    axis.text.x=element_text(size=7),
    axis.title.x=element_text(hjust=0,
                              size=7),
    axis.text.y=element_text(size=7),
    panel.spacing.x = unit(0, units="cm"),
    legend.position = "none",
    axis.title.y=element_blank(),
    panel.grid.major.x=element_blank(),
    strip.text = element_textbox_highlight(
      size = 8,
      # unnamed set (all facet windows except named sets below)
      color = "black", 
      fill = "white", 
      box.color = "white",
      halign = 0, 
      valign=0.5,
      linetype = 1, 
      # r = unit(5, "pt"), 
      # height=unit(1, "cm"),
      width = unit(1, "npc"), #expands text box to full width of facet
      #padding = margin(2, 0, 1, 0), 
      #margin = margin(3, 3, 3, 3),
      # this is new relative to element_textbox():
      # first named set
      hi.labels = c("Sebastian Kurz (55)", "Herbert Kickl (41)"),
      hi.fill = "gray90", 
      hi.box.col = "gray90", #line around box
      hi.col = "black"))+
    lemon::facet_rep_wrap(~speaker_session,
             ncol=3,
             repeat.tick.labels = "x",
             labeller=label_wrap_gen(width = 30))
```

To make things even more clear, let's get the largest shares across all respondents. There was no other respondent who dedicated a larger share of his/her answers to questions put forward by the MPs of his/her own party than Sebastian Kurz during the commission's 55th session. I thought that's quite remarkable. Whether this was a deliberate tactic, alike a filibuster, to avoid questions by other MPs I don't know. However, the mere numbers would be in line with such an approach. Your own MPs give you softball questions, you answer in a particularly verbose fashion... and deprive other MPs of asking questions.

```{r}
df_respondent_response_party %>% 
  slice_head(., n=1) %>%
  ungroup() %>%
  arrange(desc(text_length_sum_rel)) %>% 
  slice_head(., n=10) %>% 
  mutate(respondent_session=glue::glue("{respondent} ({session_no})")) %>% 
  ggplot()+
  labs(title="IBIZA-AUSSCHUSS: Auskunftspersonen und ihr größter Antwortanteil",
       subtitle=str_wrap("Wie groß und an welche Partei war der größte Anteil der Antworten einer Auskunftsperson gerichtet? Sitzungsnummer in Klammer. Top-10.", 100),
       caption="Beispiel: Bundeskanzler Kurz hat im Rahmen seiner Aussage in der 55. Sitzung rund 75 % aller Wörter seiner Antworten and Fragen von\nÖVP Abgeordneten adressiert.\n\nRoland Schmidt | @zoowalk | http://werk.statt.codes",
       x="Anteil an der Gesamtantwort der Auskunftsperson (Wortanzahl)")+
  geom_bar(
    aes(
      y=reorder(respondent_session, text_length_sum_rel),
      x=text_length_sum_rel,
      fill=respondent_questioner_party
    ),
    orientation="y",
    stat="identity"
  )+
  geom_text(
    aes(
      y=reorder(respondent_session, text_length_sum_rel),
      x=text_length_sum_rel+0.01,
      color=respondent_questioner_party,
      label=text_length_sum_rel %>% 
        scales::percent(., accuracy = .1)),
    hjust=0,
    size=3,
    stat="identity"
  )+
  scale_fill_manual(values = vec_party_col,
                    name="Partei mit größtem Antwortteil.")+
  scale_color_manual(values = vec_party_col,
                    name="Partei mit größtem Antwortteil.")+
  scale_x_continuous(labels=scales::label_percent(),
                     expand=expansion(mult=c(0, 0.1)))+
  theme_post()+
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    plot.caption.position = "plot",
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(),
    plot.caption = element_text(hjust=0)
  )

```

```{r, eval=F, include=F}
df_respondent_response_party %>% 
  select(-speaker_name) %>% 
  pivot_wider(.,
              id_cols=c(session_no, respondent, text_length_sum_total),
              names_from = c(respondent_questioner_party),
              names_sort = T,
              values_from = c(text_length_sum_rel),
              values_fill = 0) %>% 
  reactable(.,
            columns = list(
              session_no = colDef(name = "Sitzung",
                                  align = "center",
                                  sticky = "left",
                                  width = 60
                                  ),
              respondent = colDef(name = "Auskunftsperson",
                                  sticky = "left",
                                  filterable = T,
                                  width = 170),
              text_length_sum_total = colDef(name = "Antwortlänge insg. (Wörter)",
                                             format = colFormat(
                                               separators = T),
                                               sticky = "left",
                                             width=100,
                                             ),
              ÖVP = colDef(format=colFormat(percent = T, 
                                            digits = 1),
                           width = 50,
                           align = "left"),
              SPÖ = colDef(format=colFormat(percent = T, 
                                            digits = 1),
                           width = 50,
                           align = "left"),
        Grüne = colDef(format=colFormat(percent = T, 
                                            digits = 1),
                             width = 50,
                             align = "left"),
        FPÖ = colDef(format=colFormat(percent = T,
                                      digits = 1),
                           width = 75,
                           align = "left",
                           style=cell_style(
                              vertical_align = "center",
                              horizontal_align = "left"),
                            cell=data_bars(.,
                               text_position="inside-end",
                               text_size=13,
                               text_color="#000",
                               fill_color="#005DA8",
                               number_fmt = scales::percent)
                     ),
        NEOS = colDef(width = 150,
                      align="left",
                            # style=cell_style(
                            #   font_size=13,
                            #   # vertical_align = "top",
                            #   horizontal_align = "left"
                            #   ),
                            cell=data_bars(.,
                               text_position="outside-end",
                               text_size=13,
                               text_color="#000",
                               fill_color="#EA5290",
                               number_fmt=scales::percent
                               )
                      )
        )
        )
        )

```

## 'Can't remember'

(Un)Surprisingly, there were quite a few respondents who stated to have no recollection or 'perception' of the matter raised by the MPs. To be fair, there are instances in which I have some sympathy for a respondent if s/he can't recall a specific detail of an event which dates back numerous years. Nevertheless, considering the issue at stake, respondents' possibility to prepare in advance for their testimony etc. - the lack of memory by some respondents seemed unusual, to put it mildly.

To get an idea which respondents were particularly affected by a loss of memory or a lack of knowledge in the first place, let's first define some key phrases and then check for their (absolute and relative) frequency in the respondents' answers. Again, bear in mind that the subsequent results are entirely contingent on these key phrases.

Here the phrases which I thought of as being helpful indicators: 'no perception' (*'keine Wahrnehmung'*), 'not remember' (*'nicht erinnern'*), 'I do not know' (*'weiß ich nicht'*), 'no recollection' (*'keine Erinnerung'*);

To be clear, the approach is pretty crude and the results should be taken with some caution. I don't argue that these key phrases will catch every instance of a respondent claiming to have no knowledge/memory [^2]; nor do I claim that there are no other phrases in which respondents express their inability to answer. I merely think that these key phrases give us an informative general impression.

[^2]: [Reportedly](https://kurier.at/politik/inland/u-ausschuss-86-mal-keine-erinnerung-und-eine-anzeige/400951499){target="_blank"}, Finance Minister Blümel responded 86 times that he was unable to remember during his testimony in June. As you'll see below, this number differs considerably from the number obtained by means of key phrases.

Below the results. The table gives the absolute (how often did the key phrases appear in the respondent's statements) and relative frequency (the absolute frequency divided by the total number of words spoken by the respondent).

```{r}
vec_recall <- c(
  "weiß ich nicht",
  "nicht erinnern", 
  "keine Wahrnehmung",
  "keine Erinnerung"
  )

vec_recall_or <- vec_recall %>% paste(., collapse="|")

vec_recall_text <- glue::glue_collapse(glue::glue("'{vec_recall}'"), sep=", ", last=" oder ")

df_recall <- df_data %>% 
  filter(speaker_position=="Auskunftsperson") %>% 
  group_by(speaker, session_no) %>% 
  mutate(text_lengh_sum=sum(text_length, na.rm = T)) %>% 
  mutate(freq=str_count(text, regex(vec_recall_or, 
                                    ignore_case = T,
                                    dotall = T))) %>% 
  summarise(freq=sum(freq),
            text_lengh_sum=mean(text_lengh_sum)) %>% 
  ungroup() %>% 
  arrange(desc(freq)) %>% 
  mutate(freq_rel=freq/text_lengh_sum) %>% 
  arrange(desc(freq))

df_recall %>% 
  reactable(.,
          theme = nytimes(cell_padding=0),
            columns = list(
              speaker = colDef(
                name="Auskunftsperson",
                filterable = T,
                width = 200
                ),
              session_no=colDef(
                name="Sitzung",
                align = "center"
              ),
              freq=colDef(
                align="left",
                name="Häufigkeit",
                cell = data_bars(.,
                                 fill_color = "orange",
                                 text_size = 13,
                                 max_value = 150,
                                # bar_height = 15,
                                 text_position = "inside-end",
                                 number_fmt = scales::comma)
              ),
              freq_rel=colDef(
                align="left",
                name="rel. Häufigkeit",
                cell = data_bars(.,
                                 text_size = 13,
                                 fill_color = "gray30",
                                # bar_height = 15,
                                 text_position = "inside-end",
                                 number_fmt = scales::label_percent(accuracy = 0.01))#,
                # format=colFormat(
                #   percent=T,
                #   digits=2)
                ),
              text_lengh_sum=colDef(
                show=F
              )
            )
          ) %>% 
  add_title(title=glue::glue("IBIZA-AUSSCHUSS: Auskunftspersonen und die fehlende Erinnerung"),
            font_size=18
           # font_weight="bold",
          #  font_color="black"
          ) %>% 
  add_subtitle(subtitle=html(glue::glue("Wie oft wurden die Wortfolgen <i>{vec_recall_text}</i> in den Antworten einer Auskunftsperson gefunden. Relative Häufigkeit setzt Anzahl der Treffer in Relation zur Länge der Antworten (Wortanzahl).")),
               font_size=12,
               font_weight="normal") %>%
  add_source(source="Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes · Data: parlament.gv.at",
             font_size=11,
             font_weight = "normal",
             font_color = "#999999",
             font_family = "Helvetica",
             align="right") 
```

Let's have a look now at who comes out on top: First, Arno Melicharek, a member of Sebastian Kurz' staff and mainly known for having (under a false name) ordered the shredding of several harddrives of the Chancellery relativly soon after the release of the 'Ibizia Video'. Second, Bernhard Bonelli - Chancellor Kurz's chief of staff (*Kabinettschef*) and by all accounts one of his closest advisors. Third, Lisa Wieser - the Chancellor's personal assistant.

A bit further behind comes Gernot Blümel, Austria's finance minister and a long-time alley of Sebastian Kurz, who famously stated that he couldn't remember whether he actually owned a laptop during his first stint as finance minister. In short, the table's top positions are dominated by individuals closely related with the chancellor and the ÖVP. This picture is largely unaltered if one considers the relative frequency of the key phrases.

## Debate on procedural issues

At least judging from the news report I came across, debates on procedural issues erupted quite frequently and were rather heated. After all, defining the rules by which respondents have to answer (or the possibility to abstain from doing so), including which questions are deemed as admissible, can predefine the inquiry's eventual outcome. Interestingly, the inquiry's transcripts contain an annotation which indicates whether a specific statement by a participant of the inquiry is actually pertaining to the discussion of procedural issues. See eg. below for an example:

![](the-ibizia-inquiry-in-numbers_files/images/geschaeftsbehandlung.png){height="200"}

The annotation 'zur Geschäftsbehandlung' (pretty dusty wording, but essentially meaning something like 'related to procedural matters') allows to detect these statements. This further allows us to see in which sessions of which respondents procedural issues were debated and to which extent (share of words spoken during the session). Below the result:

```{r}
df_data %>% 
  group_by(session_no, respondent, procedural) %>% 
  summarise(text_sum=sum(text_length)) %>% 
  pivot_wider(.,
              id_cols=c(session_no, respondent),
              names_from = procedural,
              values_from = text_sum,
              values_fill = 0) %>% 
  mutate(procedural_share=`TRUE`/(`FALSE`+`TRUE`)) %>% 
  arrange(desc(procedural_share)) %>% 
  ungroup() %>% 
  select(-`TRUE`, -`FALSE`) %>% 
  reactable(.,
            columns=list(
              session_no=colDef(
                name="Sitzung",
                align="center",
                width = 60
              ),
              respondent=colDef(
                name="Auskunftsperson",
                width=350,
                filterable=T
              ),
              procedural_share=colDef(
                name="Anteil Geschäftsordnung",
                align="left",
                format=colFormat(percent = T,
                                 digits = 2),
                cell=data_bars(.,
                               fill_color = "#92bdb2",
                               max_value = 1,
                               text_size=13,
                               background = "lightgrey",
                               text_position = "inside-end",
                               number_fmt = scales::percent)
              )),
            theme=nytimes(cell_padding = 0)) %>% 
  add_source(source="Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes · Data: parlament.gv.at",
             font_size=11,
             font_weight = "normal",
             font_color = "#999999",
             font_family = "Helvetica",
             align="right") %>% 
  add_title(title = "IBIZA-AUSSCHUSS: Anteil der Debatte zur Geschäftsordnung",
            font_size = 18) %>% 
  add_subtitle(subtitle="Die Sitzungsprotokolle des Ausschusses zeigen an, ob ein Statement Teil einer Diskussion zur Geschäftsordnung ist (Annotation 'zur Geschäftsbehandlung'). Die Länge dieser Statements lässt sich in Relation zur Gesamtlänge aller Statements setzen und gibt damit die Größe des Anteils der Debatte zur Geschäftsordnung an.",
               font_size = 12,
               font_weight = "normal")
```

The figures presented above indicate the number of words spoken in statements pertaining to procedural issues in relation (share) to the total number of words spoken during the session in question. E.g. 27 % of all words spoken during the hearing of MoF Gernot Blümel in the commission's 55th session pertained to the discussion of procedural issues. In other words, more than a quarter of the words spoken during the hearing did not focus on substantive matters. That seems quite remarkable to me.

## Annotations

The transcripts also include bracket terms with annotations providing further details on the speakers and their statements. These annotations refer e.g. to participants' gestures (i.e. 'schüttelt den Kopf'/'shakes head'; 'hebt die Hand'/'raises hand'), facial expressions ('erheitert'/'amused'), interjections etc. Below some examples highlighted in the transcript.

![](the-ibizia-inquiry-in-numbers_files/images/annotations_example.png)

In the following section these bracket terms are extracted and analyzed. Whether the aggregate numbers indeed convey something meaningful, I'll leave up to your discretion. I thought some are at least entertaining.

```{r, code_folding=F}
df_data <- df_data %>% 
  #extract all bracket terms
  mutate(bracket_terms=str_extract_all(text, regex("\\(.*?\\)")))

df_bracket_terms <- df_data %>% 
  filter(!is.na(bracket_terms)) %>% 
  select(link, session_no, respondent, speaker, bracket_terms, text) %>% 
  #unnest list if there are multiple bracket terms in one row
  unnest_longer(bracket_terms) %>% 
  filter(!is.na(bracket_terms)) %>% 
  #add a unique identifier for each bracket term
  mutate(bracket_terms_id=row_number())

```

```{r, include=F}
#TEST: are there any bracket terms which would be indicative for an error in extracting party affiliation/line breaks
df_bracket_terms %>% 
  count(bracket_terms, sort = T )

x <- df_bracket_terms %>% 
  filter(bracket_terms=="(ÖVP)") %>% 
  slice_head(., n=5)

y <- df_data %>% 
  distinct(speaker)

z <- df_data %>% 
  filter(str_detect(text, regex("\\(ÖVP\\)"))) %>% 
  mutate(ÖVP_bracket=str_extract_all(text, regex(".{20}\\(ÖVP\\).{30}"))) %>% 
  filter(!str_detect(text, "Gödl")) %>% 
  select(ÖVP_bracket) %>% 
  unnest_longer(ÖVP_bracket) %>% 
  distinct()

```

### Interjections (*Zwischenrufe*)

As one might expect, the commission's hearings got at a number of times rather heated and participants' exchanges were by no means amicable. Interjections were not uncommon. Remarkably, the transcripts of the sessions include annotations (in brackets) recording these interjections. Whenever a participant made/shouted a comment while not having having the floor, the transcripts contain a pertaining annotation. This seemed like an interesting source of information to me. However, extracting these annotations of interest is not entirely straightforward.

Essentially, there are two types of annotations which indicate interjections. *Type A:* Those interjections which are explicitly annotated as "Zwischenruf" (interjection). They may or may not be part of a composite bracket term, which includes other additional information, and/or refer to one or multiple sources/persons interrupting. This type of annotations generally does not provide the text of the actual interjection. *Type B:* These are annotations which state the name of the person making the interjection and the actual text. However, the annotation does not include the word 'Zwischenruf'. Below some examples.

```{r}
df_example <- df_bracket_terms %>% 
  filter(str_detect(bracket_terms, "Zwischenruf") |
           str_detect(bracket_terms, regex("\\:"))) %>% 
  filter(str_detect(bracket_terms, "-")) %>% 
  filter(!str_detect(bracket_terms, regex("Auskunftsperson"))) %>% 
  slice_head(., n=5) %>% 
  select(bracket_terms_id, bracket_terms)

df_example %>% 
  reactable(.,
            columns = list(
              bracket_terms=colDef(
                width=500
              ),
              bracket_terms_id=colDef(
                width=100,
                align="center"
              )
            ))
```

To eventually obtain a consolidated dataframe which comprises one row per interjection per source (i.e. person interrupting) requires hence two separate approaches. As a common first step though, it is necessary to split bracket terms comprising multiple annotations into distinct ones, i.e. getting one row/observation for each annotation. Different annotation *within* one bracket term are separated by a hyphen, hence we use them to split the bracket terms. Below the result; note how the previously unique `brackt_terms_id` now appears in multiple rows (e.g row\_id 240 and 318).

```{r, code_folding=F}
df_bracket_terms_long <-  df_bracket_terms %>% 
  #regex splits the bracket terms when there is a special type of hyphen
  #preceded by a dot, a question mark or an exclamation mark.
  mutate(bracket_terms_split=str_split(bracket_terms, regex("(?<=[\\.\\?\\!]\\s)\\p{Pd}"))) %>% 
  #makes data frame long; one row per annotation
  unnest_longer(bracket_terms_split) %>% 
  mutate(bracket_terms_split=str_remove_all(bracket_terms_split, regex("[\\)\\(]")) %>% 
           str_trim(., side=c("both")))
```

```{r, include=T, echo=F}
df_bracket_terms_long %>% 
  filter(bracket_terms_id %in% df_example$bracket_terms_id) %>% 
  select(bracket_terms_id, bracket_terms_split) %>% 
  reactable(.,
            columns = list(
              bracket_terms_id=colDef(
                width=100,
                align="center"
              ),
              bracket_terms_split=colDef(
                width=500
              )
            ))
```

Now let's focus on the *type A* of interjections - those which are explicitly annotated with the word 'Zwischenruf'. This can be done by simply filtering the split bracket terms.

```{r, code_folding=F}
df_interjections_a <- df_bracket_terms_long %>% 
  filter(str_detect(bracket_terms_split, regex("Zwischenruf")))
```

There are `r nrow(df_interjections_a)` such instances. Below some examples. Note that there are instances in which one annotation refers to more than one source.

```{r, include=F}
df_interjections_a %>% 
  select(bracket_terms_id, bracket_terms_split) %>% 
  slice_head(., n=10)%>% 
  reactable(.,
            columns = list(
              bracket_terms_id=colDef(
                width=100,
                align="center"
              ),
              bracket_terms_split=colDef(
                width=500
              )
            ))
```

```{r, echo=F, code_folding=F}
df_interjections_a %>% 
  filter(str_detect(bracket_terms_split, regex("Stögmüller")) &
           str_detect(bracket_terms_split, regex("Matznetter"))) %>% 
  select(bracket_terms_id, bracket_terms_split)%>% 
  reactable(.,
            columns = list(
              bracket_terms_id=colDef(
                width=100,
                align="center"
              ),
              bracket_terms_split=colDef(
                width=500
              )
            ))
```

This is something we have to account for when extracting the name of the MP(s). Eventually we want one row for each interjection-source (MP) combination. This can be best done with `fuzzyjoin::regex_left_join`.

```{r, code_folding=F}
#create df with unique names of MPs
df_mps <- df_data %>% 
  select(speaker_name, speaker_party) %>% 
  #regex gets last word from name 
  mutate(speaker_name_family=str_extract(speaker_name, regex("\\S+$"))) %>% 
  distinct() %>% 
  #I am only interested in interjections from MPs
  filter(str_detect(speaker_name, regex("^Abge")))

#search for the presence of MPs names in the interjection
df_interjections_a <- df_interjections_a %>% 
  fuzzyjoin::regex_left_join(.,
                             df_mps %>% 
                               select(source_name_family=speaker_name_family,
                                      source_name=speaker_name,
                                      source_party=speaker_party),
                             by=c(bracket_terms_split="source_name_family")) 
```

Below the result for the previous examples in which multiple sources were identified. Instead of one row per interjection, we have now one row per interjection-source combination. This disentanglement will later allow us to adequately aggregate the number of interjections per source (MP).

```{r, echo=F}
df_interjections_a %>%     
  filter(str_detect(bracket_terms_split, regex("Stögmüller")) &
           str_detect(bracket_terms_split, regex("Matznetter"))) %>% 
  select(bracket_terms_id, bracket_terms_split, source_name) %>% 
  reactable(.,
            columns = list(
              bracket_terms_id=colDef(
                width=150,
                align="center"
              ),
              bracket_terms_split=colDef(
                width=300
              ),
              source_name=colDef(
                width=200
              )
            ))
```

Now let's focus on *type b* of the interjections. Here the identification is less straightforward. I included explanatory comments directly in the code chunk below.

```{r, code_folding=F}
df_interjections_b <- df_bracket_terms_long %>% 
  #keep those bracket terms where Zwischenruf is NOT  included
  filter(!str_detect(bracket_terms_split, regex("Zwischenruf"))) %>% 
  select(session_no, respondent, speaker, bracket_terms_id, bracket_terms, bracket_terms_split) %>% 
  #keep only those rows which include a colon; interjections always include colons
  filter(str_detect(bracket_terms_split, regex("\\:"))) %>% 
  #extract everything before the colon; that's were the name is located
  mutate(source=str_extract(bracket_terms_split, regex("^.*?\\:"))) %>% 
  #remove comments which are additionally included between hyphens; see. eg. id 18
  mutate(source=str_remove(source, regex("\\p{Pd}.*?\\p{Pd}")))

#extract name  
df_interjections_b <- df_interjections_b %>% 
    fuzzyjoin::regex_left_join(.,
                             df_mps %>% 
                               select(source_name_family=speaker_name_family,
                                      source_name=speaker_name,
                                      source_party=speaker_party),
                             by=c(source="source_name_family")) %>% 
  filter(str_detect(source_name, regex("^Abgeo")))
```

Here the results for our previous example.

```{r, echo=F}
df_interjections_b %>% 
  filter(bracket_terms_id %in% c(63, 811, 812)) %>% 
  select(bracket_terms_id, bracket_terms_split, source_name, source_party)

```

Now let's *combine the results for type A and type B* into a consolidated dataframe.

```{r, code_folding=F}
df_interjections_all <- df_interjections_b %>% 
  bind_rows(., df_interjections_a) %>% 
  select(session_no, respondent, source_name, source_party) %>% 
  #only MPs
  filter(str_detect(source_name, regex("^Abge")))
```

With this dataframe available, let's see which MP made most interjections during the commission's hearings:

```{r}
df_interjections_freq <- df_interjections_all %>% 
  count(source_name, source_party, sort=T) %>% 
  # tidyr::replace_na(list("source_name", "unbekannt")) %>% 
  mutate(source_name=replace_na(source_name, "unbekannt")) %>% 
  #only interjections from MPs
  filter(str_detect(source_name, regex("^Abge|^unbek"))) %>% 
  mutate(source_name=str_remove(source_name, regex("^A[a-z]+\\s"))) %>% 
  left_join(.,
            vec_party_col %>% 
              enframe(.,
                      name="source_party",
                      value="party_col"))

df_interjections_freq %>% 
  reactable(.,
            theme=nytimes(cell_padding=0),
            columns=list(
              source_name=colDef(
                name="Parlamentarier",
                filterable = T,
                width=200),
              party_col=colDef(
                show = F
              ),
              source_party=colDef(
                show = F
              ),
              n=colDef(
                name="Anzahl der Zwischenrufe",
                width=400,
                align="left",
                cell=data_bars(.,
                  text_position="inside-end",
                  text_size=13,
                  fill_color_ref="party_col"
                )))) %>% 
  add_title(title=glue::glue("IBIZA-AUSSCHUSS: Ausschussmitglieder und Anzahl ihrer Zwischenrufe"),
            font_size=18) %>% 
  add_subtitle(subtitle = "Die Protokolle des Ibiza-Ausschusses enthalten Anmerkungen, die Zwischenrufe von Mitgliedern des Ausschusses aufzeigen. Diese Anmerkungen wurden extrahiert, dem jeweiligen Mitglied zugeordnet und zu Häufigkeiten pro Abgeordneten aggregiert.",
               font_size=12,
               font_weight="normal") %>% 
  add_source(source="Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes · Data: parlament.gv.at",
             font_size=11,
             font_weight = "normal",
             font_color = "#999999",
             font_family = "Helvetica",
             align="right") 
```

The result is at least to me somewhat surprising. I have to admit, I didn't follow the proceedings in detail, but David Stögmüller's propensity to interrupt speakers was new to me.

Another angle to look at this data would be check which hearing of a respondent evoked most interjections. Below the result. No real surprise, at least as far as the top 2 are concerned.

```{r}
df_interjections_all %>% 
  count(session_no, respondent, sort=T) %>% 
  reactable(.,
            theme=nytimes(cell_padding=0),
            columns = list(
              session_no=colDef(
                name="Sitzung",
                width=75,
                align="center"),
              respondent=colDef(
                name="Auskunftsperson",
                width=300,
                filterable = T),
              n=colDef(
                name="Anzahl der Zwischenrufe",
                width=200,
                align="left",
                cell=data_bars(.,
                  text_position="outside-end",
                  text_size=13,
                  fill_color="#d8382e"
                )))) %>% 
  add_title(title=glue::glue("IBIZA-AUSSCHUSS: Auskunftspersonen und Anzahl an Zwischenrufe"),
            font_size=18) %>% 
  add_subtitle(subtitle = "Die Protokolle des Ibiza-Ausschusses enthalten Anmerkungen, die Zwischenrufen von Mitgliedern des Ausschusses aufzeigen. Diese Anmerkungen wurden extrahiert und zu Häufigkeiten pro Befragung einer Auskunftsperson aggregiert. Hier wurden nur Zwischenrufe von Parlamentariern berücksichtigt.",
               font_size=12,
               font_weight = "normal") %>% 
  add_source(source="Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes · Data: parlament.gv.at",
             font_size=11,
             font_weight = "normal",
             font_color = "#999999",
             font_family = "Helvetica",
             align="right") 
```

### 'Amused' ('erheitert')

Another annotation included in the transcripts' brackets is *'(erheitert)'*, what probably translates best as 'amused'. Admittedly, being amused - at least in the Austrian context - can have quite a number of connotations, e.g. from finding something genuinely funny to 'making fun of' something in a smug way. Having said that, I am not entirely clear what the aggregated number can actually tell us, but, well, I thought the information is somewhat entertaining.

```{r}
df_amused_speaker <- df_bracket_terms %>% 
  #filter keeps only rows referring to speaker; not composite bracket terms referring to other participants
  filter(str_detect(bracket_terms, regex("\\(erheitert\\)"))) %>% 
  group_by(speaker) %>% 
  summarise(n_obs=n()) %>% 
  arrange(desc(n_obs)) %>%
  ungroup()

df_tb_amused_speaker <- df_data %>% 
  distinct(speaker) %>% 
  left_join(.,
            df_amused_speaker) %>% 
  mutate(n_obs=case_when(
    is.na(n_obs) ~ 0,
    TRUE ~ as.numeric(n_obs))) %>% 
  arrange(desc(n_obs))

df_tb_amused_speaker %>% 
  reactable(.,
            columns = list(
              speaker=colDef(
                name="Auskunftsperson",
                width=350,
                filterable=T),
              n_obs=colDef(
                name="Anzahl",
                align="left",
                cell=data_bars(.,
                               fill_color = "#4c0d3e",
                               text_size = 13,
                               text_position = "outside-end"))
              ),
            theme=nytimes(cell_padding=0)) %>% 
  add_title(title = "IBIZA-AUSSCHUSS: Welches Ausschussmitglied war am öftesten 'erheitert'?",
            font_size=18) %>%
  add_subtitle(subtitle = "Die Protokolle des Ibiza-Ausschusses enthalten ergänzende Anmerkungen zu Statements der Ausschussmitglieder. Eine Ergänzung zeigt an, ob der Sprecher/die Sprecherin 'erheitert' war. Diese Anmerkungen wurden extrahiert, dem jeweiligen Ausschussmitglied zugeordnet und zu Häufigkeiten aggregiert.",
               font_size=12,
               font_weight="normal") %>% 
  add_source(source="Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes · Data: parlament.gv.at",
             font_size=11,
             font_weight = "normal",
             font_color = "#999999",
             font_family = "Helvetica",
             align="right") 
```

The result is somewhat 'amusing', however, if one bears in mind that the top 4 are also among the most prolific speakers of the entire inquiry, it isn't too surprising that they come out first. Nevertheless, I am somewhat struck by the lead of the commission's chair Sobotka. But again, as stated above, the word 'amused' can have different shades.... Note that Sobotka appears twice: once in his function as chair, once in his function as a respondent.

```{r, include=F, eval=F}
df_bracket_terms %>% 
  filter(str_detect(bracket_terms, regex("\\(erheitert\\)"))) %>% 
  mutate(session_no_respondent=glue::glue("{respondent} ({session_no})")) %>% 
  group_by(session_no_respondent) %>% 
  summarise(n_obs=n()) %>% 
  arrange(desc(n_obs)) %>%
  ungroup()

```

### Consultations with person of trust ('Vertrauensperson')

The final annotation I looked into was the comment highlighting that a respondent consulted with his or her counsel or trusted person (*Vertrauensperson*). Respondents are provided with this possibility to discuss their answers, potential legal ramification of their reply etc. before actually responding. On top of the results shown below is Lisa Wieser, the personal assistant of Chancellor Kurz. In some way I can feel for her. As somebody who is hardly in the political spotlight and yet - I assume - rather familiar with the inner workings of ÖVP leadership you don't want to mess up your hearing and potentially cause any trouble for your boss. Nevertheless, reaching out to your counsel 203 times during one hearing.... that's quite something. But also more generally, the table features respondents related to the ÖVP among the top ranks. Also those who are from the ÖVP's front row and could be expected to be rather savy, e.g. Melchior, Köstinger, Bonelli made rather frequent use of their counsel.

```{r}
df_vertrauensperson <- df_bracket_terms %>% 
  mutate(vertrauensperson=case_when(str_detect(bracket_terms, "Vertrauensperson") & 
                                      str_detect(bracket_terms, "berät") &
                                      str_detect(bracket_terms, "Auskunftsperson") ~ "yes",
         TRUE ~ as.character("no"))) %>% 
  filter(vertrauensperson=="yes")

df_vertrauensperson <- df_vertrauensperson %>% 
  group_by(session_no, respondent) %>% 
  summarise(n_obs=n()) %>% 
  arrange(desc(n_obs))

df_tbl_vertrauensperson <- df_data %>% 
  distinct(session_no, respondent) %>% 
  left_join(., 
            df_vertrauensperson) %>% 
  mutate(n_obs=case_when(
    is.na(n_obs) ~ 0,
    TRUE ~ as.numeric(n_obs)
  )) %>% 
  arrange(desc(n_obs))

df_tbl_vertrauensperson %>% 
  reactable(.,
            columns = list(
              session_no=colDef(
                name="Sitzung",
                align="center",
                width=60),
              respondent=colDef(
                name="Auskunftsperson",
                width=300,
                filterable=T),
              n_obs=colDef(
                name="Anzahl",
                align="left",
                cell=data_bars(.,
                               fill_color = "firebrick",
                               text_size=13,
                               max_value = 300,
                               text_position = "outside-end"))
            ),
            theme=nytimes(cell_padding=0)) %>% 
  add_title(title = "IBIZA-AUSSCHUSS: Wie oft beriet sich die Auskunftsperson mit ihrer Vertrauensperson?",
            font_size = 18) %>%
  add_subtitle(subtitle = "Die Sitzungsprotokolle des Ibiza-Ausschusses enthalten ergänzende Klammerausdrücke wie '(Die Auskunftsperson berät sich mit ihrer Vertrauensperson'). Das der Tabelle zugrundeliegende Script identifiziert diese Ausdrücke und summiert sie auf. So hat sich zB. Lisa Wieser, die persönliche Assistentin von Kanzler Kurz, während ihrer Befragung mehr als 200 mal mit ihrer Vertrauensperson beraten.",
               font_size=12,
               font_weight="normal") %>% 
  add_source(source="Analysis: Roland Schmidt | @zoowalk | https://werk.statt.codes · Data: parlament.gv.at",
             font_size=11,
             font_weight = "normal",
             font_color = "#999999",
             font_family = "Helvetica",
             align="right")

```

# Wrap-up

This post - once again - got much longer (and time consuming) than initially intended. I hope the obtained dataset and the above analysis is helpful to some. The file can be downloaded here. As already stated above, if you spot any error etc. please feel free to contact me, best via direct message on [twitter](https://twitter.com/zoowalk){target="_blank"}.

### 2021-10-07 _post/2021-07-18-the-ibizia-inquiry-in-numbers/the-ibizia-inquiry-in-numbers_files/reactable-binding-0.2.3.9000/reactable.js

@@ -1,0 +1,2 @@
+!function(){var e={860:function(e,t,r){(t=r(645)(!1)).push([e.id,".ReactTable{position:relative;display:-webkit-box;display:-ms-flexbox;display:flex;-webkit-box-orient:vertical;-webkit-box-direction:normal;-ms-flex-direction:column;flex-direction:column}.ReactTable *{-webkit-box-sizing:border-box;box-sizing:border-box}.ReactTable .rt-table{-webkit-box-flex:1;-ms-flex:auto 1;flex:auto 1;flex-direction:column;-webkit-box-align:stretch;-ms-flex-align:stretch;align-items:stretch;width:100%;border-collapse:collapse;overflow:auto}.ReactTable .rt-table,.ReactTable .rt-thead{display:-webkit-box;display:-ms-flexbox;display:flex;-webkit-box-orient:vertical;-webkit-box-direction:normal;-ms-flex-direction:column}.ReactTable .rt-thead{flex-direction:column}.ReactTable .rt-thead .rt-td,.ReactTable .rt-thead .rt-th{line-height:normal;position:relative}.ReactTable .rt-th.rt-th-resizable{overflow:visible}.ReactTable .rt-th.rt-th-resizable:last-child{overflow:hidden}.ReactTable .rt-tbody{display:-webkit-box;display:-ms-flexbox;display:flex;-webkit-box-orient:vertical;-webkit-box-direction:normal;-ms-flex-direction:column;flex-direction:column;overflow:auto}.rt-td-expandable{cursor:pointer}.ReactTable .rt-tr-group{-ms-flex:1 0 auto;flex:1 0 auto;-webkit-box-orient:vertical;-webkit-box-direction:normal;-ms-flex-direction:column;flex-direction:column;-webkit-box-align:stretch;-ms-flex-align:stretch;align-items:stretch}.ReactTable .rt-tr,.ReactTable .rt-tr-group{-webkit-box-flex:1;display:-webkit-box;display:-ms-flexbox;display:flex}.ReactTable .rt-tr{-ms-flex:1 0 auto;flex:1 0 auto}.ReactTable .rt-td,.ReactTable .rt-th{-webkit-box-flex:1;-ms-flex:1 0 0px;flex:1 0 0;overflow:hidden}.ReactTable .rt-resizer{display:inline-block;position:absolute;width:36px;top:0;bottom:0;right:-18px;cursor:col-resize;z-index:10}.ReactTable .rt-tfoot{display:-webkit-box;display:-ms-flexbox;display:fl
[…]

### 2021-10-07 _post/2021-07-18-the-ibizia-inquiry-in-numbers/the-ibizia-inquiry-in-numbers.Rmd
update index

@@
 description: In mid-June, the so-called 'Ibiza Commission of Inquiry' heard its last witness after almost one year at work.
-The commission had been established by the Austrian Parliament and set up to investigate the prevalence of corruption during the coalition government of the ÖVP and FPÖ (Dec, 2017- May, 2019).
+The commission had been established by the Austrian Parliament and set up to investigate the prevalence of corruption during the coalition government of the ÖVP and FPÖ (Dec, 2017 - May, 2019).
 This post diggs into its published transcripts.

### 2021-10-07 _post/2021-07-18-the-ibizia-inquiry-in-numbers/the-ibizia-inquiry-in-numbers.Rmd
adding link for data download

@@
 readr::write_excel_csv2(df_data %>% 
-                          select(-contains("id")),
+                          select(-id_check) %>% 
+                          arrange(link, session_no, id_statement),
   file = here::here("_blog_data", "2021_7_ibiza_inquiry", file_name)
@@
 The dataframe obtained comprises the transcripts of `r length(unique(df_data$link))` sessions, which cover `r scales::number(nrow(df_data), big.mark=",")` single statements (rows of our dataframe).
-If someone is interested in the entire dataset, it can be downloaded here.
+If someone is interested in the entire dataset, it can be downloaded [here](https://data.world/zoowalk/ibizaausschuss){target="_blank"}.
 Note that the dataframe only covers statements, and not the accompanying text (introduction etc) of the transcript (if you use the dataset, grateful if you quote this post as its source).
@@
 I hope the obtained dataset and the above analysis is helpful to some.
-The file can be downloaded here.
+The file can be downloaded [here](https://data.world/zoowalk/ibizaausschuss){target="_blank"}.
 As already stated above, if you spot any error etc.

### 2021-11-23 _post/2021-07-18-the-ibizia-inquiry-in-numbers/the-ibizia-inquiry-in-numbers.Rmd
updating with latest changes

@@
   - web scraping
-preview: the-ibizia-inquiry-in-numbers_files/images/subpage_link_to_transcript.PNG
+preview: the-ibizia-inquiry-in-numbers_files/images/amused_graph.png
 output:
@@
 
-If you are not interested in any of the coding steps generating the data, jump directly to the [Analysis] section which still contains plenty of code, but you're not burdened with how to obtain the data in the first place.
+*If you are not interested in any of the coding steps generating the data, jump directly to the [Analysis] section* which still contains plenty of code, but you're not burdened with how to obtain the data in the first place.
 

### 2021-11-23 about.Rmd
main update
updating with latest changes

@@
 
+Note that I recently moved this blog and hence a few entries are somewhat 'messy' (broken links, images out of place). I'll hope to be able to iron this out soon.
+
 Roland Schmidt #IStandWithCEU #COYG
 
-Codes are collected at [www.github.com/werkstattcodes](https://github.com/werkstattcodes){target="_blank"}
-Tweets are at @zoowalk.
+Codes are collected at [www.github.com/werkstattcodes](https://github.com/werkstattcodes){target="_blank"}.
+Tweets are at [`@zoowalk`](https://twitter.com/zoowalk){target="_blank"}.
